### 2016-04-06 docprep/draft1.tex
Added a docprep folder for the documentation

\documentclass{article}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[toc,page]{appendix}
\usepackage{listings}
\usepackage{color}
\usepackage{listings}

\numberwithin{equation}{subsection}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Octave,                 % the language of the code
  otherkeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

\begin{document}

\title{Discrete Ordinates}
\author{Edward Norris}

\maketitle

\begin{abstract}
The abstract text goes here.
\end{abstract}

\tableofcontents

\section{CODE}
Consider that we have the following variables, $\Psi$ which is a function of $x$, $y$, $z$, $E$, and $\hat{\Omega}$. Further, we have the in-flux variables $\Psi_{x-in}$, $\Psi_{y-in}$, and $\Psi_{z-in]}$.

\begin{lstlisting}[escapeinside={(*}{*)}]
  (*$\Psi(x, y, z, E, \hat{\Omega}) \leftarrow 123$*)
  list= { (*$S_1,S_2,S_3$*) }
\end{lstlisting}

\begin{lstlisting}[escapeinside={(*}{*)}]
for (*$E \in \mathcal{E}$*) in descending order
  for (*$\Omega \in Q$*) in any order
    if (*$\mu(\Omega) > 0$*) and (*$\xi(\Omega) > 0$*) and (*$\eta(\Omega) > 0$*)
      for (*$x \in X$*) in ascending order
        for (*$y \in Y$*) in ascending order
          for (*$z \in Z$*) in ascending order
            (*$\Psi_{x-in}(x, y, z, E, \Omega) \leftarrow 123$*)
            (*$\Psi_{y-in}(x, y, z, E, \Omega) \leftarrow 123$*)
            (*$\Psi_{z-in}(x, y, z, E, \Omega) \leftarrow 123$*)
            (*$\Psi(x, y, z, E, \Omega) \leftarrow 123$*)

            (*$\Phi(x, y, z, E) \leftarrow \Phi(x, y, z, E) + \omega(\Omega) \Psi(x, y, z, E, \Omega)$*)
  list= { (*$S_1,S_2,S_3$*) }
\end{lstlisting}

\section{Introduction}
This document will teach you everything you need to know about the Boltzmann Equation from its derivation to the details of its implementation in computer code.

\section{Mathematical Development}

\subsection{The Boltzmann Equation}

The time dependent gamma transport equation can be written as

\begin{equation} \label{eq_boltz_t_dep}
\begin{split}
	&\left[ \frac{1}{v(E)} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
	\psi(\boldsymbol{r}, E, \hat{\Omega}, t) = \\
	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}, t) \psi(\boldsymbol{r}, E', \hat{\Omega}', t) dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega}, t)
\end{split}
\end{equation}

However, we are typically not concerned with the transient case in medical diagnostic imaging. We are more interested in the steady state case. The time independent form of Eq. \ref{eq_boltz_t_dep} is written as

\begin{equation} \label{eq_boltz}
\begin{split}
	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
	\psi(\boldsymbol{r}, E, \hat{\Omega}) = \\
	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega})
\end{split}
\end{equation}

\subsection{Harmonic Approximation}
Equation \ref{eq_boltz} has a number of terms that cannot be solved directly. Instead, some numerical approximation must be used. The macroscopic scattering cross section, $\Sigma_s$ is typically expanded with a Legendre polynomial (for more information of Legendre polynomials, refer to Section \ref{sec_legendre}). The expansion is as follows:

\begin{equation} \label{eq_sigma_expansion}
\Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \approx \sum_{l=0}^L \frac{2l+1}{4 \pi} \Sigma_{s,l}(\boldsymbol{r}, E' \rightarrow E) P_l(\hat{\Omega}' \rightarrow \hat{\Omega})
\end{equation}

where $\Sigma_{s,l}$ is the expansion coefficients termed the "scattering moments." The Legendre polynomials $P_l(\hat{\Omega}' \rightarrow \hat{\Omega}$ are defined as

\begin{equation} \label{eq_harmonic}
P_l(\hat{\Omega}' \rightarrow \hat{\Omega}) = \frac{1}{2l+1}\sum_{m=-l}^l Y_{l,m}^*(\hat{\Omega}')Y_{l,m}(\hat{\Omega})
\end{equation}

The angular fluence ($\phi$) is also expanded as

\begin{equation}
\phi(\boldsymbol{r}, E', \hat{\Omega}') \approx \sum_{l=0}^L \sum_{m=-l}^l
\phi_{lm}(\boldsymbol{r}, E')Y_{lm}(\hat{\Omega}')
\end{equation}

\noindent
A source term suitable for numeric integration is then

\begin{equation}\label{eq_harmonic_simp1}
\begin{split}
\int_0^\infty dE' \int_{4 \pi} d\hat{\Omega}' \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \phi(\boldsymbol{r}, E', \hat{\Omega}')
\approx \\
 \sum_{l=0}^L \frac{2l+1}{4 \pi} \sum_{m=-l}^l \Sigma_{s,l}^{gg'}\phi_{i,j,k,lm}^{g'}Y_{l,m}(\hat{\Omega}_n)
\end{split}
\end{equation}

Finally, substituting Eqs. \ref{eq_harmonic_simp1} and <> into <>, we arrive at

\begin{equation}\label{eq_discretized}
\begin{split}
\hat{\Omega}_n \cdot \nabla\psi_{i,j,k,n}^g + \sigma_{i,j,k}^g \psi_{i,j,k,n}^g = \\
\sum_{g' = 0}^G \sum_{l=0}^L \frac{2l+1}{4 \pi} \sum_{m=-l}^l \sigma_{s,l}^{g,g'}\psi_{i,j,k,l,m}^{g'}Y_{l,m}(\hat{\Omega}_n) + \frac{1}{4 \pi}q_{i,j,k}^g
\end{split}
\end{equation}

We can calculate the scalar flux as

\begin{equation}
\phi_{i,j,k}^g = \sum_{n=1}^{|\Omega|} w_n \psi_{i,j,k,n}^g
\end{equation}

\subsection{Discretization}
The continuous Boltzmann Equation Approximation has to be discretized in all dimensions to run on a computer.

The gradient of $\psi$ is calculated as

\begin{equation}
\nabla \psi_{i,j,k,n}^g \approx
\left\langle  
\frac{\psi_{i+1,j,k}^g - \psi_{i-1,j,k}^g}{\Delta x},
\frac{\psi_{i,j+1,k}^g - \psi_{i,j-1,k}^g}{\Delta y},
\frac{\psi_{i,j,k+1}^g - \psi_{i,j,k-1}^g}{\Delta z}
\right\rangle
\end{equation}

There are six faces on each parallelpiped with normals $<\pm 1, 0, 0>$, $<0, \pm 1, 0>$, and $<0, 0, \pm 1>$. Taking the dot product of these normals and the evaluated gradient gives

\begin{equation}
\hat{\Omega} \cdot \nabla \psi_{i,j,k}^g \approx
2 \left(
\frac{\psi_{i+1,j,k}^g - \psi_{i-1,j,k}^g}{\Delta x} +
\frac{\psi_{i,j+1,k}^g - \psi_{i,j-1,k}^g}{\Delta y} +
\frac{\psi_{i,j,k+1}^g - \psi_{i,j,k-1}^g}{\Delta z}
\right)
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=3.0in]{myfigure}
    \caption{Simulation Results}
    \label{simulationfigure}
\end{figure}

\section{Implementation}
At this point, we have successfully developed an algorithm that can be implemented on a computer. However, there are many challenges in actually implementing this algorithm.

\subsection{Design}
Yeah, it was designed.

\subsection{Language Selection}
Yeah, it was selected.

\subsection{Salome}
We chose to use the Salome framework for numerical pre/post-processing. Salome includes a built in geometry system and meshing capabilities.

\subsection{Framework Integration}
Yeah, it was integrated

\subsection{C++ Implementation}
Yeah, it was implemented.

\begin{lstlisting}
#include <stdio.h>
#define N 10
/* Block
 * comment */

int main()
{
    int i;

    // Line comment.
    puts("Hello world!");

    for (i = 0; i < N; i++)
    {
        puts("LaTeX is also great for programmers!");
    }

    return 0;
}
\end{lstlisting}

\section{MPI Implementation}
MPI stands for Message Passing Interface and is the \textit{de Facto} standard for implementing parallel algorithms across multiple physical machines.

\section{GPU Implementation}
GPU stands for Graphical Processing Unit and is a collection of SIMD (single instruction multiple data) processors. Each processor operates on its own piece of data but performs the same operation each other processor in its warp is executing.

\section{Conclusion}
Hooray! It works!

\section{Appendix}

\subsection{Laplace's Equation} \label{sec_laplace}
Many physical phenomena can be described by Poisson's Equation given below.

\begin{equation} \label{eq_poisson}
\nabla^2 f = \psi
\end{equation}

Laplace's Equation is a special form of Poisson's Equation and is given by Eq. \ref{eq_laplace}.

\begin{equation} \label{eq_laplace}
\nabla^2 f = 0
\end{equation}

Laplace's equation in cartesian coordinates expands to

\begin{equation} \label{eq_laplace_cartesian}
\nabla^2 f(x, y, z) = \frac{\partial^2 f}{\partial x^2} + \frac{\partial^2 f}{\partial y^2} + \frac{\partial^2 f}{\partial z^2} = 0
\end{equation}

In cylindrical coordinates
\begin{equation} \label{eq_laplace_cylindrical}
\nabla^2 f(r, \phi, z) = \frac{1}{r} \frac{\partial}{\partial r}\left( r \frac{\partial f}{\partial r}\right) + \frac{1}{r^2}\frac{\partial^2 f}{\partial \phi^2} + \frac{\partial^2 f}{\partial z^2}
\end{equation}

In spherical coordinates
\begin{equation} \label{eq_laplace_spherical}
\nabla^2 f(\rho, \theta, \phi) = \frac{1}{\rho^2}\frac{\partial}{\partial \rho} \left( \rho^2 \frac{\partial f}{\partial \rho}\right) + \frac{1}{\rho^2 \sin \theta}\frac{\partial}{\partial \theta}\left( \sin \theta \frac{\partial f}{\partial \theta}\right) + \frac{1}{\rho^2 \sin^2 \theta} \frac{\partial^2 f}{\partial \phi^2}
\end{equation}

Any solution to Laplace's equation is known as a harmonic function. Further, if any two functions are a solution to Laplace's equation, then by the superposition principle, the summation of the two functions is also a solution to Laplace's equation.

Common solutions to Laplace's equation include

\begin{table}
\begin{center}
\label{tbl_laplace_solutions}
\caption{A list of solutions to Laplace's equation in different geometrical systems}
\begin{tabular}{|l|c|}
\hline
Cartesian & Exponential, Circular (a class of trig function), hyperbolic \\ \hline
Cylindrical & Bessel, Exponential, Circular \\ \hline
Spherical & Legendre polynomial, Power, Circular \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{Legendre Polynomials} \label{sec_legendre}

Legendre polynomials are solutions to
\begin{equation}
P_n(x) = \frac{1}{2 \pi i} \oint (1-2\xi x + \xi^2)^{-1/2} \xi^{-n-1} d\xi
\end{equation}

\begin{equation}
P_n(x) = \frac{1}{2^n n!}\frac{d^n}{dx^n}(x^2-1)^n
\end{equation}

The first few such solutions are \\
$P_0(x) = 1$ \\
$P_1(x) = x$ \\
$P_2(x) = \frac{1}{2}(3x^2-1)$ \\
$P_3(x) = \frac{1}{2}(5x^3 - 3x)$ \\
$P_4(x) = \frac{1}{8}(35x^4 - 30x^2 + 3)$ \\
$P_5(x) = \frac{1}{8}(63x^5 - 70x^3 + 15x)$ \\
$P_6(x) = \frac{1}{16}(231x^6 - 315x^4 + 105x^2 - 5)$ \\

Associated Legendre polynomials are a generalization of the Legendre polynomial and solve

\begin{equation}
P_l^m(x) = (-1)^m(1-x^2)^{m/2}\frac{d^m P_l(x)}{dx^m}
\end{equation}

Both Legendre polynomials and Associated Legendre Polynomials can be calculated using the Boost C++ library.

\subsection{Spherical Harmonic Function}

First, we assume that Laplace's equation is of the seperable form $f(\rho, \theta, \phi) = R(\rho)Y(\theta, \phi)$. Equation \ref{eq_laplace_spherical} can then be rewritten as

\begin{equation} \label{eq_spherical1}
\begin{split}
\nabla^2 f(\rho, \theta, \phi) = 
\frac{1}{\rho^2}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial (R(\rho)Y(\theta, \phi))}{\partial \rho}\right) 
+ \\
\frac{1}{\rho^2 \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial (R(\rho)Y(\theta, \phi))}{\partial \theta}\right) 
+ 
\frac{1}{\rho^2 \sin^2 \theta} 
\frac{\partial^2 (R(\rho)Y(\theta, \phi))}{\partial \phi^2} = 0
\end{split}
\end{equation}

\noindent
which simplifies to
\begin{equation} \label{eq_spherical2}
\begin{split}
\frac{Y(\theta, \phi)}{\rho^2}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
+ 
\frac{R(\rho)}{\rho^2 \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
+ \\
\frac{R(\rho)}{\rho^2 \sin^2 \theta} 
\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2}
\end{split} = 0
\end{equation}

\noindent
Multiplying Eq. \ref{eq_spherical2} by $\rho^2 / (R(\rho)Y(\theta, \phi))$ yields

\begin{equation}\label{eq_spherical3}
\begin{split}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
+ 
\frac{1}{Y(\theta, \phi) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
+ \\
\frac{1}{Y(\theta, \phi) \sin^2 \theta} 
\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2} = 0
\end{split}
\end{equation}

Equation \ref{eq_spherical3} can be split into two parts, a function of $\rho$ alone and a component that is a function of $\theta$ and $\phi$ alone. The component that is a function of $\theta$ and $\phi$ can be moved to the other side of the equation. Since two functions of different variables are equal, they must each be equal to some constant, $\lambda$. Therefore, Eq. \ref{eq_spherical3} can be rewritten as two equations:

\begin{equation}\label{eq_spherical_r}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) = \lambda
\end{equation}

\begin{equation}\label{eq_spherical_tp}
\frac{1}{Y(\theta, \phi) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
+ \\
\frac{1}{Y(\theta, \phi) \sin^2 \theta} 
\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2} = -\lambda
\end{equation}

$Y(\theta, \phi)$ can be further seperated into $Y(\theta, \phi) = \Theta(\theta)\Phi(\phi)$ which produces

\begin{equation}\label{eq_spherical_y}
\begin{split}
\frac{1}{\Theta(\theta)\Phi(\phi) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial (\Theta(\theta)\Phi(\phi))}{\partial \theta}\right) 
+ \\
\frac{1}{\Theta(\theta)\Phi(\phi) \sin^2 \theta} 
\frac{\partial^2 (\Theta(\theta)\Phi(\phi))}{\partial \phi^2} = -\lambda
\end{split}
\end{equation}

\noindent
which simplifies to

\begin{equation}\label{eq_spherical_y2}
\begin{split}
\frac{1}{\Theta(\theta) \sin \theta}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ 
\frac{1}{\Phi(\phi) \sin^2 \theta} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2} = -\lambda
\end{split}
\end{equation}

Multiplying Eq. \ref{eq_spherical_y2} by $\sin^2(\theta)$ and rearranging yields

\begin{equation}\label{eq_spherical_y3}
\begin{split}
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ \lambda \sin \theta
 = \frac{-1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
\end{split}
\end{equation}

As before, each side of Eq. \ref{eq_spherical_y3} is a function of a single variable, therefore, both sides are equal to some constant. In this case, we select $m^2$ to be  the constant variable. The produces

\begin{equation}\label{eq_spherical_t}
\begin{split}
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ \lambda \sin \theta
 = m^2
\end{split}
\end{equation}

\begin{equation}\label{eq_spherical_p}
\begin{split}
\frac{1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2} = -m^2
\end{split}
\end{equation}

Combining Eq. \ref{eq_spherical_r}, \ref{eq_spherical_t}, and \ref{eq_spherical_p} Laplace's equation in sperical coordinates can be expressed in a fully seperated form as

\begin{equation}\label{eq_spherical_sep}
\begin{split}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
&= \lambda \\
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ \lambda \sin \theta
&= m^2 \\
\frac{1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
&= -m^2
\end{split}
\end{equation}

It can be shown through a detailed anlysis that is outside the scope of this paper that $m$ and $\lambda$ must both be integers. $\lambda$ is further constrained in that the relation $\lambda = l(l+1)$ where $l \leq |m|$ and $l \in \mathbb{Z}$. Therefore, Eq. \ref{eq_spherical_sep} can be written as

\begin{equation}\label{eq_spherical_sep2}
\begin{split}
\frac{1}{R(\rho)}
\frac{\partial}{\partial \rho} 
\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
&= l(l+1) \\
\frac{\sin \theta}{\Theta(\theta)}
\frac{\partial}{\partial \theta}
\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
+ l(l+1) \sin \theta
&= m^2 \\
\frac{1}{\Phi(\phi)} 
\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
&= -m^2 \\
& l, m \in \mathbb{Z} \\
& l \leq |m|
\end{split}
\end{equation}

% TODO
TODO - Discontinuous

It can be shown that the solution is stated

\begin{equation}
Y_l^m(\theta, \phi) = \sqrt{\frac{2l+1}{4 \pi}
\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) e^{im\phi}
\end{equation}

whose real part can be expanded to

\begin{equation}
Y_l^0 = \sqrt{\frac{2l+1}{4 \pi}}P_l^0(1)
\end{equation}

\begin{equation}
Y_l^{m, even} = (-1)^m\sqrt{\frac{2l+1}{2 \pi} \frac{(l-m)!}{(l+m)!}}P_l^m(\cos(m \phi))
\end{equation}

\begin{equation}
Y_l^{m,odd} = (-1)^m \sqrt{\frac{2l+1}{2 \pi} \frac{(l-m)!}{(l+m)!}}P_l^m(\sin(m \phi))
\end{equation}

\subsection{Klein-Nishina Formula}

The Klein-Nishina formula gives the differential scattering cross section of photon incident on a single free electron.

\begin{equation}\label{klein_nishina}
\frac{\partial \sigma}{\partial \Omega} = \frac{1}{2}\alpha^2 r_c^2 P(E_\gamma, \theta)^2
\left( P(E_\gamma, \theta) + P(E_\gamma, \theta)^{-1} - 1 + \cos^2(\theta) \right)
\end{equation}

\begin{equation}\label{klein_nishina2}
P(E_\gamma, \theta) = \frac{1}{1 + (\frac{E_\gamma}{m_e c^2})(1-\cos(\theta))}
\end{equation}

\noindent
where $\alpha$ is the fine structure constant ($\alpha \approx 7.2971E-3$), $r_c$ is the reduced Compton wavelength ($r_c = \hbar$)

\subsection{Numeric Integration}

\begin{equation}
\tilde{H}
\end{equation}

\end{document}

### 2016-04-08 docprep/draft1.tex
Updated the code in docprep

@@
 
+Given:
+\begin{itemize}
+\item A list of all energies, $\mathcal{E}$
+\item A list of all directions in the quadrature, $Q$
+\item A mapping from $r$ to zone number, $\mathcal{Z}(r)$
+\item Lists of direction cosines, $\mu$, $\xi$, and $\eta$
+\item A map of $r$ to voxel volume, $V(r)$
+\item A list of all $x$, $y$, and $z$ voxel centers, $X$, $Y$, and $Z$
+\item A list of weights for the quadrature, $\omega$
+\item A list of scatter cross sections in zone $\mathcal{Z}(r)$ from energy $E'$ to $E$, $\Sigma_s(\mathcal{Z}(r), E', E)$
+\item A list of total cross sections in zone $\mathcal{Z}(r)$ at energy $E$, $\Sigma_T(\mathcal{Z}(r), E)$
+\item The surface area of all voxels, $A_{xy}(r)$, $A_{yz}(r)$, $A_{xz}(r)$
+\end{itemize}
+
+\pagebreak
 \begin{lstlisting}[escapeinside={(*}{*)}]
-  (*$\Psi(x, y, z, E, \hat{\Omega}) \leftarrow 123$*)
-  list= { (*$S_1,S_2,S_3$*) }
+function sweep()
+
+  (*$\phi(r, E) = \vec{0}$*)    # Scalar flux
+  (*$\Psi(r, \Omega, E) = \vec{0}$*)  # Angular flux
+
+  for (*$E \in \mathcal{E}$*) in descending order
+    while not converged
+      for (*$\Omega \in Q$*) in any order
+        if (*$\mu(\Omega) > 0$*) and (*$\xi(\Omega) > 0$*) and (*$\eta(\Omega) > 0$*)
+          octant1()
+        else if (*$\mu(\Omega) > 0$*) and (*$\xi(\Omega) > 0$*) and (*$\eta(\Omega) < 0$*)
+          octant2()
+        ...
+        else
+          octant8()
 \end{lstlisting}
 
+\pagebreak
 \begin{lstlisting}[escapeinside={(*}{*)}]
-for (*$E \in \mathcal{E}$*) in descending order
-  for (*$\Omega \in Q$*) in any order
-    if (*$\mu(\Omega) > 0$*) and (*$\xi(\Omega) > 0$*) and (*$\eta(\Omega) > 0$*)
+function octant1()
+
   for (*$x \in X$*) in ascending order
@@
       for (*$z \in Z$*) in ascending order
-            (*$\Psi_{x-in}(x, y, z, E, \Omega) \leftarrow 123$*)
-            (*$\Psi_{y-in}(x, y, z, E, \Omega) \leftarrow 123$*)
-            (*$\Psi_{z-in}(x, y, z, E, \Omega) \leftarrow 123$*)
-            (*$\Psi(x, y, z, E, \Omega) \leftarrow 123$*)
 
-            (*$\Phi(x, y, z, E) \leftarrow \Phi(x, y, z, E) + \omega(\Omega) \Psi(x, y, z, E, \Omega)$*)
-  list= { (*$S_1,S_2,S_3$*) }
+        (*$S$*) = totalScatter((*$x, y, z, \Omega, E$*))
+
+        # Get the in-flux from the previous out-flux
+        (*$\Psi_{x-in}(x, y, z, E, \Omega) \leftarrow \Psi_{x-out}(x-\Delta x, y, z, E, \Omega)$*)
+        (*$\Psi_{y-in}(x, y, z, E, \Omega) \leftarrow \Psi_{x-out}(x, y-\Delta y, z, E, \Omega)$*)
+        (*$\Psi_{z-in}(x, y, z, E, \Omega) \leftarrow \Psi_{x-out}(x, y, z-\Delta z, E, \Omega)$*)
+
+        # Calculate the angular flux
+        (*$n \leftarrow S + 2\mu(\Omega)A_{yz}(y,z)\Psi_{x-in}(x,y,z,E,\Omega) + $*)
+               (*$2\xi(\Omega)A_{xz}(x,z)\Psi_{y-in}(x,y,z,E,\Omega) + $*)
+               (*$2\eta(\Omega)A_{xy}(x,y)\Psi_{z-in}(x,y,z,E,\Omega)$*)
+        (*$d \leftarrow 2\mu(\Omega)A_{yz}(y,z) + 2\xi(\Omega)A_{xz}(x,z) + 2\eta(\Omega)A_{xy}(x,y) + \Sigma_T(\mathcal{Z}(r), E)$*)
+        (*$\Psi(x, y, z, E, \Omega) \leftarrow n/d $*)
+
+        # Calculate the out-flux
+        (*$\Psi_{x-out}(x, y, z, E, \Omega) \leftarrow 2 \Psi(x, y, z, E, \Omega) - \Psi_{x-in}(x, y, z, E, \Omega)$*)
+        (*$\Psi_{y-out}(x, y, z, E, \Omega) \leftarrow 2 \Psi(x, y, z, E, \Omega) - \Psi_{y-in}(x, y, z, E, \Omega)$*)
+        (*$\Psi_{z-out}(x, y, z, E, \Omega) \leftarrow 2 \Psi(x, y, z, E, \Omega) - \Psi_{z-in}(x, y, z, E, \Omega)$*)
+
+        # Increment the scalar flux
+        (*$\phi(x, y, z, E) \leftarrow \phi(x, y, z, E) + \omega(\Omega) \Psi(x, y, z, E, \Omega)$*)
+\end{lstlisting}
+
+\pagebreak
+\begin{lstlisting}[escapeinside={(*}{*)}]
+function totalScatter((*$r, \Omega, E$*))
+
+  # Assume a point source at (*\color{mygreen} $r_0$*) with energy (*\color{mygreen} $E_0$*) and strength (*\color{mygreen} $S_0$*) particles / sec
+  (*$S_x \leftarrow 0$*)
+  if (*$r = r_0$*) and (*$E = E_0$*)
+    (*$S_x \leftarrow (1/4\pi) \times S_0$*)
+
+  (*$S_D \leftarrow 0$*)
+  for (*$E' \in \mathcal{E} | E' \geq E$*)
+    (*$S_D = S_D + (1/4\pi) \times \phi(r, \Omega, E') \times \Sigma_s(\mathcal{Z}(r), E', E) \times V(r)$*)
+
+  (*$S_T = S_x + S_D$*)
+  return (*$S_T$*)
 \end{lstlisting}

### 2016-04-08 doctors/config.cpp
Updated the config to point to S4 quad again

@@ -54,9 +54,9 @@ void Config::loadDefaults()
     flatFilterMat = {"cu"};
 
     // Directional quadrature set
-    quadType = "special";  // Type of quadrature
+    quadType = "sn";  // Type of quadrature
     quadSpecial = 2;  // Used for special quad types
-    sn = 6;  // N in Sn quadrature order
+    sn = 4;  // N in Sn quadrature order
     m = sn * (sn + 2);  // Total number of directions in all 8 octants
 
     // Cross section data set

### 2016-04-19 doctors/config.cpp
Removed the libconfig dependency Possibly changed line endings to Windows style

@@ -1,12 +1,15 @@
+#define _USE_MATH_DEFINES
+#include <cmath>
+
 #include "config.h"
 
 #include <QDebug>
 
-#include <cmath>
+
 
 // This project relies on libconfig++ being installed. To install it, download libconfig-X.X.tar.gz from http://www.hyperrealm.com/libconfig/  unzip.
 //   Follow the directions in the INSTALL file (./configure, make, make install). Then run "sudo ldconfig" to update the LD path variables so it can be found.
-#include <libconfig.h++>
+//#include <libconfig.h++>
 
 Config::Config() //: m_flatFilterThickness(NULL), m_flatFilterMat(NULL), m_xsection(NULL)
 {
@@ -69,29 +72,31 @@ void Config::loadDefaults()
     ms = 0;   // Cross section mixing table length (ms = 0 no mixing)
     mtm = 6;  // Total number of materials including al Pns (each Pn is considered a nuclide)
     isct = 1; // Maximum order of Legendre (ie scattering) expansion of the cross section (lowest is 0)
-    xsection = {0.0, 0.0,  0.0002, 0.0002,     // Air x-section at 60 keV nearly void P0
-                  0.0, 0.0,  0.0000, 0.0002,     // Air x-section P1 expansion
-                  0.0, 0.0,  0.2059, 0.1770,     // Water x-section at 60 keV  P0
-                  0.0, 0.0,  0.0000, 0.1770,     // Water x-section P1 expansion
-                  0.0, 0.0, 71.4753, 0.0000,     // Tunsgten x-section at 60 keV P0
-                  0.0, 0.0, 71.4753, 0.0000};    // Tunsgten x-section P1 expansion
+    xsection = {0.0f, 0.0f,    0.0002f, 0.0002f,     // Air x-section at 60 keV nearly void P0
+                  0.0f, 0.0f,  0.0000f, 0.0002f,     // Air x-section P1 expansion
+                  0.0f, 0.0f,  0.2059f, 0.1770f,     // Water x-section at 60 keV  P0
+                  0.0f, 0.0f,  0.0000f, 0.1770f,     // Water x-section P1 expansion
+                  0.0f, 0.0f, 71.4753f, 0.0000f,     // Tunsgten x-section at 60 keV P0
+                  0.0f, 0.0f, 71.4753f, 0.0000f};    // Tunsgten x-section P1 expansion
 
     // Total cross section
-    xsTot = {0.0002,      // Air
-             0.2059,      // Water
-             71.4753};    // Tungster
[…]

### 2016-04-19 doctors/quadrature.cpp

@@ -2,6 +2,7 @@
 
 #include <QDebug>
 #include <string>
+#include <qmath.h>
 
 // Static allocators
 const Quadrature Quadrature::ms_sn2(2);

### 2016-04-19 doctors/solvers.cpp

@@ -1,7 +1,11 @@
 //#include "solvers.h"
+#define _USE_MATH_DEFINES
+#include <cmath>
+
 #include "mainwindow.h"
 
 #include <QDebug>
+
 #include <ctime>
 
 #include "quadrature.h"

### 2016-09-12 docprep/draft1.tex
No change
Fixed the margins on the report

@@
 \usepackage{listings}
+\usepackage{fullpage}
 
@@
 
-\begin{equation}
-Y_l^m(\theta, \phi) = \sqrt{\frac{2l+1}{4 \pi}
-\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) e^{im\phi}
+\begin{equation}\label{eq:spherical_harmonic}
+Y_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
+\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) e^{im \varphi}
+\end{equation}
+
+Using Euler's Equation (Eq. \ref{eq:euler}) to compute the imaginary exponent, the spherical harmonic can be expanded to Eq. \ref{eq:spherical_harmonic_exp}.
+
+\begin{equation}\label{eq:euler}
+e^{i \theta} = \cos(\theta) + i \sin(\theta)
+\end{equation}
+
+\begin{equation}\label{eq:spherical_harmonic_exp}
+Y_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
+\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) \left[ \cos(m \varphi) + i \sin(m \varphi) \right]
+\end{equation}
+
+\begin{equation}\label{eq:spherical_harmonic_xs}
+\Sigma_s(\Omega \cdot \Omega') \approx \sum_{l=0}^L \sigma_l \sum_{m=-l}^l Y_l^m(\Omega)\bar{Y}_l^m(\Omega')
+\end{equation}
+
+where the over-bar denotes the complex conjugate defined by Eq.~\ref{eq:spherical_harmonic_conj}.
+
+\begin{equation}\label{eq:spherical_harmonic_conj}
+\bar{Y}_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
+\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) \left[ \cos(m \varphi) - i \sin(m \varphi) \right]
+\end{equation}
+
+Expanding the inner summation of \ref{eq:spherical_harmonic_xs} with Eq.~\ref{eq:spherical_harmonic_exp} and \ref{eq:spherical_harmonic_conj} yields:
+
+\begin{equation}\label{eq:spherical_harmonic_xs_inner}
+\begin{split}
+Y_l^m(\theta, \varphi)\bar{Y}_l^m(\theta', \varphi') & = 
+(-1)^{2l}\frac{2l+1}{4 \pi}
+\frac{(l-m)!}{(l+m)!} P_l^m(\cos \theta) P_l^m(\cos \theta') \times \\
+&\left[ \cos(m \varphi) + i \sin(m \varphi) \right] \left[ \cos(m \varphi') - i \sin(m \varphi') \right] \\
+& = \frac{2l+1}{4 \pi}
+\frac{(l-m)!}{(l+m)!} P_l^m(\cos \theta) P_l^m(\cos \theta') \times \\
+& \left[ \cos(m \varphi) \cos(m \varphi') + i \sin(m \varphi)\cos(m \varphi') - i \sin(m \varphi') \cos(m \varphi) - i^2\sin(m \varphi) \sin(m \varphi') \right]
+\end{split}
 \end{equation}

### 2016-09-12 python/xsdataPlotter/main/legendrePlotter.py
Added some python xs plotting routines

@@ -90,36 +90,14 @@ def factorial(x):
         return x
     return x * factorial(x-1)
 
-
-def clm(l, m):
-    return numpy.sqrt((2 - diracdelta(m, 0)) * factorial(l-abs(m)) / factorial(l+abs(m)))
-
+def pl(mu, l):
+    legendre_function = scipy.special.legendre(l)
+    return legendre_function(mu)
 
 def plm(mu, l, m):
     #lpmv(m, v, x) 	Associated legendre function of integer order.
     #sph_harm(m, n, theta, phi) 	Compute spherical harmonics.
     return scipy.special.lpmv(abs(m), l, mu)
-
-
-def spherical(theta, phi, sigmas):
-
-    s = 0
-    for l in range(len(sigmas)):
-        coeff = (2*l + 1) / (4 * numpy.pi)
-        ss = 0
-        for m in range(-l, l+1):
-            ss += ylm(theta, phi, l, m)
-        s += coeff * sigmas[l] * ss
-    return s
-
-
-def ylm(theta, phi, l, m):
-    mu = numpy.cos(theta)
-    mphi = abs(m) * phi
-    if m >= 0:
-        return clm(l, m) * plm(mu, l, m) * numpy.cos(mphi)
-    else:
-        return clm(l, m) * plm(mu, l, m) * numpy.sin(mphi)
 
 e, eprime, nl, vals = parse_matmsh3("/media/Storage/thesis/python/xsdataPlotter/be9scatter504.dat")
 
@@ -129,29 +107,15 @@ e, eprime, nl, vals = parse_matmsh3("/media/Storage/thesis/python/xsdataPlotter/
 sh = vals.shape
 nls = sh[2]
 
-minv = vals.min()
-maxv = vals.max()
+#print("plm = " + str(plm(.6, 2, 1)))  # Expect -1.44
 
 pyplot.close("all")
 
 indexes = numpy.linspace(1, 19, 19)
 
-for i in range(nls):
-    pyplot.figure()
-    pyplot.contourf(e/1E6, eprime/1E6, vals[:, :, i], 64, cmap='viridis', vmin=minv, vmax=maxv)
-    caxis = pyplot.colorbar()
-    pyplot.contour(e/1E6, eprime/1E6, vals[:, :, i], [0], colors='k')
-    #ax.set_yscale('log')
-    #ax.set_xscale('log')
[…]

### 2016-09-12 python/xsdataPlotter/main/spharm2Plotter.py

@@ -76,13 +76,6 @@ def parse_matmsh3(filename):
 
     return numpy.array(dim1), numpy.array(dim2), numpy.array(dim3), ndata
 
-
-def diracdelta(x, y):
-    if x == y:
-        return 1.0
-    return 0.0
-
-
 def factorial(x):
     if x == 0:
         return 1.0
@@ -91,40 +84,42 @@ def factorial(x):
     return x * factorial(x-1)
 
 
-def clm(l, m):
-    return numpy.sqrt((2 - diracdelta(m, 0)) * factorial(l-abs(m)) / factorial(l+abs(m)))
-
-
 def plm(mu, l, m):
-    #lpmv(m, v, x) 	Associated legendre function of integer order.
-    #sph_harm(m, n, theta, phi) 	Compute spherical harmonics.
-    return scipy.special.lpmv(abs(m), l, mu)
+    return scipy.special.lpmv(m, l, mu)
 
 
-def spherical(theta, phi, sigmas):
-
-    s = 0
-    for l in range(len(sigmas)):
-        coeff = (2*l + 1) / (4 * numpy.pi)
-        ss = 0
-        for m in range(-l, l+1):
-            ss += ylm(theta, phi, l, m)
-        s += coeff * sigmas[l] * ss
-    return s
+#def ylm(theta, phi, l, m):
+#    mu = numpy.cos(theta)
+#    mphi = abs(m) * phi
+#    real = numpy.sqrt((2*l + 1)/(4*numpy.pi) * factorial(l-m)/factorial(l+m)) * plm(mu, l, m) * numpy.cos(mphi)
+#    img = numpy.sqrt((2*l + 1)/(4*numpy.pi) * factorial(l-m)/factorial(l+m)) * plm(mu, l, m) * numpy.sin(mphi)
+#    return real, img
 
 
-def ylm(theta, phi, l, m):
+#def ylm_conj(theta, phi, l, m):
+#    rl, im = ylm(theta, phi, l, m)
+#    return rl, -im
+
+
+def ylm_e(theta, phi, l, m):
     mu = numpy.cos(theta)
-    mphi = abs(m) * phi
-    if m >= 0:
-        return clm(l, m) * plm(mu, l, m) * numpy.cos(mphi)
-    else:
[…]

### 2016-09-12 python/xsdataPlotter/main/spharmPlotter.py

@@ -96,35 +96,25 @@ def clm(l, m):
 
 
 def plm(mu, l, m):
-    #lpmv(m, v, x) 	Associated legendre function of integer order.
-    #sph_harm(m, n, theta, phi) 	Compute spherical harmonics.
     return scipy.special.lpmv(abs(m), l, mu)
-
-
-def spherical(theta, phi, sigmas):
-
-    s = 0
-    for l in range(len(sigmas)):
-        coeff = (2*l + 1) / (4 * numpy.pi)
-        ss = 0
-        for m in range(-l, l+1):
-            ss += ylm(theta, phi, l, m)
-        s += coeff * sigmas[l] * ss
-    return s
 
 
 def ylm(theta, phi, l, m):
     mu = numpy.cos(theta)
     mphi = abs(m) * phi
-    if m >= 0:
-        return clm(l, m) * plm(mu, l, m) * numpy.cos(mphi)
-    else:
-        return clm(l, m) * plm(mu, l, m) * numpy.sin(mphi)
+
+    real = numpy.sqrt((2*l + 1)/(4*numpy.pi) * factorial(l-m)/factorial(l+m)) * plm(mu, l, m) * numpy.cos(mphi)
+    img = numpy.sqrt((2*l + 1)/(4*numpy.pi) * factorial(l-m)/factorial(l+m)) * plm(mu, l, m) * numpy.sin(mphi)
+
+    return real, img
+
+
+def ylm_conj(theta, phi, l, m):
+    rl, im = ylm(theta, phi, l, m)
+    return rl, -im
+
 
 e, eprime, nl, vals = parse_matmsh3("/media/Storage/thesis/python/xsdataPlotter/be9scatter504.dat")
-
-#e = e[::-1]
-#eprime = eprime[::-1]
 
 sh = vals.shape
 nls = sh[2]
@@ -133,21 +123,6 @@ minv = vals.min()
 maxv = vals.max()
 
 pyplot.close("all")
-
-indexes = numpy.linspace(1, 19, 19)
-
-for i in range(nls):
-    pyplot.figure()
-    pyplot.contourf(e/1E6, eprime/1E6, vals[:, :, i], 64, cmap='viridis', vmin=minv, vmax=maxv)
-    caxis = pyplot.colorbar()
-    pyplot.contour(e/1E6, eprime/1E6, vals[:, :, i], [0], colors='k')
-    #ax.set_yscale('log')
[…]

### 2016-09-23 exnih/AbstractCrossSection1d.h
First commit of the exnih project

@@ -1,0 +1,34 @@
+#ifndef Interface_ABSTRACTCROSSSECTION1D_H
+#define Interface_ABSTRACTCROSSSECTION1D_H
+
+#include<iostream>
+#include<vector>
+#include<string>
+
+namespace Standard {
+
+/**
+ * @class AbstractCrossSection1d
+ * @brief Simple interface to a 1d cross section
+ */
+class AbstractCrossSection1d {
+public:
+
+    virtual float * getValues()=0;
+    virtual void setValues(float * values, int size)=0;
+
+    virtual int getSize()=0;
+    virtual void setSize(int size)=0;
+
+    virtual int getMt()=0;
+    virtual void setMt(int mt)=0;
+
+    virtual float getAt(int index)=0;
+    virtual void setAt(int index, float value)=0;
+
+private:
+
+};
+}//namespace Standard
+
+#endif	/* Interface_ABSTRACTCROSSSECTION1D_H */

### 2016-09-23 exnih/AbstractDataNode.cpp

@@ -1,0 +1,21 @@
+/* 
+ * @file:   AbstractDataNode.cpp
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov. Robert A. Lefebvre, lefebvrera@ornl.gov
+ *
+ * Created on November 13, 2012, 10:19 AM
+ */
+
+#include "AbstractDataNode.h"
+namespace ScaleData {
+
+  AbstractDataNode::AbstractDataNode() {
+  }
+
+  AbstractDataNode::AbstractDataNode(const AbstractDataNode& orig) {
+  }
+
+  AbstractDataNode::~AbstractDataNode() {
+  }
+
+}//namespace ScaleData
+

### 2016-09-23 exnih/AbstractDataNode.h

@@ -1,0 +1,45 @@
+/* 
+ * @file:   AbstractDataNode.h
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov. Robert A. Lefebvre, lefebvrera@ornl.gov
+ *
+ * Created on November 13, 2012, 10:19 AM
+ */
+
+#ifndef ABSTRACTDATANODE_H
+#define	ABSTRACTDATANODE_H
+#include <string>
+namespace ScaleData {
+
+  /**
+   * @class AbstractDataNode
+   * @brief Abstract data node for dynamic type storage
+   */
+  class AbstractDataNode {
+  public:
+    /**
+     * Constructor
+     */
+    AbstractDataNode();
+    /**
+     * Copy constructor
+     * @param orig const AbstractDataNode& orig
+     */
+    AbstractDataNode(const AbstractDataNode& orig);
+    /**
+     * Destructor
+     */
+    virtual ~AbstractDataNode();
+    /**
+     * @brief getCopy returns a deep copy of the node
+     * @return deep copy of the node
+     */
+    virtual AbstractDataNode* getCopy() const = 0;
+    /**
+     * @brief produce a string representing this object
+     * @return 
+     */
+    virtual std::string toString()const=0;
+  };
+}//namespace ScaleData
+#endif	/* ABSTRACTDATANODE_H */
+

### 2016-09-23 exnih/AmpxDataHelper.cpp

@@ -1,0 +1,1285 @@
+/*
+ * File:   AmpxDataHelper.cpp
+ * Author: raq
+ *
+ * Created on April 2, 2012, 1:36 PM
+ */
+
+#include "AmpxDataHelper.h"
+#include "CrossSection2d.h"
+#include "CrossSection1d.h"
+#include "AmpxLibrary.h"
+#include "LibraryNuclide.h"
+#include <QDebug>
+//#include "Resource/AmpxLib/ampxlib_config.h"
+AmpxDataHelper::AmpxDataHelper() {
+}
+
+AmpxDataHelper::AmpxDataHelper(const AmpxDataHelper& orig) {
+}
+
+AmpxDataHelper::~AmpxDataHelper() {
+}
+/**
+    * @brief Obtain the nuclide description
+    * @param nuclide - the nuclide for which to obtain the NumRecords
+    * @param library - the library that backs up the nuclide
+    * @return char * - the character array containing the description
+    * Note: This checks the nuclide to determine if the nuclide has the description
+    * if the nuclide's description is NULL, it attempts to retrieve the value from the nuclide
+    * on the ampxlibrary. If the nuclide does not exist on the library, the value (probably NULL) from the
+    * given nuclide will be returned.
+    */
+char * AmpxDataHelper::getDescription(LibraryNuclide * nuclide, AmpxLibrary * library){
+    //Require(nuclide !=NULL);
+    char * description = nuclide->getDescription();
+    if( description !=NULL ){
+        return description;
+    }
+    if( library == NULL ) return description;
+
+    LibraryNuclide * libraryNuclide = library->getNuclideById(nuclide->getId());
+    if( libraryNuclide == NULL ) return description;
+    return libraryNuclide->getDescription();
+}
+/**
+    * @brief Obtain the NumRecords
+    * @param nuclide - the nuclide for which to obtain the NumRecords
+    * @param library - the library that backs up the nuclide
+    * @return int - the NumRecords
+    * Note: This check the nuclide to determine if the nuclide has the NumRecords
+    * if the nuclide's NumRecords is zero, it attempts to retrieve the value from the nuclide
+    * on the ampxlibrary. If the nuclide does not exist on the library, the value (probably zero) from the
+    * given nuclide will be returned.
+    */
+int AmpxDataHelper::getNumRecords(LibraryNuclide * nuclide, AmpxLibrary * library){
+    //Require(nuclide !=NULL);
+    int numRecords = nuclide->getNumRecords();
[…]

### 2016-09-23 exnih/AmpxDataHelper.h

@@ -1,0 +1,444 @@
+/*
+ * File:   AmpxDataHelper.h
+ * Author: raq
+ *
+ * Created on April 2, 2012, 1:36 PM
+ */
+
+#ifndef AMPXDATAHELPER_H
+#define	AMPXDATAHELPER_H
+#include <QList>
+class LibraryNuclide;
+class CrossSection1d;
+class CrossSection2d;
+class BondarenkoData;
+class BondarenkoGlobal;
+class AmpxLibrary;
+
+/**
+ * @class AmpxDataHelper
+ * @brief This class is to assist with retrieval of data from an ampx data component( nuclide, etc)
+ * Most methods provided have take a data component (nuclide, etc) and an AmpxLibrary.
+ * If the data requested is not on the nuclide, the methods will consult the ampx library.
+ */
+class AmpxDataHelper {
+public:
+    AmpxDataHelper();
+    AmpxDataHelper(const AmpxDataHelper& orig);
+    virtual ~AmpxDataHelper();
+private:
+
+public:
+/**
+    * @brief Obtain the nuclide description
+    * @param nuclide - the nuclide for which to obtain the NumRecords
+    * @param library - the library that backs up the nuclide
+    * @return char * - the character array containing the description
+    * Note: This checks the nuclide to determine if the nuclide has the description
+    * if the nuclide's description is NULL, it attempts to retrieve the value from the nuclide
+    * on the ampxlibrary. If the nuclide does not exist on the library, the value (probably NULL) from the
+    * given nuclide will be returned.
+    */
+    static char * getDescription(LibraryNuclide * nuclide, AmpxLibrary * library);
+    /**
+     * @brief Obtain the NumRecords
+     * @param nuclide - the nuclide for which to obtain the NumRecords
+     * @param library - the library that backs up the nuclide
+     * @return int - the NumRecords
+     * Note: This check the nuclide to determine if the nuclide has the NumRecords
+     * if the nuclide's NumRecords is zero, it attempts to retrieve the value from the nuclide
+     * on the ampxlibrary. If the nuclide does not exist on the library, the value (probably zero) from the
+     * given nuclide will be returned.
+     */
+    static int getNumRecords(LibraryNuclide * nuclide, AmpxLibrary * library);
+    /**
+     * @brief Obtain the GammaProdData
+     * @param nuclide - the nuclide for which to obtain the GammaProdData
+     * @param library - the library that backs up the nuclide
+     * @return int - the GammaProdData
[…]

### 2016-09-23 exnih/AmpxLibRegistrar.cpp

@@ -1,0 +1,43 @@
+#include <stdio.h>
+#include "Standard/Interface/SerialFactory.h"
+#include "AmpxLibRegistrar.h"
+#include "BondarenkoData.h"
+#include "BondarenkoInfiniteDiluted.h"
+#include "BondarenkoFactors.h"
+#include "BondarenkoGlobal.h"
+#include "SinkGroup.h"
+#include "ScatterMatrix.h"
+#include "CrossSection1d.h"
+#include "CrossSection2d.h"
+#include "LibraryNuclide.h"
+#include "WorkingLibraryNuclide.h"
+#include "LibraryHeader.h"
+#include "LibraryEnergyBounds.h"
+#include "AmpxLibrary.h"
+#include "SubGroupData.h"
+/**
+ * @brief AmpxLibrary registrar for all serializable AmpxLib objects
+ * @param SerialFactory * factory - the factory to which AmpxLib objects are registered
+ */
+void AmpxLibRegistrar( Standard::SerialFactory * factory){
+    if( factory == NULL) return;
+    factory->registerSerializable(new BondarenkoData());
+    factory->registerSerializable(new BondarenkoInfiniteDiluted());
+    factory->registerSerializable(new BondarenkoFactors());
+    factory->registerSerializable(new BondarenkoGlobal());
+    factory->registerSerializable(new SinkGroup());
+    factory->registerSerializable(new ScatterMatrix());
+    factory->registerSerializable(new CrossSection1d());
+    factory->registerSerializable(new CrossSection2d());
+    factory->registerSerializable(new LibraryNuclide());
+    factory->registerSerializable(new WorkingLibraryNuclide());
+    factory->registerSerializable(new LibraryHeader());
+    factory->registerSerializable(new LibraryEnergyBounds());
+    factory->registerSerializable(new AmpxLibrary());
+    
+    // subgroup data
+    factory->registerSerializable(new SubGroupWeight());
+    factory->registerSerializable(new SubGroupSetGrp());
+    factory->registerSerializable(new SubGroupSet());
+    factory->registerSerializable(new SubGroupData());
+}

### 2016-09-23 exnih/AmpxLibRegistrar.h

@@ -1,0 +1,21 @@
+/* 
+ * File:   AmpxLibRegistrar.h
+ * Author: raq
+ *
+ * Created on November 4, 2011, 10:36 AM
+ */
+
+#ifndef AMPXLIBREGISTRAR_H
+#define	AMPXLIBREGISTRAR_H
+namespace Standard {
+class SerialFactory;
+}
+
+/**
+ * @brief AmpxLibrary registrar for all serializable AmpxLib objects
+ * @param SerialFactory * factory - the factory to which AmpxLib objects are registered
+ */
+void AmpxLibRegistrar( Standard::SerialFactory * factory);
+
+#endif	/* AMPXLIBREGISTRAR_H */
+

### 2016-09-23 exnih/AmpxLibrary.cpp

@@ -1,0 +1,183 @@
+//#include "Standard/Interface/AbstractStream.h"
+#include <QDebug>
+#include "AmpxLibrary.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+#include "LibraryHeader.h"
+#include "LibraryNuclide.h"
+#include "LibraryEnergyBounds.h"
+#include "BondarenkoData.h"
+#include "BondarenkoGlobal.h"
+#include "NuclideResonance.h"
+#include "CrossSection1d.h"
+#include "CrossSection2d.h"
+#include "resources.h"
+#include "AmpxReader.h"
+//#include "Standard/Interface/SerialFactory.h"
+#include "FileStream.h"
+#include "AmpxLibRegistrar.h"
+AmpxLibrary::AmpxLibrary()
+{
+    initialize();
+}
+AmpxLibrary::AmpxLibrary(const AmpxLibrary & orig)
+{
+    initialize();
+    if( orig.header != NULL ) header = orig.header->getCopy();
+    if( orig.neutronEnergyBounds != NULL ) neutronEnergyBounds = orig.neutronEnergyBounds->getCopy();
+    if( orig.gammaEnergyBounds != NULL ) gammaEnergyBounds = orig.gammaEnergyBounds->getCopy();
+
+    for( int i = 0; i < orig.libraryNuclides.size(); i++){
+        libraryNuclides.append(orig.libraryNuclides[i]->getCopy());
+    }
+
+    formatId = orig.formatId;
+}
+AmpxLibrary * AmpxLibrary::getCopy() const {
+    return new AmpxLibrary(*this);
+}
+AmpxLibrary::~AmpxLibrary()
+{
+    if( isOpen() ) close();
+    if( header != NULL ) delete header;
+    if( neutronEnergyBounds != NULL ) delete neutronEnergyBounds;
+    if( gammaEnergyBounds != NULL ) delete gammaEnergyBounds;
+    if( file != NULL ) delete file;
+    for( int i = 0; i < libraryNuclides.size(); i++)
+        delete libraryNuclides.value(i);
+    libraryNuclides.clear();
+}
+
+bool AmpxLibrary::operator==(AmpxLibrary & a){
+    if( header != NULL && a.header != NULL ){
+        bool equal = *header == *a.header;
+        if( !equal ){
+            qDebug() << qPrintable("AmpxHeaders not equal!");
+            return false;
+        }
+    }else if( header != a.header){
[…]

### 2016-09-23 exnih/AmpxLibrary.h

@@ -1,0 +1,301 @@
+#ifndef AMPXLIBRARY_H
+#define AMPXLIBRARY_H
+#include <memory>
+#include <iostream>
+#include <fstream>
+#include <QString>
+#include <QList>
+#include "Resource.h"
+#include "LibraryEnergyBounds.h"
+#include "LibraryNuclide.h"
+#include "LibraryHeader.h"
+#include "CrossSection1d.h"
+#include "CrossSection2d.h"
+//#include "Standard/Interface/Serializable.h"
+using namespace std;
+
+
+class AmpxLibrary : public Resource{
+protected:
+   LibraryHeader * header;
+   LibraryEnergyBounds * neutronEnergyBounds;
+   LibraryEnergyBounds * gammaEnergyBounds;
+
+   QList<LibraryNuclide*> libraryNuclides;
+
+   fstream * file;
+
+   int formatId;
+
+   void initialize();
+   /// Remove a nuclide with a given Id from the given list of library nuclides
+   /// @param int id - The id of the specified nuclide to remove
+   /// @param QList<LibraryNuclide*> list - The list of LibraryNuclides from which to remove
+   /// @param bool deleteItem=false - indicate whether to delete the object removed
+   /// @return int count - the count of items removed.
+   int removeNuclide(int id, QList<LibraryNuclide*>& list, int mixid=0, bool deleteItem=false)
+   {
+       int count = 0;
+       for( int i = 0; i < list.size(); i++){
+            if( list.at(i)->getId() == id && list.at(i)->getMixture() == mixid ){
+                LibraryNuclide * nuclide = list.value(i);
+                list.removeAt(i);
+                count++;
+                i --;
+                if( deleteItem ) delete nuclide;
+            }
+       }
+       return count;
+   }// end removeNuclide
+
+public:
+   AmpxLibrary();
+   AmpxLibrary(const AmpxLibrary & orig);
+   virtual ~AmpxLibrary();
+
+   bool operator==(AmpxLibrary & a);
+
+   virtual AmpxLibrary * getCopy() const;
+
+   void setFormatId(int id){
+       this->formatId = id;
+   }
+
+   int getFormatId(){
+       return formatId;
+   }
+
+   virtual int open();
+   virtual int close();
+   bool isOpen();
+   QString getTitle() const {return "AmpxLibrary Resource";}
+   ///This operation returns the name of the Resource.
+   QString getResourceName(){return "AmpxLibrary";}
+
[…]

### 2016-09-23 exnih/AmpxReader.cpp

@@ -1,0 +1,1177 @@
+#include <stdio.h>
+#include <string.h>
+#include <istream>
+#include <QDebug>
+#include "AmpxReader.h"
+#include "LibraryHeader.h"
+#include "LibraryEnergyBounds.h"
+#include "LibraryNuclide.h"
+#include "NuclideResonance.h"
+#include "BondarenkoGlobal.h"
+#include "BondarenkoData.h"
+#include "BondarenkoInfiniteDiluted.h"
+#include "BondarenkoFactors.h"
+#include "CrossSection1d.h"
+#include "CrossSection2d.h"
+#include "ScatterMatrix.h"
+#include "resources.h"
+#include "EndianUtils.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+
+//using namespace ScaleUtils::IO;
+
+AmpxReader::AmpxReader()
+{
+    initialize();
+}
+
+void AmpxReader::initialize()
+{
+    swapEndian = false;
+}
+
+int AmpxReader::footprintLibraryEnergyBounds(int numGrps)
+{
+    return 2*(numGrps+1)*sizeof(float);
+}
+
+int AmpxReader::footprintBondarenkoGlobal(int numSig0, int numTemp)
+{
+    return ((numSig0+numTemp)*sizeof(float))
+           + (sizeof(float)*2); /// elo ehi
+}// end of footprint
+
+int AmpxReader::footprintLibraryNuclide(){
+    return AMPX_NUCLIDE_DESCR_LENGTH+
+         (sizeof(int)*AMPX_NUMBER_NUCLIDE_INTEGER_OPTIONS);
+}
+
+int AmpxReader::footprintBondarenkoInfiniteDiluted(int startGrp, int endGrp){
+        return ((endGrp-startGrp + 1)*sizeof(float));
+}// end of footprint
+
+int AmpxReader::footprintBondarenkoFactors(int numSig, int numTemp, int startGrp, int endGrp){
+    return (numSig*numTemp*(endGrp-startGrp + 1)*sizeof(float));
+}// end of footprint
+
+int AmpxReader::footprintCrossSection1d(int numGrps){
+        return ((numGrps + 1)*sizeof(float));
+}// end of footprint
+
+int AmpxReader::footprintCrossSection2d(){
+    return CROSS_SECTION_2D_INTEGER_OPTIONS * sizeof(int);
+}
+
+int AmpxReader::footprintNuclideResonance(int numResolved, int numUnresolved){
[…]

### 2016-09-23 exnih/AmpxReader.h

@@ -1,0 +1,170 @@
+#ifndef AMPXREADER_H
+#define AMPXREADER_H
+#include <QList>
+#include <iostream>
+#include <fstream>
+#include <QString>
+#include "AmpxLibrary.h"
+
+using namespace std;
+
+class LibraryHeader;
+class LibraryEnergyBounds;
+class LibraryNuclide;
+class NuclideResonance;
+class BondarenkoGlobal;
+class BondarenkoData;
+class BondarenkoInfiniteDiluted;
+class BondarenkoFactors;
+class CrossSection1d;
+class CrossSection2d;
+class ScatterMatrix;
+
+class AmpxReader
+{
+protected:
+    bool swapEndian;
+    LibraryHeader * readLibraryHeader(fstream & is,
+                               int * pos=NULL,
+                               bool verbose=false);
+    LibraryEnergyBounds * readEnergyBounds(fstream & is,
+                                            int numGroups,
+                                            int * pos=NULL,
+                                            bool verbose=false);
+
+    LibraryNuclide* readLibraryNuclide(fstream & is,
+                                int * pos=NULL,
+                                bool verbose=false);
+    BondarenkoGlobal * readBondarenkoGlobal(fstream & is,
+                                            int numSig0,
+                                            int numTemps,
+                                            int * pos=NULL,
+                                            bool verbose=false);
+
+    /// reads a nuclide's bondarenko data sets
+    /// and returns a list of objects
+    /// @param fstream &is - The file from which to read the data
+    /// @param int numBondSets - The number of bondarenko data sets to read
+    /// @param int numBondSig0- The number of bondarenko Sig0 data sets to read
+    /// @param int numBondTemps - The number of bondarenko Temp data sets to read
+    /// @param int * pos=NULL - The optional position from which to read
+    /// @param bool verbose - be verbose about activity
+    QList<BondarenkoData*> readBondarenkoData(fstream & is,
+                        int numBondSets,
+                        int numBondSig0,
+                        int numBondTemps,
+                        int * pos=NULL, bool verbose=false);
+
+    BondarenkoInfiniteDiluted * readBondarenkoInfiniteDiluted(fstream & is,
+                                                            int startGrp,
+                                                            int endGrp,
+                                                            int * pos=NULL,
+                                                            bool verbose=false);
+
+    BondarenkoFactors * readBondarenkoFactors(fstream & is,
+                                                int numSig,
+                                                int numTemp,
+                                                int startGrp,
[…]

### 2016-09-23 exnih/BondarenkoData.cpp

@@ -1,0 +1,75 @@
+//#include "Standard/Interface/AbstractStream.h"
+#include "BondarenkoData.h"
+#include "BondarenkoInfiniteDiluted.h"
+#include "BondarenkoFactors.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+
+BondarenkoData::BondarenkoData()
+{
+    initialize();
+} // end
+BondarenkoData::BondarenkoData(const BondarenkoData & orig){
+    d = orig.d;
+    d->ref.ref();
+    setInfiniteDiluted(NULL);
+    setFactors(NULL);    
+    if( orig.infiniteDiluted != NULL )
+        setInfiniteDiluted(orig.infiniteDiluted->getCopy());
+    if( orig.factors != NULL )
+        setFactors(orig.factors->getCopy());
+}
+void BondarenkoData::initialize()
+{
+   setInfiniteDiluted(NULL);
+   setFactors(NULL);
+   d = new BondarenkoDataData();
+}// end of initialize
+BondarenkoData::~BondarenkoData()
+{
+    if( getInfiniteDiluted() != NULL ) delete getInfiniteDiluted();
+    if( getFactors() != NULL ) delete getFactors();
+    if( !d->ref.deref() ){
+        delete d;
+    }
+}
+bool BondarenkoData::operator==(BondarenkoData & a){
+    if( infiniteDiluted != NULL && a.infiniteDiluted != NULL ){
+        bool equal = *infiniteDiluted == *a.infiniteDiluted;
+        if( !equal ) return false;
+    }else if( infiniteDiluted != a.infiniteDiluted ){
+        return false;
+    }
+    if( factors != NULL && a.factors != NULL ){
+        bool equal = *factors == *a.factors;
+        if( !equal ) return false;
+    }else if( factors != NULL && a.factors == NULL){
+        return false;
+    }else if( factors == NULL && a.factors != NULL){
+        return false;
+    }
+    if( d == a.d ) return true;
+    
+    for( int i = 0; i < AMPX_BONDARENKO_DATA_INTEGER_OPTIONS; i++ ){
+        if( getData(i) != a.getData(i) ){
+            return false;
+        }
+    }
+    return true;
+}
+BondarenkoData * BondarenkoData::getCopy() const
+{
+    BondarenkoData * copy = new BondarenkoData(*this);
+    return copy;
+}
+
+void BondarenkoData::setMt(int MT){
[…]

### 2016-09-23 exnih/BondarenkoData.h

@@ -1,0 +1,117 @@
+#ifndef BONDARENKODATA_H
+#define BONDARENKODATA_H
+#include "LibraryItem.h"
+#include "BondarenkoInfiniteDiluted.h"
+#include "BondarenkoFactors.h"
+#include <QList>
+#include <QString>
+#include <QAtomicInt>
+//#include "Standard/Interface/Serializable.h"
+
+/// Indicates the number of integer options 
+/// in a bondarenko set. MT, NF, NL,... etc.
+#define AMPX_BONDARENKO_DATA_INTEGER_OPTIONS 6
+/// indicates the index of the MT variable
+#define AMPX_BONDARENKO_MT 0
+/// indicates the index of the NF variable
+#define AMPX_BONDARENKO_NF 1
+/// indicates the index of the NL variable
+#define AMPX_BONDARENKO_NL 2
+/// indicates the index of the ORDER variable
+#define AMPX_BONDARENKO_ORDER 3
+/// indicates the index of the IOFF variable
+#define AMPX_BONDARENKO_IOFF 4
+/// indicates the index of the NZ variable
+#define AMPX_BONDARENKO_NZ 5
+class BondarenkoDataData{
+public:
+    /// contains bondarenko items
+    /// mt, nf, nl, order, ioff, and nz
+    int data[AMPX_BONDARENKO_DATA_INTEGER_OPTIONS]; 
+    QAtomicInt ref;
+    BondarenkoDataData():ref(1){
+        for( int i = 0; i < AMPX_BONDARENKO_DATA_INTEGER_OPTIONS; i++)
+            data[i] = 0;
+    }
+    BondarenkoDataData(const BondarenkoDataData & orig):ref(1){
+        for( int i = 0; i < AMPX_BONDARENKO_DATA_INTEGER_OPTIONS; i++)
+            data[i] = orig.data[i];
+    }
+};
+class BondarenkoData : public LibraryItem //, public Standard::Serializable{
+{
+private:
+    BondarenkoInfiniteDiluted * infiniteDiluted;
+    BondarenkoFactors * factors;
+    BondarenkoDataData * d;
+       
+    void initialize();
+public:
+    BondarenkoData();
+    BondarenkoData(const BondarenkoData & orig);
+    ~BondarenkoData();
+    
+    bool operator==(BondarenkoData & a);
+    
+    /// Obtain a copy of this bondarenkoData
+    /// BondarenkoData * - The copy
[…]

### 2016-09-23 exnih/BondarenkoFactors.cpp

@@ -1,0 +1,72 @@
+#include "BondarenkoFactors.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+//#include "Standard/Interface/AbstractStream.h"
+
+#include <QDebug>
+
+BondarenkoFactors::BondarenkoFactors()
+{
+    initialize();
+}
+void BondarenkoFactors::initialize()
+{
+    d = new BondarenkoFactorsData();
+}
+BondarenkoFactors::BondarenkoFactors(const BondarenkoFactors & factors){
+    d = factors.d;
+    d->ref.ref();
+}
+BondarenkoFactors::~BondarenkoFactors()
+{
+    if( !d->ref.deref() ){
+        delete d;
+    }
+}
+bool BondarenkoFactors::operator==(BondarenkoFactors & a){
+    if( d == a.d ) return true;
+    if( d->nSig0 != a.d->nSig0 ) return false;
+    if( d->nTemp != a.d->nTemp ) return false;
+    if( d->nGrps != a.d->nGrps ) return false;
+    for( int i = 0; i < getSig0Size(); i++){
+        for( int t = 0; t < getTempSize(); t++){
+            for( int g = 0; g < getGrpSize(); g++){
+                if( getAt(i,t,g) != a.getAt(i,t,g) ) return false;
+            }
+        }
+    }
+    return true;
+}
+BondarenkoFactors * BondarenkoFactors::getCopy() const
+{
+    BondarenkoFactors * copy = new BondarenkoFactors(*this);    
+    return copy;
+}
+void BondarenkoFactors::setSizes(int numSig0, int numTemps, int numGrps){
+    int size = numSig0*numTemps*numGrps;
+    //Require( size > 0);
+    //Ensure( this->d != NULL );
+    qDebug() << "Setting bondarenko sizes nSig0 "<<numSig0<<", numTemps "<<numTemps<<", numGrps "<<numGrps;
+    qDebug() << "Existing bondarenko sizes nSig0 "<<d->nSig0<<", numTemps "<<d->nTemp<<", numGrps "<<d->nGrps;
+    if( this->d->nGrps != numGrps ){
+        qAtomicDetach(d);
+    }else if( this->d->nTemp != numTemps ){
+        qAtomicDetach(d);
+    }else if( this->d->nSig0 != numSig0 ){
+        qAtomicDetach(d);
+    }else{
+        // the sizes do not need to be changed because they are the same            
+        return;
+    }
[…]

### 2016-09-23 exnih/BondarenkoFactors.h

@@ -1,0 +1,95 @@
+#ifndef BONDARENKOFACTORS_H
+#define BONDARENKOFACTORS_H
+#include "LibraryItem.h"
+#include "resources.h"
+#include <QString>
+#include <QAtomicInt>
+//#include "Standard/Interface/Serializable.h"
+class BondarenkoFactorsData{
+public:
+    QAtomicInt ref;
+    int nSig0, nTemp, nGrps;
+    float * values;
+
+    BondarenkoFactorsData(): ref(1), nSig0(0),nTemp(0),nGrps(0),values(NULL){}
+    BondarenkoFactorsData(const BondarenkoFactorsData & orig)
+        :ref(1),nSig0(orig.nSig0), nTemp(orig.nTemp), nGrps(orig.nGrps),values(NULL){
+        int size = nSig0* nTemp*nGrps;
+        if( size == 0) return;
+        values = new float[size];
+        for( int i = 0; i < size; i ++)
+            values[i] = orig.values[i];
+    };
+    virtual ~BondarenkoFactorsData(){
+        if( values != NULL) delete [] values;
+    }
+};
+class BondarenkoFactors: public LibraryItem //, public Standard::Serializable
+{
+private:
+    BondarenkoFactorsData * d;
+    void initialize();
+public:
+    BondarenkoFactors();
+    BondarenkoFactors(const BondarenkoFactors & factors);
+    ~BondarenkoFactors();
+    
+    bool operator==(BondarenkoFactors & a);
+
+    /// Obtain a copy of the factors
+    /// @return BondarenkoFactors * - the copy
+    BondarenkoFactors * getCopy() const;
+
+    float * getValues(){return d->values;}
+    float getAt(int sig0, int temp, int grp){
+        if( sig0 < 0 || sig0 >= d->nSig0 ) return 0.0;
+        if( temp < 0 || temp >= d->nTemp ) return 0.0;
+        if( grp < 0 || grp >= d->nGrps ) return 0.0;
+        return d->values[(grp*d->nTemp*d->nSig0)+(temp*d->nSig0)+(sig0)];
+    }
+    void setAt(int sig0, int temp, int grp, float value){
+        if( sig0 < 0 || sig0 >= d->nSig0 ) return ;
+        if( temp < 0 || temp >= d->nTemp ) return ;
+        if( grp < 0 || grp >= d->nGrps ) return ;
+        qAtomicDetach(d);
[…]

### 2016-09-23 exnih/BondarenkoGlobal.cpp

@@ -1,0 +1,46 @@
+//#include "Standard/Interface/AbstractStream.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+#include "BondarenkoGlobal.h"
+BondarenkoGlobal::BondarenkoGlobal()
+{
+    initialize();
+}
+void BondarenkoGlobal::initialize()
+{
+    d = new BondarenkoGlobalData();
+}
+BondarenkoGlobal::~BondarenkoGlobal()
+{
+    if( !d->ref.deref() ){
+        delete d;
+    }
+}
+BondarenkoGlobal::BondarenkoGlobal(const BondarenkoGlobal & orig){
+    d = orig.d;
+    d->ref.ref();    
+}
+bool BondarenkoGlobal::operator==(BondarenkoGlobal & a){
+    if( d == a.d ) return true;
+    if( d->elo != a.d->elo ) return false;
+    if( d->ehi != a.d->ehi ) return false;
+    if( d->sig0Size != a.d->sig0Size ) return false;
+    if( d->tempSize != a.d->tempSize ) return false;
+    for( int i = 0; i < d->sig0Size; i++){
+        if( d->sig0[i] != a.d->sig0[i] ) return false;
+    }
+    for( int i = 0; i < d->tempSize; i++){
+        if( d->temp[i] != a.d->temp[i] ) return false;
+    }
+    return true;
+}
+BondarenkoGlobal * BondarenkoGlobal::getCopy() const
+{
+    BondarenkoGlobal * copy = new BondarenkoGlobal(*this);
+    return copy;
+}
+
+// Serialization interfaces
+
+const long BondarenkoGlobal::uid = 0xdeaaf95200140511;
+
+

### 2016-09-23 exnih/BondarenkoGlobal.h

@@ -1,0 +1,108 @@
+#ifndef BONDARENKO_GLOBAL_H
+#define BONDARENKO_GLOBAL_H
+#include "LibraryItem.h"
+#include "resources.h"
+#include <QString>
+#include <QAtomicInt>
+//#include "Standard/Interface/Serializable.h"
+
+class BondarenkoGlobalData{
+public:
+    QAtomicInt ref;
+    float * sig0, * temp;
+    float elo, ehi;
+    int sig0Size, tempSize;
+    BondarenkoGlobalData(): ref(1), sig0(NULL),temp(NULL),elo(0), ehi(0)
+                        ,sig0Size(0),tempSize(0) {}
+    BondarenkoGlobalData(const BondarenkoGlobalData & orig)
+        :ref(1), sig0(NULL), temp(NULL), elo(0), ehi(0),sig0Size(0),tempSize(0)
+    {
+        elo = orig.elo;
+        ehi = orig.ehi;
+        if( orig.sig0Size > 0 ){
+            sig0 = new float[orig.sig0Size];
+            sig0Size = orig.sig0Size;
+            for( int i = 0; i < orig.sig0Size; i ++)
+                sig0[i] = orig.sig0[i];
+        }
+        if( orig.tempSize > 0 ){
+            temp = new float[orig.tempSize];
+            tempSize = orig.tempSize;
+            for( int i = 0; i < orig.tempSize; i++)
+                temp[i] = orig.temp[i];
+        }
+    }
+    ~BondarenkoGlobalData(){
+        if( sig0 != NULL ) delete [] sig0;
+        if( temp != NULL ) delete [] temp;
+    }
+};
+
+class BondarenkoGlobal : public LibraryItem
+{
+private:
+    BondarenkoGlobalData * d;
+    void initialize();
+public:
+    BondarenkoGlobal();
+    BondarenkoGlobal(const BondarenkoGlobal & orig);
+    ~BondarenkoGlobal();
+    bool operator==(BondarenkoGlobal & a);
+    /// Obtain a copy of this BondarenkoGlobal data
+    /// @return BondarenkoGlobal * - The copy
+    BondarenkoGlobal * getCopy() const;
+
+    /// obtain the lower energy  for which factors can apply in the
+    /// case where they do not span all energy groups
+    float getElo(){ return d->elo;}
+    void setElo(float elo){
+        qAtomicDetach(d);
+        this->d->elo = elo;
+    }
+
[…]

### 2016-09-23 exnih/BondarenkoInfiniteDiluted.cpp

@@ -1,0 +1,123 @@
+#include <stdio.h>
+#include <string.h>
+//#include "Standard/Interface/AbstractStream.h"
+#include "BondarenkoInfiniteDiluted.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+BondarenkoInfiniteDiluted::BondarenkoInfiniteDiluted()
+{
+    initialize();
+}
+BondarenkoInfiniteDiluted::BondarenkoInfiniteDiluted(const BondarenkoInfiniteDiluted& orig){
+    d = orig.d;
+    d->ref.ref();
+}
+void BondarenkoInfiniteDiluted::initialize()
+{
+    d = new BondarenkoInfiniteDilutedData();
+}
+BondarenkoInfiniteDiluted::~BondarenkoInfiniteDiluted()
+{
+    if( !d->ref.deref() ){
+        delete d;
+    }
+}
+bool BondarenkoInfiniteDiluted::operator==(BondarenkoInfiniteDiluted & a){
+    if( d == a.d ) return true;
+    if( d->size != a.d->size ) return false;
+    for( int i = 0; i < d->size; i++ ){
+        if( d->values[i] != a.d->values[i] ) return false;
+    }
+    return true;
+}
+BondarenkoInfiniteDiluted * BondarenkoInfiniteDiluted::getCopy() const
+{
+    BondarenkoInfiniteDiluted * copy = new BondarenkoInfiniteDiluted(*this);
+    return copy;
+}
+
+// Serialization interfaces
+
+const long BondarenkoInfiniteDiluted::uid = 0x6baff95172442622;
+
+/**
+ * @brief Serialize the object into a contiguous block of data
+ * @param AbstractStream * stream - the stream into which the contiguous data will be stored
+ * @return int - 0 upon success, error otherwise
+ */
+/*
+int BondarenkoInfiniteDiluted::serialize(Standard::AbstractStream * stream) const {
+    if( stream == NULL ) return -1;    
+    long classUID = this->getUID();
+    stream->write((char*)&classUID, sizeof(classUID));
+    
+    // write the size of this object
+    unsigned long serializedSize = getSerializedSize();
+    stream->write((char*)&serializedSize, sizeof(serializedSize));
+    
+    // DBC=4 - capture the oldWriteHead for checksumming later
+    Remember(long oldWriteHead = stream->getWriteHead());
+    
+    // write the data size  
[…]

### 2016-09-23 exnih/BondarenkoInfiniteDiluted.h

@@ -1,0 +1,129 @@
+#ifndef BONDARENKOINFINITEDILUTED_H
+#define BONDARENKOINFINITEDILUTED_H
+
+#include "LibraryItem.h"
+#include "resources.h"
+#include <QString>
+#include <QAtomicInt>
+//#include "Standard/Interface/Serializable.h"
+class BondarenkoInfiniteDilutedData{
+public:
+    QAtomicInt ref;
+    int size;
+    float * values;
+    BondarenkoInfiniteDilutedData():ref(1), size(0), values(NULL){}
+    BondarenkoInfiniteDilutedData(const BondarenkoInfiniteDilutedData & orig)
+            : ref(1),size(0), values(NULL){
+        if( orig.size <= 0 ) return;
+        this->size = orig.size;
+        this->values = new float[this->size];
+        for( int i = 0; i < this->size; i++)
+            this->values[i] = orig.values[i];
+    }
+    ~BondarenkoInfiniteDilutedData(){
+        if( this->values != NULL ) delete [] this->values;
+    }
+};
+
+class BondarenkoInfiniteDiluted: public LibraryItem //, public Standard::Serializable
+{
+private:
+    BondarenkoInfiniteDilutedData * d;
+    void initialize();
+public:
+    BondarenkoInfiniteDiluted();
+    BondarenkoInfiniteDiluted(const BondarenkoInfiniteDiluted & orig);
+    ~BondarenkoInfiniteDiluted();
+    
+    bool operator==(BondarenkoInfiniteDiluted & a);
+    
+    float * getValues(){return d->values;};
+    int getSize(){return d->size;};
+    void setValues(float * values, int size){
+        qAtomicDetach(d);
+        this->d->values = values;
+        this->d->size=size;
+    }
+    void setSize(int size){
+        if( size <= 0 || size == d->size) return;
+        qAtomicDetach(d);
+
+        float * newValues = new float[size];
+        for( int i = 0; i < size; i++){
+            newValues[i] = 0; // zero out memory
+        }
+        setValues(newValues,size);
+    }
+    /// Obtain a copy of the infinite diluted
+    /// @return BondarenkoInfiniteDiluted * - the copy
+    BondarenkoInfiniteDiluted * getCopy() const;
+   
+
+    float getAt(int index){
+        if( index < 0 || index >= getSize()) return 0.0;
+        return getValues()[index];
+    }
+    void setAt(int index, float value){
+        if( index < 0 || index >= getSize()) return ;
[…]

### 2016-09-23 exnih/CrossSection1d.cpp

@@ -1,0 +1,49 @@
+#include "CrossSection1d.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+//#include "Standard/Interface/AbstractStream.h"
+CrossSection1d::CrossSection1d()
+{
+    initialize();
+}
+CrossSection1d::CrossSection1d(const CrossSection1d & orig){
+    d = orig.d;
+    d->ref.ref();
+}
+void CrossSection1d::initialize()
+{
+    d = new CrossSection1dData();
+}
+
+bool CrossSection1d::operator==(CrossSection1d & a){
+    if( d == a.d ) return true;
+    if( getMt() != a.getMt()) return false;
+    if( getSize() != a.getSize() ) return false;
+    for( int i = 0; i < getSize(); i++){
+        if( getAt(i) != a.getAt(i) ) return false;
+    }
+    return true;
+}
+CrossSection1d::~CrossSection1d()
+{
+    if( !d->ref.deref() ){
+        delete d;
+    }
+}
+
+CrossSection1d * CrossSection1d::getCopy() const
+{
+    CrossSection1d * copy = new CrossSection1d(*this);
+    return copy;
+}
+
+void CrossSection1d::setMt(int mt)
+{
+    int oldMt = getMt();
+    if( oldMt == mt ) return;
+    qAtomicDetach(d);
+    this->d->mt=mt;
+}
+// Serialization interfaces
+
+const long CrossSection1d::uid = 0xabeef95102040602;
+

### 2016-09-23 exnih/CrossSection1d.h

@@ -1,0 +1,92 @@
+#ifndef CROSSSECTION1D_H
+#define CROSSSECTION1D_H
+#include "LibraryItem.h"
+#include "resources.h"
+#include <QString>
+#include <QList>
+#include <QAtomicInt>
+//#include "Standard/Interface/Serializable.h"
+#include "AbstractCrossSection1d.h"
+
+class CrossSection1dData{
+public:
+    QAtomicInt ref;
+    float mt;
+    int size;
+    float * values;
+    CrossSection1dData(): ref(1), mt(0), size(0), values(NULL){}
+    CrossSection1dData(const CrossSection1dData & orig): ref(1),mt(orig.mt),size(0), values(NULL){
+        if( orig.size <= 0 ) return;
+        values = new float[orig.size];
+        for( int i = 0; i < orig.size; i ++){
+            values[i] = orig.values[i];
+        }
+        size = orig.size;
+    }
+    ~CrossSection1dData(){
+        if( values != NULL) delete [] values;
+    }
+};
+class CrossSection1d: public LibraryItem,
+                      public Standard::AbstractCrossSection1d
+{
+private:
+    CrossSection1dData * d;
+    void initialize();
+public:
+    CrossSection1d();
+    CrossSection1d(const CrossSection1d & orig);
+    ~CrossSection1d();
+
+    bool operator==(CrossSection1d & a);
+
+    /// Obtain a copy of the CrossSection1d
+    /// CrossSection1d * - the copy
+    virtual CrossSection1d * getCopy() const;
+    float * getValues(){return d->values;};
+    int getSize(){return d->size;};
+    void setValues(float * values, int size){
+        qAtomicDetach(d);
+        if( this->d->values != NULL) delete [] this->d->values;
+        this->d->values = values;
+        this->d->size=size;
+    }
+    int getMt(){return (int)d->mt;}
+    void setMt(int mt);
+    float getAt(int index){
+        if( index < 0 || index >= getSize()) return 0.0;
+        return getValues()[index];
+    }
+    void setAt(int index, float value){
+        if( index < 0 || index >= getSize()) return ;
+        qAtomicDetach(d);
+        getValues()[index]=value;
+    }
+    void setSize(int size){
+        if( size <= 0 ) return;
+        qAtomicDetach(d);
+
+        float * newValues = new float[size];
+        for( int i = 0; i < size; i++){
+            newValues[i] = 0; // zero out memory
+        }
+        setValues(newValues,size);
+    }
[…]

### 2016-09-23 exnih/CrossSection2d.cpp

@@ -1,0 +1,128 @@
+#include "CrossSection2d.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+#include <QDebug>
+//#include "Standard/Interface/AbstractStream.h"
+
+
+CrossSection2d::CrossSection2d()
+{
+    initialize();
+}
+CrossSection2d::CrossSection2d(const CrossSection2d & orig){
+    d = orig.d;
+    d->ref.ref();
+    for( int i = 0; i < orig.scatterMatrices.size(); i++)
+        this->scatterMatrices.append(orig.scatterMatrices[i]->getCopy());
+}
+void CrossSection2d::initialize()
+{
+    d = new CrossSection2dData();
+}
+bool CrossSection2d::operator==(CrossSection2d & a){
+    if( scatterMatrices.size() != a.scatterMatrices.size() ) return false;
+    for( int i = 0; i < scatterMatrices.size(); i++){
+        bool equal = *scatterMatrices[i] == *a.scatterMatrices[i];
+        if( !equal ) {
+            qDebug() << "CrossSection2d ScatterMatrix not equal for mt="<<getMt()<<" at index "<<i;
+            return false;
+        }
+    }
+    if( d == a.d ) return true;
+    for( int i = 0; i < CROSS_SECTION_2D_INTEGER_OPTIONS;i ++)
+        if( d->data[i] != a.d->data[i] )
+        {
+            qDebug() << "CrossSection2d option not equal for mt="<<getMt()<<" at index "<<i;
+            return false;
+        }
+    return true;
+}
+CrossSection2d::~CrossSection2d()
+{
+    if( !d->ref.deref() ){
+        delete d;
+    }
+    for( int i = 0; i < scatterMatrices.size(); i++){
+        delete scatterMatrices.value(i);
+    }
+    scatterMatrices.clear();
+}
+int CrossSection2d::getLength()
+{
+    int maxLength = -1;
+    if( scatterMatrices.size() == 0 ){
+        cerr<<"***Warning: CrossSection2d contains no ScatterMatrices!"<<endl;
+        return 0;
+    }
+    for( int i = 0; i < scatterMatrices.size(); i++)
+    {
+        int length;
+        /// delete the newly allocated collapsed matrix
+        /// only care about the length of the collapsed data
+        float * data = scatterMatrices.value(i)->getCollapsedMatrix(length);
+        if( data == NULL ){
[…]

### 2016-09-23 exnih/CrossSection2d.h

@@ -1,0 +1,188 @@
+#ifndef CROSSSECTION2D_H
+#define CROSSSECTION2D_H
+#include "LibraryItem.h"
+#include "ScatterMatrix.h"
+#include <QList>
+#include <QString>
+#include <QtAlgorithms>
+#include <QAtomicInt>
+//#include "Standard/Interface/Serializable.h"
+
+#define CROSS_SECTION_2D_INTEGER_OPTIONS 4
+#define CROSS_SECTION_2D_MT 0
+#define CROSS_SECTION_2D_LENGTH 1
+#define CROSS_SECTION_2D_NL 2
+#define CROSS_SECTION_2D_NT 3
+
+class CrossSection2dData{
+public :
+    int data[CROSS_SECTION_2D_INTEGER_OPTIONS];
+    QAtomicInt ref;
+    int type;
+    CrossSection2dData(): ref(1), type(0){
+        for( int i = 0; i < CROSS_SECTION_2D_INTEGER_OPTIONS;i ++)
+            data[i] = 0;
+    };
+    CrossSection2dData(const CrossSection2dData & orig): ref(1){
+        for( int i = 0; i < CROSS_SECTION_2D_INTEGER_OPTIONS;i ++)
+            data[i] = orig.data[i];
+        type = orig.type;
+    }
+};
+
+class CrossSection2d : public LibraryItem//, public Standard::Serializable
+{
+private:
+
+    QList<ScatterMatrix *> scatterMatrices;
+    CrossSection2dData * d;
+    void initialize();
+
+public:
+    CrossSection2d();
+    CrossSection2d(const CrossSection2d & orig);
+    const CrossSection2d & operator=(const CrossSection2d & orig);
+    ~CrossSection2d();
+
+    bool operator==(CrossSection2d & a);
+
+    /// Obtain a copy of the CrossSection2d
+    /// @return CrossSection2d * - the copy
+    CrossSection2d * getCopy() const;
+
+    /// obtain the type of cross section data.
+    ///  neutron2d, gamma2d, or gammaproduction
+    /// AMPXLIB_NEUTRON2D_DATA,_GAMMA2D_DATA, or GAMMAPRODUCTION_DATA
+    int getType(){return d->type;}
+
+    /// Obtain a list of ascending temperature data
+    /// for all ScatterMatrices associated with CrossSection2d
+    QList<float> getTemperatureList();
+    QList<ScatterMatrix *> * getScatterMatrices(){return &scatterMatrices;}
[…]

### 2016-09-23 exnih/DataNode.h

@@ -1,0 +1,62 @@
+/* 
+ * @file:   DataNode.h
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov. Robert A. Lefebvre, lefebvrera@ornl.gov
+ *
+ * Created on November 13, 2012, 10:19 AM
+ */
+
+#ifndef DATANODE_H
+#define	DATANODE_H
+#include <string>
+
+#include "AbstractDataNode.h"
+
+namespace ScaleData {
+
+  /**
+   * @class DataNode
+   * @brief  data node for dynamic type storage
+   */
+  template<class T>
+  class DataNode : public AbstractDataNode {
+  private:
+    T value;
+  public:
+    /**
+     * Constructor
+     */
+    DataNode(T& v);
+    /**
+     * Copy constructor
+     * @param orig const DataNode& orig
+     */
+    DataNode(const DataNode& orig);
+    /**
+     * Destructor
+     */
+    virtual ~DataNode();
+    
+    AbstractDataNode* getCopy() const
+    {
+        return new DataNode(*this);
+    }
+
+    /**
+     * Get the value of this type
+     */
+    T& getRef();
+    const T& get() const;
+    /**
+     * Set the value of this type
+     * @param value template<T>
+     */
+    void set(T& value);
+    virtual std::string toString()const;
+  }; // end pf DataNode
+}//namespace ScaleData
+//
+// Include inline and template memebers
+//
+#include "DataNode.i.h"
+#endif	/* DATANODE_H */
+

### 2016-09-23 exnih/DataNode.i.h

@@ -1,0 +1,50 @@
+/* 
+ * @file:   DataNode.i.h
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov. Robert A. Lefebvre, lefebvrera@ornl.gov
+ *
+ * Created on November 13, 2012, 10:19 AM
+ */
+
+#ifndef DATANODE_I_H
+#define	DATANODE_I_H
+
+#include <sstream>
+#include <typeinfo>
+
+namespace ScaleData {
+  template<class T>
+  DataNode<T>::DataNode(T& v)
+  : value(v) {
+  }
+
+  template<class T>
+  DataNode<T>::DataNode(const DataNode& orig)
+  : value(orig.value) {
+  }
+  template<class T>
+  DataNode<T>::~DataNode()
+  {
+  }
+
+  template<class T>
+  T& DataNode<T>::getRef() {
+    return value;
+  }
+  template<class T>
+  const T& DataNode<T>::get() const {
+    return value;
+  }
+
+  template<class T>
+  void DataNode<T>::set(T& v) {
+    this->value = v;
+  }
+  template<class T>
+  std::string DataNode<T>::toString() const{
+      std::stringstream stream;
+      stream<<"type("<<typeid(T).name()<<")";
+      return stream.str();
+  }
+}//namespace ScaleData
+#endif	/* DATANODE_I_H */
+

### 2016-09-23 exnih/DataSet.cpp

@@ -1,0 +1,87 @@
+/*
+ * File:   DataSet.cpp
+ * Author: raq
+ *
+ * Created on November 21, 2012, 12:49 PM
+ */
+#include <sstream>
+
+#include "DataSet.h"
+
+namespace ScaleData{
+  DataSet::DataSet() : AbstractDataNode() {
+  }
+
+  DataSet::DataSet(const DataSet& orig) : AbstractDataNode(orig){
+
+      for( DataMap::const_iterator i = orig.data.begin(), ie = orig.data.end(); i != ie; i++ ) {          
+          data.insert(std::make_pair(i->first,i->second->getCopy()));
+      }
+  }
+
+  DataSet::~DataSet() {
+      DataMap::const_iterator iter;
+      //
+      // Loop over map deleting infrastructure nodes
+      //
+      for (iter = data.begin(); iter != data.end(); iter++) {
+        delete iter->second;
+      }
+//      data.clear();
+  }
+
+  size_t DataSet::size() const {
+    return data.size();
+  }
+
+  std::vector<std::string> DataSet::keys() const {
+    std::vector<std::string> key_list(data.size());
+    DataMap::const_iterator iter;
+    //
+    // Loop over map compiling list of keys
+    //
+    int i = 0;
+    for (iter = data.begin(); iter != data.end(); iter++) {
+      key_list[i++] = iter->first;
+    }
+    return key_list;
+  }//keys
+
+  bool DataSet::containsKey(const std::string& key) const {
+    DataMap::const_iterator iter = data.find(key);
+    if (iter == data.end()) return false;
+    return true;
+  }//containsKey
+
+  AbstractDataNode* DataSet::getCopy() const {
+      return new DataSet( *this );
+  }
+
+  std::string DataSet::toString() const{
+      std::stringstream stream;
+      stream <<"DataSet: "<<std::endl;
+      DataMap::const_iterator iter;
+      //
+      // Loop over map
+      //
+      for (iter = data.begin(); iter != data.end(); iter++) {
+          stream<<"'"<<iter->first<<"':"<<iter->second->toString()<<std::endl;
+      }
+      return stream.str();
+  }
+  DataSet& DataSet::operator =(const DataSet& orig)
+  {
+      if (this != &orig)
+      { // self-assignment check expected
+
+          for( DataMap::const_iterator i = orig.data.begin(), ie = orig.data.end();
+               i != ie;
+               i++ )
+          {
[…]

### 2016-09-23 exnih/DataSet.h

@@ -1,0 +1,126 @@
+/*
+ * @file:   DataSet.h
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov. Robert A. Lefebvre, lefebvrera@ornl.gov
+ *
+ * Created on November 13, 2012, 10:19 AM
+ */
+
+#ifndef DATASET_H
+#define DATASET_H
+#include <string>
+#include <map>
+#include <vector>
+#include <typeinfo>
+#include "AbstractDataNode.h"
+
+namespace ScaleData {
+
+  /**
+   * @class DataSet
+   * @brief  data set for dynamic type storage and relations
+   */
+  class DataSet : public AbstractDataNode {
+  private:    
+    typedef std::map<std::string, AbstractDataNode*> DataMap;
+    /**
+     * std::map<string,AbstractDataSet*>
+     */
+    DataMap data;
+  public:
+    /**
+     * Constructor
+     */
+    DataSet();
+    /**
+     * Copy constructor
+     * @param orig const DataSet& orig
+     */
+    DataSet(const DataSet& orig);
+    /**
+     * Destructor
+     */
+    virtual ~DataSet();
+    /**
+     * Determine if this dataset is empty
+     * @return true, if and only if the dataset contains data, false otherwise
+     */
+    bool empty()const{return data.empty();}
+    /**
+     * Get the value of this type
+     */
+    template<class T>
+    const T& get(const std::string& key) const;
+    template<class T>
+    T& get(const std::string& key);
+    /**
+     * Get the reference to value of this type
+     */
+    template<class T>
+    T& operator[](const std::string& key);
+    /**
+     * Get the value of this type with a default value (if there is no value in the dataset)
+     */
+    template<class T>
+    T get(const std::string& key, T default_value) const;
+    /**
+     * Set the key,value
+     * @param key std::string
+     * @param value template<T>
+     */
+    template<class T>
+    void put(const std::string& key, T value);
+    /**
+     * Set the key,value as a reference
+     * @param key std::string
+     * @param value template<T>
+     */
+    template<class T>
+    void putRef(const std::string& key, T&value);
+    /**
+     * does this data set contain key value
+     * @param key std::string
+     * @return true if the data set contains it as is of type T, false otherwise
+     */
+    template<class T>
+    bool contains(const std::string& key) const;
+
+    /**
+     * does this data set contain key
+     * @param key std::string
+     * @return true if the data set contains it, false otherwise
[…]

### 2016-09-23 exnih/DataSet.i.h

@@ -1,0 +1,92 @@
+/*
+ * @file:   DataSet.i.h
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov. Robert A. Lefebvre, lefebvrera@ornl.gov
+ *
+ * Created on November 13, 2012, 10:19 AM
+ */
+
+#ifndef DATASET_I_H
+#define DATASET_I_H
+
+//#include "DataSet.h"
+#include "DataNode.h"
+#include <iostream>
+#include <stdexcept>
+
+namespace ScaleData {
+
+  template<class T>
+  const T& DataSet::get(const std::string& key) const {
+    DataMap::const_iterator iter = data.find(key);
+    if (iter == data.end()) throw std::runtime_error("***Error: DataSet::get("+key+") has no results!");
+    DataNode<T>* dataNode = dynamic_cast<DataNode<T>*> (iter->second);
+    return dataNode->get();
+  }//get
+  template<class T>
+  T& DataSet::get(const std::string& key){
+    DataMap::iterator iter = data.find(key);
+    if (iter == data.end()) throw std::runtime_error("***Error: DataSet::get("+key+") has no results!");
+    DataNode<T>* dataNode = dynamic_cast<DataNode<T>*> (iter->second);
+    return dataNode->getRef();
+  }//get
+  template<class T>
+  T& DataSet::operator[](const std::string& key) {
+    DataMap::iterator iter = data.find(key);
+    if (iter == data.end()) throw "No results!";
+    DataNode<T>* dataNode = dynamic_cast<DataNode<T>*> (iter->second);
+    return dataNode->getRef();
+  }//get
+
+  template<class T>
+  T DataSet::get(const std::string& key, T default_value) const {
+    if (contains<T>(key)) {
+      return get<T>(key);
+    } else {
+      return default_value;
+    }
+  }//get
+
+  template<class T>
+  void DataSet::put(const std::string& key, T value) {
+    //
+    // Check if key,value already exists
+    //
+    DataMap::iterator iter = data.find(key);
+    if (iter != data.end()) {
+      //
+      // Delete previous node
+      //
+      delete iter->second;
+    }
+    AbstractDataNode* dataNode = new DataNode<T> (value);
+    data[key] = dataNode;
+  }//put
+ 
+  template<class T>
+  bool DataSet::contains(const std::string& key) const {
+    DataMap::const_iterator iter = data.find(key);
+    if (iter == data.end()) return false;
[…]

### 2016-09-23 exnih/EndianUtils.cpp

@@ -1,0 +1,139 @@
+/*!
+ ********************************************************
+ * @author: Jordan P. Lefebvre
+ * @brief: Endianess utilities for reversing bytes
+ **********************************************************/
+#include <stdio.h>
+#include <cstdlib>
+#include <stdint.h>
+
+namespace EndianUtils {
+
+void reverse_bytes( char * original, void * returnVal, int length )
+{
+    char *reversed = (char*) returnVal;
+    switch(length) {
+    case 2:
+        reversed[0]=original[1];
+        reversed[1]=original[0];
+        return;
+    case 4:
+        reversed[0]=original[3];
+        reversed[1]=original[2];
+        reversed[2]=original[1];
+        reversed[3]=original[0];
+        return;
+    case 8:
+        reversed[0]=original[7];
+        reversed[1]=original[6];
+        reversed[2]=original[5];
+        reversed[3]=original[4];
+        reversed[4]=original[3];
+        reversed[5]=original[2];
+        reversed[6]=original[1];
+        reversed[7]=original[0];
+        return;
+    }
+
+}//reverse_bytes
+
+/*!
+ * Reverses the bytes of an integer
+ * */
+int reverse_int_bytes(int value)
+{
+    int returnVal;
+    reverse_bytes((char*)&value, (void*)&returnVal, sizeof(int));
+    return returnVal;
+}//reverse_int_bytes
+
+/*!
+ * Reverses the bytes of all integers in an array
+ * */
+void reverse_array_int_bytes(int * array, int size)
+{
+    for(int i = 0; i < size; i++ ) {
+        array[i] = reverse_int_bytes(array[i]);
+    }
+}//reverse_array_int_bytes
+
+/*!
+* Reverses the bytes of a long
+ * */
+long reverse_long_bytes(long value)
+{
+    long returnVal;
+    reverse_bytes((char*)&value, (void*)&returnVal, sizeof(long));
+    return returnVal;
+}//reverse_long_bytes
+
+/*!
+* Reverses the bytes of a float
+ * */
+float reverse_float_bytes(float value)
+{
+    float returnVal;
+    reverse_bytes((char*)&value, (void*)&returnVal, sizeof(float));
+    return returnVal;
+}//reverse_float_bytes
+
+/*!
+ * Reverses the bytes of all float in an array
+ * */
+void reverse_array_float_bytes(float * array, int size)
+{
+    for(int i = 0; i < size; i++ ) {
[…]

### 2016-09-23 exnih/EndianUtils.h

@@ -1,0 +1,18 @@
+/*********************************************************
+ * Author: Jordan P. Lefebvre
+ * Description: Endianess utilities for reversing bytes
+ **********************************************************/
+
+namespace EndianUtils{
+
+	void reverse_bytes( char * original, void * returnVal, int length );
+	int reverse_int_bytes(int value);
+	void reverse_array_int_bytes(int * array, int size);
+	void reverse_array_float_bytes(float * array, int size);
+	long reverse_long_bytes(long value);
+	float reverse_float_bytes(float value);
+	double reverse_double_bytes(double value);
+	void reverse_array_double_bytes(double * array, int size);
+    bool system_is_big_endian();
+    bool file_is_big_endian(bool file_is_native_endian);
+}//end namespace

### 2016-09-23 exnih/EnergyBounds.h

@@ -1,0 +1,62 @@
+/*
+ * @file:   EnergyBounds.h
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov
+ *
+ * Created on August 22, 2013, 7:37 AM
+ */
+
+#ifndef ScaleData_ENERGYBOUNDS_H
+#define	ScaleData_ENERGYBOUNDS_H
+#include <vector>
+
+namespace ScaleData {
+
+    /**
+     * @class EnergyBounds
+     * @brief Basic class which wraps energy bounds with default type=float
+     */
+    template<class T = float>
+    class EnergyBounds {
+    public:
+        /**
+         * Default constructor
+         */
+        EnergyBounds();
+        /**
+         * Copy constructor
+         * @param orig <b>const EnergyBounds&</b>
+         */
+        EnergyBounds(const EnergyBounds& orig);
+        /**
+         * Destructor
+         */
+        virtual ~EnergyBounds();
+        /**
+         * Returns the number of energy groups
+         * @return <b>size_t</b>
+         */
+        size_t getSize() const;
+        /**
+         * @brief Get pointer to the bounds.
+         * @return <b>T*</b>
+         */
+        const T* getBounds() const;
+        /**
+         * @brief set the energy bounds
+         * @param new_bounds <b>const std::vector&lt;T&gt;</b>
+         */
+        void setBounds(const std::vector<T> new_bounds);
+    protected:
+        std::vector<T> bounds;
+    private:
+
+    };//class EnergyBounds
+
+}//namespace ScaleData
+
+//
+// Templated code include
+//
+#include "EnergyBounds.i.h"
+#endif	/* ScaleData_ENERGYBOUNDS_H */
+

### 2016-09-23 exnih/EnergyBounds.i.h

@@ -1,0 +1,46 @@
+/*
+ * @file:   EnergyBounds.cpp
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov
+ *
+ * Created on August 22, 2013, 7:37 AM
+ */
+#include <algorithm>
+
+#include "EnergyBounds.h"
+namespace ScaleData {
+
+    template<class T>
+    EnergyBounds<T>::EnergyBounds() {
+    }
+
+    template<class T>
+    EnergyBounds<T>::EnergyBounds(const EnergyBounds& orig) : bounds(orig.bounds) {
+    }
+
+    template<class T>
+    EnergyBounds<T>::~EnergyBounds() {
+    }
+
+    template<class T>
+    size_t EnergyBounds<T>::getSize() const {
+        return bounds.size();
+    }//getSize
+
+    template<class T>
+    const T* EnergyBounds<T>::getBounds() const {
+        return &(bounds[0]);
+    }//getBounds;
+
+    template<class T>
+    void EnergyBounds<T>::setBounds(const std::vector<T> new_bounds) {
+        //
+        // Resize my bounds according to new bounds
+        //
+        this->bounds.resize(new_bounds.size());
+        //
+        // Copy new bounds
+        //
+        std::copy(new_bounds.begin(), new_bounds.end(), bounds.begin());
+    }//setBounds
+}//namespace ScaleData
+

### 2016-09-23 exnih/FileStream.cpp

@@ -1,0 +1,203 @@
+/*
+ * File:   FileStream.cpp
+ * Author: raq
+ *
+ * Created on November 12, 2011, 2:00 PM
+ */
+
+//#include "Standard/Interface/Serializable.h"
+//#include "Standard/Interface/SerialFactory.h"
+#include "FileStream.h"
+#include <QDebug>
+
+//#include <Standard/Interface/standard_interface_config.h>
+
+FileStream::FileStream() {
+    stream = NULL;
+    //factory = NULL;
+    ownsStream = true;
+}
+FileStream::FileStream(std::fstream * stream) {
+    this->stream = stream;
+    ownsStream = false;
+    //factory = NULL;
+}
+
+
+FileStream::~FileStream() {
+    if( ownsStream ){
+        if( stream != NULL ){
+            if( stream->is_open() ){
+                stream->close();
+            }
+            delete stream;
+        }
+    }
+}
+
+/**
+ * @brief Open the FileStream with the given file path
+ * @param const char * file - the file path from which to open this stream
+ * @param std::ios_base::openmode mode - indicates how to open the fstream
+ * @return bool - true, if the file was successfully opened, false otherwise
+ */
+bool FileStream::open(const char * file, std::ios_base::openmode mode) {
+    qDebug() << "opening "<<file;
+    if( file == NULL ){
+        return false;
+    }
+    if( stream != NULL ){
+        delete stream;
+        stream = NULL;
+    }
+    stream = new std::fstream(file, mode);
+
+    if( !stream->is_open() || stream->fail() ){
+        qDebug() << "Is open "<<stream->is_open()<<", has failed "<<stream->fail();
+        return false;
+    }
+    ownsStream = true;
+    // get file size
+    stream->seekg(0, std::ios::end);
+    file_size = stream->tellg();
+    stream->seekg(0, std::ios::beg);
+
+    return true;
+}
+
+/**
+ * @brief Close the FileStream
+ */
+void FileStream::close() {
+    if( stream != NULL){
+        if( stream->is_open() ){
+            stream->close();
+        }
+    }
+}
+
+/**
+ * @brief Obtain the size of this stream
+ */
+long FileStream::getSize() {
+    long writeHead = stream->tellp();
+    if( writeHead > file_size ) file_size = writeHead;
+    return file_size;
+}
+
+/**
+ * @brief Write the given size byte block into this SerialStream
+ * Upon completion of this call, the write head will be position size bytes
[…]

### 2016-09-23 exnih/FileStream.h

@@ -1,0 +1,128 @@
+/*
+ * File:   FileStream.h
+ * Author: raq
+ *
+ * Created on November 12, 2011, 2:00 PM
+ */
+
+#ifndef FILESTREAM_H
+#define	FILESTREAM_H
+
+#include <iostream>
+#include <fstream>
+
+//#include "Standard/Interface/AbstractStream.h"
+
+namespace Standard {
+   class SerialFactory;
+   class Serializable;
+}
+
+class FileStream {
+public:
+    FileStream();
+    FileStream(std::fstream * stream);
+    virtual ~FileStream();
+private:
+    //Standard::SerialFactory * factory;
+    std::fstream * stream;
+    long file_size;
+    bool ownsStream;
+public:
+
+    /**
+     * @brief Open the FileStream with the given file path
+     * @param const car * file - the file path from which to open this stream
+     * @param std::ios_base::openmode mode - indicates how to open the fstream
+     * @return bool - true, if the file was successfully opened, false otherwise
+     */
+    virtual bool open(const char * file, std::ios_base::openmode mode = (std::ios_base::in | std::ios_base::out | std::ios_base::binary | std::ios_base::trunc) );
+
+    /**
+     * @brief Close the FileStream
+     */
+    virtual void close();
+
+    /**
+     * @brief Obtain the size of this stream
+     */
+    virtual long getSize();
+
+    /**
+     * @brief Write the given size byte block into this SerialStream
+     * Upon completion of this call, the write head will be position size bytes
+     * forward, immediately following the data block just written
+     * @param char * data - the memory block containing the data to be written
+     * @param long size - the size of the memory block in bytes
+     */
+    virtual void write(const char * data, long size);
+    /**
+     * @brief Read into the given byte block size bytes
+     * Upon complete of this call, the read head will have been moved size bytes
+     * forward.
+     * @param char * data - the memory block to which the data will be written
+     * @param long size - the size of the memory block in bytes
+     */
+    virtual void read(char* data, long size);
+
+    /**
+     * @brief Obtain the universal version identifier located under the read head
+     * This will not increment the read head, but provides convenience for
+     * objects to conduct a look-ahead into the buffer to determine
+     * what type of object was serialized, and if they are that type of object
+     * they know how to inflate the given object under the read head
+     * @return long - the UID of the next object in the stream, -1 if end of stream
+     */
+    virtual long getNextUID();
+
+    /**
+     * @brief ignores count bytes
[…]

### 2016-09-23 exnih/Filter.cpp

@@ -1,0 +1,17 @@
+/*
+ * @file:   Filter.cpp
+ * @author: Jordan P. Lefebvre, lefebvrejp@ornl.gov
+ *
+ * Created on September 10, 2013, 8:22 AM
+ */
+#include "Filter.h"
+
+namespace ScaleData {
+
+    Filter::Filter(Filter::Policy mpolicy) : policy(mpolicy) {
+    }
+
+    Filter::Filter(const Filter& filter) : data(filter.data), policy(filter.policy) {
+    }
+
+}//namespace ScaleData

### 2016-09-23 exnih/Filter.h

@@ -1,0 +1,83 @@
+#ifndef CORE_FILTER_H
+#define CORE_FILTER_H
+/**
+ * @Author : Robert A. Lefebvre
+ * @Date   : 04-16-2013
+ * Filter class to allow for object component filtering.
+ */
+
+#include <string>
+#include <stdexcept>
+#include "DataSet.h"
+
+namespace ScaleData{
+/**
+ * Filter class provides basic constructs for allowing sets of data to be
+ * kept or ignored.
+ * For use of this class see tests/ExampleFilter.h.
+ *
+ * Typical use will include writing a subclass and providing domain specific
+ * filter constructs as seen in the tests/ExampleFilter.h class
+ */
+class Filter{
+
+public:
+
+    /**
+     * How do we handle objects
+     */
+    enum Policy{
+        Keep,
+        Ignore
+    };
+private:
+    /**
+     * The generic data container for storing filter descriptions
+     */
+    DataSet data;
+    /**
+     * Policy for dealing with unspecified components
+     */
+    Policy policy;
+
+public:
+    /**
+     * Construct a filter object
+     */
+    Filter(Policy mpolicy=Keep);
+    Filter(const Filter& filter);
+    virtual ~Filter(){};
+
+    /**
+     * Determine if this filters a certain given named component
+     * @param key - the key/name of the component
+     * @param class T - the type of the component that the key maps to
+     * @return bool - true, if and only if, the component with the given key and type T are present
+     */
+    template<class T>
+    bool contains(const std::string& key) const;
+
+    /**
+     * Obtain the object describing what to filter for the given key
+     * @param key - the name of the object describing what to filter
+     * @param class T - the type of the object describing what to filter
+     * @return T - the object describing what to filter
+     * @Note - if the object T is not contained, a std::runtime_error is thrown
+     */
+    template<class T>
+    T  get(const std::string& key) const;
+
+    /**
+     * Set the description data structure for the given key filter
+     * @param key - the key/name of the filter
+     * @param filterDescription - the data/ data structure describing what to filter
+     */
+    template<class T>
+    void set(const std::string& key, T filterDescription);
+
+    Policy getGlobalPolicy()const { return policy;}
+
+}; // end of class Filter
+} // end of namespace
+#include "Filter.i.h"
+#endif // end of include gaurd

### 2016-09-23 exnih/Filter.i.h

@@ -1,0 +1,18 @@
+namespace ScaleData {
+
+    template<class T>
+    bool Filter::contains(const std::string& key) const {
+        return data.contains<T>(key);
+    } // end of contains
+
+    template<class T>
+    T Filter::get(const std::string& key) const {
+        if (!data.contains<T>(key)) throw std::runtime_error("Filter does not contain the appropriately typed description of '" + key + "'");
+        return data.get<T>(key);
+    } // end of get
+
+    template<class T>
+    void Filter::set(const std::string& key, T filterDescription) {
+        data.put(key, filterDescription);
+    }
+}//namespace ScaleData

### 2016-09-23 exnih/LibraryEnergyBounds.cpp

@@ -1,0 +1,41 @@
+//#include "Resource/AmpxLib/ampxlib_config.h"
+//#include "Standard/Interface/AbstractStream.h"
+#include "LibraryEnergyBounds.h"
+#include "resources.h"
+LibraryEnergyBounds::LibraryEnergyBounds()
+{
+    initialize();
+}
+void LibraryEnergyBounds::initialize()
+{
+}
+
+LibraryEnergyBounds::~LibraryEnergyBounds()
+{
+}
+LibraryEnergyBounds * LibraryEnergyBounds::getCopy() const
+{
+    LibraryEnergyBounds * copy = new LibraryEnergyBounds();
+    if( this->getBoundsSize() > 0 ) copy->setBounds(getBounds(), getBoundsSize());
+    if( this->getLethargySize() > 0 ) copy->setLethargyBounds(getLethargyBounds(), getLethargySize());
+
+    return copy;
+}
+bool LibraryEnergyBounds::operator==(LibraryEnergyBounds & a){
+    if( bounds.size() != a.bounds.size() ) return false;
+
+    for( size_t i = 0; i < bounds.size(); i++){
+        if( bounds[i] != a.bounds[i] ) return false;
+    }
+
+    if( lethargyBounds.size() != a.lethargyBounds.size() ) return false;
+
+    for( size_t i = 0; i < lethargyBounds.size(); i++){
+        if( lethargyBounds[i] != a.lethargyBounds[i] ) return false;
+    }
+    return true;
+}
+// serializable interface
+const long LibraryEnergyBounds::uid = 0x931523456f0bdc6a;
+
+

### 2016-09-23 exnih/LibraryEnergyBounds.h

@@ -1,0 +1,75 @@
+#ifndef LIBRARYENERGYBOUNDS_H
+#define LIBRARYENERGYBOUNDS_H
+#include <iostream>
+#include <fstream>
+#include <vector>
+//#include "Standard/Interface/Serializable.h"
+#include "EnergyBounds.h"
+#include "LibraryItem.h"
+using namespace std;
+
+class LibraryEnergyBounds : public ScaleData::EnergyBounds<float>,
+public LibraryItem { //, public Standard::Serializable {
+private:
+    vector<float> lethargyBounds;
+    void initialize();
+public:
+    LibraryEnergyBounds();
+    ~LibraryEnergyBounds();
+
+    bool operator==(LibraryEnergyBounds & a);
+
+    LibraryEnergyBounds * getCopy() const;
+
+    void resizeBounds(int size) {
+        if (size <= 0) return;
+        bounds.resize(size);
+    }
+
+    void resizeLethargy(int size) {
+        if (size <= 0) return;
+        lethargyBounds.resize(size);
+    }
+
+    float * getBounds() {
+        return &(bounds[0]);
+    }
+
+    const float * getBounds() const {
+        return &(bounds[0]);
+    }
+
+    void setBounds(const float * bounds, int size) {
+        this->bounds.resize(size);
+        for (int i = 0; i < size; i++)this->bounds[i] = bounds[i];
+    }
+
+    int getBoundsSize() const {
+        return bounds.size();
+    }
+
+    float * getLethargyBounds() {
+        return &(lethargyBounds[0]);
+    }
+
+    const float * getLethargyBounds() const {
+        return &(lethargyBounds[0]);
+    }
+
+    void setLethargyBounds(const float * bounds, int size) {
+        lethargyBounds.resize(size);
+        for (int i = 0; i < size; i++)lethargyBounds[i] = bounds[i];
+    }
+
+    int getLethargySize() const {
+        return lethargyBounds.size();
+    }
+
+
+public:
+    /**
+     * @brief the universal version identifier for this class
+     */
+    static const long uid;
+};
+#endif

### 2016-09-23 exnih/LibraryHeader.cpp

@@ -1,0 +1,44 @@
+//#include "Resource/AmpxLib/ampxlib_config.h"
+#include "LibraryHeader.h"
+#include <stdio.h>
+#include <string.h>
+#include <QDebug>
+#include <QString>
+//#include "Standard/Interface/AbstractStream.h"
+LibraryHeader::LibraryHeader()
+{
+    initialize();
+}
+
+void LibraryHeader::initialize()
+{
+     idtape = 0;
+     nnuc = 0;
+     igm = 0;
+     iftg = 0;
+     msn = 0;
+     ipm = 0;
+     i1 = 0;
+     i2 = 0;
+     i3 = 0;
+     i4 = 0;
+     strcpy(itm,"");
+}
+
+bool LibraryHeader::operator==(LibraryHeader & a){
+    if( QString(itm).trimmed() !=  QString(a.itm).trimmed() ) return false;
+    if( idtape != a.idtape )return false;
+    if( nnuc != a.nnuc ) return false;
+    if( igm != a.igm ) return false;
+    if( iftg != a.iftg ) return false;
+    if( msn != a.msn ) return false;
+    if( ipm != a.ipm ) return false;
+    if( i1 != a.i1 ) return false;
+    if( i2 != a.i2 ) return false;
+    if( i3 != a.i3 ) return false;
+    if( i4 != a.i4 ) return false;
+    return true;
+}
+// serializable interface
+const long LibraryHeader::uid = 0x39513254f6b0dc7e;
+

### 2016-09-23 exnih/LibraryHeader.h

@@ -1,0 +1,99 @@
+#ifndef LIBRARY_HEADER_H
+#define LIBRARY_HEADER_H
+#include <QString>
+#include <string.h>
+//#include "Standard/Interface/Serializable.h"
+#include "LibraryItem.h"
+
+using namespace std;
+
+#define AMPX_HEADER_TITLE_LENGTH (4*100)
+#define AMPX_HEADER_NUMBER_INTEGER_OPTIONS 10
+
+class LibraryHeader : public LibraryItem{
+
+private:
+  int idtape;    /// identification number of tape
+  int nnuc;     /// number of data sets on the tape
+  int igm;       /// number of neutron groups
+  int iftg;      /// the first thermal neutron group on the library
+  int msn;       /// 0 - pre-NITWAL-III library 1 - library for use with NITWAL-II        I
+  int ipm;       /// number of gamma groups
+  int i1;        /// zero
+  int i2;        /// working library flag
+  int i3;        /// zero
+  int i4;        /// zero
+  char itm[AMPX_HEADER_TITLE_LENGTH+1]; /// title
+
+  void initialize();
+public:
+  LibraryHeader();
+
+  bool operator==(LibraryHeader & a);
+
+  LibraryHeader * getCopy() const{
+      LibraryHeader * copy = new LibraryHeader();
+      copy->idtape = idtape;
+      copy->nnuc = nnuc;
+      copy->igm = igm;
+      copy->iftg = iftg;
+      copy->msn = msn;
+      copy->ipm = ipm;
+      copy->i1 = i1;
+      copy->i2 = i2;
+      copy->i3 = i3;
+      copy->i4 = i4;
+      strncpy(copy->itm, itm, AMPX_HEADER_TITLE_LENGTH);
+      return copy;
+  }
+
+  int getIdTape(){return idtape;}
+  void setIdTape(int idtape){this->idtape = idtape;}
+  int getNNuc(){return nnuc;}
+  void setNNuc(int nnuc){this->nnuc = nnuc;}
+  int getIGM(){return igm;}
+  void setIGM(int igm){this->igm = igm;}
+  int getIFTG(){return iftg;}
+  void setIFTG(int iftg){this->iftg = iftg;}
+  int getMSN(){return msn;}
+  void setMSN(int msn){this->msn = msn;}
+  int getIPM(){return ipm;}
+  void setIPM(int ipm){this->ipm = ipm;}
+  int getI1(){return i1;}
+  void setI1(int i1){this->i1 = i1;}
+  int getI2(){return i2;}
[…]

### 2016-09-23 exnih/LibraryItem.cpp

@@ -1,0 +1,15 @@
+#include "LibraryItem.h"
+LibraryItem::LibraryItem()
+{
+    initialize();
+} // end of constructor
+
+void LibraryItem::initialize()
+{
+}// end of initiliaze
+
+LibraryItem * LibraryItem::getCopy() const
+{
+    LibraryItem * copy = new LibraryItem();
+    return copy;
+}// end of getCopy

### 2016-09-23 exnih/LibraryItem.h

@@ -1,0 +1,23 @@
+#ifndef LIBRARY_ITEM_H
+#define LIBRARY_ITEM_H
+#include <fstream>
+#include <iostream>
+using namespace std;
+
+/// This class references a library item
+/// and assisted with keeping track of where
+/// and if a library item (nuclideInfo, header, energybound, etc)
+/// has been read.
+class LibraryItem {
+    void initialize();
+public:
+    LibraryItem();
+
+    // virtual destructor.
+    virtual ~LibraryItem() {}
+
+    /// Obtain a copy of the libraryItem
+    /// return LibraryItem * - The copy
+    virtual LibraryItem * getCopy() const;
+};
+#endif

### 2016-09-23 exnih/LibraryNuclide.cpp

@@ -1,0 +1,869 @@
+//#include "Standard/Interface/AbstractStream.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+#include "LibraryNuclide.h"
+#include "BondarenkoGlobal.h"
+#include "NuclideResonance.h"
+#include "AmpxLibrary.h"
+#include <string>
+#include <QDebug>
+#include "AmpxDataHelper.h"
+
+LibraryNuclide::LibraryNuclide() {
+    initialize();
+}
+
+LibraryNuclide::LibraryNuclide( const LibraryNuclide & orig, NuclideFilter * f){
+    d = orig.d;
+    d->ref.ref();
+    bondarenkoGlobalData = NULL;
+    subGroups = NULL;
+    if( orig.subGroups  != NULL) subGroups = orig.subGroups->getCopy();
+
+
+    setResonance(NULL);
+    setTotal2dXS(NULL);
+    NuclideFilter filter(f!=NULL ? *f : NuclideFilter(NuclideFilter::Keep)); // default global policy is Keep
+    AmpxLibrary * library = filter.getAmpxLib();
+    qDebug() << qPrintable("Nuclide ")<<getId()<<" filtered copy with library at address ";
+    if (orig.totalXS != NULL) {
+        //qDebug()<<"Copying total xs...";
+        //if( filter.acceptsTotal() )
+        this->setTotal2dXS(orig.totalXS->getCopy());
+    }
+    if( orig.resonance != NULL){
+        this->resonance = new NuclideResonance(*orig.resonance);
+    }
+    this->setNumBondarenkoSig0Data(0);
+    this->setNumBondarenkoTempData(0);
+    this->setMaxNumBondGrp(0);
+    this->setNumBondarenkoData(0);
+    if (orig.bondarenkoGlobalData != NULL)
+        for (int i = 0; i < orig.bondarenkoData.size(); i++){
+            if( filter.acceptsBondarenko(orig.bondarenkoData.at(i)->getMt() ) ) {
+                this->bondarenkoGlobalData = orig.bondarenkoGlobalData->getCopy();
+                this->setNumBondarenkoSig0Data(const_cast<LibraryNuclide*>(&orig)->getNumBondarenkoSig0Data());
+                this->setNumBondarenkoTempData(const_cast<LibraryNuclide*>(&orig)->getNumBondarenkoTempData());
[…]

### 2016-09-23 exnih/LibraryNuclide.h

@@ -1,0 +1,786 @@
+#ifndef LIBRARYNUCLIDE_H
+#define LIBRARYNUCLIDE_H
+
+#include <QAtomicInt>
+#include <QString>
+#include <QList>
+#include <QDebug>
+#include <string.h>
+//#include "Standard/Interface/Serializable.h"
+
+#include "LibraryItem.h"
+#include "resources.h"
+#include "BondarenkoData.h"
+#include "BondarenkoGlobal.h"
+#include "CrossSection2d.h"
+#include "CrossSection1d.h"
+#include "NuclideFilter.h"
+#include "SubGroupData.h"
+#include "LibrarySourceDefs.h"
+
+using namespace std;
+
+#define AMPX_NUCLIDE_INTEGER_OPTION_START 19
+#define AMPX_NUMBER_NUCLIDE_INTEGER_OPTIONS  32
+#define AMPX_NUCLIDE_DESCR_LENGTH (4*18)
+#define AMPX_MAST_NUCLIDE_ID   0
+#define AMPX_NUCLIDE_WORD_19   0
+#define AMPX_MAST_NUCLIDE_NUM_RESOLVED_RESDATA   1
+#define AMPX_WORK_NUCLIDE_SET_DERIVED_FROM 1
+#define AMPX_NUCLIDE_WORD_20   1
+#define AMPX_MAST_NUCLIDE_NUM_ENERGIES_UNRESOLVED_RESDATA   2
+#define AMPX_WORK_NUCLIDE_ZONE_NUMBER 2
+#define AMPX_NUCLIDE_WORD_21   2
+#define AMPX_MAST_NUCLIDE_NUM_NEU_1D_DATA   3
+#define AMPX_WORK_NUCLIDE_NUM_ZONES_FROM_PRODUCER_PROBLEM 3
+#define AMPX_NUCLIDE_WORD_22   3
+#define AMPX_MAST_NUCLIDE_NUM_NEU_2D_DATA   4
+#define AMPX_WORK_NUCLIDE_P0_TOTAL_LENGTH 4
+#define AMPX_NUCLIDE_WORD_23   4
+#define AMPX_WORK_NUCLIDE_TOTAL_ORDER_OF_EXPANSION 5
+#define AMPX_NUCLIDE_WORD_24   5
[…]

### 2016-09-23 exnih/LibrarySourceDefs.h

@@ -1,0 +1,74 @@
+
+
+#ifndef LIBRARYSOURCEDEFS_H
+#define	LIBRARYSOURCEDEFS_H
+
+#include <string>
+
+namespace ScaleData {
+class LibrarySourceDefs {
+public:
+    LibrarySourceDefs(){}
+    LibrarySourceDefs(const LibrarySourceDefs& ){}
+    virtual ~LibrarySourceDefs(){}
+    
+    static void getLibrarySourceAsString(int source, std::string & name) {
+            switch (source) {
+                case 0:
+                case 1:
+                case 4:
+                    name.resize(0);
+                    name.append("endf");                  
+                    break;             
+                case 2:
+                case 3:
+                    name.resize(0);
+                    name.append("jeff");                    
+                    break;
+                case 5:
+                    name.resize(0);
+                    name.append("cendl");                      
+                    break;
+                case 6:
+                    name.resize(0);
+                    name.append("jendl");                       
+                    break;
+                case 21:
+                    name.resize(0);
+                    name.append("sg-33");                       
+                    break;
+                case 31:                  
+                case 32:
+                    name.resize(0);
+                    name.append("indl");                      
+                    break;
+                case 33:
+                case 37:
+                    name.resize(0);
+                    name.append("fendl");                      
+                    break;
+                case 34:
+                    name.resize(0);
+                    name.append("irdf");                     
+                    break;
+                case 35:
+                case 41:
+                    name.resize(0);
+                    name.append("brond");                      
+                    break;
+                case 36:
+                    name.resize(0);
+                    name.append("ingdb");                       
+                    break;
+                default:
+                    name.resize(0);
+                    name.append("unkown");                         
+                    break;
+            }
+        }
+    private:
+
+};
+}
+#endif	/* LIBRARYSOURCEDEFS_H */
+

### 2016-09-23 exnih/NuclideFilter.cpp

@@ -1,0 +1,15 @@
+#include "NuclideFilter.h"
+#include <string>
+const std::string NuclideFilter::BondarenkoMtListKey = "BondarenkoMts";
+const std::string NuclideFilter::BondarenkoMtListPolicyKey = "BondarenkoMtsPolicy";
+const std::string NuclideFilter::Neutron1dMtListKey = "Neutron1dMts";
+const std::string NuclideFilter::Neutron1dMtListPolicyKey = "Neutron1dMtsPolicy";
+const std::string NuclideFilter::Neutron2dMtListKey = "Neutron2dMts";
+const std::string NuclideFilter::Neutron2dMtListPolicyKey = "Neutron2dMtsPolicy";
+const std::string NuclideFilter::Gamma1dMtListKey = "Gamma1dMts";
+const std::string NuclideFilter::Gamma1dMtListPolicyKey = "Gamma1dMtsPolicy";
+const std::string NuclideFilter::Gamma2dMtListKey = "Gamma2dMts";
+const std::string NuclideFilter::Gamma2dMtListPolicyKey = "Gamma2dMtsPolicy";
+const std::string NuclideFilter::GammaProdMtListKey = "GammaProdMts";
+const std::string NuclideFilter::GammaProdMtListPolicyKey = "GammaProdMtsPolicy";
+const std::string NuclideFilter::AmpxHelperLibKey= "AmpxHelperLib";

### 2016-09-23 exnih/NuclideFilter.h

@@ -1,0 +1,216 @@
+#ifndef AMPXLIB_NUCLIDE_FILTER_H
+#define AMPXLIB_NUCLIDE_FILTER_H
+
+#include "Filter.h"
+#include <algorithm>
+class AmpxLibrary;
+
+/**
+ * NuclideFilter - Intended to allow for filtering nuclide components
+ */
+class NuclideFilter : public ScaleData::Filter
+{
+public:
+    /**
+     * TypeDef the types of data structures that describe our filter
+     */
+    typedef std::vector<int> MtList;
+
+    /**
+     * TypeDef for convenience the policy
+     */
+    typedef ScaleData::Filter::Policy Policy;
+private:
+    /**
+     * Convenience method to set a filter description with policy 
+     * and including the data key and policy key. The data key (aKey) and 
+     * policy key (pKey) cannot be the same
+     */
+    template<class T> 
+    void set(T& a, const Policy& policy, const std::string& aKey, const std::string& pKey)
+    {
+        std::sort(a.begin(), a.end()); // sort it for future binary searches
+        Filter::set(aKey, a); // set the filter description
+        Filter::set(pKey, policy); // set the filter policy
+        
+    }
+    template<class C,class I> 
+    bool accepts(const I& a, const std::string& aKey, const std::string& pKey) const
+    {
+        Policy policy = Filter::getGlobalPolicy();
+        // Check for key specific/local policy
+        if( Filter::contains<Policy>(pKey) ){
+            policy = Filter::get<Policy>(pKey);
+        }
+        // if specific ids are not specified, produce result based on provided policy
+        if( !Filter::contains<C>(aKey) ){
+            return policy == Keep;
+        }
+        // acquire ids
+        const C & ids = Filter::get<C>(aKey);
+        // conduct a binary search for object id
+        bool listed = std::binary_search(ids.begin(), ids.end(), a);
+
+        return (listed && policy == Keep ) || (!listed && policy==Ignore);
+    }
+
+public:
+    /**
+     * Construct a nuclide filter with default global policy of ignoring 
+     * everything not specified
+     * */
+    NuclideFilter(Policy globalPolicy=Ignore)
+    : Filter(globalPolicy)
+    {
+    }
+    NuclideFilter(const NuclideFilter& orig)
+    : Filter(orig)
+    {
+    }
+    /**
[…]

### 2016-09-23 exnih/NuclideResonance.cpp

@@ -1,0 +1,30 @@
+#include "NuclideResonance.h"
+NuclideResonance::NuclideResonance()
+{
+    initialize();
+}
+void NuclideResonance::initialize()
+{
+    setResolved(NULL,0);
+    setUnresolved(NULL,0);
+}
+NuclideResonance::~NuclideResonance()
+{
+    if( getResolved() != NULL ){
+        delete[] getResolved();
+        setResolved(NULL,0);
+    }
+    if( getUnresolved() != NULL ){
+        delete[] getUnresolved();
+        setUnresolved(NULL,0);
+    }
+}
+
+NuclideResonance::NuclideResonance(const NuclideResonance & orig){
+    resolvedSize = orig.resolvedSize;
+    unresolvedSize  = orig.unresolvedSize;
+    resolved = new float[resolvedSize];
+    for( int i = 0; i < resolvedSize; i++) resolved[i] = orig.resolved[i];
+    unresolved = new float[unresolvedSize];
+    for( int i = 0; i < unresolvedSize; i++) unresolved[i] = orig.unresolved[i];
+}

### 2016-09-23 exnih/NuclideResonance.h

@@ -1,0 +1,40 @@
+#include "LibraryItem.h"
+#include "resources.h"
+#include <QString>
+
+class NuclideResonance : public LibraryItem
+{
+private:
+    float * resolved, * unresolved;
+    int resolvedSize, unresolvedSize;
+    void initialize();
+public:
+    NuclideResonance();
+    ~NuclideResonance();
+    NuclideResonance(const NuclideResonance & orig);
+    
+    float * getResolved(){return resolved;}
+    int getResolvedSize(){return resolvedSize;}
+    void setResolved(float * resolved, int size){this->resolved=resolved;resolvedSize=size;}
+    float * getUnresolved(){return unresolved;}
+    int getUnresolvedSize(){return unresolvedSize;}
+    void setUnresolved(float * unresolved, int size){this->unresolved=unresolved;unresolvedSize=size;}
+    
+    QString toQString(){
+        QString results = QString("Nuclide Resonance\n");
+        if( getResolved() != NULL ){
+            int length = getResolvedSize();
+            for( int i = 0; i < length; i++ ){
+                results += QString("  resolved[%1]=%2\n").arg(i).arg(getResolved()[i]);
+            }
+        }
+        if( getUnresolved() != NULL ){
+            int length = getUnresolvedSize();
+            for( int i = 0; i < length; i++ ){
+                results += QString("  unresolved[%1]=%2\n").arg(i).arg(getUnresolved()[i]);
+            }
+        }
+        return results;
+    }// end of toQString   
+    
+}; 

### 2016-09-23 exnih/Resource.h

@@ -1,0 +1,114 @@
+#ifndef RESOURCE_H
+#define RESOURCE_H
+
+//#include "TableOfContents.h"
+#include <QString>
+//#include "Standard/Interface/Serializable.h"
+///SCALE requires several types of data, depending on the running modules, to
+///run. This data is represented by the Resource class. This includes nuclear
+///data files, the standard composition library, fluxes and others. The Resource
+///base class contains all of the operations that are common across these
+///different types of data and deals mostly with management, not technical,
+///activities.
+class Resource //: public Standard::Serializable
+{
+    protected:
+
+        ///The version number of the resource
+        int version;
+
+        ///The name of the resource
+        QString resourceName;
+
+        /// The table of contents of the resource
+        //TableOfContents * tableOfContents;
+	
+	/// The resource's filename
+	QString filename;
+
+    public:
+
+        ///This operation opens the resource so that it may be used to retrieve
+        ///or store information according to its subtype.
+        virtual int open() = 0; 
+
+        ///This operation closes the resource and it may no longer be used.
+        virtual int close() = 0; 
+	
+	/// Determine if this resource is currently open
+	/// @return bool - true, if the resource is open, false otherwise
+	virtual bool isOpen() = 0;
+
+        ///This operation returns the title of the resource.
+        virtual QString getTitle() const = 0;
+
+        ///This operation returns the version of the resource.
+        int getVersion() const {return version;}
+
+        ///This operation returns the table of contents for the resource.
+        //TableOfContents * getTableOfContents(){return tableOfContents;}
+	
+	/// Set this resource's name
+	/// @param QString resourceName - The name of this resource        
+        void setResourceName(QString resourceName){this->resourceName=resourceName;}
+	
+	/// Obtain the name of this resource
+	/// @return QString - The name of the resource
+        QString getResourceName() const {return resourceName;}
+        
+	/// Set the filename, absolulte or relative to this resource on disk
+	/// @param QString filename - The filename of this resource
+	void setFileName(QString filename){this->filename = filename;}
+
+	/// Obtain the filename of this resource
+	/// @return QString - the filename of this resource
+	QString getFileName() const {return filename;}
+// serialization interface
+    virtual Resource * getCopy() const {throw "Unimplemented!";}
+    /**
+     * @brief Serialize the object into a contiguous block of data
+     * @param Standard::AbstractStream * stream - the stream into which the contiguous data will be stored
[…]

### 2016-09-23 exnih/ScatterMatrix.cpp

@@ -1,0 +1,173 @@
+#include "ScatterMatrix.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+#include <QDebug>
+#include <limits>
+//#include "Standard/Interface/AbstractStream.h"
+
+namespace {
+//! Union to allow reinterpreting of int -> float without breaking strict aliasing
+union FloatIntUnion {
+    int   magic_word;
+    float magic_float;
+};
+} // end anonymous namespace
+
+ScatterMatrix::ScatterMatrix()
+{
+    initialize();
+}
+ScatterMatrix::ScatterMatrix(const ScatterMatrix & orig){
+    legendreOrder = orig.legendreOrder;
+    temp = orig.temp;
+    for( int i = 0; i < orig.sinkGroups.size(); i++){
+        this->sinkGroups.append(orig.sinkGroups[i]->getCopy());
+    }
+}
+const ScatterMatrix& ScatterMatrix::operator=(const ScatterMatrix & orig){
+    legendreOrder = orig.legendreOrder;
+    temp = orig.temp;
+    for( int i = 0; i < orig.sinkGroups.size(); i++){
+        this->sinkGroups.append(orig.sinkGroups[i]->getCopy());
+    }
+    return * this;
+}
+bool ScatterMatrix::operator==(ScatterMatrix & a){
+    if( sinkGroups.size() != a.sinkGroups.size() ){
+        qDebug() << "ScatterMatrix sizes not equal for legendreOrder="<<getLegendreOrder()<<" temperature="<<getTemp();
+        return false;
+    }
+    for( int i = 0; i < sinkGroups.size(); i++){
+        bool equal = *sinkGroups[i] == *a.sinkGroups[i];
+        if( !equal ){
+            qDebug() << "ScatterMatrix sinkgroup not equal for legendreOrder="<<getLegendreOrder()<<" temperature="<<getTemp()<<" at index "<<i;
+            return false;
+        }
+    }
+    if( legendreOrder != a.legendreOrder ){
+        qDebug() << "ScatterMatrix legendreOrder not equal for legendreOrder="<<getLegendreOrder()<<" temperature="<<getTemp();
+        return false;
+    }
+    if( temp != a.temp ){
+        qDebug() << "ScatterMatrix temperature not equal for legendreOrder="<<getLegendreOrder()<<" temperature="<<getTemp()<<" vs a.l="<<a.getLegendreOrder()<<" a.t="<<a.getTemp();
[…]

### 2016-09-23 exnih/ScatterMatrix.h

@@ -1,0 +1,124 @@
+#ifndef SCATTERMATRIX_H
+#define SCATTERMATRIX_H
+#include "LibraryItem.h"
+#include "resources.h"
+#include "SinkGroup.h"
+#include <QString>
+#include <QList>
+//#include "Standard/Interface/Serializable.h"
+
+class ScatterMatrix: public LibraryItem//, public Standard::Serializable
+{
+private:
+    float temp;
+    int legendreOrder;    
+    QList<SinkGroup*> sinkGroups;
+    void initialize();
+public:
+    ScatterMatrix();
+    ScatterMatrix(const ScatterMatrix & orig);
+    const ScatterMatrix & operator=(const ScatterMatrix & orig);
+    ~ScatterMatrix();
+    
+    bool operator==(ScatterMatrix & a);
+    
+    /// Obtain a copy of the ScatterMatrix
+    /// @return ScatterMatrix * - The copy
+    ScatterMatrix * getCopy() const;
+    QList<SinkGroup*> * getSinkGroups(){return & sinkGroups;}
+    /// Obtain the first and last groups
+    /// that contain sink data.
+    /// @param int & first - the first group to populate
+    /// @param int & last - the last group to populate
+    /// @return bool - true, if first and last were succesfully found
+    ///                  false, otherwise.
+    bool getFirstLastGroup(int & first, int & last);
+    
+    /// Obtain a collapsed data set 
+    /// where sink groups are represented in 
+    /// magic word form
+    /// @param int & length - variable to populate with the 
+    ///                      length of the array being returned
+    /// @return float * - the collapsed data in array form
+    float * getCollapsedMatrix(int & length);
+    
+    /// Remove a sink group with the given grp
+    /// All sink groups with a group of grp are removed and deleted
+    /// @param int grp
+    /// @return int - the number of sinkgroups removed
+    int removeSinkGroup(int grp){ 
+        int count=0;
+        for( int i = 0 ; i < sinkGroups.size(); i++)
+        {
+            if( sinkGroups.at(i)->getGrp()==grp ){
+                SinkGroup * sinkGroup = sinkGroups.value(i);
+                delete sinkGroup;
+                sinkGroups.removeAt(i);
+                i--;
+                count++;
+            }
+        }
+        return count;
+    } // end of removeSinkGroup
+
+    /// Add a sink group to this scatter matrix
[…]

### 2016-09-23 exnih/SinkGroup.cpp

@@ -1,0 +1,116 @@
+//#include "Nemesis/harness/DBC.hh"
+#include "SinkGroup.h"
+#include "resources.h"
+#include <iostream>
+#include <QDebug>
+//#include "Standard/Interface/AbstractStream.h"
+//#include "Resource/AmpxLib/ampxlib_config.h"
+using namespace std;
+
+SinkGroup::SinkGroup()
+{
+    initialize();
+}
+SinkGroup::SinkGroup(const SinkGroup & orig){
+    start = orig.start;
+    end = orig.end;
+    grp = orig.grp;
+    data = orig.data;
+}
+const SinkGroup& SinkGroup::operator=(const SinkGroup & orig){
+
+    data = orig.data;
+    start = orig.start;
+    end = orig.end;
+    grp = orig.grp;
+    return *this;
+}
+bool SinkGroup::operator==(SinkGroup & a){
+    if( data.size() != a.data.size())
+    {
+        qDebug() << "SinkGroup sizes not equal for grp="<<getGrp()<<" s="<<getStart()<<" e="<<getEnd()<<" vs a.grp="<<a.getGrp()<<" a.s="<<a.getStart()<<" a.e="<<a.getEnd();
+        return false;
+    }
+    for( int i = 0; i < data.size(); i++){
+        if( data[i] != a.data[i] )
+        {
+            qDebug() << "SinkGroup data not equal for grp"<<getGrp()<<" s="<<getStart()<<" e="<<getEnd()<<" at index "<<i<<" "<<data[i]<<" vs "<<a.data[i];
+            return false;
+        }
+    }
+    //if( this->d == d) return true;
+    if( start != a.start )
+    {
+        qDebug() << "SinkGroup start not equal for grp"<<getGrp()<<" s="<<getStart()<<" e="<<getEnd();
+        return false;
+    }
+    if( end != a.end ){
+        qDebug() << "SinkGroup end not equal for grp"<<getGrp()<<" s="<<getStart()<<" e="<<getEnd();
+        return false;
+    }
+    if( grp != a.grp )
+    {
+        qDebug() << "SinkGroup grp not equal for grp"<<getGrp()<<" s="<<getStart()<<" e="<<getEnd();
+        return false;
+    }
+    return true;
+}
+void SinkGroup::initialize()
+{
+    start = 0;
+    end = 0;
+    grp = 0;
+}
+SinkGroup::~SinkGroup()
+{
+}
+
[…]

### 2016-09-23 exnih/SinkGroup.h

@@ -1,0 +1,119 @@
+#ifndef SINKGROUP_H
+#define SINKGROUP_H
+#include <iostream>
+#include <QVector>
+//#include "Standard/Interface/Serializable.h"
+namespace Standard{
+   class AbstractStream;
+}
+
+using namespace std;
+
+class SinkGroup //: public Standard::Serializable{
+{
+private:
+    QVector<float> data;
+    int start, end, grp;
+    void initialize();
+public: 
+    SinkGroup();
+    SinkGroup(const SinkGroup& orig);
+    const SinkGroup& operator=(const SinkGroup & orig);
+    ~SinkGroup();
+    
+    /**
+     * @brief equality operator
+     * Determine if this SinkGroup and SinkGroup 'a' are equal
+     * @return bool - true, if this and 'a' are equal, false otherwise
+     */
+    bool operator==(SinkGroup & a);
+         
+    /// Obtain a copy of the sinkgroup
+    /// @return SinkGroup * - The copy
+    virtual SinkGroup * getCopy() const;
+
+    inline void setStart(int newStart){
+        if(0==this->end){ this->end = newStart; }
+        else if(newStart > this->end){ return; }
+        else if( newStart == this->start ) return;
+        // case 1, new start is greater than old start
+        // we need to remove existing preceeding values
+        for( ;this->start < newStart ; this->start++){
+            if( data.size() > 0 )
+                data.pop_front();
+        }
+        // case 2, new start is less than old start
+        // we need to add values preceeding values
+        for( ; this->start > newStart; this->start--){
+            data.prepend(0.0);
+        }
+    }
+    inline int getStart(){return this->start;}
+    inline void setEnd(int newEnd){
+        if(newEnd < this->start) return;
+        if( this->end == newEnd ) return;
+        int newSize = newEnd - this->start;
+        data.resize(newSize); 
+        this->end = newEnd;
+//        for( ;this->end > newEnd; this->end--) data.pop_front();
+//        for( ;this->end < newEnd; this->end++) data.append(0.0);
+    }
+    inline int getEnd(){return this->end;}
+     
+    void setGrp(int grp){
+        this->grp=grp;
+    }
+    int getGrp(){return this->grp;}
+  
+    /// resize the sinkgroup to encompass the given index
+    /// @param int index - the index to encompass by this sink group  
+    bool resize(int index){
+        int newStart=getStart(), newEnd=getEnd();
+        if( index < 0 ) return false;
+        if( index >= getStart() && index < getEnd() ) return true;
+        
[…]

### 2016-09-23 exnih/SubGroupData.cpp

@@ -1,0 +1,278 @@
+#include "SubGroupData.h"
+//#include "Nemesis/harness/DBC.hh"
+//#include "Standard/Interface/jdebug.h"
+
+bool SubGroupWeight::operator==(SubGroupWeight & a){
+    if( temp != a.temp) return false;
+    CrossSection1d *c1,*c2;
+    c1 = (CrossSection1d *)this;
+    c2 = (CrossSection1d *)&a;
+    return *c1 == *c2;
+}
+
+
+
+const long SubGroupWeight::uid = 0xbaeef95102041612;
+
+SubGroupSetGrp::SubGroupSetGrp(const SubGroupSetGrp& orig){
+    ig = orig.ig;
+
+    for( int i = 0; i < orig.cross.size(); i++)
+        this->cross.append(orig.cross[i]->getCopy());
+
+     for( int i = 0; i < orig.weight.size(); i++)
+        this->weight.append(orig.weight[i]->getCopy());
+}
+
+
+SubGroupSetGrp::~SubGroupSetGrp(){
+    for( int i = 0; i < cross.size(); i++){
+        delete cross.value(i);
+    }
+    cross.clear();
+    for( int i = 0; i < weight.size(); i++){
+        delete weight.value(i);
+    }
+    weight.clear();
+}
+
+void  SubGroupSetGrp::addTemperatures(QList<float> * temps) const{
+    if( temps == 0) return;
+    for( int i = 0; i < weight.size(); i++ ) {
+        SubGroupWeight * w = weight.at(i);
+        if( !temps->contains(w->getTemp()) ) temps->append(w->getTemp());
+    }
+    qSort(*temps);
+}
+
+void SubGroupSetGrp::addMts(QList<int> * mts) const{
+    if( mts == 0) return;
+    for( int i = 0; i < weight.size(); i++ ) {
+        SubGroupWeight * w = weight.at(i);
+        if( !mts->contains(w->getMt()) ) mts->append(w->getMt());
+    }
+
+    for( int i = 0; i < cross.size(); i++ ) {
+        CrossSection1d * w = cross.at(i);
+        if( !mts->contains(w->getMt()) ) mts->append(w->getMt());
+    }
+
+    qSort(*mts);
+}
+
+
+
+CrossSection1d * SubGroupSetGrp::getCrossByMt(int mt){
+    for( int i = 0; i < cross.size(); i++ ) {
+        if( cross[i]->getMt() == mt) return cross[i];
+    }
+    return 0;
+}
+
[…]

### 2016-09-23 exnih/SubGroupData.h

@@ -1,0 +1,393 @@
+#ifndef SUBGROUPDATA_H
+#define	SUBGROUPDATA_H
+
+#include "CrossSection1d.h"
+#include <QObject>
+#include <QList>
+//#include "Standard/Interface/AbstractStream.h"
+
+class SubGroupWeight:public CrossSection1d{
+  public:
+    SubGroupWeight():CrossSection1d(),temp(0.0){}
+    SubGroupWeight(const SubGroupWeight& orig):CrossSection1d(orig){
+        temp = orig.temp;
+    }
+    virtual ~SubGroupWeight(){}
+
+     virtual SubGroupWeight * getCopy() const {
+         return new SubGroupWeight(*this);
+     }
+
+    /**
+     * Get the temperature for this weight set
+     * @return  the temperature of for this weight set
+     */
+    float getTemp()const{
+        return temp;
+    }
+
+    /**
+     * Set the temperature for this weight set
+     * @return  the temperature of for this weight set
+     */
+    void setTemp(double temp){
+        this->temp = temp;
+    }
+
+    bool operator==(SubGroupWeight & a);
+
+    static const long uid;
+
+
+
+  private:
+      /** The temperature for this weight set */
+      float temp;
+};
+
+
+
+
+class SubGroupSetGrp { //Standard::Serializable{
+   public:
+      SubGroupSetGrp(){}//:Serializable(){}
+      SubGroupSetGrp(const SubGroupSetGrp& orig);
+      virtual ~SubGroupSetGrp();
+
+      SubGroupSetGrp * getCopy() const {
+         return new SubGroupSetGrp(*this);
+     }
+
+      /**
+       * Get the group number for this subgroup
+       * @return  the group number for this subgroup
+       */
+      int getGrp() const{
+          return ig;
+      }
+
+      /**
+       * Set the group number for this subgroup
+       * @param  grp the group number for this subgroup
+       */
+      void setGrp(int grp) {
+          ig = grp;
+      }
+
+
+      /**
+       * Get the list of available temperatures.
+       * This includes the weight data only
+       * @return  the list of available temperatures.
+       */
+      void  addTemperatures(QList<float> * temps) const;
+
+
+      /**
+       * Get the list of available reactions.
+       * This includes cross section data as well as weight data
+       * @return  the list of available reaction.
+       */
+      void addMts(QList<int> * mts) const;
+
+
+
+      /**
+       * Get the cross section data  for the indicated mt
+       * @param mt the desired mt value
[…]

### 2016-09-23 exnih/mainwindow.cpp

@@ -1,0 +1,82 @@
+#include "mainwindow.h"
+#include "ui_mainwindow.h"
+
+#include "AmpxLibrary.h"
+#include "AmpxReader.h"
+
+#include <fstream>
+#include <QTime>
+#include <QDebug>
+
+MainWindow::MainWindow(QWidget *parent) :
+    QMainWindow(parent),
+    ui(new Ui::MainWindow)
+{
+    ui->setupUi(this);
+
+    AmpxLibrary lib;
+    AmpxReader reader;
+
+    std::fstream fin;
+    //fin.open("/media/Storage/scale/scale6.2/data/scale.rev11.xn28g19v7.1");
+    fin.open("/media/Storage/scale/Scale6.1/data/scale.rev10.xn27g19v7");
+    //fin.open("/media/Storage/scale/Scale6.1/data/scale.rev10.xn200g47v7");
+
+    QTime t;
+    t.start();
+    reader.read(&fin, lib, true, false);
+    int millis = t.elapsed();
+
+    QList<LibraryNuclide*> nucs = lib.getConstNuclides();
+
+    LibraryNuclide *nuc = nucs[11];
+
+    qDebug() << "Nuclide: " << qPrintable(nuc->getDescription());
+
+    qDebug() << "Number of 2d datasets: " << nuc->getGamma2dList().size();
+
+    CrossSection2d *xs = nuc->getGamma2dDataByMt(504);
+
+    qDebug() << "Legendre Order for MT=502: " << xs->getLegendreOrder();
+
+    qDebug() << "Temperatures: " << xs->getNumberTemperatures();
+    for(int i = 0; i < xs->getNumberTemperatures(); i++)
+        qDebug() << "T: " << xs->getTemperatureList()[i];
+
+    QList<ScatterMatrix*> *scats = xs->getScatterMatrices();
+
+    qDebug() << "Size of scats: " << scats->size();
+
+    ScatterMatrix *scat = (*scats)[1];
+    SinkGroup *g = scat->getSinkGroup(0);
+
+    float x = g->get(19);
+
+    qDebug() << "Size of sink: " << (g->getStart() - g->getEnd());
+    qDebug() << "P(0,0) = " << x;
+
+    for(int snk = 1; snk <= 19; snk++)
+    {
+        SinkGroup *s = scat->getSinkGroupByGrp(snk);
+        qDebug() << "Sinks: " << s->getStart() << " - " << s->getEnd();
[…]

### 2016-09-23 exnih/resources.cpp

@@ -1,0 +1,50 @@
+
+
+
+int headerSize(){ return sizeof(int);}
+int footerSize(){ return sizeof(int);}
+
+/// Explode a magic word into its appropriate parts
+/// @param const int magicWord - the magic word
+/// @param int &grp - the sink grp number to be populated
+/// @param int & start - the start of the data to be populated
+/// @param int & end - the end of the data to be populated
+/// @param int & length - the length of the data to be polulated
+/// @return bool - true, if the grp value non-negative
+///                false, otherwise 
+bool explodeMagicWord(const int magicWord, int &grp, int &start, int &end, int &length)
+{
+    const int million = 1000000, thousand = 1000;
+    start = magicWord / million;
+    end = (magicWord-(start*million))/thousand;
+    grp = (magicWord-(start*million))-(end*thousand);
+    start -=1; /// decrement start due to fortran starting at index 1
+    length = end - start;
+    if( grp < 0 ) return false;
+    return true;
+}// end of explodeMagic
+
+/// Construct a magic word from its appropriate parts
+/// @param int grp - the number of groups
+/// @param int start - the starting group
+/// @param int end - the ending group
+/// @param bool & error - error flag to indicate error occured
+///                      and magicword failed to be constructed.
+/// NOTE: This constructs a fortran indexed magic word
+/// @return int magicWord - the constructed magic word
+int getMagicWord(int grp,int start,int end, bool & error)
+{
+    if( grp > 999 || start > 999 || end > 999 ){
+        error = true;
+        return 0;
+    }
+    const int million = 1000000, thousand = 1000;
+    start += 1; /// fortran indexing start at 1 not zero
+    start *= million; /// start is first component and is in the millions place
+    end *= thousand;
+    
+    ///i.e., start = 1, end = 2, grp = 3
+    /// magicword = 001002003  = 1,002,003
+    /// magic word is ssseeeggg
+    return start+end+grp;    
+}// end of getMagicWord

### 2016-09-23 exnih/resources.h

@@ -1,0 +1,18 @@
+#ifndef AMPXLIB_RESOURCES_H
+#define AMPXLIB_RESOURCES_H
+
+/// enumerated types for parsing data
+#define AMPXLIB_NEUTRON1D_DATA 0
+#define AMPXLIB_GAMMA1D_DATA 1
+#define AMPXLIB_NEUTRON2D_DATA 2
+#define AMPXLIB_GAMMA2D_DATA 3
+// Both GammaProduction and TOTAL_XS have the same self defining format
+#define AMPXLIB_GAMMAPRODUCTION_DATA 4
+#define AMPXLIB_TOTAL_XS_DATA 4
+
+
+int headerSize();///{return sizeof(int);}
+int footerSize();///{return sizeof(int);}
+bool explodeMagicWord(const int magicWord, int &grp, int &start, int &end, int & length);
+int getMagicWord(int grp, int start, int end, bool & error);
+#endif

### 2016-09-29 exnih/AmpxReader.cpp
Made minor changes to the output

@@ -1010,6 +1010,7 @@ bool AmpxReader::readNuclideData(fstream &file, AmpxLibrary & library)
 
     for( int i = 0; i < numNuclides; i++ )
     {
+        std::cout << "Reading nuclide " << i << std::endl;
         LibraryNuclide * nuclide = readLibraryNuclide(file, NULL, verbose);
         if( nuclide == NULL ){
             if(this->printErrors)cerr<<"***Error: unable to read the nuclide at index "<<i+1<<"!"<<endl;

### 2016-09-29 exnih/mainwindow.cpp

@@ -37,7 +37,7 @@ MainWindow::MainWindow(QWidget *parent) :
 
     CrossSection2d *xs = nuc->getGamma2dDataByMt(504);
 
-    qDebug() << "Legendre Order for MT=502: " << xs->getLegendreOrder();
+    qDebug() << "Legendre Order for MT=504: " << xs->getLegendreOrder();
 
     qDebug() << "Temperatures: " << xs->getNumberTemperatures();
     for(int i = 0; i < xs->getNumberTemperatures(); i++)
@@ -47,7 +47,7 @@ MainWindow::MainWindow(QWidget *parent) :
 
     qDebug() << "Size of scats: " << scats->size();
 
-    ScatterMatrix *scat = (*scats)[1];
+    ScatterMatrix *scat = (*scats)[0];
     SinkGroup *g = scat->getSinkGroup(0);
 
     float x = g->get(19);

### 2016-11-16 docprep/draft1.tex
Added a whopping 4 words
Added the Plm recurrance equations
Replaced \cdot with x in the double factorial equation because it was hard to see.
Finished the sweep() routine

@@
 \begin{itemize}
-\item A list of all energies, $\mathcal{E}$
-\item A list of all directions in the quadrature, $Q$
+\item A list of all energies, $\{E\}$
+\item A list of all directions in the quadrature, $\{\Omega\}$
 \item A mapping from $r$ to zone number, $\mathcal{Z}(r)$
-\item Lists of direction cosines, $\mu$, $\xi$, and $\eta$
+\item Lists of direction cosines, $\mu(\Omega)$, $\xi(\Omega)$, and $\eta(\Omega)$
 \item A map of $r$ to voxel volume, $V(r)$
-\item A list of all $x$, $y$, and $z$ voxel centers, $X$, $Y$, and $Z$
-\item A list of weights for the quadrature, $\omega$
-\item A list of scatter cross sections in zone $\mathcal{Z}(r)$ from energy $E'$ to $E$, $\Sigma_s(\mathcal{Z}(r), E', E)$
+%\item A list of all $x$, $y$, and $z$ voxel centers, $X$, $Y$, and $Z$
+\item A mapping of weights for the quadrature, $\omega(\Omega)$
+\item A mapping of scatter cross sections in zone $\mathcal{Z}(r)$ from energy $E'$ to $E$, $\Sigma_s(\mathcal{Z}(r), E' \rightarrow E)$
 \item A list of total cross sections in zone $\mathcal{Z}(r)$ at energy $E$, $\Sigma_T(\mathcal{Z}(r), E)$
@@
 
+  # Knowns
+  (*$V(r)$*)  # [cm^3] volume of cell ijk
+  (*$A_{xy}(\Omega,i,j), A_{yz}(\Omega,j,k), A_{xz}(\Omega,i,k)$*)  # [cm^2] Area of each face of cell ijk
+  # Unknowns
   (*$\phi(r, E) = \vec{0}$*)    # Scalar flux
@@
 
-  for (*$E \in \mathcal{E}$*) in descending order
+  for every (*$E \in \{E\}$*) in descending order
     while not converged
-      for (*$\Omega \in Q$*) in any order
-        if (*$\mu(\Omega) > 0$*) and (*$\xi(\Omega) > 0$*) and (*$\eta(\Omega) > 0$*)
-          octant1()
-        else if (*$\mu(\Omega) > 0$*) and (*$\xi(\Omega) > 0$*) and (*$\eta(\Omega) < 0$*)
-          octant2()
-        ...
+      (*$\phi^{i-1}(r) = \phi^{i+1}(r)$*)
+      (*$S(r) = \vec{0}$*)
+      for every (*$E' >= E$*)
+        for every cell (*$ijk$*)
+          (*$S(ijk) = S(ijk) + 1/4\pi \phi(E', ijk)*\Sigma_s(\mathcal{Z}(ijk), E')*V(ijk)$*)
+      for every cell (*$ijk$*)
+        (*$S(ijk) = S(ijk) + S_{x}(ijk)$*)
+      (*$\phi^{i+1} = \vec{0}$*)
+      for every (*$\Omega \in Q$*) in any order
+        (*$i_0 = f(\mu(\Omega))$*)
+        (*$\Delta i = f(\mu(\Omega))$*)
+        (*$j_0 = f(\xi(\Omega))$*)
+        (*$\Delta j = f(\xi(\Omega))$*)
+        (*$k_0 = f(\eta(\Omega))$*)
+        (*$\Delta k = f(\eta(\Omega))$*)
+        for every cell (*$ijk$*) starting at (*$(i_0, j_0, k_0)$*) in direction (*$<\Delta i, \Delta j, \Delta k>$*)
+          if (*$i \in \partial$*)
+            (*$\phi_{x+} = 0$*)
           else
-          octant8()
+            (*$\phi_{x+} = \phi_{x-}(i\pm 1, j, k)$*)    
+          if (*$j \in \partial$*)
+            (*$\phi_{y+} = 0$*)
+          else
+            (*$\phi_{y+} = \phi_{y-}(i, j \pm 1, k)$*) 
+          if (*$k \in \partial$*)
+            (*$\phi_{z+} = 0$*)
+          else
+            (*$\phi_{z+} = \phi_{z-}(i, j, k \pm 1)$*)
+
+          (*$n = S(ijk) + A_{yz}(\Omega, j,k)\phi_{x+} + A_{xz}(\Omega, i, k) \phi_{y+} + A_{xy}(\Omega, i,j)\phi_{z+}$*)
+          (*$d = V(ijk)*\Sigma_t(\mathcal{Z}(ijk), E) + A_{yz}(\Omega, j, k) + A_{xz}(\Omega, i, k) + A_{xy}(\Omega, i, j)$*)
+          (*$\Psi_0 = n/d$*)
+          (*$\Psi(E, \Omega, ijk) = \Psi_0$*)
+          (*$\phi_{x-}(ijk) = 2\Psi_0 - \phi_{x+}$*)
+          (*$\phi_{y-}(ijk) = 2\Psi_0 - \phi_{y+}$*)
+          (*$\phi_{z-}(ijk) = 2\Psi_0 - \phi_{z+}$*)
+          (*$\phi^{i+1}(ijk) = \phi^{i+1} + \omega(\Omega)\Psi(e, \Omega, ijk)$*)
+        for every cell (*$ijk$*)
+          (*$\phi(E, ijk) = \phi^{i+1}(ijk)$*)
+        EMIT (*$\phi(E, ijk)$*)
+      (*$\epsilon = max((\phi^{i+1} - phi^{i-1})/\phi^{i+1})$*)
+
+      for every cell (*$ijk$*)
+        (*$\phi(E, ijk) = \phi^{i+1}(ijk)$*)
+  output time to completion
 \end{lstlisting}
@@
 
+The real part of the
+
 whose real part can be expanded to
@@
 \end{equation}
+
+Some useful recurrence relations used to calculate the Associated Legendre Polynomials are:
+
+\begin{equation}
+P_l^l(x) = (-1)^l(2l-1)!!(1-x^2)^(l/2)
+\end{equation}
+
+\begin{equation}
+P_{l+1}^l(x) = x(2l+1)P_l^l(x)
+\end{equation}
+
+\begin{equation}
+P_l^l(x) = (-1)^l(2l-1)!!(1-x^2)^(l/2)
+\end{equation}
+
+where $x!!$ is the double factorial function defined by:
+
+\begin{equation}
+x!! = \prod_{k=0}^{\lceil x/2 \rceil - 1} x-2k
+\end{equation}
+
+For example, $5!! = 5 \times 3 \times 1 = 15$.
 

### 2016-11-16 doctors/mcnpwriter.cpp
Added a MCNP tally mesh and fine/course density options

@@ -63,9 +63,33 @@ std::string McnpWriter::generateSurfaceString(Mesh *m)
     return surfString;
 }
 
-std::string McnpWriter::generateCellString(Mesh *m)
+std::string McnpWriter::generateCellString(Mesh *m, bool fineDensity)
 {
     std::string allCellString = "c ============================== CELL DECK ==============================\n";
+
+    std::vector<float> coarseDensity;
+    if(!fineDensity)
+    {
+        // Calculate the average density for each material
+        std::vector<int> matVoxels;
+        coarseDensity.resize(MaterialUtils::hounsfieldRangePhantom19.size(), 0.0f);
+        matVoxels.resize(MaterialUtils::hounsfieldRangePhantom19.size(), 0);
+
+        for(int i = 0; i < m->atomDensity.size(); i++)
+        {
+            if(m->zoneId[i] > matVoxels.size())
+            {
+                qDebug() << "matVoxels wasn't big enough!";
+            }
+            matVoxels[m->zoneId[i]]++;
+            coarseDensity[m->zoneId[i]] += m->atomDensity[i];
+        }
+
+        for(int i = 0; i < coarseDensity.size(); i++)
+        {
+            coarseDensity[i] /= matVoxels[i];
+        }
+    }
 
     for(int xi = 0; xi < m->xElemCt; xi++)
     {
@@ -97,8 +121,8 @@ std::string McnpWriter::generateCellString(Mesh *m)
                     continue;
                 }
 
-                if(mindx == 208311)
-                    qDebug() << "halt";
+                //if(mindx == 208311)
+                //    qDebug() << "halt";
 
                 std::string matstr = "";
 
@@ -109,16 +133,30 @@ std::string McnpWriter::generateCellString(Mesh *m)
                 }
                 else
                 {
-                    float atmden = m->atomDensity[mindx];
+                    float atmden;
+                    if(fineDensity)
+                        atmden = m->atomDensity[mindx];
+                    else
+                        atmden = coarseDensity[m->zoneId[mindx]];
                     matstr = std::to_string(m->zoneId[mindx]+1) + " " + std::to_string(m->atomDensity[mindx]);
                 }
+
+                int importance = 1;
[…]

### 2016-11-16 doctors/mcnpwriter.h

@@ -14,11 +14,12 @@ public:
     const static int MAX_BOUNDS = 100000;
 
     std::string generateSurfaceString(Mesh *m);
-    std::string generateCellString(Mesh *m);
-    std::string generateDataCards();
+    std::string generateCellString(Mesh *m, bool fineDensity);
+    std::string generateDataCards(Mesh *m);
     std::string generatePhantom19MaterialString();
+    std::string generateMeshTally(Mesh *m);
 
-    void writeMcnp(std::string filename, Mesh *m);
+    void writeMcnp(std::string filename, Mesh *m, bool fineDensity);
 
 protected:
     bool m_failFlag;

### 2016-11-16 doctors/solver.cpp

@@ -332,10 +332,12 @@ void Solver::gssolver(const Quadrature *quad, const Mesh *mesh, const XSection *
     std::vector<float> errMaxList;
     std::vector<std::vector<float> > errList;
     std::vector<int> converganceIters;
+    std::vector<float> converganceTracker;
 
     errMaxList.resize(xs->groupCount());
     errList.resize(xs->groupCount());
     converganceIters.resize(xs->groupCount());
+    converganceTracker.resize(xs->groupCount());
 
     const XSection &xsref = *xs;
 
@@ -378,7 +380,7 @@ void Solver::gssolver(const Quadrature *quad, const Mesh *mesh, const XSection *
 
     qDebug() << "Solver::gssolver(): 379: Solving " << mesh->voxelCount() * quad->angleCount() * xs->groupCount() << " elements in phase space";
 
-    for(unsigned int ie = xs->groupCount()-3; ie < xs->groupCount(); ie++)  // for every energy group
+    for(unsigned int ie = xs->groupCount()-1; ie < xs->groupCount(); ie++)  // for every energy group
     {
 
         // TODO - remove check
@@ -570,7 +572,7 @@ void Solver::gssolver(const Quadrature *quad, const Mesh *mesh, const XSection *
                                     mesh->Ayz[ie*quad->angleCount()*mesh->yElemCt*mesh->zElemCt + iang*mesh->yElemCt*mesh->zElemCt + iy*mesh->zElemCt + iz] * influxX +  // [cm^2 * #/cm^2]  The 2x is already factored in
                                     mesh->Axz[ie*quad->angleCount()*mesh->xElemCt*mesh->zElemCt + iang*mesh->xElemCt*mesh->zElemCt + ix*mesh->zElemCt + iz] * influxY +
                                     mesh->Axy[ie*quad->angleCount()*mesh->xElemCt*mesh->yElemCt + iang*mesh->xElemCt*mesh->yElemCt + ix*mesh->yElemCt + iy] * influxZ;
-                            float denom = mesh->vol[ix*xjmp+iy*yjmp+iz]*xsref.totXs1d(zid, ie)*mesh->atomDensity[ix*xjmp + iy*yjmp + iz]*100 +                               // [cm^3] * [b] * [1/b-cm]
[…]

### 2016-12-01 docprep/draft1.tex
Added the external source to the sweep code

@@
   # Knowns
-  (*$V(r)$*)  # [cm^3] volume of cell ijk
+  (*$V(r)$*)  # [cm^3] Volume of cell ijk
   (*$A_{xy}(\Omega,i,j), A_{yz}(\Omega,j,k), A_{xz}(\Omega,i,k)$*)  # [cm^2] Area of each face of cell ijk
+  (*$S_x(E, r)$*)  # [#] External source
   # Unknowns

### 2017-01-09 docprep/draft1.tex
Added a sequence diagram and added a section on the raytracer math

@@
 \end{figure}
+
+\subsection{Raytracer}
+Discrete ordinate solutions greatly benefit from the presence of a raytracing algorithm to provide the uncollided flux to the solver.
+
+The raytracer was implemented as...
+
+When the solver is changed from isotropic to anisotropic, the raytracer must be able to produce a Pn expansion needed by the anisotropic solver. We defined the spherical harmonic function as
+
+\begin{equation}
+    Y_l^m(\theta, \phi)= 
+\begin{cases}
+\sqrt{2} K_l^m \cos(m \phi) P_l^m(\cos(\theta)),      & m > 0 \\
+K_l^0 P_l^0(\cos(\theta)),              & m = 0  \\
+\sqrt{2} K_l^m \sin(-m \phi) P_l^{-m}(\cos(\theta)),           & m < 0
+\end{cases}
+\end{equation}
+
+where $K_l^m$ is defined as
+
+\begin{equation}
+K_l^m = \sqrt{\frac{2l+1}{4 \pi} \frac{(l-|m|)!}{(l+|m|)!}}
+\end{equation}
+
+and the associated Legendre functions are computed by the recursion
+
+\begin{equation}
+\begin{aligned}
+P_0^0(x) &= 1 \\
+P_m^m(x) &= (2m-1)!!(1-x^2)^{m/2} \\
+P_{m+1}^m(x) &= x(2m+1)P_m^m(x) \\
+P_l^m(x) &= \frac{x(2l-1)}{l-m} P_{l-1}^m (x) - \frac{l+m-1}{l-m} P_{l-2}^m (x)
+\end{aligned}
+\end{equation}
+
+The spherical harmonic expansion is used to approximate a function $f$ by
+
+\begin{equation}
+f(\theta, \phi) = \sum_{l=0}^\infty \sum_{m=-l}^l y_l^m(\theta, \phi) f_l^m
+\end{equation}
+
+which is exact when the expansion continues to infinity. However, it is trunctated at L yielding
+
+\begin{equation}
+f(\theta, \phi) \approx \sum_{l=0}^L \sum_{m=-l}^l y_l^m(\theta, \phi) f_l^m
+\end{equation}
+
+However, the value of each $f_l^m$ must be computed before the approximation can be used.
+
+These values are computed by
+
+\begin{equation}\label{eq:flm}
+f_l^m = \int_{4 \pi} y_l^m(\theta, \phi) f(\theta, \phi) d\hat{\Omega}
+\end{equation}
+
+In the case of the raytracer solution, since \textit{only} the uncollided source is considered, the direction at any point is always precisely known, the particles are traveling away from the source point.
+Therefore, the function to be approximated is a Dirac delta function in the two dimensional surface of the unit sphere.
+Equation~\ref{eq:flm} can be reduced to
+
+\begin{equation}
+f_l^m = \int_{4 \pi} y_l^m(\theta, \phi)f(\theta_0, \phi_0) d\hat{\Omega}
+\end{equation}
+
+The property of the Dirac delta function
+
+\begin{equation}
+\int_{-\infty}^{\infty} \delta(x_0) f(x) dx = f(x_0)
+\end{equation}
+
+can now be used which reduces Equation~\ref{eq:flm} to
+
+\begin{equation}
+f_l^m = y_l^m(\theta_0, \phi_0)
+\end{equation}
+
+The moments are related to the angular flux by
+
+\begin{equation}
+\phi_l^m(\vec{r}, E) = \int Y_l^m(\hat{\Omega} ') \psi(\vec{r}, \hat{\Omega}', E) d\hat{\Omega}'
+\end{equation}
+
+which becomes
+
+\begin{equation}
+\phi_l^m(\vec{r}, E) = \Psi Y_l^m(\hat{\Omega}_0)
+\end{equation}
+
+where $\Psi$ is equal to the magnitude of $\psi_{unc}$ in the $\hat{\Omega}_0$ direction.
 

### 2017-01-09 doctors/legendre.cpp
Added the Spherical Harmonic functionality

@@ -1,11 +1,20 @@
-#include "legendre.h"
-
+#define _USE_MATH_DEFINES
 #include <cmath>
+
+#include "legendre.h"
 #include <limits>
+
+double factorial(double x)
+{
+    if(x <= 1.001)
+        return 1.0;
+
+    return x * factorial(x - 1.0);
+}
 
 double doubleFactorial(double x)
 {
-    if(x < 2.0)
+    if(x < 2.001)
         return 1.0;
 
     return x * doubleFactorial(x-2.0);
@@ -21,6 +30,7 @@ AssocLegendre::~AssocLegendre()
 {
 
 }
+
 
 float AssocLegendre::operator()(const int l, const int m, const float x)
 {
@@ -46,5 +56,63 @@ float AssocLegendre::operator()(const int l, const int m, const float x)
     }
     return pml;
 }
+
+
+
+/*
+float AssocLegendre::operator ()(const int l, const int m, const float x)
+{
+    if(l == 0 && m == 0)
+    {
+        return 1.0;
+    }
+    else if(l == m)
+    {
+        return ((m%2==0)? (1.0) : (-1.0)) * doubleFactorial(2*m - 1) * pow(1-x*x, m/2.0);
+    }
+    else if(l == m+1)
+    {
+        return x * (2*m+1) * operator ()(m, m, x);
+    }
+    else
+    {
+        return x*(2*l-1)/(l-m) * operator()(l-1, m, x) - (l+m-1.0)/(l-m)*operator()(l-2, m, x);
+    }
+}
+*/
+
+SphericalHarmonic::SphericalHarmonic()
+{
+
+}
+
+SphericalHarmonic::~SphericalHarmonic()
+{
+
+}
+
+float SphericalHarmonic::normConst(const int l, const int m)
+{
+    float t1 = static_cast<float>(2*l+1) / (4 * M_PI);
+    float t2 = factorial(l-fabs(m)) / factorial(l + fabs(m));
+    return sqrt(t1 * t2);
+}
+
+float SphericalHarmonic::operator()(const int l, const int m, const float theta, const float phi)
+{
+    if(m > 0)
+    {
+        return sqrt(2.0) * normConst(l, m) * cos(m * phi) * m_assoc(l, m, cos(theta));
+    }
+    else if(m == 0)
+    {
+        return normConst(l, 0) * m_assoc(l, 0, cos(theta));
+    }
+    else
+    {
[…]

### 2017-01-19 docprep/draft1.tex
un/collided flux clarification
Added a sentence about the indexing scheme
Clarified the Legendre part and omitted a bunch of old stuff

@@
 \usepackage{fullpage}
-
+\usepackage{mathtools}
 \numberwithin{equation}{subsection}
@@
 \end{lstlisting}
+
+The indexing scheme is as follows: E, A, x, y, z, l, m
 
@@
 
-which becomes
+which, for the uncollided flux, becomes
 
@@
 \begin{equation}
-P_n(x) = \frac{1}{2 \pi i} \oint (1-2\xi x + \xi^2)^{-1/2} \xi^{-n-1} d\xi
-\end{equation}
-
-\begin{equation}
-P_n(x) = \frac{1}{2^n n!}\frac{d^n}{dx^n}(x^2-1)^n
+P_n(x) = \frac{1}{2 \pi i} \oint \frac{1}{\sqrt{1-2\xi x + \xi^2}} \frac{1}{\xi^{n+1}} d\xi
 \end{equation}
@@
 
+Legencre polynomials are useful for approximating functions truncated at some point. Legendre polynomials are mutually orthogonal which gives them the property
+
+\begin{equation}
+f(x) = \sum_{n=0}^\infty P_n(x)
+\end{equation}
+
+where $f(x)$ is any arbitrary function.
+Often, like a Fourier series which uses sin and cos instead of Legendre functions, the series is cut off at some threshold as an approximation.
+As an example of this, consider the function $1.75x^3 + 0.6x^2+.21x-.06$ over the domain $[2,5]$ using the first four Legendre polynomials.
+
+The coefficients are computed
+
+\begin{align}
+k_0 &= \int_2^5 P_0(x) f(x) dx = \int_2^5  f(x) dx\\
+k_1 &= \int_2^5 P_1(x) f(x) dx = \int_2^5 x f(x) dx\\
+k_2 &= \int_2^5 P_2(x) f(x) dx = \int_2^5 \frac{1}{2}(3x^2-1) f(x) dx\\
+k_3 &= \int_2^5 P_3(x) f(x) dx = \int_2^5 \frac{1}{2}(5x^3 - 3x) f(x) dx\\
+\end{align}
+
+These values are then used to reconstruct the original function by
+
+\begin{equation}
+f(x) \approx k_0 P_0(x) + k_1 P_1(x) + k_2 P_2(x) + k_3 P_3(x)
+\end{equation}
+
 Associated Legendre polynomials are a generalization of the Legendre polynomial and solve
+
+\subsection{Associated Legendre Polynomials}\label{sec:assoc_leg}
+
+The associated Legendre polynomials are solutions to the equation
 
@@
 
-Both Legendre polynomials and Associated Legendre Polynomials can be calculated using the Boost C++ library.
+The associated Legendre polynomials can be computed recusively by
+
+\begin{align}
+P_m^m(x) &= (-1)^m(2m-1)!!(1-x^2)^(m/2) \\
+P_{m+1}^m(x) &= x(2m+1)P_m^m(x) \\
+(l-m) P_l^m(x) &= x(2l-1)P_{l-1}^m(x) - (l+m-1)P_{l-2}^m(x)
+\end{align}
+
+where $x!!$ is the double factorial function defined by:
+
+\begin{equation}
+x!! = \prod_{k=0}^{\lceil x/2 \rceil - 1} x-2k
+\end{equation}
+
+For example, $5!! = 5 \times 3 \times 1 = 15$ and $6!! = 6 \times 4 \times 2 = 48$.
 
@@
 
-First, we assume that Laplace's equation is of the seperable form $f(\rho, \theta, \phi) = R(\rho)Y(\theta, \phi)$. Equation \ref{eq_laplace_spherical} can then be rewritten as
-
-\begin{equation} \label{eq_spherical1}
-\begin{split}
-\nabla^2 f(\rho, \theta, \phi) = 
-\frac{1}{\rho^2}
-\frac{\partial}{\partial \rho} 
-\left( \rho^2 \frac{\partial (R(\rho)Y(\theta, \phi))}{\partial \rho}\right) 
-+ \\
-\frac{1}{\rho^2 \sin \theta}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial (R(\rho)Y(\theta, \phi))}{\partial \theta}\right) 
-+ 
-\frac{1}{\rho^2 \sin^2 \theta} 
-\frac{\partial^2 (R(\rho)Y(\theta, \phi))}{\partial \phi^2} = 0
-\end{split}
-\end{equation}
-
-\noindent
-which simplifies to
-\begin{equation} \label{eq_spherical2}
-\begin{split}
-\frac{Y(\theta, \phi)}{\rho^2}
-\frac{\partial}{\partial \rho} 
-\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
-+ 
-\frac{R(\rho)}{\rho^2 \sin \theta}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
-+ \\
-\frac{R(\rho)}{\rho^2 \sin^2 \theta} 
-\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2}
-\end{split} = 0
-\end{equation}
-
-\noindent
-Multiplying Eq. \ref{eq_spherical2} by $\rho^2 / (R(\rho)Y(\theta, \phi))$ yields
-
-\begin{equation}\label{eq_spherical3}
-\begin{split}
-\frac{1}{R(\rho)}
-\frac{\partial}{\partial \rho} 
-\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
-+ 
-\frac{1}{Y(\theta, \phi) \sin \theta}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
-+ \\
-\frac{1}{Y(\theta, \phi) \sin^2 \theta} 
-\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2} = 0
-\end{split}
-\end{equation}
-
-Equation \ref{eq_spherical3} can be split into two parts, a function of $\rho$ alone and a component that is a function of $\theta$ and $\phi$ alone.
-The component that is a function of $\theta$ and $\phi$ can be moved to the other side of the equation.
-Since two functions of different variables are equal, they must each be equal to some constant, $\lambda$.
-Therefore, Eq.
-\ref{eq_spherical3} can be rewritten as two equations:
-
-\begin{equation}\label{eq_spherical_r}
-\frac{1}{R(\rho)}
-\frac{\partial}{\partial \rho} 
-\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) = \lambda
-\end{equation}
-
-\begin{equation}\label{eq_spherical_tp}
-\frac{1}{Y(\theta, \phi) \sin \theta}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial Y(\theta, \phi)}{\partial \theta}\right) 
-+ \\
-\frac{1}{Y(\theta, \phi) \sin^2 \theta} 
-\frac{\partial^2 Y(\theta, \phi)}{\partial \phi^2} = -\lambda
-\end{equation}
-
-$Y(\theta, \phi)$ can be further seperated into $Y(\theta, \phi) = \Theta(\theta)\Phi(\phi)$ which produces
-
-\begin{equation}\label{eq_spherical_y}
-\begin{split}
-\frac{1}{\Theta(\theta)\Phi(\phi) \sin \theta}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial (\Theta(\theta)\Phi(\phi))}{\partial \theta}\right) 
-+ \\
-\frac{1}{\Theta(\theta)\Phi(\phi) \sin^2 \theta} 
-\frac{\partial^2 (\Theta(\theta)\Phi(\phi))}{\partial \phi^2} = -\lambda
-\end{split}
-\end{equation}
-
-\noindent
-which simplifies to
-
-\begin{equation}\label{eq_spherical_y2}
-\begin{split}
-\frac{1}{\Theta(\theta) \sin \theta}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
-+ 
-\frac{1}{\Phi(\phi) \sin^2 \theta} 
-\frac{\partial^2 \Phi(\phi)}{\partial \phi^2} = -\lambda
-\end{split}
-\end{equation}
-
-Multiplying Eq. \ref{eq_spherical_y2} by $\sin^2(\theta)$ and rearranging yields
-
-\begin{equation}\label{eq_spherical_y3}
-\begin{split}
-\frac{\sin \theta}{\Theta(\theta)}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
-+ \lambda \sin \theta
- = \frac{-1}{\Phi(\phi)} 
-\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
-\end{split}
-\end{equation}
-
-As before, each side of Eq.
-\ref{eq_spherical_y3} is a function of a single variable, therefore, both sides are equal to some constant.
-In this case, we select $m^2$ to be  the constant variable.
-The produces
-
-\begin{equation}\label{eq_spherical_t}
-\begin{split}
-\frac{\sin \theta}{\Theta(\theta)}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
-+ \lambda \sin \theta
- = m^2
-\end{split}
-\end{equation}
-
-\begin{equation}\label{eq_spherical_p}
-\begin{split}
-\frac{1}{\Phi(\phi)} 
-\frac{\partial^2 \Phi(\phi)}{\partial \phi^2} = -m^2
-\end{split}
-\end{equation}
-
-Combining Eq. \ref{eq_spherical_r}, \ref{eq_spherical_t}, and \ref{eq_spherical_p} Laplace's equation in sperical coordinates can be expressed in a fully seperated form as
-
-\begin{equation}\label{eq_spherical_sep}
-\begin{split}
-\frac{1}{R(\rho)}
-\frac{\partial}{\partial \rho} 
-\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
-&= \lambda \\
-\frac{\sin \theta}{\Theta(\theta)}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
-+ \lambda \sin \theta
-&= m^2 \\
-\frac{1}{\Phi(\phi)} 
-\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
-&= -m^2
-\end{split}
-\end{equation}
-
-It can be shown through a detailed anlysis that is outside the scope of this paper that $m$ and $\lambda$ must both be integers.
-$\lambda$ is further constrained in that the relation $\lambda = l(l+1)$ where $l \leq |m|$ and $l \in \mathbb{Z}$.
-Therefore, Eq.
-\ref{eq_spherical_sep} can be written as
-
-\begin{equation}\label{eq_spherical_sep2}
-\begin{split}
-\frac{1}{R(\rho)}
-\frac{\partial}{\partial \rho} 
-\left( \rho^2 \frac{\partial R(\rho)}{\partial \rho}\right) 
-&= l(l+1) \\
-\frac{\sin \theta}{\Theta(\theta)}
-\frac{\partial}{\partial \theta}
-\left( \sin \theta \frac{\partial \Theta(\theta)}{\partial \theta}\right) 
-+ l(l+1) \sin \theta
-&= m^2 \\
-\frac{1}{\Phi(\phi)} 
-\frac{\partial^2 \Phi(\phi)}{\partial \phi^2}
-&= -m^2 \\
-& l, m \in \mathbb{Z} \\
-& l \leq |m|
-\end{split}
-\end{equation}
-
-% TODO
-TODO - Discontinuous
-
-It can be shown that the solution is stated
+Combining the associated Legendre polynomials and the sin/cos series, the spherical harmonics equations can be derived. They are orthogonal in both directions.
 
 \begin{equation}\label{eq:spherical_harmonic}
-Y_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
-\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) e^{im \varphi}
+Y_l^m(\theta, \phi) = N_l^{|m|} P_l^{|m|}(\cos \theta) e^{i m \phi}
 \end{equation}
@@
 \begin{equation}\label{eq:spherical_harmonic_exp}
-Y_l^m(\theta, \varphi) = (-1)^{l}\sqrt{\frac{2l+1}{4 \pi}
-\frac{(l-m)!}{(l+m)!}} P_l^m(\cos \theta) \left[ \cos(m \varphi) + i \sin(m \varphi) \right]
+Y_l^m(\theta, \varphi) = N_l^{|m|} P_l^m(\cos \theta) \left[ \cos(m \phi) + i \sin(m \phi) \right]
+\end{equation}
+
+The normalization constant $N_l^m$ is defined as
+
+\begin{equation}
+N_l^m = \sqrt{\frac{2l+1}{4 \pi} \frac{(l-m)!}{(l+m)!}}
+\end{equation}
+
+In many cases, only the real part ($y_l^m$) of the complex spherical harmonics ($Y_l^m$) are necessary.
+
+\begin{equation}
+y_l^m =
+\begin{cases*}
+\sqrt{2} N_l^m \cos(m\phi) P_l^m(\cos(\theta)) \\
+N_l^0 P_l^0(\cos(\theta)) \\
+\sqrt{2} N_l^{|m|} \sin(|m| \phi) P_l^{|m|}(\cos(\theta))
+\end{cases*}
 \end{equation}

### 2017-01-24 docprep/draft1.tex
Added more details on Legendre polys
Added more details on the raytracer

@@
 
-When the solver is changed from isotropic to anisotropic, the raytracer must be able to produce a Pn expansion needed by the anisotropic solver. We defined the spherical harmonic function as
+When the solver is changed from isotropic to anisotropic, the raytracer must be able to not only compute the uncollided flux but compute the direction of flux in each voxel and map that onto the discrete quadrature.
+
+To map an arbitrary direction onto a discete quadrature set, the direction is split amonst the three nearest directions in the quadrature set.
+Each quadrature direction is weighted.
+The closeness of each direction computed by calcualting the cosine of the angle between a direction $\hat{\Omega}$ and the quadrature direction $\hat{\Omega}_a$.
+The cosine is computed
+
+\begin{equation}
+\cos(\theta) = \hat{\Omega} \cdot \hat{\Omega}_a
+\end{equation}
+
+The three nearest points are then taken to form a triangle with the $\hat{\Omega}$ direction between them as shown in Fig. XYZ.
+
+The area of $\triangle \mathbf{A} \mathbf{B} \mathbf{\Omega}$ is
+
+\begin{equation}
+A = |\mathbf{A}-\mathbf{B}| |\mathbf{A}-\mathbf{C}| \sqrt{1-\left[ (\mathbf{A}-\mathbf{B}) \cdot (\mathbf{A}-\mathbf{C}) \right]}
+\end{equation}
+
+Remove what is below this!
 
@@
 
+Some references omit the Condon-Shortley phase $(-1)^m$ while most others include it. These two are distinguished by the nomenclature
+
+\begin{equation}
+P_{lm}(x) = (-1)^m P_l^m(x)
+\end{equation}
+
+Only the more commonly found $P_l^m$ form will be used in this work.
+
 \subsection{Spherical Harmonic Function}
@@
 \begin{equation}\label{eq:spherical_harmonic_exp}
-Y_l^m(\theta, \varphi) = N_l^{|m|} P_l^m(\cos \theta) \left[ \cos(m \phi) + i \sin(m \phi) \right]
+Y_l^m(\theta, \varphi) = N_l^{|m|} P_l^{|m|}(\cos \theta) \left[ \cos(m \phi) + i \sin(m \phi) \right]
 \end{equation}
@@
 \begin{equation}
-N_l^m = \sqrt{\frac{2l+1}{4 \pi} \frac{(l-m)!}{(l+m)!}}
+N_l^m = \sqrt{(2-\delta_{m,0}) \frac{2l+1}{4 \pi} \frac{(l-m)!}{(l+m)!}}
+\end{equation}
+
+where $\delta$ is the Kronecker delta defined by
+
+\begin{equation}
+\delta_{x,y} =
+\begin{cases*}
+1,  \quad \text{if } x = y \\
+0, \quad \text{otherwise} \\
+\end{cases*}
 \end{equation}
@@
 \begin{cases*}
-\sqrt{2} N_l^m \cos(m\phi) P_l^m(\cos(\theta)) \\
-N_l^0 P_l^0(\cos(\theta)) \\
-\sqrt{2} N_l^{|m|} \sin(|m| \phi) P_l^{|m|}(\cos(\theta))
+N_l^m P_l^m(\cos(\theta)) \cos(m\phi) & m \textgreater \, 0  \\
+N_l^0 P_l^0(\cos(\theta)) & m = 0 \\
+N_l^{|m|} P_l^{|m|}(\cos(\theta)) \sin(|m| \phi) & m \textless \, 0 
 \end{cases*}

### 2017-01-24 doctors/gui/xsectiondialog.h
Updated the XSection dialog so it now has tabs to display raw data and evaluated data

@@ -19,12 +19,16 @@ public:
     explicit XSectionDialog(QWidget *parent = 0);
     ~XSectionDialog();
 
-    void updateXs(XSection *xs);
-    void setXs(AmpxParser *p);
+
+
 
 private:
     Ui::XSectionDialog *ui;
     XSection *m_xs;
+
+public slots:
+    void updateXs(XSection *xs);
+    void setXs(AmpxParser *p);
 };
 
 #endif // XSECTIONDIALOG_H

### 2017-02-23 docprep/draft1.tex
Updated the original dissertation draft

@@
 
-\begin{equation} \label{eq_boltz_t_dep}
+\begin{equation} \label{eq:boltz_t_dep}
 \begin{split}
@@
 The time independent form of Eq.
-\ref{eq_boltz_t_dep} is written as
+\ref{eq:boltz_t_dep} is written as
 
-\begin{equation} \label{eq_boltz}
+\begin{equation} \label{eq:boltz}
 \begin{split}
@@
 \end{equation}
+
+The following sections show how Eq.~\ref{eq:boltz} is discretized.
+
+\subsection{Angular Discretization}
+The coordinate system used in DOCTORS is shown in Fig.~\ref{fig:coord_sys}.
+A discrete set of $N_a$ angles ($\Omega_{a} \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space.
+Particles are transported only along these discrete directions form one voxel to adjacent voxels.
+
+The $S_N$ quadrature is implemented in DOCTORS.
+There are many different quadrature sets, the $S_N$ quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
+This enables simple, direct comparison to other solvers.
+Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3.0in]{figs/coord_sys}
+    \caption{The coordinate system used in DOCTORS.
+Given an arbitrary direction, $\Omega$, $\mu$, $\eta$, and $\xi$ are its direction cosines with respect to the $x$, $y$, and $z$ axes respectively.
+$\varphi$ is the azimuthal angle (with respect to $x$ and the d $\theta$ is the polar angle (with respect to $z$).}
+    \label{fig:coord_sys}
+\end{figure}
+
+Each discrete angle is associated a weight.
+The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by Eq.~\ref{eq:disc_int}.
+
+\begin{equation} \label{eq:disc_int}
+\int_{4 \pi} f(\hat{\Omega}) d\Omega \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a
+\end{equation}
+
+Discretizing Eq.~\ref{eq:boltz} in angular space gives Eq.~\ref{eq:boltz_a} where the angle is denoted by subscript $a$.
+
+\begin{equation} \label{eq:boltz_a}
+\begin{split}
+	&\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+	\psi_{a}(\boldsymbol{r}, E) = \\
+	&\sum_{a=0}^{N_a-1} \int_0^\infty \Sigma_{s, a, a'}(\boldsymbol{r}, E' \rightarrow E) \psi_{a'}(\boldsymbol{r}, E') dE' \omega_a + S_a(\boldsymbol{r}, E)
+\end{split}
+\end{equation}
+
+\subsection{Energy Discretization}
+
+Continuous energy is discretized into $G$ groups indexed from 0 to $G-1$.
+Upscatter is assumed to be negligible.
+Therefore, Eq.~\ref{eq:boltz_a} becomes the energy discretized Eq.~\ref{eq:boltz_e} where the group number is indexed by superscript $g$.
+The group structure is shown in Fig.~\ref{fig:energy_groups}.
+
+\begin{equation} \label{eq:boltz_e}
+	\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t^g(\boldsymbol{r}) \right]
+	\psi_{a}^{g}(\boldsymbol{r}) = 
+	\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
+\end{equation}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3.0in]{figs/energy_groups}
+    \caption{The energy grid structure used in DOCTORS. The highest energy group is group 0 and the lowest energy is group $G-1$.}
+    \label{fig:energy_groups}
+\end{figure}
+
+\subsection{Spatial Discretization}
+
+Finally, Eq.~\ref{eq:boltz_e} is discretized in space to give Eq.~\ref{eq:boltz_r} which is fully discretized.
+The problem domain is split into an evenly spaced Cartesian grid.
+The grid spacing between each of the spatial dimensions does not necessarily need to be uniform, but it often is in the case of CT voxel phantoms.
+
+The problem domain must be a regular rectangular parallelpiped of size $D_x$, $D_y$, and $D_z$ in the $x$, $y$, and $z$ directions respectively.
+The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each direction.
+The total number of voxels, $N_V$ is easily computed as the multiple of each of the component directions given in Eq.~\ref{eq:n_v}.
+
+\begin{equation} \label{eq:n_v}
+	N_V = N_x N_y N_z
+\end{equation}
+
+Voxel indices are flattened according to the rule given in Eq.~\ref{eq:indx_flat}. Note that all indeces always start at zero.
+
+\begin{equation} \label{eq:indx_flat}
+	i = i_x (N_z + N_y) + i_y N_z + i_z
+\end{equation}
+
+The spatial mesh is shown in Fig.~\ref{fig:spatial_disc}.
+The size of each voxel in each dimension is computed by dividing the size of the problem domain along in that dimension by the number of mesh elements along the same direction.
+Eq.~\ref{eq:mesh_x} shows this computation for the $x$ direction.
+Analagous equations apply in the $y$ and $z$ directions as well.
+
+\begin{equation} \label{eq:mesh_x}
+\Delta x = \frac{D_x}{N_x}
+\end{equation}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3.0in]{figs/spatial_disc}
+    \caption{The spatial mesh imposed on the problem domain.}
+    \label{fig:spatial_disc}
+\end{figure}
 

### 2017-02-28 docprep/chapintro.tex
Added all of the thesis chapters to the Git repo

%%
%% This is file `chapmin.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapmin.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

Computed tomography (CT) is becoming increasingly pervasive in medical diagnostics because improved algorithms are giving doctors more access to patient information. Increased information results in faster and more accurate diagnosis. However, the CT operation itself gives a radiation dose to the patient which carries some associated risk. A more detailed measure of that risk would help doctors make informed decisions regarding whether a CT scan is warannted, and if so, which type. This can also help doctors estimate spatial dose distribution to the patient to ensure no specific organ received more dose than permissible.

Another application is in treatment.

Currently, no algorithm exists to provide this information.

The goal of this work is to provide such an algorithm.

There are two primary avaialble for the methodologies: Monte Carlo and deterministic methods. Monte Carlo has proven to be very slow in these areas. Furhter, the 3D Cartesian mesh generated from CT reconstruction is well suited to deterministic computation.

This work forcuses on the discrete ordinates method and presents an implementation specially developed for dose estimation from a medical CT scan.

\endinput
%%
%% End of file `chapmin.tex'.

### 2017-02-28 docprep/chapmeth.tex
Added all of the thesis chapters to the Git repo

%%
%% This is file `chapmin.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapmin.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

\section{Mathmatic Development}

Where does this text go?

\subsection{The Boltzmann Equation}

The time dependent gamma transport equation can be written as

\begin{equation} \label{eq:boltz_t_dep}
\begin{split}
	&\left[ \frac{1}{v(E)} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
	\psi(\boldsymbol{r}, E, \hat{\Omega}, t) = \\
	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}, t) \psi(\boldsymbol{r}, E', \hat{\Omega}', t) dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega}, t)
\end{split}
\end{equation}

However, we are typically not concerned with the transient case in medical diagnostic imaging. We are more interested in the steady state case. The time independent form of Eq. \ref{eq:boltz_t_dep} is written as

\begin{equation} \label{eq:boltz}
\begin{split}
	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
	\psi(\boldsymbol{r}, E, \hat{\Omega}) = \\
	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega})
\end{split}
\end{equation}

The following sections show how Eq.~\ref{eq:boltz} is discretized.

\subsection{Angular Discretization}
The coordinate system used in DOCTORS is shown in Fig.~\ref{fig:coord_sys}. A discrete set of $N_a$ angles ($\Omega_{a} \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space. Particles are transported only along these discrete directions form one voxel to adjacent voxels.

The $S_N$ quadrature is implemented in DOCTORS. There are many different quadrature sets, the $S_N$ quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers. This enables simple, direct comparison to other solvers. Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/coord_sys}
  \end{center}
  \caption{The coordinate system used in DOCTORS. Given an arbitrary direction, $\Omega$, $\mu$, $\eta$, and $\xi$ are its direction cosines with respect to the $x$, $y$, and $z$ axes respectively. $\varphi$ is the azimuthal angle (with respect to $x$ and the d $\theta$ is the polar angle (with respect to $z$).}
\label{fig:coord_sys}
\end{figure}%

A given direction, $\hat{\Omega}$ is determined by its three cosine components. This relation is given in Eq.~\ref{eq:omega_cos} where $\hat{i}$, $\hat{j}$, and $\hat{k}$ are the unit directions in the $x$, $y$, and $z$ directions respectively.

\begin{equation} \label{eq:omega_cos}
	\hat{\Omega} = \mu \hat{i} + \eta \hat{j} + \xi \hat{k}
\end{equation}

Each discrete angle is associated a weight. The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by Eq.~\ref{eq:disc_int}.

\begin{equation} \label{eq:disc_int}
\int_{4 \pi} f(\hat{\Omega}) d\Omega \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a
\end{equation}

Discretizing Eq.~\ref{eq:boltz} in angular space gives Eq.~\ref{eq:boltz_a} where the angle is denoted by subscript $a$.

\begin{equation} \label{eq:boltz_a}
\begin{split}
&\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
\psi_{a}(\boldsymbol{r}, E) = \\
&\sum_{a=0}^{N_a-1} \int_0^\infty \Sigma_{s, a, a'}(\boldsymbol{r}, E' \rightarrow E) \psi_{a'}(\boldsymbol{r}, E') dE' \omega_a + S_a(\boldsymbol{r}, E)
\end{split}
\end{equation}

\subsection{Energy Discretization}

Continuous energy is discretized into $G$ groups indexed from 0 to $G-1$. Upscatter is assumed to be negligible. Therefore, Eq.~\ref{eq:boltz_a} becomes the energy discretized Eq.~\ref{eq:boltz_e} where the group number is indexed by superscript $g$. The group structure is shown in Fig.~\ref{fig:energy_groups}.

\begin{equation} \label{eq:boltz_e}
\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t^g(\boldsymbol{r}) \right]
\psi_{a}^{g}(\boldsymbol{r}) = 
\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
\end{equation}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/energy_groups}
  \end{center}
  \caption{The energy grid structure used in DOCTORS. The highest energy group is group 0 and the lowest energy is group $G-1$.}
\label{fig:energy_groups}
\end{figure}%

\subsection{Spatial Discretization}

Finally, Eq.~\ref{eq:boltz_e} is discretized in space to give Eq.~\ref{eq:boltz_r} which is fully discretized. The problem domain is split into an evenly spaced Cartesian grid. The grid spacing between each of the spatial dimensions does not necessarily need to be uniform, but it often is in the case of CT voxel phantoms.

The problem domain must be a regular rectangular parallelpiped of size $D_x$, $D_y$, and $D_z$ in the $x$, $y$, and $z$ directions respectively. The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each direction. The total number of voxels, $N_V$ is easily computed as the multiple of each of the component directions given in Eq.~\ref{eq:n_v}.

\begin{equation} \label{eq:n_v}
	N_V = N_x N_y N_z
\end{equation}

Voxel indices are flattened according to the rule given in Eq.~\ref{eq:indx_flat}. Note that all indeces always start at zero.

\begin{equation} \label{eq:indx_flat}
	i = i_x (N_z + N_y) + i_y N_z + i_z
\end{equation}

The spatial mesh is shown in Fig.~\ref{fig:spatial_disc}. The size of each voxel in each dimension is computed by dividing the size of the problem domain along in that dimension by the number of mesh elements along the same direction. Eq.~\ref{eq:mesh_x} shows this computation for the $x$ direction. Analagous equations apply in the $y$ and $z$ directions as well.

\begin{equation} \label{eq:mesh_x}
\Delta x = \frac{D_x}{N_x}
\end{equation}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/spatial_disc}
  \end{center}
  \caption{The spatial mesh imposed on the problem domain.}
\label{fig:spatial_disc}
\end{figure}%

The index of interest is denoted as a subscript $i$ which makes the fully discretized form of the steady state Linear Boltzmann Equation given in Eq.~\ref{eq:boltz} into the fully discretized form given in Eq.~\ref{eq:boltz_i}

\begin{equation} \label{eq:boltz_i}
\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_{t,i}^g \right]
\psi_{i,a}^{g} = 
\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
\end{equation}

\subsection{Solution}

To solve the fully discretized form of the Linear Boltzmann Equation given in Eq.~\ref{eq:boltz_i}, the gradient operator must first be computed numerically. Recall that $\Omega_a$ can be written in vector notation using its cosine components as defined in Eq.~\ref{eq:omega_cos}. Also recall the definition of the gradient, given in Eq.~\ref{eq:grad}.

\begin{equation} \label{eq:grad}
\nabla f(x, y, z) = \frac{\partial f(x, y, z)}{\partial x} \hat{i} + \frac{\partial f(x, y, z)}{\partial y} \hat{j} + \frac{\partial f(x, y, z)}{\partial z} \hat{k}
\end{equation}

To the first order, a partial derivitive can be computed using Eq.~\ref{eq:deriv_1}.

\begin{equation} \label{eq:deriv_1}
\frac{\partial f(x)}{\partial x} \approx \frac{f(x) - f(x + \Delta x)}{\Delta x}
\end{equation}

Using the definition of $\Omega$ and Eq.~\ref{eq:grad} the $\hat{\Omega} \cdot \nabla \psi$ term can be rewritten using Eq.~\ref{eq:spatial_1}

\begin{equation} \label{eq:deriv_1}
\begin{split}
\Omega \cdot \nabla \psi & \approx 
\left\langle \mu, \eta, \xi \right\rangle \cdot
\left\langle \frac{\psi(x) - \psi(x + \Delta x)}{\Delta x},
\frac{\psi(y + \Delta y) - \psi(y)}{\Delta y},
\frac{\psi(z + \Delta z) - \psi(z)}{\Delta z} \right\rangle \\
& \approx 
\frac{\psi(x + \Delta x) - \psi(x)}{\Delta x} \mu + 
\frac{\psi(y + \Delta y) - \psi(y)}{\Delta y} \eta + 
\frac{\psi(z + \Delta z) - \psi(z)}{\Delta z} \xi
\end{split}
\end{equation}

Along some direction, $\Omega$, as it enters an arbitrary voxel, it must pass through either of the sides indicated in red and exit via the other as shown in Fig.~\ref{fig:gradient}. Fig.~\ref{fig:gradient} uses the $yz$ faces of the voxel for illustration. If $\eta$ is positive, $\Omega$ will enter through the $x_i$ surface and exit through the $x_{i+1}$ surface. If $\eta$ is negative, this order will be reversed. A limitation to this methodology is that $\mu$, $\eta$, and $\xi$ are assumed to be nonzero. This is ensured as long as none of the directions in the quadrature are parallel to a major axis.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/gradient}
  \end{center}
  \caption{The caption.}
\label{fig:gradient}
\end{figure}%

Depending on the sign of $\eta$, the first order approximation of the a derivitive along the $y$ direction is given by Eq.~\ref{eq:grad_y}.

\begin{equation} \label{eq:grad_y}
\begin{split}
\frac{\partial f}{\partial y} &= \frac{f(x_i) - f(x_{i+1})}{\Delta x}, \quad \eta>0 \\
\frac{\partial f}{\partial y} &= \frac{f(x_{i+1}) - f(x_{i})}{\Delta x}, \quad \eta <0 \\
\frac{\partial f}{\partial y} &= 0, \quad \eta=0
\end{split}
\end{equation}

Equation~\ref{eq:grad_y} can be generalized to use $f_{in}$ and $f_{out}$ instead of relying on the $x$-coordinate. This is given in Eq.~\ref{eq:grad_y2}

\begin{equation} \label{eq:grad_y2}
\begin{split}
\frac{\partial f}{\partial y} &= \frac{f_{in} - f_{out}}{\Delta x}, \quad \eta \neq 0 \\
\frac{\partial f}{\partial y} &= 0, \quad \eta=0
\end{split}
\end{equation}

The generalization of Eq.~\ref{eq:grad_y2} is applied to Eq.~\ref{eq:deriv_1} to yield Eq.~\ref{eq:deriv_2} which is generally valid for all octants.

\begin{equation} \label{eq:deriv_2}
\Omega \cdot \nabla \psi \approx 
\frac{\psi_{x,in} - \psi_{x,out}}{\Delta x} \mu + 
\frac{\psi_{y,in} - \psi_{y,out}}{\Delta y} \eta + 
\frac{\psi_{z,in} - \psi_{z,out}}{\Delta z} \xi
\end{equation}

Using the approximation given in Eq.~\ref{eq:deriv_2}, Eq.~\ref{eq:boltz_i} can be rewritten as Eq.~\ref{eq:boltz_i2}.

\begin{equation} \label{eq:boltz_i2}
\left[ 
\frac{\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
\frac{\psi_{i,a,y,in}^g - \psi_{i,a,y,out}^g}{\Delta y} \eta + 
\frac{\psi_{i,a,z,in}^g - \psi_{i,a,z,out}^g}{\Delta z} \xi
\right]
+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
\end{equation}

It is useful to point out that each of $\psi_{i,a,x,in}^g$, $\psi_{i,a,x,out}^g$, $\psi_{i,a,y,in}^g$, $\psi_{i,a,y,out}^g$, $\psi_{i,a,z,in}^g$, and $\psi_{i,a,z,out}^g$ are \textit{surface} flux values while $\psi_{i,a}^{g}$ is a \textit{volume averaged} flux value.

In order for the discretized Linear Boltzmann Equation to be solved using Eq.~\ref{eq:boltz_i2}, the incoming flux must always be known. Either the boundary conditions or previously calculated values are used to determine each incoming flux. Therefore, in order to compute the average flux in a cell, some relationship between the average flux and the outgoing surface fluxes must be known. These values can be related by any one of many different mechanisms, the most common of which is the diamond difference approximation.

\subsection{Diamond Difference Approximation}

In the diamond difference approximation, it is assumed that the incoming and outgoing flux are both known. The average flux is assumed to be the average of the two surface fluxes. This is assumed to be the case in all three spatial dimensions. This leads to Eq.~\ref{eq:dd}.

\begin{equation} \label{eq:dd}
\begin{split}
\frac{\psi_{i,a,x,in}^g + \psi_{i,a,x,out}^g}{2} &= \psi_{i,a}^{g} \\
\frac{\psi_{i,a,y,in}^g + \psi_{i,a,y,out}^g}{2} &= \psi_{i,a}^{g} \\
\frac{\psi_{i,a,z,in}^g + \psi_{i,a,z,out}^g}{2} &= \psi_{i,a}^{g}
\end{split}
\end{equation}

Multiplying both sides of Eq.~\ref{eq:dd} and subtracting twice the outgoing surface flux from each yields Eq.~\ref{eq:dd2}.

\begin{equation} \label{eq:dd2}
\begin{split}
\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g \\
\psi_{i,a,y,in}^g - \psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,out}^g \\
\psi_{i,a,z,in}^g - \psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,out}^g
\end{split}
\end{equation}

The form of the left hand side of Eq.~\ref{eq:dd2} is the same as in Eq.~\ref{eq:boltz_i2}. This allows Eq.~\ref{eq:dd2} to be used to make a substitution in Eq.~\ref{eq:boltz_i2} that removes the dependence on the outgoing flux from discretized equation. After the substitution, Eq.~\ref{eq:boltz_i3} will arise.

\begin{equation} \label{eq:boltz_i3}
\left[ 
\frac{2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
\frac{2\psi_{i,a}^{g} - \psi_{i,a,y,out}^g}{\Delta y} \eta + 
\frac{2\psi_{i,a}^{g} - \psi_{i,a,z,out}^g}{\Delta z} \xi
\right]
+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
\end{equation}

The diamond difference approximation defined in Eq.~\ref{eq:dd} can also be rearranged to solve for each outgoing flux once the volume averaged flux is computed. The outgoing flux is given by Eq.~\ref{eq:dd3}.

\begin{equation} \label{eq:dd3}
\begin{split}
\psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,in}^g \\
\psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,in}^g \\
\psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,in}^g
\end{split}
\end{equation}

Rearranging Eq.~\ref{eq:boltz_i3} to factor out the volume averaged flux term yields Eq.~\ref{eq:boltz_i4}.

\begin{equation} \label{eq:boltz_i4}
\left[ 
\frac{2\psi_{i,a}^{g}}{\Delta x} \mu + 
\frac{2\psi_{i,a}^{g}}{\Delta y} \eta + 
\frac{2\psi_{i,a}^{g}}{\Delta z} \xi
\right] - 
\left[ 
\frac{2\psi_{i,a,x,out}^g}{\Delta x} \mu + 
\frac{2\psi_{i,a,y,out}^g}{\Delta y} \eta + 
\frac{2\psi_{i,a,z,out}^g}{\Delta z} \xi
\right]
+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
\end{equation}

Equation~\ref{eq:boltz_i4} can be directly solved for the volume averaged flux. The result is given in Eq.~\ref{eq:boltz_i5}.

\begin{equation} \label{eq:boltz_i5}
\psi_{i,a}^{g} = 
\frac{
  \sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i,     a'}^{g'} \omega_a + 
  \left[ 
    \frac{2\psi_{i,a,x,out}^g}{\Delta x} \mu + 
    \frac{2\psi_{i,a,y,out}^g}{\Delta y} \eta + 
    \frac{2\psi_{i,a,z,out}^g}{\Delta z} \xi
  \right] + S_{i,a}^g
}{
  \frac{2\mu}{\Delta x}  + 
  \frac{2\eta}{\Delta y} + 
  \frac{2\xi}{\Delta z} + 
  \Sigma_{t,i}^g
}
\end{equation}

\subsection{Anisotropy Treatment}

Scatter is assumed to be a function of only the scatter angle between the initial and scattered directions, $\theta_s$. Often, the cosine of the scatter angle, $\mu_s$ is used.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/scat_ang}
  \end{center}
  \caption{The scatter angle. A photon at energy $E$ traveling in direction $\Omega$, hits a stationary atom, indicated by the green circle, and scatters into a new direction $\Omega'$ with a new energy $E'$.}
\label{fig:scat_ang}
\end{figure}%

The data files containing the scatter cross sections are distributed using a Legendre polynomial expansion. The use of a Legendre expansion removes the dependence on the quadrature from the data. This allows any quadrature to work with any dataset. A $l$-order Legendre polynomial is donoted by $P_l$. The anisotropic scatter cross section can then be rewritten as a Legendre expansion given in Eq.~\ref{eq:leg_1} where the scatter cosine is defined in Eq.~\ref{eq:scat_cos}

\begin{equation} \label{eq:leg_1}
\Sigma_{s, i, a, a'}^{g, g'} = \sum_l=0^L \frac{2l+1}{4 \pi}\Sigma_{s, i, l}^{g, g'} P_l(\mu_s)
\end{equation}

\begin{equation} \label{eq:scat_cos}
\mu_s = \Omega_a \cdot \Omega_{a'}
\end{equation}

\endinput
%%
%% End of file `chapmin.tex'.

### 2017-02-28 docprep/colocation.tex
Added all of the thesis chapters to the Git repo

\documentclass{article}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[toc,page]{appendix}
\usepackage{listings}
\usepackage{color}
\usepackage{listings}
\usepackage{fullpage}
\usepackage{mathtools}
\numberwithin{equation}{subsection}

\begin{document}

%%---------------------------------------------------------------------------%%
%% collocation_methods.tex
%%---------------------------------------------------------------------------%%

%%---------------------------------------------------------------------------%%
\chapter{Collocation Angular Methods}
\label{cha:coll-angul-meth}
%%---------------------------------------------------------------------------%%

Many deterministic transport methods treat the angular terms on the left-hand
side of Eq.~(\ref{eq:mg-transport}) using a finite-element collocation method.
Foremost among these are discrete ordinates (\Sn), method of characteristics
(MOC), and short characteristics schemes. In each of these methods, the
right-hand side scattering kernel is expanded using Spherical Harmonics.  A
finite element collocation discretization is then applied to the angular
terms. The expansion of the scattering and source terms are described in
\S\S~\ref{sec:angul-discr-scatt-1}--\ref{sec:angul-discr-extern}.  The angular
collocation scheme is explained in \S~\ref{sec:angul-coll-discr}.

%%---------------------------------------------------------------------------%%
\section{Angular Discretization of Scattering Terms}
\label{sec:angul-discr-scatt-1}
%%---------------------------------------------------------------------------%%

Equation~(\ref{eq:mg-transport}) is still an integro-differential equation in
which the principal unknown, $\psi$, is present in the scattering integral.
Additionally, the multigroup scattering cross section is a function of angle.
Accordingly, we need some method for dealing with the angular dependence of
the scattering source.  We tackle this problem by recognizing that
$\sigma_{\text{s}}$ is only a function of the cosine between the incoming and
outgoing angles (the polar scattering angle assuming azimuthal symmetry).
Therefore, we can expand the scattering cross section in Legendre polynomials,
\begin{equation}
  \sigg{}(\vOmega'\cdot\vOmega)=\sum_{l=0}^{N}
  \frac{2l+1}{4\pi}P_{l}(\vOmega'\cdot\vOmega)
  \sigg{l}\:.
  \label{eq:legendre-expansion-scattering}
\end{equation}
Integrating over all angles we calculate the total scattering cross section as
follows:
\begin{equation}
  \begin{split}
    \sigma_{\text{s}}^{gg'} &=
    \int_{4\pi}\sigg{}(\vOmega'\cdot\vOmega)\:d\vOmega \equiv
    2\pi\int_{-1}^{1}\sigg{}(\mu_o)\:d\mu_o\\
    %%
    &=\sum_{l=0}^{N}\frac{2l+1}{2}\sigg{l}\int_{-1}^{1}
    P_0(\mu_o)P_l(\mu_o)\:d\mu_o\\
    %%
    &=\sigg{0}\:,
  \end{split}
\end{equation}
where we have used the orthogonality of Legendre polynomials listed in
Chap.~\ref{cha:math-prop-ident}.

Applying Eq.~(\ref{eq:legendre-expansion-scattering}) in
Eq.~(\ref{eq:mg-transport}) gives a scattering source defined by
\begin{equation}
  q^g_{s}(\vOmega) = \sum_{g'=0}^G \int_{4\pi}
  \sum_{l=0}^N\frac{2l+1}{4\pi}P_{l}(\vOmega'\cdot\vOmega)\sigg{l}
  \psi^{g'}(\vOmega')\:d\vOmega'\:,
  \label{eq:scattering-source-before-addition}
\end{equation}
where we have suppressed the spatial dependence, $\ve{r}$.  The
addition theorem of Spherical Harmonics can be used to evaluate the
Legendre function, $P_l(\vOmega'\cdot\vOmega)$,
\begin{equation}
  P_l(\vOmega'\cdot\vOmega) = \frac{4\pi}{2l+1}\sum_{m=-l}^l
  Y_{lm}(\vOmega)Y^{\ast}_{lm}(\vOmega')\:,
\end{equation}
where the $Y_{lm}$ are defined in Eq.~(\ref{eq:complete-spherical-harmonics}).
The scattering must be real; therefore, we can follow a methodology similar to
the techniques described in Chap.~\ref{cha:spher-harm-expans} that shows how
to expand a real-valued function using complex Spherical Harmonics.  First,
the expansion is split into positive and negative components of $m$,
\begin{equation}
  P_l(\vOmega'\cdot\vOmega) = \frac{4\pi}{2l+1}
  \Bigl[
  Y_{l0}(\vOmega)Y_{l0}(\vOmega') +
  \sum_{m=1}^l
  \bigl(Y_{lm}(\vOmega)Y^{\ast}_{lm}(\vOmega') +
  Y_{l-m}(\vOmega)Y^{\ast}_{l-m}(\vOmega')\bigr)\Bigr]\:.
\end{equation}
Examining the $m=0$ term gives the following result
\begin{equation}
  Y_{l0} = \sqrt{\frac{2l+1}{4\pi}}\,P_{l0} = Y^e_{l0}\:,
\end{equation}
where the $Y^e$ are defined in Eq.~(\ref{eq:Ye}).

Expanding the Spherical Harmonics into real and imaginary components as shown
in Eq.~(\ref{eq:complex-Y}), the sum over $m>0$ becomes
\begin{equation}
  \sum_{m=1}^l\Bigl(
  \hat{Y}^e_{lm}(\vOmega)\hat{Y}^e_{lm}(\vOmega') +
  \hat{Y}^o_{lm}(\vOmega)\hat{Y}^o_{lm}(\vOmega') +
  \hat{Y}^e_{l-m}(\vOmega)\hat{Y}^e_{l-m}(\vOmega') +
  \hat{Y}^o_{l-m}(\vOmega)\hat{Y}^o_{l-m}(\vOmega')\Bigr)\:,
\end{equation}
where the imaginary terms have been set to zero because the scattering must be
real.  Using Eqs.~(\ref{eq:Yem-Ye-m}) and (\ref{eq:Yom-Yo-m}), the summation
becomes
\begin{equation}
  \sum_{m=1}^l\Bigl(
  2\hat{Y}^e_{lm}(\vOmega)\hat{Y}^e_{lm}(\vOmega') +
  2\hat{Y}^o_{lm}(\vOmega)\hat{Y}^o_{lm}(\vOmega')\Bigr)
\end{equation}
Comparing Eqs.~(\ref{eq:hat-Ye}) and (\ref{eq:hat-Yo}) with Eqs.~(\ref{eq:Ye})
and (\ref{eq:Yo}) leads to the following relationships,
\begin{equation}
  \hat{Y}^e_{lm} = \frac{1}{\sqrt{2}}Y^e_{lm}\:,\quad
  \hat{Y}^o_{lm} = \frac{1}{\sqrt{2}}Y^o_{lm}\:.
\end{equation}
After applying these equations in the $m>0$ terms and combining with the $m=0$
term described above, the expression for $P_l(\vOmega\cdot\vOmega')$ is
\begin{equation}
  P_l(\vOmega'\cdot\vOmega) = \frac{4\pi}{2l+1}
  \Bigl[
  Y^e_{l0}(\vOmega)Y^e_{l0}(\vOmega') +
  \sum_{m=1}^l
  \bigl(Y^e_{lm}(\vOmega)Y^e_{lm}(\vOmega') +
  Y^o_{lm}(\vOmega)Y^o_{lm}(\vOmega')\bigr)\Bigr]\:,
  \label{eq:P_l(mu_o)}
\end{equation}
where, as shown in Chap.~\ref{cha:spher-harm-expans}, the $Y^e$ and $Y^o$ form
an orthonormal basis.

Returning to the scattering source defined in
Eq.~(\ref{eq:scattering-source-before-addition}), Eq.~(\ref{eq:P_l(mu_o)})
provides the Legendre polynomial for the cosine of the scattering angle, and
\begin{equation}
    q^g_{s}(\vOmega) = \sum_{g'=0}^G \int_{4\pi}
    \sum_{l=0}^N
    \Bigl[Y^e_{l0}(\vOmega)Y^e_{l0}(\vOmega')
    +\sum_{m=1}^l
    \bigl(Y^e_{lm}(\vOmega)Y^e_{lm}(\vOmega') +
    Y^o_{lm}(\vOmega)Y^o_{lm}(\vOmega')\bigr)\Bigr]
    \sigg{l}
    \psi^{g'}(\vOmega')\:d\vOmega'\:.
\end{equation}
Rearranging terms and defining
\begin{alignat}{3}
  \even_{lm} &= \int_{4\pi}Y^e_{lm}(\vOmega)\psi^g(\vOmega)\:d\vOmega\:,
  \quad& m\ge 0\:,\label{eq:even-flux}\\
  %%
  \odd_{lm} &= \int_{4\pi}Y^o_{lm}(\vOmega)\psi^g(\vOmega)\:d\vOmega\:,
  \quad& m>0\:,\label{eq:odd-flux}
\end{alignat}
which follows directly from Eqs.~(\ref{eq:even_moment}) and
(\ref{eq:odd_moment}), the scattering source becomes
\begin{equation}
  q^g_{s}(\ve{r},\vOmega) = \sum_{g'=0}^G
  \sum_{l=0}^N
  \sigg{l}(\ve{r})
  \Bigl[
  Y^e_{l0}(\vOmega)\evenp_{l0}(\ve{r}) +
  \sum_{m=1}^l
  \bigl(
  Y^e_{lm}(\vOmega)\evenp_{lm}(\ve{r}) +
  Y^o_{lm}(\vOmega)\oddp_{lm}(\ve{r})
  \bigr)\Bigr]\:.
  \label{eq:mg-scattering-source}
\end{equation}
Equation~(\ref{eq:mg-scattering-source}) is the multigroup anisotropic
scattering source that is defined by the order of the Legendre expansion,
$P_N$, of the scattering.  For a given $P_N$ order, $(N+1)^2$ moments are
required to integrate the scattering operator.  The moments in
Eqs.~(\ref{eq:even-flux}) and (\ref{eq:odd-flux}) are the \textit{angular flux
  moments} or, simply, flux moments.

The scalar flux is defined in Eq.(\ref{eq:scalar_flux}) as the zeroth moment
of the angular flux.  Therefore, we have
\begin{equation}
  \phi^g = \int_{4\pi}\psi^g\:d\vOmega = \sqrt{4\pi}\int_{4\pi}Y^e_{00}
  \psi^g\:d\vOmega = \sqrt{4\pi}\even_{00}\:.
\end{equation}
The current is defined
\begin{equation}
  \begin{split}
    \ve{J}^g &= \int_{4\pi}\Bigl[\mu\psi^g\hat{\ve{e}}_x +
    \eta\psi^g\hat{\ve{e}}_y +
    \xi\psi^g\hat{\ve{e}}_z\Bigr]\:d\vOmega\\
    %%
    &=-\sqrt{\frac{4\pi}{3}}\even_{11}\hat{\ve{e}}_x -
    \sqrt{\frac{4\pi}{3}}\odd_{11}\hat{\ve{e}}_y +
    \sqrt{\frac{4\pi}{3}}\even_{10}\hat{\ve{e}}_z\:.
  \end{split}
\end{equation}
More concisely,
\begin{equation}
  J_x^g = -\sqrt{\frac{4\pi}{3}}\even_{11}\:,\quad
  J_y^g = -\sqrt{\frac{4\pi}{3}}\odd_{11}\:,\quad
  J_z^g = \sqrt{\frac{4\pi}{3}}\even_{10}\:,
\end{equation}
for $\ve{J}= J_x\hat{\ve{e}}_x + J_y\hat{\ve{e}}_y
+J_z\hat{\ve{e}}_z$.

%%---------------------------------------------------------------------------%%
\subsection{Scattering Expansions in 2D Geometry}
\label{sec:scatt-expans-2d}
%%---------------------------------------------------------------------------%%

To expand the scattering kernel in 2D geometry, we must define a symmetry
plane.  We choose the $XY$-plane as the symmetry plane; although the
$XZ$-plane could be chosen as well.  The resulting 2D $XY$ coordinate system
is shown in Fig.~\ref{fig:2D_coord}.
\begin{figure}
  \begin{center}
    \input{2D_coord.pdftex_t}
  \end{center}
  \caption{2D coordinate system used in \denovo.  Symmetry is defined about
    the $XY$ plane.}
  \label{fig:2D_coord}
\end{figure}
The anisotropic scattering source is defined in
Eq.~(\ref{eq:mg-scattering-source}).  Using symmetry, the number of terms
required to define this source can be reduced.  The symmetry illustrated in
Fig.~\ref{fig:2D_coord} yields the following constraint equation
\begin{equation}
  \psi(\theta, \varphi) = \psi(\pi - \theta, \varphi)\:,\quad
  0\le\theta\le\frac{\pi}{2}\:.
\end{equation}
With this constraint in mind, we expand the integral in
Eq.~(\ref{eq:even-flux}) as follows
\begin{equation}
  \phi_{lm} = \int_{0}^{2\pi}d\varphi\int_{0}^{\pi}
  Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)\sin\theta\,d\theta\:.
  \label{eq:expanded-even-moment}
\end{equation}
The inner integral over the polar angle can be split into 2 components,
\begin{equation}
    \int_{0}^{\pi}
    Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)\sin\theta\,d\theta =
    \int_{0}^{\pi/2}Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)
    \sin\theta\,d\theta +
    \int_{\pi/2}^{\pi}Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)
    \sin\theta\,d\theta\:.
\end{equation}
Transforming variables to $\theta'=\pi-\theta$ in the second integral on the
right-hand side gives
\begin{equation}
  \int_{\pi/2}^{\pi}Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)
  \sin\theta\,d\theta = \int_{0}^{\pi/2}
  Y^e_{lm}(\pi-\theta,\varphi)\psi(\pi-\theta,\varphi)
  \sin(\pi-\theta)\,d\theta\:.
\end{equation}
Applying $\sin(\pi-\theta) = \sin\theta$ and the symmetry constraint to the
preceding two equations and substituting into
Eq.~(\ref{eq:expanded-even-moment}), the even angular flux moments become
\begin{equation}
  \phi_{lm} = \int_{0}^{2\pi}d\varphi
  \int_{0}^{\pi/2}\bigl(Y^e_{lm}(\theta,\varphi) +
  Y^e_{lm}(\pi-\theta,\varphi)\bigr)\psi(\theta,\varphi)
    \sin\theta\,d\theta\:.
\end{equation}
Substituting Eq.~(\ref{eq:Ye}) into the preceding expression gives
\begin{equation}
  \phi_{lm} = \int_{0}^{2\pi}d\varphi
  \int_{0}^{\pi/2}
  D_{lm}\cos m\varphi\Bigl(
  P_{lm}(\cos\theta) +
  P_{lm}\bigl(\cos(\pi-\theta)\bigr)
  \Bigr)
  \psi(\theta,\varphi)\sin\theta\,d\theta\:.
\end{equation}
Using $\cos(\pi-\theta)=-\cos\theta$ and applying the following identity
\cite{arfken},
\begin{equation}
  P_{lm}(-x) = (-1)^{l+m}P_{lm}(x)\:,
\end{equation}
we derive the following formula for calculating the even angular flux moments,
\begin{equation}
  \phi_{lm} = \int_{0}^{2\pi}d\varphi
  \int_{0}^{\pi/2}
  \bigl(1 + (-1)^{l+m}\bigr)
  Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)\sin\theta\,d\theta\:.
\end{equation}
Using the same procedure for the odd moments, the angular flux moments are
defined
\begin{align}
  \phi_{lm} &= \int_{0}^{2\pi}d\varphi
  \int_{0}^{1}
  \bigl(1 + (-1)^{l+m}\bigr)
  Y^e_{lm}(\theta,\varphi)\psi(\theta,\varphi)\,d\xi\:,
  \label{eq:2D-even-moments}\\
  \vartheta_{lm} &= \int_{0}^{2\pi}d\varphi
  \int_{0}^{1}
  \bigl(1 + (-1)^{l+m}\bigr)
  Y^o_{lm}(\theta,\varphi)\psi(\theta,\varphi)\,d\xi\:,
  \label{eq:2D-odd-moments}
\end{align}
where we have written the integrals over $d\theta$ using the substitution
$\xi=\cos\theta$.  Equations~(\ref{eq:2D-even-moments}) and
(\ref{eq:2D-odd-moments}) are the 2D equivalents of Eqs.~(\ref{eq:even-flux})
and (\ref{eq:odd-flux}) when symmetry is defined about the $XY$ plane.

A brief examination of Eqs.~(\ref{eq:2D-even-moments}) and
(\ref{eq:2D-odd-moments}) reveals that for all odd sums of $l+m$ (ie.
$l+m\in\{2k+1; \forall k\in Z\}$) the angular moments vanish. This results in
$(N+1)(N+2)/2$ angular moments for $P_N$ scattering whereas 3D calculations
require $(N+1)^2$ moments.  Furthermore, we only need to consider quadrature
angles in the $0\le\xi\le 1$ directions and the factor of 2 that results when
$l+m$ is even preserves the integration over $4\pi$; this is a natural result
of symmetry in the polar direction.  Three-dimensional space contains 8
octants that consist of 4 octants for $+\xi$ and 4 octants for $-\xi$.  In 2D
space, only the 4 octants in $+\xi$ are required.  In 2D geometry these 4
octants are referred to as quadrants.

%%---------------------------------------------------------------------------%%
\section{Angular Discretization of External Sources}
\label{sec:angul-discr-extern}
%%---------------------------------------------------------------------------%%

In most cases, the external source will be a known function of angle.  For
example, an isotropic external source is
\begin{equation}
  q^g_e(\vOmega) = \frac{s^g}{4\pi}\:.
\end{equation}
However, in certain cases (e.g. coupling to a $k$-eigenvalue calculation) the
external source is defined in moments.
Equation~(\ref{eq:Spherical_Harmonic_Expansion_f}) gives the Spherical
Harmonics expansion of a real-valued function.  Applying the same methodology
gives the expansion of the external source
\begin{equation}
  q^g_e(\vOmega) = \sum_{l=0}^{N}\Bigl[
  Y^e_{l0}(\vOmega)\qe_{l0} +
  \sum_{m=1}^{l}
  \bigr(
  Y^e_{lm}(\vOmega)\qe_{lm} + Y^o_{lm}(\vOmega)\qo_{lm}\bigr)
  \Bigr]\:,
  \label{eq:mg-external-source}
\end{equation}
where the spatial dependence has been suppressed.  The even and odd source
moments are defined
\begin{alignat}{3}
  \qe_{lm} &= \int_{4\pi}Y^e_{lm}(\vOmega)q^g_e(\vOmega)\:d\vOmega\:,
  \quad&m\ge 0\:,\label{eq:even-source}\\
  \qo_{lm} &= \int_{4\pi}Y^o_{lm}(\vOmega)q^g_e(\vOmega)\:d\vOmega\:,
  \quad&m>0\:.\label{eq:odd-source}
\end{alignat}
Now, using Eqs.(\ref{eq:mg-scattering-source}) and
(\ref{eq:mg-external-source}), we write the entire source as
\begin{equation}
  \begin{split}
    Q^g(\ve{r},\vOmega) &= q^g_s(\ve{r},\vOmega) +
    q^g_e(\ve{r},\vOmega)\\
    %%
    &=\sum_{g'=0}^G
    \sum_{l=0}^N
    \sigg{l}(\ve{r})
    \Bigl[
    Y^e_{l0}(\vOmega)\evenp_{l0}(\ve{r}) +
    \sum_{m=1}^l
    \bigl(
    Y^e_{lm}(\vOmega)\evenp_{lm}(\ve{r}) +
    Y^o_{lm}(\vOmega)\oddp_{lm}(\ve{r})
    \bigr)\Bigr]\\
    %%
    &\quad+
    \sum_{l=0}^{N}\Bigl[
    Y^e_{l0}(\vOmega)\qe_{l0}(\ve{r}) +
    \sum_{m=1}^{l}
    \bigl(
    Y^e_{lm}(\vOmega)\qe_{lm}(\ve{r}) +
    Y^o_{lm}(\vOmega)\qo_{lm}(\ve{r})\bigr)
    \Bigr]\:.
  \end{split}
  \label{eq:mg-source}
\end{equation}

%%---------------------------------------------------------------------------%%
\section{Angular Collocation Discretization}
\label{sec:angul-coll-discr}
%%---------------------------------------------------------------------------%%

Combining Eqs.~(\ref{eq:mg-transport}) and (\ref{eq:mg-source}) gives the
multigroup transport equation with the scattering expanded in Spherical
Harmonics,
\begin{equation}
  \hOmega\cdot\grad\psi^g(\vOmega) +
  \sigma^g\psi^g(\vOmega) = Q^g(\vOmega)\:
  \label{eq:mg-transport-scattering+source}
\end{equation}
where the spatial dependence has been suppressed.  While Spherical Harmonics
have been used to expand the scattering sources (and possibly the external
source), we still have a dependence on $\vOmega$ that needs to be resolved.
We apply the discrete ordinates ($S_N$) approximation, which is a collocation
method in angle.  Solving Eq.~(\ref{eq:mg-transport-scattering+source}) at
discrete angular locations requires the following equation,
\begin{equation}
    \hOmega_a\cdot\grad\psi^g_a +
    \sigma^g\psi^g_a
    =\sum_{g'=0}^G
    \sum_{l=0}^N
    \sigg{l}
    \Bigl[
    Y^e_{l0}(\vOmega_a)\evenp_{l0} +
    \sum_{m=1}^l
    \bigl(
    Y^e_{lm}(\vOmega_a)\evenp_{lm} +
    Y^o_{lm}(\vOmega_a)\oddp_{lm}
    \bigr)\Bigr]\\
    + q_e^g(\vOmega_a)\:,
    \label{eq:MG-SN-Equation}
 \end{equation}
where $\psi^g_a \equiv\psi^g(\vOmega_a)$.  The angles are integrated by a
quadrature rule such that
\begin{equation}
  \int_{4\pi}d\vOmega = \sum_{a=1}^{n}w_a = 4\pi\:,
\end{equation}
where $w_a$ are the quadrature weights, and $n$ is the total number of angles.
Different quadrature sets have different numbers of unknowns.  The
Level-Symmetric quadrature set has $n = N(N+2)$ unknowns for an $S_N$
approximation.

Given that many angles result from even low-order $S_N$ approximations, we see
why the source has been expanded in Spherical Harmonics.  Consider, an $S_8$
calculation has 80 angles per unknown location per group.  For a $P_3$
expansion 16 moments are required to define the source, a factor of 5
reduction in memory storage.  An $S_{16}$ calculation could be used for more
accuracy with a $P_3$ calculation resulting in a factor of 18 savings in
memory.

Using the quadrature integration rule the flux moments in
Eqs.~(\ref{eq:even-flux}) and (\ref{eq:odd-flux}) are evaluated using
\begin{align}
  \even_{lm} &= \sum_{a=1}^{n}Y^e_{lm}(\vOmega_a)\psi^g_a w_a\:,
  \label{eq:even-flux-quad-int}\\
  \odd_{lm} &= \sum_{a=1}^{n}Y^o_{lm}(\vOmega_a)\psi^g_a w_a\:.
  \label{eq:odd-flux-quad-int}
\end{align}
Similarly, the source moments in Eqs.~(\ref{eq:even-source}) and
(\ref{eq:odd-source}) are calculated using
\begin{align}
  \qe_{lm} &= \sum_{a=1}^{n}Y^e_{lm}(\vOmega_a)q^g_e(\vOmega_a) w_a\:,\\
  \qo_{lm} &= \sum_{a=1}^{n}Y^o_{lm}(\vOmega_a)q^g_e(\vOmega_a) w_a\:.
\end{align}
The $S_N$ method will be conservative if the quadrature set effectively
integrates the even and odd Spherical Harmonics.  If the orthogonality
conditions in Eqs.~(\ref{eq:Ye-orthogonality}) and (\ref{eq:Yo-orthogonality})
are preserved, then integrating the anisotropic scattering should yield,
\begin{equation}
  \int_{4\pi}q_s^g(\vOmega)\,d\vOmega =
  \sum_{a=1}^{n}q_s^g(\vOmega_a)w_a = \sqrt{4\pi}\sigg{0}\evenp_{00}
    = \sigma^{gg'}_{s}\phi^{g'}\:,
\end{equation}
which will yield a conservative particle balance equation.

%%---------------------------------------------------------------------------%%
\section{Operator Form of the Discrete Ordinates Equation}
\label{sec:oper-form-discr-1}
%%---------------------------------------------------------------------------%%

The multigroup discrete ordinates, or \Sn\, equation is (see
Eq.~(\ref{eq:MG-SN-Equation}))
\begin{multline}
    \hOmega_a\cdot\grad\psi^g_a(\ve{r}) +
    \sigma^g(\ve{r})\psi^g_a(\ve{r})
    \\=\sum_{g'=0}^G
    \sum_{l=0}^N
    \sigg{l}(\ve{r})
    \Bigl[
    Y^e_{l0}(\vOmega_a)\evenp_{l0}(\ve{r})
    +
    \sum_{m=1}^l
    \bigl(
    Y^e_{lm}(\vOmega_a)\evenp_{lm}(\ve{r}) +
    Y^o_{lm}(\vOmega_a)\oddp_{lm}(\ve{r})
    \bigr)\Bigr]
    +
    q_e^g(\ve{r}, \vOmega_a)
    \:.
    \tag{\ref{eq:MG-SN-Equation}}
\end{multline}
We defer the spatial treatment of Eq.~(\ref{eq:MG-SN-Equation}) until
Chap.~\ref{cha:spat-discr-sn}.  This equation can be written using a concise
operator notation that helps illuminate numerical solution techniques.  The
operator form of Eq.~(\ref{eq:MG-SN-Equation}) is
\begin{equation}
  \ve{L}\Psi = \ve{M}\ve{S}\Phi + Q\:.
  \label{eq:operator-transport}
\end{equation}
Here we use the convention that bold letters represent discrete operators or
matrices and script symbols and letters represent vectors.

The sizes of the operators in Eq.~(\ref{eq:operator-transport}) are determined
from the following dimensions:
\begin{equation}
  \begin{aligned}
    N_g &= \text{number of groups}\:,\\
    t   &= \text{number of moments}\:,\\
    n   &= \text{number of angles}\:,\\
    N   &= \text{$P_N$ order}\:,\\
    N_c &= \text{number of cells}\:,\\
    N_e &= \text{number of unknowns per cell}\:.
  \end{aligned}
\end{equation}
Now, we define
\begin{align}
  a &= N_g\times n\times N_c\times N_e\:,\\
  f &= N_g\times t\times N_c\times N_e\:.
\end{align}
Equation~(\ref{eq:operator-transport}) can then be defined in terms of
the sizes of the operators,
\begin{equation}
  (a\times a)(a\times 1) = (a\times f)(f\times f)(f\times 1) +
  (a\times 1)\:.
  \label{eq:operator-sizes}
\end{equation}
More specifically, with the groups defined over the range $g\in[0,G]$,
at each spatial unknown we can write
\begin{equation}
    \ve{L}
    \begin{pmatrix}
      \Psi_0 \\
      \Psi_1 \\
      \Psi_2 \\
      \vdots   \\
      \Psi_G
    \end{pmatrix} =
    \begin{pmatrix}
      \ve{M} & 0 & 0 & 0 & 0 \\
      0 & \ve{M} & 0 & 0 & 0 \\
      0 & 0 & \ve{M} & 0 & 0 \\
      0 & 0 & 0 & \ddots & 0 \\
      0 & 0 & 0 & 0 & \ve{M} \\
    \end{pmatrix}
    %%
    \begin{pmatrix}
      \ve{S}_{00} & \ve{S}_{01} & \ve{S}_{02} & \cdots &
      \ve{S}_{0G} \\
      \ve{S}_{10} & \ve{S}_{11} & \ve{S}_{12} & \cdots &
      \ve{S}_{1G} \\
      \ve{S}_{20} & \ve{S}_{21} & \ve{S}_{22} & \cdots &
      \ve{S}_{2G} \\
      \vdots & \vdots & \vdots & \ddots & \vdots \\
      \ve{S}_{G0} & \ve{S}_{G1} & \ve{S}_{G2} & \cdots &
      \ve{S}_{GG}
    \end{pmatrix}
    \begin{pmatrix}
      \Phi_0 \\
      \Phi_1 \\
      \Phi_2 \\
      \vdots   \\
      \Phi_G
    \end{pmatrix}
    %%
    +
    \begin{pmatrix}
      Q_0 \\
      Q_1 \\
      Q_2 \\
      \vdots   \\
      Q_G
    \end{pmatrix}\:,
    \label{eq:matrix-transport}
\end{equation}
Here, $\Psi_g$ and $Q_g$ are vectors of size $n$,
\begin{align}
  \Psi_g &= \begin{pmatrix}
    \psi^g_1 & \psi^g_2 & \psi^g_3 & \cdots \psi^g_n
  \end{pmatrix}^T\:,\\
  Q_g &= \begin{pmatrix}
    Q^g_1 & Q^g_2 & Q^g_3 & \cdots Q^g_n
  \end{pmatrix}^T\:.
\end{align}
The spatial unknowns are implicit in the above matrices and will be
investigated in Chaps.~\ref{cha:spat-discr-sn} and
\ref{cha:transp-solut-meth}.

The operator $\ve{M}$ is the moment-to-discrete matrix. It is used to project
harmonic moments onto discrete angle space, and it is defined
\begin{equation}
  {\small
    \ve{M} = \begin{pmatrix}
      \Ye{00}{1} & \Ye{10}{1} & \Yo{11}{1} & \Ye{11}{1} &
      \Ye{20}{1} & \cdots & \Yo{NN}{1} & \Ye{NN}{1} \\
      \Ye{00}{2} & \Ye{10}{2} & \Yo{11}{2} & \Ye{11}{2} &
      \Ye{20}{2} & \cdots & \Yo{NN}{2} & \Ye{NN}{2} \\
      \Ye{00}{3} & \Ye{10}{3} & \Yo{11}{3} & \Ye{11}{3} &
      \Ye{20}{3} & \cdots & \Yo{NN}{3} & \Ye{NN}{3} \\
      \vdots     & \vdots     & \vdots     & \vdots     &
      \vdots     &        & \vdots     & \vdots     \\
      \Ye{00}{n} & \Ye{10}{n} & \Yo{11}{n} & \Ye{11}{n} &
      \Ye{20}{n} & \cdots & \Yo{NN}{n} & \Ye{NN}{n}
  \end{pmatrix}}\:.
\end{equation}
The moments of the angular flux are calculated from discrete angular fluxes
using the moment-to-discrete matrix,
\begin{equation}
  \phi = \ve{D}\psi\:.
\end{equation}
Clearly, the discrete form of $\ve{D}$ is defined by the quadrature
integration rules in Eqs.~(\ref{eq:even-flux-quad-int}) and
(\ref{eq:odd-flux-quad-int}),
\begin{equation}
    \ve{D} = \ve{M}^{T}\ve{W}\:.
\end{equation}
where $\ve{W}$ is an $(a\times a)$ diagonal matrix of the quadrature weights.
Using the definition of $\ve{M}$ from Eq.~(\ref{eq:operator-sizes}), the size
of $\ve{D}$ can be determined from
\begin{equation}
  \ve{D} \equiv (f\times a)(a\times a) = (f\times a)\:.
\end{equation}
The $\ve{D}$ matrix has the following form
\begin{equation}
  \ve{D} = \begin{pmatrix}
    w_1\Ye{00}{1} & w_2\Ye{00}{2} & w_3\Ye{00}{3} & \cdots & w_n\Ye{00}{n} \\
    w_1\Ye{10}{1} & w_2\Ye{10}{2} & w_3\Ye{10}{3} & \cdots & w_n\Ye{10}{n} \\
    w_1\Yo{11}{1} & w_2\Yo{11}{2} & w_3\Yo{11}{3} & \cdots & w_n\Yo{11}{n} \\
    w_1\Ye{11}{1} & w_2\Ye{11}{2} & w_3\Ye{11}{3} & \cdots & w_n\Ye{11}{n} \\
    w_1\Ye{20}{1} & w_2\Ye{20}{2} & w_3\Ye{20}{3} & \cdots & w_n\Ye{20}{n} \\
    \vdots & \vdots & \vdots & & \vdots \\
    w_1\Yo{NN}{1} & w_2\Yo{NN}{2} & w_3\Yo{NN}{3} & \cdots & w_n\Yo{NN}{n} \\
    w_1\Ye{NN}{1} & w_2\Ye{NN}{2} & w_3\Ye{NN}{3} & \cdots & w_n\Ye{NN}{n}
  \end{pmatrix}\:.
\end{equation}
Also, even though $\ve{M}$ projects angular flux moments onto discrete angular
flux space, in general $\psi\ne\ve{M}\phi$ unless $\ve{M} = \ve{D}^{-1}$.
This condition is met when using the \textit{Galerkin} quadrature set, but
most quadrature sets do not satisfy this requirement. The moments of the
angular flux vector are defined
\begin{equation}
  \Phi_g = \begin{pmatrix}
    \even_{00} & \even_{10} & \odd_{11} & \even_{11} & \even_{20}
    & \cdots & \odd_{NN} & \even_{NN}
  \end{pmatrix}^T\:,
\end{equation}

The scattering cross sections are defined
\begin{equation}
  \ve{S}_{gg'} = \begin{pmatrix}
    \sigg{0} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    0 & \sigg{1} & 0 & 0 & 0 & 0 & 0 & 0 \\
    0 & 0 & \sigg{1} & 0 & 0 & 0 & 0 & 0 \\
    0 & 0 & 0 & \sigg{1} & 0 & 0 & 0 & 0 \\
    0 & 0 & 0 & 0 & \sigg{2} & 0 & 0 & 0 \\
    0 & 0 & 0 & 0 & 0 & \ddots   & 0 & 0 \\
    0 & 0 & 0 & 0 & 0 & 0 & \sigg{N} & 0 \\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & \sigg{N}
  \end{pmatrix}\:.
\end{equation}
Here the block-matrix $\ve{S}_{gg'}$ defines scattering cross sections for
particles that scatter from group $g'$ into group $g$.  The lower triangular
part of $\ve{S}$ represents down-scattering, the diagonal represents in-group
scattering, and the upper diagonal is up-scattering.

%%---------------------------------------------------------------------------%%
%% end of collocation_methods.tex
%%---------------------------------------------------------------------------%%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "exnihilo"
%%% End:

\end{document}

### 2017-03-01 docprep/appdx_a.tex
Updated the literature review section

--- 2017-02-28 docprep/chapmeth.tex
@@
 
-\section{Mathmatic Development}
+ %% ... sample chapter ...
 
-Where does this text go?
+This appendix covers the math for Legendre polynomials and spherical harmonics
 
-\subsection{The Boltzmann Equation}
+\section{Legendre Polynomials}
 
-The time dependent gamma transport equation can be written as
+They exist.
 
-\begin{equation} \label{eq:boltz_t_dep}
-\begin{split}
-	&\left[ \frac{1}{v(E)} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
-	\psi(\boldsymbol{r}, E, \hat{\Omega}, t) = \\
-	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}, t) \psi(\boldsymbol{r}, E', \hat{\Omega}', t) dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega}, t)
-\end{split}
-\end{equation}
+\subsection{Spherical Harmonics}
 
-However, we are typically not concerned with the transient case in medical diagnostic imaging.
-We are more interested in the steady state case.
-The time independent form of Eq.
-\ref{eq:boltz_t_dep} is written as
+They also exist.
 
-\begin{equation} \label{eq:boltz}
-\begin{split}
-	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-	\psi(\boldsymbol{r}, E, \hat{\Omega}) = \\
-	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega})
-\end{split}
-\end{equation}
+The real harmonics. The even/odd harmonics.
 
-The following sections show how Eq.~\ref{eq:boltz} is discretized.
-
-\subsection{Angular Discretization}
-The coordinate system used in DOCTORS is shown in Fig.~\ref{fig:coord_sys}.
-A discrete set of $N_a$ angles ($\Omega_{a} \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space.
-Particles are transported only along these discrete directions form one voxel to adjacent voxels.
-
-The $S_N$ quadrature is implemented in DOCTORS.
-There are many different quadrature sets, the $S_N$ quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
-This enables simple, direct comparison to other solvers.
-Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/coord_sys}
-  \end{center}
-  \caption{The coordinate system used in DOCTORS.
-Given an arbitrary direction, $\Omega$, $\mu$, $\eta$, and $\xi$ are its direction cosines with respect to the $x$, $y$, and $z$ axes respectively.
-$\varphi$ is the azimuthal angle (with respect to $x$ and the d $\theta$ is the polar angle (with respect to $z$).}
-\label{fig:coord_sys}
-\end{figure}%
-
-A given direction, $\hat{\Omega}$ is determined by its three cosine components.
-This relation is given in Eq.~\ref{eq:omega_cos} where $\hat{i}$, $\hat{j}$, and $\hat{k}$ are the unit directions in the $x$, $y$, and $z$ directions respectively.
-
-\begin{equation} \label{eq:omega_cos}
-	\hat{\Omega} = \mu \hat{i} + \eta \hat{j} + \xi \hat{k}
-\end{equation}
-
-Each discrete angle is associated a weight.
-The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by Eq.~\ref{eq:disc_int}.
-
-\begin{equation} \label{eq:disc_int}
-\int_{4 \pi} f(\hat{\Omega}) d\Omega \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a
-\end{equation}
-
-Discretizing Eq.~\ref{eq:boltz} in angular space gives Eq.~\ref{eq:boltz_a} where the angle is denoted by subscript $a$.
-
-\begin{equation} \label{eq:boltz_a}
-\begin{split}
-&\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-\psi_{a}(\boldsymbol{r}, E) = \\
-&\sum_{a=0}^{N_a-1} \int_0^\infty \Sigma_{s, a, a'}(\boldsymbol{r}, E' \rightarrow E) \psi_{a'}(\boldsymbol{r}, E') dE' \omega_a + S_a(\boldsymbol{r}, E)
-\end{split}
-\end{equation}
-
-\subsection{Energy Discretization}
-
-Continuous energy is discretized into $G$ groups indexed from 0 to $G-1$.
-Upscatter is assumed to be negligible.
-Therefore, Eq.~\ref{eq:boltz_a} becomes the energy discretized Eq.~\ref{eq:boltz_e} where the group number is indexed by superscript $g$.
-The group structure is shown in Fig.~\ref{fig:energy_groups}.
-
-\begin{equation} \label{eq:boltz_e}
-\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t^g(\boldsymbol{r}) \right]
-\psi_{a}^{g}(\boldsymbol{r}) = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
-\end{equation}
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/energy_groups}
-  \end{center}
-  \caption{The energy grid structure used in DOCTORS. The highest energy group is group 0 and the lowest energy is group $G-1$.}
-\label{fig:energy_groups}
-\end{figure}%
-
-\subsection{Spatial Discretization}
-
-Finally, Eq.~\ref{eq:boltz_e} is discretized in space to give Eq.~\ref{eq:boltz_r} which is fully discretized.
-The problem domain is split into an evenly spaced Cartesian grid.
-The grid spacing between each of the spatial dimensions does not necessarily need to be uniform, but it often is in the case of CT voxel phantoms.
-
-The problem domain must be a regular rectangular parallelpiped of size $D_x$, $D_y$, and $D_z$ in the $x$, $y$, and $z$ directions respectively.
-The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each direction.
-The total number of voxels, $N_V$ is easily computed as the multiple of each of the component directions given in Eq.~\ref{eq:n_v}.
-
-\begin{equation} \label{eq:n_v}
-	N_V = N_x N_y N_z
-\end{equation}
-
-Voxel indices are flattened according to the rule given in Eq.~\ref{eq:indx_flat}. Note that all indeces always start at zero.
-
-\begin{equation} \label{eq:indx_flat}
-	i = i_x (N_z + N_y) + i_y N_z + i_z
-\end{equation}
-
-The spatial mesh is shown in Fig.~\ref{fig:spatial_disc}.
-The size of each voxel in each dimension is computed by dividing the size of the problem domain along in that dimension by the number of mesh elements along the same direction.
-Eq.~\ref{eq:mesh_x} shows this computation for the $x$ direction.
-Analagous equations apply in the $y$ and $z$ directions as well.
-
-\begin{equation} \label{eq:mesh_x}
-\Delta x = \frac{D_x}{N_x}
-\end{equation}
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/spatial_disc}
-  \end{center}
-  \caption{The spatial mesh imposed on the problem domain.}
-\label{fig:spatial_disc}
-\end{figure}%
-
-The index of interest is denoted as a subscript $i$ which makes the fully discretized form of the steady state Linear Boltzmann Equation given in Eq.~\ref{eq:boltz} into the fully discretized form given in Eq.~\ref{eq:boltz_i}
-
-\begin{equation} \label{eq:boltz_i}
-\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_{t,i}^g \right]
-\psi_{i,a}^{g} = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
-\end{equation}
-
-\subsection{Solution}
-
-To solve the fully discretized form of the Linear Boltzmann Equation given in Eq.~\ref{eq:boltz_i}, the gradient operator must first be computed numerically.
-Recall that $\Omega_a$ can be written in vector notation using its cosine components as defined in Eq.~\ref{eq:omega_cos}.
-Also recall the definition of the gradient, given in Eq.~\ref{eq:grad}.
-
-\begin{equation} \label{eq:grad}
-\nabla f(x, y, z) = \frac{\partial f(x, y, z)}{\partial x} \hat{i} + \frac{\partial f(x, y, z)}{\partial y} \hat{j} + \frac{\partial f(x, y, z)}{\partial z} \hat{k}
-\end{equation}
-
-To the first order, a partial derivitive can be computed using Eq.~\ref{eq:deriv_1}.
-
-\begin{equation} \label{eq:deriv_1}
-\frac{\partial f(x)}{\partial x} \approx \frac{f(x) - f(x + \Delta x)}{\Delta x}
-\end{equation}
-
-Using the definition of $\Omega$ and Eq.~\ref{eq:grad} the $\hat{\Omega} \cdot \nabla \psi$ term can be rewritten using Eq.~\ref{eq:spatial_1}
-
-\begin{equation} \label{eq:deriv_1}
-\begin{split}
-\Omega \cdot \nabla \psi & \approx 
-\left\langle \mu, \eta, \xi \right\rangle \cdot
-\left\langle \frac{\psi(x) - \psi(x + \Delta x)}{\Delta x},
-\frac{\psi(y + \Delta y) - \psi(y)}{\Delta y},
-\frac{\psi(z + \Delta z) - \psi(z)}{\Delta z} \right\rangle \\
-& \approx 
-\frac{\psi(x + \Delta x) - \psi(x)}{\Delta x} \mu + 
-\frac{\psi(y + \Delta y) - \psi(y)}{\Delta y} \eta + 
-\frac{\psi(z + \Delta z) - \psi(z)}{\Delta z} \xi
-\end{split}
-\end{equation}
-
-Along some direction, $\Omega$, as it enters an arbitrary voxel, it must pass through either of the sides indicated in red and exit via the other as shown in Fig.~\ref{fig:gradient}.
-Fig.~\ref{fig:gradient} uses the $yz$ faces of the voxel for illustration.
-If $\eta$ is positive, $\Omega$ will enter through the $x_i$ surface and exit through the $x_{i+1}$ surface.
-If $\eta$ is negative, this order will be reversed.
-A limitation to this methodology is that $\mu$, $\eta$, and $\xi$ are assumed to be nonzero.
-This is ensured as long as none of the directions in the quadrature are parallel to a major axis.
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/gradient}
-  \end{center}
-  \caption{The caption.}
-\label{fig:gradient}
-\end{figure}%
-
-Depending on the sign of $\eta$, the first order approximation of the a derivitive along the $y$ direction is given by Eq.~\ref{eq:grad_y}.
-
-\begin{equation} \label{eq:grad_y}
-\begin{split}
-\frac{\partial f}{\partial y} &= \frac{f(x_i) - f(x_{i+1})}{\Delta x}, \quad \eta>0 \\
-\frac{\partial f}{\partial y} &= \frac{f(x_{i+1}) - f(x_{i})}{\Delta x}, \quad \eta <0 \\
-\frac{\partial f}{\partial y} &= 0, \quad \eta=0
-\end{split}
-\end{equation}
-
-Equation~\ref{eq:grad_y} can be generalized to use $f_{in}$ and $f_{out}$ instead of relying on the $x$-coordinate. This is given in Eq.~\ref{eq:grad_y2}
-
-\begin{equation} \label{eq:grad_y2}
-\begin{split}
-\frac{\partial f}{\partial y} &= \frac{f_{in} - f_{out}}{\Delta x}, \quad \eta \neq 0 \\
-\frac{\partial f}{\partial y} &= 0, \quad \eta=0
-\end{split}
-\end{equation}
-
-The generalization of Eq.~\ref{eq:grad_y2} is applied to Eq.~\ref{eq:deriv_1} to yield Eq.~\ref{eq:deriv_2} which is generally valid for all octants.
-
-\begin{equation} \label{eq:deriv_2}
-\Omega \cdot \nabla \psi \approx 
-\frac{\psi_{x,in} - \psi_{x,out}}{\Delta x} \mu + 
-\frac{\psi_{y,in} - \psi_{y,out}}{\Delta y} \eta + 
-\frac{\psi_{z,in} - \psi_{z,out}}{\Delta z} \xi
-\end{equation}
-
-Using the approximation given in Eq.~\ref{eq:deriv_2}, Eq.~\ref{eq:boltz_i} can be rewritten as Eq.~\ref{eq:boltz_i2}.
-
-\begin{equation} \label{eq:boltz_i2}
-\left[ 
-\frac{\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
-\frac{\psi_{i,a,y,in}^g - \psi_{i,a,y,out}^g}{\Delta y} \eta + 
-\frac{\psi_{i,a,z,in}^g - \psi_{i,a,z,out}^g}{\Delta z} \xi
-\right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
-\end{equation}
-
-It is useful to point out that each of $\psi_{i,a,x,in}^g$, $\psi_{i,a,x,out}^g$, $\psi_{i,a,y,in}^g$, $\psi_{i,a,y,out}^g$, $\psi_{i,a,z,in}^g$, and $\psi_{i,a,z,out}^g$ are \textit{surface} flux values while $\psi_{i,a}^{g}$ is a \textit{volume averaged} flux value.
-
-In order for the discretized Linear Boltzmann Equation to be solved using Eq.~\ref{eq:boltz_i2}, the incoming flux must always be known.
-Either the boundary conditions or previously calculated values are used to determine each incoming flux.
-Therefore, in order to compute the average flux in a cell, some relationship between the average flux and the outgoing surface fluxes must be known.
-These values can be related by any one of many different mechanisms, the most common of which is the diamond difference approximation.
-
-\subsection{Diamond Difference Approximation}
-
-In the diamond difference approximation, it is assumed that the incoming and outgoing flux are both known.
-The average flux is assumed to be the average of the two surface fluxes.
-This is assumed to be the case in all three spatial dimensions.
-This leads to Eq.~\ref{eq:dd}.
-
-\begin{equation} \label{eq:dd}
-\begin{split}
-\frac{\psi_{i,a,x,in}^g + \psi_{i,a,x,out}^g}{2} &= \psi_{i,a}^{g} \\
-\frac{\psi_{i,a,y,in}^g + \psi_{i,a,y,out}^g}{2} &= \psi_{i,a}^{g} \\
-\frac{\psi_{i,a,z,in}^g + \psi_{i,a,z,out}^g}{2} &= \psi_{i,a}^{g}
-\end{split}
-\end{equation}
-
-Multiplying both sides of Eq.~\ref{eq:dd} and subtracting twice the outgoing surface flux from each yields Eq.~\ref{eq:dd2}.
-
-\begin{equation} \label{eq:dd2}
-\begin{split}
-\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g \\
-\psi_{i,a,y,in}^g - \psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,out}^g \\
-\psi_{i,a,z,in}^g - \psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,out}^g
-\end{split}
-\end{equation}
-
-The form of the left hand side of Eq.~\ref{eq:dd2} is the same as in Eq.~\ref{eq:boltz_i2}.
-This allows Eq.~\ref{eq:dd2} to be used to make a substitution in Eq.~\ref{eq:boltz_i2} that removes the dependence on the outgoing flux from discretized equation.
-After the substitution, Eq.~\ref{eq:boltz_i3} will arise.
-
-\begin{equation} \label{eq:boltz_i3}
-\left[ 
-\frac{2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
-\frac{2\psi_{i,a}^{g} - \psi_{i,a,y,out}^g}{\Delta y} \eta + 
-\frac{2\psi_{i,a}^{g} - \psi_{i,a,z,out}^g}{\Delta z} \xi
-\right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
-\end{equation}
-
-The diamond difference approximation defined in Eq.~\ref{eq:dd} can also be rearranged to solve for each outgoing flux once the volume averaged flux is computed.
-The outgoing flux is given by Eq.~\ref{eq:dd3}.
-
-\begin{equation} \label{eq:dd3}
-\begin{split}
-\psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,in}^g \\
-\psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,in}^g \\
-\psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,in}^g
-\end{split}
-\end{equation}
-
-Rearranging Eq.~\ref{eq:boltz_i3} to factor out the volume averaged flux term yields Eq.~\ref{eq:boltz_i4}.
-
-\begin{equation} \label{eq:boltz_i4}
-\left[ 
-\frac{2\psi_{i,a}^{g}}{\Delta x} \mu + 
-\frac{2\psi_{i,a}^{g}}{\Delta y} \eta + 
-\frac{2\psi_{i,a}^{g}}{\Delta z} \xi
-\right] - 
-\left[ 
-\frac{2\psi_{i,a,x,out}^g}{\Delta x} \mu + 
-\frac{2\psi_{i,a,y,out}^g}{\Delta y} \eta + 
-\frac{2\psi_{i,a,z,out}^g}{\Delta z} \xi
-\right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
-\end{equation}
-
-Equation~\ref{eq:boltz_i4} can be directly solved for the volume averaged flux. The result is given in Eq.~\ref{eq:boltz_i5}.
-
-\begin{equation} \label{eq:boltz_i5}
-\psi_{i,a}^{g} = 
-\frac{
-  \sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i,     a'}^{g'} \omega_a + 
-  \left[ 
-    \frac{2\psi_{i,a,x,out}^g}{\Delta x} \mu + 
-    \frac{2\psi_{i,a,y,out}^g}{\Delta y} \eta + 
-    \frac{2\psi_{i,a,z,out}^g}{\Delta z} \xi
-  \right] + S_{i,a}^g
-}{
-  \frac{2\mu}{\Delta x}  + 
-  \frac{2\eta}{\Delta y} + 
-  \frac{2\xi}{\Delta z} + 
-  \Sigma_{t,i}^g
-}
-\end{equation}
-
-\subsection{Anisotropy Treatment}
-
-Scatter is assumed to be a function of only the scatter angle between the initial and scattered directions, $\theta_s$. Often, the cosine of the scatter angle, $\mu_s$ is used.
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/scat_ang}
-  \end{center}
-  \caption{The scatter angle.
-A photon at energy $E$ traveling in direction $\Omega$, hits a stationary atom, indicated by the green circle, and scatters into a new direction $\Omega'$ with a new energy $E'$.}
-\label{fig:scat_ang}
-\end{figure}%
-
-The data files containing the scatter cross sections are distributed using a Legendre polynomial expansion.
-The use of a Legendre expansion removes the dependence on the quadrature from the data.
-This allows any quadrature to work with any dataset.
-A $l$-order Legendre polynomial is donoted by $P_l$.
-The anisotropic scatter cross section can then be rewritten as a Legendre expansion given in Eq.~\ref{eq:leg_1} where the scatter cosine is defined in Eq.~\ref{eq:scat_cos}
-
-\begin{equation} \label{eq:leg_1}
-\Sigma_{s, i, a, a'}^{g, g'} = \sum_l=0^L \frac{2l+1}{4 \pi}\Sigma_{s, i, l}^{g, g'} P_l(\mu_s)
-\end{equation}
-
-\begin{equation} \label{eq:scat_cos}
-\mu_s = \Omega_a \cdot \Omega_{a'}
-\end{equation}
+The addition theorem.
 

### 2017-03-01 docprep/chapbib.tex
Updated the literature review section

%%
%% This is file `chapbib.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin,addbib')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapbib.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

 %% ... sample chapter ...

This section is the literature review. I do these things.

\section{Introduction}
Many different things are required.

\section{CT Dose Estimation}
How is this done now?

\section{CT Phantom Generation}
The CT phantom is populated with CT numbers, also known as Hounsfield numbers. These values are related to the attenuation coefficient of the material represented by the voxel.

Many authors have provided correlations that map the CT number to both a material and a density.

First ones.
Saw \citep{ref:sawc} did stuff. And \citep{ref:plessisf}. Then \citet{ref:schneideru} said stuff. And \citet{ref:kimh} did stuff with GEANT4.

Another Schneider did some stuff too \citep{ref:schneiderw}. And Vanderstraeten has a long name \citep{ref:vanderstraetenb}.

The new guy provides 19 materials. The density is measured.

Two phantom datasets are used in this work.

\section{Discrete Ordinates}
Discrete ordinates are old. The book is \citet{ref:lewise}.

The Exnihilo thing by \citet{ref:evanst}. And DeHart wrote in \citep{ref:dehartm}

The Ibrahim did stuff \citep{ref:ibrahima} and Lee did stuff too \citep{ref:leeb}.

With regards to angles, stuff has been done \citep{ref:ahrensc}

Attila had a 1st col src added by \citep{ref:wareingt}. Some rays are better than others \citep{ref:mathewsk}.

Dort and Tort \citep{ref:rhoadesw}.

Denovo.

\section{Raytracing}

Raytracing is important too. Woo came up with on \citep{ref:wooa} which was improved by \citet{ref:liuy} and \citet{ref:hel}.

\section{GPU Acceleration}
GPUs are cool.

CUDA is cool.

\endinput
%%
%% End of file `chapbib.tex'.

### 2017-03-01 docprep/norris_phd_dissertation.tex
Updated the literature review section

 % -*- mode: LaTeX -*-
 % ... preamble ...
\documentclass[times,12pt,titlepage]{mstthesis}
\doublespacing
\usepackage{threeparttable}
\usepackage[final]{graphicx}
\usepackage{psfrag}
\usepackage{listings}
\usepackage{minted}
\usepackage[noprefix]{nomencl}
\makenomenclature
\renewcommand{\nomgroup}[1]{
   \ifthenelse{\equal{#1}{A}}{\medskip \item \textbf{Roman}}{}%
   \ifthenelse{\equal{#1}{G}}{\medskip \item \textbf{Greek}}{}%
   \ifthenelse{\equal{#1}{S}}{\medskip \item \textbf{Subscripts}}{}%
}
\usepackage[nopar]{lipsum} % ... provides dummy text ...

 % ... end preamble ...

\begin{document}

 % ... specify: ms or phd ...

\begin{ThesisTitlePage}{phd}

 % ... title page info ...

\author{\MakeUppercase{Edward T. Norris}}

\thesistitle{\MakeUppercase{Discrete Ordinates CT Organ Dose Simulator (DOCTORS)}}

\department{Nuclear Engineering}

 % ... thesis committee ...

\ThesisAdviser{Xin Liu}

 % If you have a co-advisor, enter the name in the
 % curly braces below and uncomment.
 % Otherwise, leave commented out.
 %\cothadviser{} % If you have 2 thesis advisers

\memberone{Ayodeji Alajo}
\membertwo{Hyoung Koo Lee}
\memberthree{Gary Mueller}
\memberfour{Fikret Ercal}
%\memberfive{Prandtl}

 % ... Graduation date - NOT your submission date ...
\graddate{2017}

\end{ThesisTitlePage}
 % ... copyright page - true|false ...

\copyrightyear{2017}
\ThesisCopyrightPage{true}

 % ... front matter - thesis abstract ...

\begin{ThesisAbstract}
This is the text for the abstract. Here, I explain what I did and how cool I am for doing it. Because I am awesome.
\end{ThesisAbstract}

 % ... front matter - thesis acknowledgements ...

\begin{ThesisAcknowledgment}
I would like to thank these people. They are all so super cool. So much coolness. So much so that this is now over a line long.

And these funding agencies.
\end{ThesisAcknowledgment}
\begin{ThesisFrontMatter}
\tableofcontents
\listoffigures
\listoftables
\listofsymbols
\end{ThesisFrontMatter}

\begin{ThesisBody}

 % ... introduction chapter ...
\ThesisBodyChapter{Introduction}\input{chapintro}
 % ... chapter ...
\ThesisBodyChapter{Literature Review}\input{chapbib}
 % ... chapter ...
\ThesisBodyChapter{Methods}\input{chapmeth}

\ThesisBodyChapter{Implementation}\input{chapimpl}
 % ... conclusions chapter ...
\ThesisBodyChapter{Conclusions}\input{chapconc}

\end{ThesisBody}
 % ... appendix - specify number of appendix chapters ...

\begin{ThesisAppendix}{two}
\ThesisAppendixChapter{Legendre Polynomials and Spherical Harmonics}\input{appdx_a}
\ThesisAppendixChapter{Two}\input{appdx_b}
\end{ThesisAppendix}

 % ... references - comma separated list of bib files ...

\begin{ThesisBibliography}{BIBLIOGRAPHY}
\bibliographystyle{plainnat}
\bibliography{refs}
\end{ThesisBibliography}

 % ... vita ...

\begin{Vita}
Edward Norris graduated from Missouri University of Science and Technology in 2013 with a B.S. in both nuclear engineering and computer science.

He has worked...

He is a member of...
\end{Vita}

\end{document}

### 2017-03-07 doctors/mainwindow.cpp
Added an empy file for the CUDA code

@@ -89,9 +89,7 @@ MainWindow::MainWindow(QWidget *parent):
     m_configSelectDialog->setFileMode(QFileDialog::ExistingFile);
 
     m_parser = new AmpxParser;
-
     m_solver = new Solver;
-
     m_xs = new XSection();
 
     // Connect explore buttons

### 2017-04-06 docprep/chapmeth.tex
Fixed an equation and added a section about vector flattening
Updated the literature review section
Added a section on the uncollided raytracer

@@
 
-Finally, Eq.~\ref{eq:boltz_e} is discretized in space to give Eq.~\ref{eq:boltz_r} which is fully discretized.
+Finally, Eq.~\ref{eq:boltz_e} is discretized in space to give Eq.~\ref{eq:boltz_i} which is fully discretized.
 The problem domain is split into an evenly spaced Cartesian grid.
@@
 
-\begin{equation} \label{eq:deriv_1}
+\begin{equation} \label{eq:spatial_1}
 \begin{split}
@@
 \begin{equation} \label{eq:leg_1}
-\Sigma_{s, i, a, a'}^{g, g'} = \sum_l=0^L \frac{2l+1}{4 \pi}\Sigma_{s, i, l}^{g, g'} P_l(\mu_s)
+\Sigma_{s, i, a, a'}^{g, g'} = \sum_{l=0}^L \frac{2l+1}{4 \pi}\Sigma_{s, i, l}^{g, g'} P_l(\mu_s)
 \end{equation}
@@
 
+\section{Uncollided Flux}
+To avoid ray effect, the flux is split into collided and uncollided components as shown in Eq.~\ref{eq:totflux}.
+Equation~\ref{eq:boltz} can be rewritten as two equations solving for the uncollided flux, $\psi_u$, and collided flux, $\psi_c$ independently.
+Equation~\ref{eq:unc} has no scatter term since scattered particles are not considered in the uncollided flux.
+The source term, $S_u$, in Eq.~\ref{eq:col} is the first collision source.
+
+\begin{equation} \label{eq:totflux}
+\psi = \psi_u + \psi_c
+\end{equation}
+
+\begin{equation} \label{eq:unc}
+\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+\psi_u(\boldsymbol{r}, E, \hat{\Omega}) =  S(\boldsymbol{r}, E, \hat{\Omega})
+\end{equation}
+
+\begin{equation} \label{eq:col}
+\begin{split}
+	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
+	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + S_{u}(\boldsymbol{r}, E, \hat{\Omega})
+\end{split}
+\end{equation}
+
+The uncollided flux in Eq.~\ref{eq:unc} can be analytically solved.
+The solution is given in Eq.~\ref{eq:uncsol} for a known isotropic point source located at position $\boldsymbol{s}$.
+The $\delta$ symbol in Eq.~\ref{eq:uncsol} is the Kronecker delta given by Eq.~\ref{eq:kronecker}.
+
+\begin{equation} \label{eq:uncsol}
+\psi_u(\boldsymbol{r}, E, \hat{\Omega}) = 
+S(\boldsymbol{s}, E)
+\frac{e^{-\sum_i \Sigma_{t,i} x_i}}{4\pi |\boldsymbol{r}-\boldsymbol{s}|^2}
+\delta\left( \hat{\Omega}, \frac{\boldsymbol{r}-\boldsymbol{s}}{|\boldsymbol{r}-\boldsymbol{s}|^2}\right)
+\end{equation}
+
+\begin{equation} \label{eq:kronecker}
+\delta(x, y) = 
+\begin{cases}
+1, \quad \mathrm{if} \ x=y \\
+0, \quad \mathrm{otherwise}
+\end{cases}
+\end{equation}
+
+Equation~\ref{eq:uncsol} can be expanded to accomodate any number of spatially distributed sources of arbitrary directionality given in Eq.~\ref{eq:uncsol2} by integrating over the entire spatial domain, $V$.
+
+\begin{equation} \label{eq:uncsol2}
+\psi_u(\boldsymbol{r}, E, \hat{\Omega}) = \iiint_{V}
+S(\boldsymbol{s}, E, \hat{\Omega})
+\frac{e^{-\sum_i \Sigma_{t,i} x_i}}{|\boldsymbol{r}-\boldsymbol{s}|^2}
+\delta\left( \hat{\Omega}, \frac{\boldsymbol{r}-\boldsymbol{s}}{|\boldsymbol{r}-\boldsymbol{s}|^2}\right)
+d \boldsymbol{s}
+\end{equation}
+
+The uncollided source term, $S_u$ in Eq.~\ref{eq:col} is computed from the uncollided flux using Eq.~\ref{eq:uncsrc}.
+
+\begin{equation} \label{eq:uncsrc}
+S_u(\boldsymbol{r}, E, \hat{\Omega}) = \int_{4\pi} \int_{0}^{\infty} 
+\Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_u(\boldsymbol{r}, E', \hat{\Omega}') 
+dE' d\hat{\Omega}'
+\end{equation}
+
+Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~{eq:col} gives Eq.~\ref{eq:col2} which simplifies to Eq.~\ref{eq:col3} by using Eq.~\ref{eq:totflux} to replace the collided and uncollided components with the total angular flux.
+
+\begin{equation} \label{eq:col2}
+\begin{split}
+	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
+	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + \\
+	&\int_{4\pi} \int_{0}^{\infty} 
+\Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_u(\boldsymbol{r}, E', \hat{\Omega}') 
+dE' d\hat{\Omega}'
+\end{split}
+\end{equation}
+
+\begin{equation} \label{eq:col3}
+\begin{split}
+	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
+	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'
+\end{split}
+\end{equation}
+
 \endinput

### 2017-04-24 docprep/chapmeth.tex
Updated the thesis chapters and the beamshapes figures

@@
 
-The $S_N$ quadrature is implemented in DOCTORS.
-There are many different quadrature sets, the $S_N$ quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
+The rotation symmetrical quadrature is implemented in DOCTORS.
+There are many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
 This enables simple, direct comparison to other solvers.
@@
 
+\subsection{Cross Section Parsing}
+AMPX cross sections are read in binary. They are formatted like so. They are read like so. They were obtained from SCALE6.2. These files can be generated by other codes.
+
 \subsection{Spatial Discretization}
@@
 
-Voxel indices are flattened according to the rule given in Eq.~\ref{eq:indx_flat}. Note that all indeces always start at zero.
+Voxel indices are reordered into a one dimensional vector (flattened) according to the rule given in Eq.~\ref{eq:indx_flat}. Note that indices start at zero.
 
@@
 \begin{equation} \label{eq:kronecker}
-\delta(x, y) = 
+\delta(\vec{A}, \vec{B}) = 
 \begin{cases}
-1, \quad \mathrm{if} \ x=y \\
+1, \quad \mathrm{if} \ \vec{A}=\vec{B} \\
 0, \quad \mathrm{otherwise}
@@
 
-Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~{eq:col} gives Eq.~\ref{eq:col2} which simplifies to Eq.~\ref{eq:col3} by using Eq.~\ref{eq:totflux} to replace the collided and uncollided components with the total angular flux.
+Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~\ref{eq:col} gives Eq.~\ref{eq:col2} which simplifies to Eq.~\ref{eq:col3} by using Eq.~\ref{eq:totflux} to replace the collided and uncollided components with the total angular flux.
 
@@
 
+\section{Source Generation}
+There are many kinds of external sources implemented in this work.
+
+\subsection{Point Source}
+The simplest source implemented is an isotropic point source.
+This source is defined only by its position and energy distribution.
+The center of the cell whose flux is being computed is located at position $<C_x, C_y, C_z>$ and the point source is located at position $<S_x, S_y, S_z>$.
+
+\subsection{Fan Source}
+In addition to the values defining a point source (position $\vec{S}$ and the energy distribution) a fan source is defined by two additional parameters, $\varphi$ and $\theta$ which describe the azimuthal and polar angles subtented by the fan beam respectively.
+Figure~\ref{fig:fanbeam} shows $\varphi$ and $\theta$ in reference to the geometry setup.
+The raytracing algorithm used for the fan beam is identical to the point source raytracer except for one difference: before the raytracing is done, whether the ray falls within the fan beam is determined and any ray falling outside the fan beam is not traced, its uncollided flux is immediately set to zero.
+
+Some voxels would be partially enclosed inside the beam. A partial acceptance value is computed for these voxels and multiplied by the final result of the raytrace.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/fan_rejection}
+  \end{center}
+  \caption{The fan.}
+\label{fig:fan_rejection}
+\end{figure}
+
+First, the ray direction is projected onto the $xy$ plane giving $<S_x, S_y, 0>$ and $C_x, C_y, 0$.
+Then the vector from the source to the geometry center and the vector from the source to the cell center are dotted to give the cosine of the angle between them as shown in Eq.~\ref{eq:phicos}.
+Once $\zeta$ is computed, the condition given by Eq.~\ref{eq:phicoscon} is evaluated.
+If the condition is met, the angle is accepted and raytraced normally, otherwise, it is rejected immediately.
+Figure~\ref{fig:phi}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/phi}
+  \end{center}
+  \caption{How $\varphi$ is computed.}
+\label{fig:phi}
+\end{figure}
+
+\begin{equation}\label{eq:phicos}
+\cos(\zeta_\varphi) = \frac{<S_x, S_y, 0>}{\sqrt{S_x^2 + S_y^2}} \cdot \frac{<C_x, C_y, 0>}{\sqrt{C_x^2 + C_y^2}} = \frac{S_x C_x + S_y C_y}{\sqrt{S_x^2 + S_y^2} \sqrt{C_x^2 + C_y^2}}
+\end{equation}
+
+\begin{equation}\label{eq:phicoscon}
+|\zeta_\varphi| < \frac{\varphi}{2}
+\end{equation}
+
+Treatment for the $\theta$ variable is similar to that of $\varphi$ except the polar angle is required instead of the azimuthal angle.
+This is computed using Eq.~\ref{eq:thetacos} for an arbitrary vector $\vec{\xi}$.
+This is used to compute the polar angle for both $\vec{S}$ and $\vec{C}$.
+The difference between the polar angles of the two vectors is computed using Eq.~\ref{eq:thetacos2} and then the condition given in Eq.~\ref{eq:thetacoscon} is evaluated.
+If the condition is not met, the ray is rejected and no raytrace is performed.
+
+\begin{equation}\label{eq:thetacos}
+\cos(\theta_\xi) = \frac{\xi_z}{\sqrt{\xi_z^2 + \xi_y^2 + \xi_z^2}}
+\end{equation}
+
+\begin{equation}\label{eq:thetacos2}
+\zeta_\theta = \theta_S - \theta_C
+\end{equation}
+
+\begin{equation}\label{eq:thetacoscon}
+|\zeta_\theta| < \frac{\vartheta}{2}
+\end{equation}
+
+\subsection{Multi-fan Source}
+A focal point is determined. The focal spot is at the center of each fan beam. The focal spot is always assumed to be at the center of the geometry and at the same $z$ elevation as the source point.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/proj6}
+        \caption{$N=6$}
+        \label{fig:proj6}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/proj12}
+        \caption{$N=12$}
+        \label{fig:proj12}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+    %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/proj16}
+        \caption{$N=16$}
+        \label{fig:proj16}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/proj24}
+        \caption{$N=24$}
+        \label{fig:proj24}
+    \end{subfigure}
+    \caption{Different numbers of fan beam projections.}\label{fig:fanproj}
+\end{figure}
+
+\begin{equation}\label{eq:xp}
+x' = x \cos \theta - y \sin \theta
+\end{equation}
+
+\begin{equation}\label{eq:yp}
+y' = y \cos \theta + x \sin \theta
+\end{equation}
+
+To rotate a point $(x,y)$ about an arbitrary point $(x_f, y_f)$, Eq.~\ref{eq:xp2} and~\ref{eq:yp2} are used.
+
+\begin{equation}\label{eq:xp2}
+x' = (x-x_f) \cos \theta - (y-y_f) \sin \theta + x_f
+\end{equation}
+
+\begin{equation}\label{eq:yp2}
+x' = (y - y_f) \cos \theta + (x - x_f) \sin \theta + y_f
+\end{equation}
+
+\subsection{Cone Source}
+Pointy!
+
+\subsection{Multi-cone Source}
+All of the pointy!
+
+\subsection{Multileaf Collimator}
+How fancy.
+
 \endinput

### 2017-04-24 doctors/mainwindow.cpp
Added fan beam support

@@ -162,6 +162,14 @@ MainWindow::MainWindow(QWidget *parent):
     ui->tabWidget->tabBar()->setTabTextColor(3, Qt::red);
     ui->tabWidget->tabBar()->setTabTextColor(4, Qt::red);
 
+    // Add unicode Greek text
+    const QChar phySymbol(0x03C6);
+    const QChar thetaSymbol(0x03B8);
+    ui->sourcePhiLabel->setText(phySymbol);
+    ui->sourceThetaLabel->setText(thetaSymbol);
+    ui->sourcePhiLabel->setFont(QFont("Times New Roman"));
+    ui->sourceThetaLabel->setFont(QFont("Times New Roman"));
+
     quadDialog->updateQuad(m_quad);
 }
 
@@ -542,7 +550,10 @@ void MainWindow::on_actionMCNP6_Generation_triggered()
         return;
     }
 
-    if(!m_srcParams->update(energyDialog->getUserIntensity(), ui->sourceXDoubleSpinBox->value(), ui->sourceYDoubleSpinBox->value(), ui->sourceZDoubleSpinBox->value()))
+    if(!m_srcParams->update(energyDialog->getUserIntensity(), ui->sourceXDoubleSpinBox->value(), ui->sourceYDoubleSpinBox->value(), ui->sourceZDoubleSpinBox->value(),
+                            ui->sourcePhiSpinBox->value(), ui->sourceThetaSpinBox->value(), ui->sourceNSpinBox->value(), ui->sourceDegreesCheckBox->isChecked(),
+                            ui->sourceDistanceSpinBox->value(), ui->sourceWidthSpinBox->value(), ui->sourceHeightSpinBox->value(),
+                            ui->sourceTypeComboBox->currentIndex()))
     {
         QString errmsg = QString("Could not update the energy intensity data. The energy structure may have changed.");
         QMessageBox::warning(this, "Invalid Vector Size", errmsg, QMessageBox::Close);
@@ -637,7 +648,10 @@ void MainWindow::energyGroupsUpdated()
 {
     if(m_srcParams != NULL)
     {
-        if(!m_srcParams->update(energyDialog->getUserIntensity(), ui->sourceXDoubleSpinBox->value(), ui->sourceYDoubleSpinBox->value(), ui->sourceZDoubleSpinBox->value()))
[…]

### 2017-04-24 doctors/mesh.cpp

@@ -72,6 +72,10 @@ void Mesh::uniform(const int xelems, const int yelems, const int zelems, const f
 
     vol.resize(xElemCt * yElemCt * zElemCt);
 
+    m_maxX = xLen;
+    m_maxY = yLen;
+    m_maxZ = zLen;
+
     // The coordinates between mesh elements
     for(unsigned int i = 0; i < xElemCt+1; i++)  // iterate the xMesh+1 to get the last bin
         xNodes[i] = i * (xLen / xElemCt);

### 2017-04-24 doctors/mesh.h

@@ -53,6 +53,10 @@ public:
     /** atom density in [atom/b-cm] */
     std::vector<float> atomDensity;
 
+    float getMaxX() const {return m_maxX;}
+    float getMaxY() const {return m_maxY;}
+    float getMaxZ() const {return m_maxZ;}
+
     void load(const Config *config, const Quadrature *quad);
 
     unsigned int voxelCount() const;
@@ -75,6 +79,10 @@ public:
 private:
     bool insideBox(int x, int y, int z, int xmin, int xmax, int ymin, int ymax, int zmin, int zmax);
     bool insideTightBox(int x, int y, int z, int xmin, int xmax, int ymin, int ymax, int zmin, int zmax);
+
+    float m_maxX;
+    float m_maxY;
+    float m_maxZ;
 };
 
 #endif // MESH_H

### 2017-04-24 doctors/solver.cpp

@@ -116,12 +116,88 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *quad, const Mesh 
     const unsigned short DIRECTION_Y = 2;
     const unsigned short DIRECTION_Z = 3;
 
-    const RAY_T sx = static_cast<RAY_T>(srcPar->sourceX);
-    const RAY_T sy = static_cast<RAY_T>(srcPar->sourceY);
-    const RAY_T sz = static_cast<RAY_T>(srcPar->sourceZ);
+    std::vector<RAY_T> sx;
+    std::vector<RAY_T> sy;
+    std::vector<RAY_T> sz;
 
+    RAY_T xt;
+    RAY_T yt;
+
+    RAY_T focusX = mesh->getMaxX()/2;
+    RAY_T focusY = mesh->getMaxY()/2;
+    RAY_T focusZ = mesh->getMaxZ()/2;
+
+    // Build a list of point sources
+    switch(srcPar->sourceType)
+    {
+    case 0:  // Isotropic
+        sx.push_back(static_cast<RAY_T>(srcPar->sourceX));
+        sy.push_back(static_cast<RAY_T>(srcPar->sourceY));
+        sz.push_back(static_cast<RAY_T>(srcPar->sourceZ));
+        break;
+
+    case 1:  // Fan
+        sx.push_back(static_cast<RAY_T>(srcPar->sourceX));
+        sy.push_back(static_cast<RAY_T>(srcPar->sourceY));
+        sz.push_back(static_cast<RAY_T>(srcPar->sourceZ));
+        break;
+
+    case 2:  // Multifan
+        sx.push_back(static_cast<RAY_T>(srcPar->sourceX));
+        sy.push_back(static_cast<RAY_T>(srcPar->sourceY));
+        sz.push_back(static_cast<RAY_T>(srcPar->sourceZ));
+
+        RAY_T dt = 2*m_pi / srcPar->sourceN;
+
+        for(unsigned int i = 0; i < srcPar->sourceN-1; i++)
+        {
+            xt = sx[sx.size()-1];
+            yt = sy[sy.size()-1];
+            sx.push_back((xt-focusX)*cos(dt) - (yt-focusY)*sin(dt) + focusX);
+            sy.push_back((yt-focusY)*cos(dt) + (xt-focusX)*sin(dt) + focusY);
+            sz.push_back(srcPar->sourceZ);
+        }
+        break;
+
+    case 3:  // Cone
[…]

### 2017-04-24 doctors/sourceparams.cpp

@@ -2,7 +2,7 @@
 
 #include "xs_reader/ampxparser.h"
 
-SourceParams::SourceParams(AmpxParser *parser)
+SourceParams::SourceParams(AmpxParser *parser) : sourceType(-1)
 {
     //spectraEnergyLimits.resize(parser->getGammaEnergyGroups()+1);
     spectraIntensity.resize(parser->getGammaEnergyGroups(), 0);
@@ -40,7 +40,7 @@ bool SourceParams::normalize()
     return true;
 }
 
-bool SourceParams::update(std::vector<float> e, double x, double y, double z)
+bool SourceParams::update(std::vector<float> e, double x, double y, double z, double phi, double theta, int n, bool degrees, double d, double w, double h, int type)
 {
     if(spectraIntensity.size() != e.size())
     {
@@ -50,5 +50,13 @@ bool SourceParams::update(std::vector<float> e, double x, double y, double z)
     sourceX = x;
     sourceY = y;
     sourceZ = z;
+    sourcePhi = phi;
+    sourceTheta = theta;
+    sourceN = n;
+    this->degrees = degrees;
+    sourceD = d;
+    sourceW = w;
+    sourceH = h;
+    sourceType = type;
     return normalize();
 }

### 2017-04-24 doctors/sourceparams.h

@@ -11,13 +11,20 @@ public:
     SourceParams(AmpxParser *parser);
     ~SourceParams();
 
+    int sourceType;
+
     float sourceX, sourceY, sourceZ;
+    float sourcePhi, sourceTheta;
+    int sourceN;
+    bool degrees;
+
+    float sourceD, sourceW, sourceH;
 
     std::vector<float> spectraEnergyLimits;
     std::vector<float> spectraIntensity;
 
     bool normalize();
-    bool update(std::vector<float> e, double x, double y, double z);
+    bool update(std::vector<float> e, double x, double y, double z, double phi, double theta, int n, bool degrees, double d, double w, double h, int type);
 };
 
 #endif // SOURCEPARAMS_H

### 2017-05-01 docprep/chapmeth.tex
Added the AMPX parsing section and a figure

@@
 \label{fig:energy_groups}
-\end{figure}%
+\end{figure}
+
+The energy group structure should be selected appropriately for the problem.
+
+The cross sections currently availabl eare those available from the SCALE 6.2 distribution.
+Those cross sections are designed for light water reactor analysis and are not well suited to medical physics applications.
+This is currently one of the greatest limitaitons of DOCOTRS.
+The SCALE cross section data also contains neutron data which is useless for the CT dosimetry DOCTORS performs.
+
+Once the energy group structure is selected, cross section data must be available.
+Group averaged cross section values from energy $E_1$ to $E_2$ are computed using Eq.~\ref{eq:groupxs} where $\sigma(E)$ is the continuous microscopic cross section.
+A weighting function, $f$, is used to weight some energies.
+The weighting function can be any of a number commonly used functions.
+The following section discusses the data format used by the DOCTORS code.
+
+\begin{equation}\label{eq:groupxs}
+\sigma_G = \frac{\int_{E_1}^{E_2}f(E')\sigma(E') dE'}{\int_{E_1}^{E_2}}
+\end{equation}
 
@@
 AMPX cross sections are read in binary. They are formatted like so. They are read like so. They were obtained from SCALE6.2. These files can be generated by other codes.
+
+DOCTORS reads AMPX fromatted binary data files. Figure~\ref{fig:ampx} shows the data format of the overall file.
+
+Bondarenko data and resonance parameters.
+
+The next block of data contains the average cros section data.
+
+The 2D data is stored in a special format optimized for scatter matrix data.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/ampx1}
+        \caption{The AMPX file format}
+        \label{fig:ampx1}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/ampx2}
+        \caption{The structure of each individual record.}
+        \label{fig:ampx2}
+    \end{subfigure}
+    \caption{The format of the AMPX file.}\label{fig:ampx}
+\end{figure}
+
+The binary data file is composed of multiple records, each of one of thirteen types. Each record is formatted differently to hold different data.
 

### 2017-05-01 doctors/mainwindow.cpp
Added energy presets

@@ -230,19 +230,32 @@ void MainWindow::on_launchSolverPushButton_clicked()
             return;
     }
 
+    unsigned int srcType = ui->sourceTypeComboBox->currentIndex();
+    if((srcType == 2 || srcType == 4) && ui->sourceNSpinBox->value() < 2)
+    {
+        QString errmsg = QString("You don't have at least 2 fan beams on a multi-fan source. ");
+        errmsg += "You may either abort the run or ignore this warning (which may produce incorrect results).";
+        int resp = QMessageBox::warning(this, "Internal Error", errmsg, QMessageBox::Abort | QMessageBox::Ignore);
+        if(resp == QMessageBox::Abort)
+            return;
+    }
+
     if(m_srcParams == NULL)
     {
-        QString errmsg = QString("The energy spectra must be loaded before a MCNP6 file can be generated.");
+        QString errmsg = QString("The energy spectra must be loaded before the solver can be run");
         QMessageBox::warning(this, "Insufficient Data", errmsg, QMessageBox::Close);
         return;
     }
 
-    //if(!m_srcParams->update(energyDialog->getUserIntensity(), ui->sourceXDoubleSpinBox->value(), ui->sourceYDoubleSpinBox->value(), ui->sourceZDoubleSpinBox->value()))
-    //{
-    //    QString errmsg = QString("Could not update the energy intensity data. The energy structure may have changed.");
-    //    QMessageBox::warning(this, "Invalid Vector Size", errmsg, QMessageBox::Close);
-    //    return;
-    //}
+    if(!m_srcParams->update(energyDialog->getUserIntensity(), ui->sourceXDoubleSpinBox->value(), ui->sourceYDoubleSpinBox->value(), ui->sourceZDoubleSpinBox->value(),
+                            ui->sourcePhiSpinBox->value(), ui->sourceThetaSpinBox->value(), ui->sourceNSpinBox->value(), ui->sourceDegreesCheckBox->isChecked(),
+                            ui->sourceDistanceSpinBox->value(), ui->sourceWidthSpinBox->value(), ui->sourceHeightSpinBox->value(),
+                            ui->sourceTypeComboBox->currentIndex()))
+    {
+        QString errmsg = QString("Could not update the energy intensity data. The energy structure may have changed.");
+        QMessageBox::warning(this, "Invalid Vector Size", errmsg, QMessageBox::Close);
+        return;
+    }
 
     if(!m_srcParams->normalize())
     {

### 2017-05-01 doctors/solver.cpp

@@ -108,7 +108,7 @@ void Solver::gsSolverHarmonic(const Quadrature *quad, const Mesh *mesh, const XS
  * ========================================= CPU Code ========================================= *
  ************************************************************************************************/
 
-std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SourceParams *srcPar)
+std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *, const Mesh *mesh, const XSection *xs, const SourceParams *srcPar)
 {
     unsigned int groups = xs->groupCount();
 
@@ -119,9 +119,11 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *quad, const Mesh 
     std::vector<RAY_T> sx;
     std::vector<RAY_T> sy;
     std::vector<RAY_T> sz;
+    int sourceCt = 1;
 
     RAY_T xt;
     RAY_T yt;
+    RAY_T dt;
 
     RAY_T focusX = mesh->getMaxX()/2;
     RAY_T focusY = mesh->getMaxY()/2;
@@ -147,9 +149,9 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *quad, const Mesh 
         sy.push_back(static_cast<RAY_T>(srcPar->sourceY));
         sz.push_back(static_cast<RAY_T>(srcPar->sourceZ));
 
-        RAY_T dt = 2*m_pi / srcPar->sourceN;
+        dt = 2*m_pi / srcPar->sourceN;
 
-        for(unsigned int i = 0; i < srcPar->sourceN-1; i++)
+        for(int i = 0; i < srcPar->sourceN-1; i++)
         {
             xt = sx[sx.size()-1];
             yt = sy[sy.size()-1];
@@ -157,6 +159,8 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *quad, const Mesh 
             sy.push_back((yt-focusY)*cos(dt) + (xt-focusX)*sin(dt) + focusY);
             sz.push_back(srcPar->sourceZ);
         }
+
+        sourceCt = srcPar->sourceN;
         break;
 
     case 3:  // Cone
@@ -173,7 +177,7 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *quad, const Mesh 
 
     default:
[…]

### 2017-05-01 docprep/chapmeth.tex
Added more description of the AMPX format

@@
 
-DOCTORS reads AMPX fromatted binary data files. Figure~\ref{fig:ampx} shows the data format of the overall file.
+The file is split into a series of records.
+Each record is a string of binary bytes between a four byte header and four byte footer.
+The header and footer are identical and, when interpreted as a signed four byte integer, give the size in words of the record.
+A word is a 32-bit chunk (4 bytes) of binary data.
+The data is stored in Big Endian format which must be converted to Little Endian for most Intel and AMD processors.
+An example record is shown in Fig.~\ref{fig:ampxbytes1} and the endianess rearrangement is shown in Fig.~\ref{fig:ampxbytes2}.
+Once the bytes are properly ordered, the entire string can be reinterpreted as a 32-bit integer.
+In the case of the header bytes given in Fig.~\ref{fig:ampxbytes1} and~\ref{fig:ampxbytes2}, the bytes correspond to the integer 440 which is the length in bytes of the header record.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/ampxbytes1}
+  \end{center}
+  \caption{The bytes}
+\label{fig:ampxbytes1}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/ampxbytes2}
+  \end{center}
+  \caption{The byte ordering.}
+\label{fig:ampxbytes2}
+\end{figure}
+
+The size always reports the number of 32-bit words required to store the data. However, some data types, such as characters, are only 8 bits so each word represents four distinct characters.
+
+DOCTORS reads AMPX fromatted binary data files.
+Figure~\ref{fig:ampx} shows the data format of the overall file.
+The header section contains information about each individual nuclide such as which reaction types are available as well as information the same such as the group structure information.
+
+The first entry in a nuclide is the directory record. The directory contains general information regarding the nuclide and information necessary to parse the proceeding records.
 
@@
 
-The next block of data contains the average cros section data.
+The next block of data contains the average cross section data.
+This data is averaged over all energies and directions using Eq.~\ref{eq:groupxs}.
+The only data used in this section in DOCTORS is the total cross section values necessary for implementing the fully discretized form of the LBE given later.
 
@@
 \label{fig:spatial_disc}
-\end{figure}%
+\end{figure}
 

### 2017-05-01 docprep/chapmeth.tex
Added a note about cone beams and deleted and old image from beamshapes

@@
 
-\subsection{Cone Source}
-Pointy!
+\subsection{Cone Source and Multi-cone Source}
+Cone sources are identical to fan beam sources except for the rejection criteria used to determine whether or not a voxel is inside the beam.
+The rejection crieria is a function of only a single parameter, $\theta$, the maximum permissible angle between the center of the beam and a ray toward the point of interest.
 
-\subsection{Multi-cone Source}
-All of the pointy!
+The actual angle is $\zeta$ which is computed similarly to Eq.~\ref{eq:phicos} without the projection onto the $xy$ plane in Eq.~\ref{eq:zetacos} can be simplified to a simple summation since both $\vec{S}$ and $\vec{C}$ are unit vectors.
+
+\begin{equation}\label{eq:zetacos}
+\cos(\zeta) = \frac{<S_x, S_y, S_z>}{\sqrt{S_x^2 + S_y^2 + S_z^2}} \cdot \frac{<C_x, C_y, C_z>}{\sqrt{C_x^2 + C_y^2 + C_z^2}} = S_x C_x + S_y C_y + S_z C_z
+\end{equation}
+
+\begin{equation}\label{eq:zetacon}
+|\zeta| < \theta
+\end{equation}
 

### 2017-05-02 doctors/gui/energydialog.cpp
.spec files can now be imported

@@ -4,6 +4,8 @@
 #include <QDebug>
 #include <QMouseEvent>
 #include <QGraphicsRectItem>
+#include <QFileDialog>
+#include <QMessageBox>
 
 #include "xs_reader/ampxparser.h"
 #include "gui/colormappable.h"
@@ -340,7 +342,49 @@ void EnergyDialog::on_energyPresetComboBox_activated(int indx)
     }
 
     update();
-    //repaint();
+}
+
+void EnergyDialog::on_energyOpenPushButton_clicked()
+{
+    QString filename = QFileDialog::getOpenFileName(this, "Open Spectrum Formatted Data File", "/media/data/thesis/doctors/data/", "SPEC (*.spec);;All Files (*)");
+
+    if(filename.isEmpty())
+    {
+        qDebug() << "Error, failed to load spectrum data file";
+        return;
+    }
+
+    std::vector<float> values;
+    std::ifstream fin(filename.toStdString().c_str());
+
+    if(fin.good())
+    {
+        float f;
+        while(fin >> f)
+        {
+            values.push_back(f);
+        }
+    }
+
+    // Check to make sure the correct number of groups were loaded
+    if(values.size() != m_energyBins.size()-1)
+    {
+        QString errmsg = QString("The XS Data file has ") + QString::number(m_energyBins.size()) + " energy groups, but " + filename + " contained " + QString::number(values.size()) + " data points";
+        QMessageBox::warning(this, "Data Size Mismatch", errmsg, QMessageBox::Close);
+        return;
+    }
+
+    float sum = 0.0f;
+    for(unsigned int iv = 0; iv < values.size(); iv++)
+        sum += values[iv];
+
+    for(unsigned int iv = 0; iv < values.size(); iv++)
+    {
+        QRectF old = m_rects[iv]->rect();
+        m_rects[iv]->setRect(old.x(), old.y(), old.width(), values[iv]/sum);
+    }
+
+    return;
 }
 
 

### 2017-05-02 doctors/gui/energydialog.h

@@ -73,6 +73,7 @@ public slots:
 
     void on_energyOkPushButton_clicked();
     void on_energyPresetComboBox_activated(int);
+    void on_energyOpenPushButton_clicked();
 };
 
 #endif // ENERGYDIALOG_H

### 2017-05-05 docprep/chapbib.tex
Added a new chapter - results.

@@
 
-This section is the literature review. I do these things.
-
-\section{Introduction}
-Many different things are required.
+This chapter summarizes the literature reviewed in preparation of this work.
+The literature review is split into multiple sections, each highlighting relevant works pertaining to a particular component or section of this work.
+The first sections cover introductory work and steps necessary for preprocessing data.
+Section~\ref{sec:discordlit} covers historical and recent work in discrete ordinate methodology, the methodology used by DOCTORS.
+The remaining sections cover GPU hardware and other implementation facets.
 
 \section{CT Dose Estimation}
-How is this done now?
+The dose a patient receives form both diagnostic CT procedures and radiation therapy is of great concern to the medical community [CITE].
+
+Monte Carlo codes have been used in the past to compute flux-to-dose conversion factors for a reference person [CITE ICRP].
 
@@
 
-Two phantom datasets are used in this work.
+Two phantom datasets are used in this work [CITE].
 
-\section{Discrete Ordinates}
+\section{Discrete Ordinates}\label{sec:discordlit}
 Discrete ordinates are old. The book is \citet{ref:lewise}.

### 2017-05-05 docprep/chapconc.tex
Updated the thesis chapters and the beamshapes figures
Added a new chapter - results.

--- 2017-03-01 docprep/appdx_a.tex
@@
 
-This appendix covers the math for Legendre polynomials and spherical harmonics
+Conclusions.
 
-\section{Legendre Polynomials}
+\section{First conclusion}
 
-They exist.
+I concluded many things.
 
-\subsection{Spherical Harmonics}
+\section{Future Work}
 
-They also exist.
+A number of simple modifications to DOCTORS could be made that would greatly increase the code's usability and robustness.
+One of these modifications, generation of more refined group structures, is not a change to DOCTORS \textit{per se} but rather a change in the input cross section data.
+The other changes are additions to DOCTORS that would add new capabilities to the software.
 
-The real harmonics. The even/odd harmonics.
+In addition to specific modifications, some additional, broader future goals can also be identified.
 
-The addition theorem.
+\subsection{Group Structure}
+
+The cross section data currently used by DOCTORS are taken from SCALE6.2.
+While these cross sections have been found to be sufficient to produce flux distributions in medical CT imaging, a more refined group structure designed for medical applications would be worth investigation.
+
+This work can be done using either NJOY [CITE] or the newly released AMPX [CITE] code.
+Either code has the capability to collapse an ENDF formatted data file into a group structure of the user's choosing in the format DOCTORS reads.
+
+NJOY or AMPX.
+
+\subsection{Temporal Extension}
+
+Use 4D CT to do temporal stuff as well as spatial stuff.
+
+\subsection{Multileaf Collimation}
+Collimators!
 

### 2017-05-05 docprep/chapimpl.tex
Added a new chapter - results.

%%
%% This is file `chapall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin,addfig,addtbl,addbib,addglo')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

 %% ... sample chapter ...

This chapter covers the details of the implementation of the different components of the DOCTORS code base.

The code is available from the GitHub repository. At the time of this writing, the repository is located at \texttt{https://github.com/eNorris/thesis}.

The code is implemented in C++ and requires a C0x11 complianct compiler. The user interface is written using the Qt5 framework. GPU computation is performed via CUDA which must be compiled seperatly by Nvidia's proprietary compiler, \texttt{nvcc}.

\section{Vector Flattening}

Data is stored as \texttt{std::vector<>} objects.

Figure~\ref{fig:indx_ex} gives an example of the spatial indexing scheme using an $8 \times 8 \times 8$ mesh. The indicated voxel has $i_x$, $i_y$, and $i_z$ indices of 1, 5, and 6 respectively. Therefore, the index value of the highlighted voxel is $1 \times 8 \times 8 + 5 \times 8 + 6$ = 110. This flattening pattern continues to higher dimensional phase space to encompas energy and direction.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/indx_ex}
  \end{center}
  \caption{The indexing scheme.}
\label{fig:indx_ex}
\end{figure}

A variable that is dependent on $x$, $y$, $z$, $E$, and $\Omega$ would be written as $\phi(x, y, z, E, \Omega)$ would be discretized to $\phi(i_x, i_y, i_z, G, i_a)$. Rather than storing $\phi$ as a 5-dimensional matrix, $\phi$ is stored as a 1-dimensional vector of the same number of elements.

\section{Qt5 Framework}
The Qt5 framework was used for implementation of the user interface. Qt5 enables asynchronous calls through its ssignal/slot mechanism. A signal can be emitted which will execute all connected slots. Signals and slots can be connected manually by the user except for ones automatically generated by Qt. Qt prvides a user interface for building user interfaces. Components such a sbuttons, drop boxes, radio buttons, etc. are offered to the user. Qt makes extensive use of polymorphism. All user components inherit from the base \texttt{QWidget} class which inherits from \texttt{QObject}. Any class that utilizes the signal/sloot mechanism must extend the \texttt{QOjbect} class.

Listing~\ref{lst:sync1} gives a pseudocode snippet that has a long function that will block the user interface. A corresponding dequence diagram is given in Fig.~\ref{fig:sync1}. When the user interacts with the GUI, the writer object begins executing the \texttt{doWrite()} function. Until this function completes, the UI thread will be busy an dunable to handle additionaly user interaction or updates. This results in the GUI becoming unresponsive and potentially issuing a warning to the user from the operating system.

\begin{listing}
\begin{minted}[frame=lines,linenos]{python}
import numpy as np
def myfcn(hello,world):
   return hello*world
\end{minted}
\caption{Blah blah blah.}\label{lst:sync1}
\end{listing}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/writer_sync}
  \end{center}
  \caption{Sequence diagram for the synchronous call.}
\label{fig:sync1}
\end{figure}%

The code in Listing~\ref{lst:sync1} is updated to use the Qt signal/slot mechanism.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/writer_async}
  \end{center}
  \caption{Sequence diagram for the synchronous call.}
\label{fig:async1}
\end{figure}%

\section{CUDA}
CUDA code is compoiled with the Nvidia complier nvcc. Qt uses the gcc compiler and its MOC generator for meta code. In order to connect CUDA code to the Qt MOC, the CUDA code is compiled by nvcc to produce a .o file. Qt then compiles all other files into corresponding.o files. The linker than automactically picks up all of the .o files gneerated by nvcc. The final result is an executable that has a Qt generarted user interface that can communicat with an Nvidia graphics card through the CUDA language.

The sweep through the mesh is different.

\begin{table}[ht]
\caption{Subsweep Indices}
\centering 
\begin{tabular}{l | c | c | c | c}
  \hline \hline   
  i  & ix & iy & iz & ix+iy+iz \\ [0.5ex] % inserts table 
  \hline
  0  & 4 & 0 & 0 & 4\\
  1  & 3 & 1 & 0 & 4\\
  2  & 3 & 0 & 1 & 4\\
  3  & 2 & 2 & 0 & 4\\
  4  & 2 & 1 & 1 & 4\\
  5  & 2 & 0 & 2 & 4\\
  6  & 1 & 3 & 0 & 4\\
  7  & 1 & 2 & 1 & 4\\
  8  & 1 & 1 & 2 & 4\\
  9  & 1 & 0 & 3 & 4\\
  10 & 0 & 4 & 0 & 4\\
  11 & 0 & 3 & 1 & 4\\
  12 & 0 & 2 & 2 & 4\\
  13 & 0 & 1 & 3 & 4\\
  14 & 0 & 0 & 4 & 4\\ [1ex]      % [1ex] adds vertical space
  \hline
\end{tabular}
\label{table:subsweep}
\end{table}

We first split the sweep process into $N_L$ subsweeps. For a $N_x \times N_x \times N_x$ mesh, the subsweep process is straightforward. Some of the initial subsweeps are shown for a $8 \times 8 \times 8$ mesh in Fig.~\ref{fig:subsweep_cube}. This is valid for any direction, $\hat{\Omega}$, for which all three of its direction cosines ($\mu$, $\xi$, and $\eta$) are positive.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_cube1}
        \caption{Subsweep 0 ($S=0$)}
        \label{fig:subsweep_cube1}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_cube2}
        \caption{Subsweep 1 ($S=1$)}
        \label{fig:subsweep_cube2}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_cube3}
        \caption{Subsweep 6 ($S=6$)}
        \label{fig:subsweep_cube3}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_cube4}
        \caption{Subsweep 6 ($S=6$) with labels}
        \label{fig:subsweep_cube4}
    \end{subfigure}
    \caption{The progression of subsweeps throughout a sweep. Each subsweep must complete before those after it. Each voxel within a subsweep can be solved in parallel with all others in its subsweep.}\label{fig:subsweep_cube}
\end{figure}

In general, blah blah.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general1}
        \caption{$S=0$}
        \label{fig:subsweep_general1}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general2}
        \caption{$S=1$}
        \label{fig:subsweep_general2}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general3}
        \caption{$S=2$}
        \label{fig:subsweep_general3}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general4}
        \caption{$S=3$}
        \label{fig:subsweep_general4}
    \end{subfigure}

    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general5}
        \caption{$S=4$}
        \label{fig:subsweep_general5}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general6}
        \caption{$S=5$}
        \label{fig:subsweep_general6}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general7}
        \caption{$S=6$}
        \label{fig:subsweep_general7}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general8}
        \caption{$S=7$}
        \label{fig:subsweep_general8}
    \end{subfigure}

    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general9}
        \caption{$S=8$}
        \label{fig:subsweep_general9}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general10}
        \caption{$S=9$}
        \label{fig:subsweep_general10}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general11}
        \caption{$S=10$}
        \label{fig:subsweep_general11}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general12}
        \caption{$S=11$}
        \label{fig:subsweep_general12}
    \end{subfigure}

    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general13}
        \caption{$S=12$}
        \label{fig:subsweep_general13}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general14}
        \caption{$S=13$}
        \label{fig:subsweep_general14}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/subsweep_general15}
        \caption{$S=14$}
        \label{fig:subsweep_general15}
    \end{subfigure}
    \caption{The generalized subsweep.}\label{fig:subsweep_general}
\end{figure}

The number of parallel tasks, $P$, that can be done on subsweep $S$ of an $N_x \times N_y \times N_z$ mesh is given by Eq.~\ref{eq:taskspersub}.

\begin{equation}\label{eq:taskspersub}
P = C_S - L_x - L_y - L_z + G_{xy} + G_{yz} + G_{xz}
\end{equation}

\begin{equation}
C_S = \frac{(S+1)(S+2)}{2}
\end{equation}

\begin{equation}
L_x = \frac{d_x(d_x+1)}{2}
\end{equation}

\begin{equation}
L_y = \frac{d_y(d_y+1)}{2}
\end{equation}

\begin{equation}
L_z = \frac{d_z(d_z+1)}{2}
\end{equation}

\begin{equation}
G_{xy} = \frac{d_{xy}(d_{xy}+1)}{2}
\end{equation}

\begin{equation}
G_{yz} = \frac{d_{yz}(d_{yz}+1)}{2}
\end{equation}

\begin{equation}
G_{xz} = \frac{d_{xz}(d_{xz}+1)}{2}
\end{equation}

\begin{equation}
d_x = \max(S+1-N_x, 0)
\end{equation}

\begin{equation}
d_y = \max(S+1-N_y, 0)
\end{equation}

\begin{equation}
d_z = \max(S+1-N_z, 0)
\end{equation}

\begin{equation}
d_{xy} = \max(S+1-N_x - Ny, 0)
\end{equation}

\begin{equation}
d_{yz} = \max(S+1-N_y - Nz, 0)
\end{equation}

\begin{equation}
d_{xz} = \max(S+1-N_x - Nz, 0)
\end{equation}

Each voxel in a subsweep can be computed in parallel. Mathematically, the $i^{th}$ subsweep from all directions can be computed in parallel. However, in practice, this results in a race condition on the GPU hardware.

DELETE BELOW HERE.

The number of elements before the 4$^{th}$ level is 1 + 2 + 3 = 6. This is simply the sum of all integers less than the level number given by Eq.~\ref{eq:levelsum}. This is conincidentally the index of the first element of that level. However, there is a special case of $N_0 = 1$.

\begin{equation} \label{eq:levelsum}
N_L = \sum_{i=1}^{L} = \frac{L(L+1)}{2}
\end{equation}

Given an index, $i$, the level to which it belongs can be computed by substituting $i$ for $N_L$ in Eq.~\ref{eq:levelsum} and solving for $L$. The index value is then floored.

\begin{equation} \label{eq:levelquadratic}
i = \frac{L(L+1)}{2} \rightarrow L^2 + L - 2i = 0
\end{equation}

\begin{equation} \label{eq:levelquadsol1}
L = \left \lfloor{\frac{-1 \pm \sqrt{1+8i}}{2}} \right \rfloor
\end{equation}

Since the index must be a positive value, the $\pm$ sign can be removed from Eq.~\ref{eq:levelquadsol1} yielding Eq.~\ref{eq:levelquadsol2}

\begin{equation} \label{eq:levelquadsol2}
L = \left \lfloor{\frac{-1 + \sqrt{1+8i}}{2}} \right \rfloor
\end{equation}

The maximum index of any level in a the $N$ subsweep is $N$. Each subsweep adds an additional level. Therefore, the ix, iy, and iz indices can be computed using Eq.~\ref{eq:levelix},~\ref{eq:leveliy}, and~\ref{eq:leveliz} respectively.

\begin{equation} \label{eq:levelix}
ix = N - N_L
\end{equation}

\begin{equation} \label{eq:leveliy}
iy = L + N_L - i
\end{equation}

\begin{equation} \label{eq:leveliz}
iz = L + i - N_L
\end{equation}

\section{MCNP Generation}\label{sec:mcnpgen}
DOCTORS has the capability to generate MCNP6 input files from the CT mesh data and source specification provided by the user.

\section{Hardware}
For this work, a computer with an Intel i7-5960X 8 core (16 hyperthreads) processor with a base clock speed of 3.X GHz and an Nvidia Titan Z graphics card was used. Currently, if the problem requires more memory than is available on the GPU, the problem can still be solved, but much more time will be required to to copy overhead between the CPU and GPU. If the problem requires more memory than either the GPU or CPU can provide, an error is thrown and the simulation is not run.

CUDA is Nvidia only so AMD is not considered.

\endinput
%%
%% End of file `chapall.tex'.

### 2017-05-05 docprep/chapmeth.tex
Added a note about convergence and a figure for beam angles.
Added a new chapter - results.

@@
 
+\subsection{Convergence Criteria}
+The typical converence criteria used to determine whether or to terminate further iteration is the maximum relative error from the $i-1^{th}$ iteration to the $i^{th}$ iteration given in Eq.~\ref{eq:conv}.
+Once the relative error in the parameter of interest, $\phi$ is below some threshold, $\epsilon$, the iteration stops and the solution is considered converged.
+
+\begin{equation}\label{eq:conv}
+\frac{\phi_i - \phi_{i-1}}{\phi_i} < \epsilon
+\end{equation}
+
+In some problems, the convergence criteria given in Eq.~\ref{eq:conv} is not reached for many iterations.
+In these cases, terminating the solution early rather than waiting for full convergence is appropriate.
+This is enforced by a second criterial.
+Whenever the number of iterations, $i$, exceeds the total permissible number of iterations, $I$, the iteration is terminated.
+
+\begin{equation}\label{eq:conv2}
+i > I
+\end{equation}
+
+Under further investigation, the cause of the convergence failure that results in Eq.~\ref{eq:conv2} being exercised often stems from floating point error.
+The flux magnitude fluctuates wildly in regions of very flow flux, such as the gantry.
+In these regions, the relative uncertainty will remain much higher than in regions with smoother flux distributions.
+FIGURE XYZ SHOWS.
+
+In order to measure the uncertainty due to floating-point error, another convergence criteria was added.
+Rather than looking at the relative change from iteration to iteration, the total absolute change is measured, computed by Eq.~\ref{eq:conv3}.
+The termination criteria is given by Eq.~\ref{eq:conv4}.
+Computation of a value for $\epsilon$ requires two previous iteration, thus this condition cannot be met until the third iteration and thereafter.
+
+\begin{equation}\label{eq:conv3}
+\epsilon = \sum_{\vec{r}} |\phi_i - \phi_{i-1}|
+\end{equation}
+
+\begin{equation}\label{eq:conv4}
+\epsilon_i > \epsilon_{i-1}
+\end{equation}
+
 \subsection{Anisotropy Treatment}
@@
 
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/beamconexy}
+        \caption{Cone xy}
+        \label{fig:beamconexy}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/beamfanxy}
+        \caption{Fan xy}
+        \label{fig:beamfanxy}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/beamconeyz}
+        \caption{Cone yz}
+        \label{fig:subsweep_general3}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/beamfanyz}
+        \caption{Fan yz}
+        \label{fig:subsweep_general4}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/beamconexz}
+        \caption{Cone xz}
+        \label{fig:subsweep_general5}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/beamfanxz}
+        \caption{Fan xz}
+        \label{fig:subsweep_general6}
+    \end{subfigure}
+    \caption{The beam shapes.}\label{fig:beamfancone}
+\end{figure}
+
 \subsection{Multileaf Collimator}
 How fancy.
+
+Talk about the MCNP generation in the comparison to results part.
 

### 2017-05-05 docprep/chapresults.tex
Added a new chapter - results.

%%
%% This is file `chapmin.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapmin.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

\section{Preprocessing}
Two computational phantoms were used for verification of the DOCTORS code.

Some preprocessing was done to each phantom before it was used in DOCTORS to provide results.

The results from DOCTORS were compared to analagous results from MCNP6. The MCNP6 reference compared to is automatically generated by DOCTORS to guarantee that the geometry and source parameters are identical. This process is described in detail in Section~\ref{sec:mcnpgen}.

\subsection{Artifact Removal}
At the periphery of the computational mesh, artifacts can appear as a byproduct of the reconstruction algorithm used. Figure~\ref{fig:phantomWaterUnmodified} shows the artifacts in question in a $sy$ slice at the lowest ($z = 0$) level in the phantom. Two hundred sixty one (261) artifacts were detected in the phantom. To remove these artifacts, the CT number of any voxels whose CT number was greater than or equal to 65500 was set to 0. Figure~\ref{fig:phantomWaterModified} shows the same cross sectional plot after the artifacts are removed.

The water phantom in Fig.~\ref{fig:phantomWaterModified} can be broken into four regions: the water phantom, the container, the air, and the corner artifacts. The water phantom is the centermost region of the slice

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/phantomWaterUnmodified}
  \end{center}
  \caption{The unmodified phantom. An $xy$ slice at $z$ index = 0.}
\label{fig:phantomWaterUnmodified}
\end{figure}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/phantomWaterModified}
  \end{center}
  \caption{The phantom after artifacts are removed. An $xy$ slice at $z$ index = 0.}
\label{fig:phantomWaterModified}
\end{figure}

\begin{table}[ht]
\caption{Water Regions}
\centering 
\begin{tabular}{l c r r}
\hline \hline   
Region    & CT Range & Voxels & Fraction (\%)\\ [0.5ex] 
\hline
Water     & $0 \leq x < 67$ & 1581199 & 37.70 \\
Container & $67 \leq x < 600$ & 81502 & 1.94 \\
Air       & $600 \leq x < 1080$ & 1650372  & 39.35 \\
Artifact  & $1080 \leq x < 65535$ & 881231 & 21.01 \\  [1ex]
\hline
\end{tabular}
\label{table:nonlin}
\end{table}

Once the artifacts are removed, the histogram is plotted. Figure~\ref{fig:waterHist} shows two versions of the water histogram, \ref{fig:waterHistLin} on a linear scale and \ref{fig:waterHistLog} on a logarithmic scale.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/waterHistogramLin2}
        \caption{Linear}
        \label{fig:waterHistLin}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/waterHistogramLog}
        \caption{Logarithmic}
        \label{fig:waterHistLog}
    \end{subfigure}
    \caption{The histogram.}\label{fig:waterHist}
\end{figure}

\subsection{Gantry Handling}
The gantry is there.

\subsection{Geometry Simplification}
The original phantom data for both the water phantom and the liver phantom are $256\times256\times64$ meshes which result in 4,194,304 total voxels. Though MCNP6 is capable of running such a mesh, the overhead of loading the mesh into memory and initiating the Monte Carlo solver can take on the order of hours. Therefore, the benchmarks were run using a simplified geometry of $64\times64\times16$ which has only 65,536 voxels. The smaller files can run to completion of $10^9$ particles within two hours.

\section{Benchmark}

MCNP is good.

\section{Comparison}

They agree.

\endinput
%%
%% End of file `chapmin.tex'.

### 2017-05-12 python/ct_data_reader/main/reader.py
Font is now set to Times New Roman

@@ -1,8 +1,10 @@
 import numpy
 import matplotlib
 import matplotlib.pyplot as pyplot
+import matplotlib.font_manager
 
 __author__ = 'etnc6d'
+
 
 def parse(filename):
     with open(filename, 'rb') as f:
@@ -10,17 +12,18 @@ def parse(filename):
 
         return a
 
+
 def write(filename, mtrx):
     print("Writing a matrix of size: " + str(mtrx.shape))
     with open(filename, 'wb') as f:
         f.write(mtrx.tobytes())
-        #numpy.ndarray.tofile(f, "")
 
-def simplify(mtrx):
-    outmtrx = numpy.zeros((16, 64, 64), dtype=numpy.uint16)
 
-    print(outmtrx.shape)
-    print(mtrx.shape)
+def simplify(matrix):
+    out_matrix = numpy.zeros((16, 64, 64))  # , dtype=numpy.uint16)  # Setting the dtype is required for saving
+
+    # print(out_matrix.shape)
+    # print(matrix.shape)
 
     for ox in range(64):
         for oy in range(64):
@@ -29,26 +32,30 @@ def simplify(mtrx):
                 for ix in range(4):
                     for iy in range(4):
                         for iz in range(4):
-                            v += mtrx[4*oz+iz, 4*oy+iy, 4*ox+ix]
-                outmtrx[oz, oy, ox] = v / (4*4*4)
+                            v += matrix[4*oz+iz, 4*oy+iy, 4*ox+ix]
+                out_matrix[oz, oy, ox] = v / (4*4*4)
 
-    return outmtrx
+    return out_matrix
 
+# Set the font
 font = {'family': 'Times New Roman', 'size': 16}
 matplotlib.rc('font', **font)
 
+# Fixes issue where Times New Roman is bold by default
+del matplotlib.font_manager.weight_dict['roman']
+matplotlib.font_manager._rebuild()
+
 mtrx = parse("/media/Storage/thesis/doctors/data/water35_volume.bin")
-
-
-#mtrx[mtrx > 256] = 256
-
 
 print("Size should be: " + str(256*256*64))
 print(mtrx.shape)
 
-#mtrx[mtrx==65535] = 0
[…]

### 2017-05-12 docprep/chapbib.tex
Worked on the intro and results

@@
 
- %% ... sample chapter ...
-
 This chapter summarizes the literature reviewed in preparation of this work.
 The literature review is split into multiple sections, each highlighting relevant works pertaining to a particular component or section of this work.
-The first sections cover introductory work and steps necessary for preprocessing data.
-Section~\ref{sec:discordlit} covers historical and recent work in discrete ordinate methodology, the methodology used by DOCTORS.
+The first two sections cover introductory work and steps necessary for preprocessing data.
+Section~\ref{sec:discordlit} covers historical and recent work in discrete ordinate solution methodology, the methodology used by DOCTORS.
 The remaining sections cover GPU hardware and other implementation facets.
@@
 
+MOTIVATION
+
 Monte Carlo codes have been used in the past to compute flux-to-dose conversion factors for a reference person [CITE ICRP].
 
+COMPUTATION
+
 \section{CT Phantom Generation}
-The CT phantom is populated with CT numbers, also known as Hounsfield numbers. These values are related to the attenuation coefficient of the material represented by the voxel.
+The CT phantom is populated with CT numbers, also known as Hounsfield numbers.
+These values are related to the attenuation coefficient of the material represented by the voxel.
+Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air CITE though some variations do exist CITE.
 
@@
 The new guy provides 19 materials. The density is measured.
+The DOCTORS code relies on the CT nuber to material mapping of \citet{ref:ottossonr} who developed a 19-group BLAH.
+Table~\ref{table:ctmap} shows the mapping of CT number to material composition (reproduced from \citep{ref:ottossonr}).
+
+\begin{table}[ht]
+\caption{Water Regions}
+\centering 
+\begin{tabular}{l c r r r r r r r r r r r r}
+\hline \hline   
+Media       & CT Range      & H    & C    & N    & 0    & Na  & Mg & P    & S   & Cl  & Ar  & K   & Ca \\ [0.5ex] 
+\hline
+Air         & -950 to -100  &      &      & 75.7 & 23.3 &     &    &      &     &     & 1.3 &     &       \\
+Lung        & -1000 to -950 & 10.3 & 10.5 & 3.1  & 74.9 & 0.2 &    & 0.2  & 0.3 & 0.3 &     & 0.2 &       \\
+Adipose     & -100 to 15    & 11.2 & 50.8 & 1.2  & 36.4 & 0.1 &    &      & 0.1 & 0.1 &     &     &       \\
+Connective  & 15 to 129     & 10.0 & 16.3 & 4.3  & 68.4 & 0.4 &    &      & 0.4 & 0.3 &     &     &       \\
+Bone        & 129 to 200    & 9.7  & 44.7 & 2.5  & 35.9 &     &    & 2.3  & 0.2 & 0.1 &     & 1.0 & 4.5   \\  [1ex]
+\hline
+\end{tabular}
+\label{table:ctmap}
+\end{table}
 
@@
 \section{Discrete Ordinates}\label{sec:discordlit}
-Discrete ordinates are old. The book is \citet{ref:lewise}.
+The dsicrete ordinates method dates back to CITE but remains one of the most prominent solution modalities for ration transport in use today.
+
+A comprehensive review of discrete ordinate methods is given by \citet{ref:lewise}. Additional references include
 

### 2017-05-12 docprep/chapintro.tex
Added a new chapter - results.
Worked on the intro and results

@@
 
-Another application is in treatment.
+Currently, dose estimation relies on \textit{a priori} computation verified with a standardized benchmark.
+No methodology currently exists to verify that the dosimetry evaluation was accurate after the patient has undergone the procedure.
+This work proposes a methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam.
 
-Currently, no algorithm exists to provide this information.
+Another, similar, application is dose estimation of patients receiving radiation therapy.
+Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment.
 
-The goal of this work is to provide such an algorithm.
+ESTIMATE. PHANTOM. CORRELATE.
 
-There are two primary avaialble for the methodologies: Monte Carlo and deterministic methods.
+BASED ON REFERENCE PHANTOM.
+
+ADD FIGURE
+
+Once the procedure (whether diagnostic or treatment) is completed, the dose distribution is assumed to follow the empirical model.
+
+There are two primary methodologies available to provide the transport solution: Monte Carlo and deterministic methods.
 Monte Carlo has proven to be very slow in these areas.
-Furhter, the 3D Cartesian mesh generated from CT reconstruction is well suited to deterministic computation.
+Further, the 3D Cartesian mesh generated from CT reconstruction is well suited to deterministic computation.
 
 This work forcuses on the discrete ordinates method and presents an implementation specially developed for dose estimation from a medical CT scan.
+
+Computational mesh from a sinogram. Reconstruction.
+The DOCTORS code takes a computational reconstruction mesh of CT numbers, cross section data, and other parameters related to the solution methodology as input.
+The output is the collided and uncollided flux as well as the dose in each computational cell.
+These values can be compared to the empirical correlation results as a verification that the patient received an appropriate dose.
+
+SHOULD BE FAST - GPU
+
+USER INTERFACE
 

### 2017-05-12 docprep/chapmeth.tex
Worked on the intro and results

@@
 
-\section{Mathmatic Development}
+This chapter summarizes the solution methodology employed by DOCTORS.
+The techniques used to compute the collided and uncollided fluxes are described in seperate sections.
+After the flux methodology is explained, the methods used for source generation and covered as well as the flux to dose conversion.
 
-Where does this text go?
+\section{Collided Solution Methods}
+
+The discrete ordinate solver computes the collided flux distribution inside the CT mesh given an input uncollided flux, physical geometry information, and other associated solver parameters such as quadrature and energy discretization.
 
@@
 
-The time dependent gamma transport equation can be written as
+The time dependent gamma transport equation is
 
@@
 \begin{split}
-	&\left[ \frac{1}{v(E)} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
+	&\left[ \frac{1}{v(\boldsymbol{r}), t} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
 	\psi(\boldsymbol{r}, E, \hat{\Omega}, t) = \\
@@
 
-However, we are typically not concerned with the transient case in medical diagnostic imaging.
-We are more interested in the steady state case.
-The time independent form of Eq.
-\ref{eq:boltz_t_dep} is written as
+where $v$ is the velocity of the photon in the medium at $\boldsymbol{r}$, $\Sigma_t$ is the total macroscopic cross section, $\Sigma_s$ is the macroscopic scattering cross section, $S$ is an external photon source, and $\phi$ is the total photon angular flux.
+However, the transient case is not typically of interest in medical diagnostic imaging, rather the steady state is more important.
+The time dependence in Eq.
+\ref{eq:boltz_t_dep} is removed to yield
 
@@
 
-The following sections show how Eq.~\ref{eq:boltz} is discretized.
+The following sections show how Eq.~\ref{eq:boltz} is discretized and solved computationally.
 
@@
 The coordinate system used in DOCTORS is shown in Fig.~\ref{fig:coord_sys}.
-A discrete set of $N_a$ angles ($\Omega_{a} \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space.
-Particles are transported only along these discrete directions form one voxel to adjacent voxels.
-
-The rotation symmetrical quadrature is implemented in DOCTORS.
-There are many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
-This enables simple, direct comparison to other solvers.
-Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.
+A discrete set of $N_a$ angles ($\Omega_{a}, \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space.
+Particles are transported only along these discrete directions from one voxel to adjacent voxels.
 
@@
 \label{fig:coord_sys}
-\end{figure}%
+\end{figure}
+
+The rotation symmetrical quadrature is implemented in DOCTORS.
+There are many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
+This enables simple, direct comparison to other solvers.
+Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.
+
+The plain text file should be formatted such that it contains four columns similar to Table~\ref{tab:quad_format} (without any headings).
+No checks are performed to ensure that octants are covered, this allows simple testing quadratures that utilize only a single direction to be used for testing purposes.
+Values for $\mu$, $\eta$, and $\xi$ do not necessarily need to add to unity, they will be automatically normalized by DOCTORS.
+Likewise, the weights will be normalized such that their summation is unity.
+
+\begin{table}[ht]
+\caption{User Defined Quadrature Format}
+\centering 
+\begin{tabular}{c c c c}
+\hline \hline   
+$\mu$    & $\eta$ & $\xi$ & Weight\\ [0.5ex] 
+\hline
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\
+5.12345  & 5.12345 & 5.12345 & 5.12345 \\ [1ex]
+\hline
+\end{tabular}
+\label{table:quad_format}
+\end{table}
 
@@
 The binary data file is composed of multiple records, each of one of thirteen types. Each record is formatted differently to hold different data.
+
+Once the data is parsed, weighted combinations of elemental data is used to create material cross sections.
+
+WATER EXAMPLE.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/airwaterxs1_19group}
+        \caption{Data}
+        \label{fig:ampx1}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
+        \caption{Same data.}
+        \label{fig:ampx2}
+    \end{subfigure}
+    \caption{Comparison to NIST data.}\label{fig:airwaterxs_19group}
+\end{figure}
 
@@
 
-\section{Uncollided Flux}
+\section{Uncollided Solution Methods}
 To avoid ray effect, the flux is split into collided and uncollided components as shown in Eq.~\ref{eq:totflux}.
@@
 
+\section{Dose Computation}
+Dose is computed.
+
 \endinput

### 2017-05-12 docprep/chapresults.tex
Worked on the intro and results

@@
 
+First subtract 1024 to account for the unsigned 16-bit int data structure.
+
 \subsection{Artifact Removal}
@@
 
+\subsection{Gantry Handling}
+In some phantom models, a gantry is present. The gantry poses a challenge since it is a highly attenuating material relative to the patient which reduces the available flux.
+
+The CT number to material conversion developed by~\cite{ref:ottossonr} which is implemented in this work does not have an equivalent for the gantry region.
+
+\subsection{Geometry Simplification}
+The original phantom data for both the water phantom and the liver phantom are $256\times256\times64$ meshes which result in 4,194,304 total voxels.
+Though MCNP6 is capable of running such a mesh, the overhead of loading the mesh into memory and initiating the Monte Carlo solver can take on the order of hours.
+Therefore, the benchmarks were run using a simplified geometry of $64\times64\times16$ which has only 65,536 voxels.
+The smaller files can run to completion of $10^9$ particles within two hours.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/waterPhantom256}
+  \end{center}
+  \caption{Raw CT number data for the $256 \times 256 \times 64$ water phantom.}
+\label{fig:waterPhantom256}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/waterPhantom64}
+  \end{center}
+  \caption{The small one}
+\label{fig:waterPhantom64}
+\end{figure}
+
+\section{Cone-Beam Water Benchmark}
+
+The first benchmark, the cone with 19 groups is given in Fig. 1. The cross section data for the water and air regions are shown in Fig.~\ref{fig:airwater19}.
+
+The water phantom is a 35 cm diameter water cylinder that stands 12.5 inches tall. The water is encased in a 0.5 cm thick plastic container.
+
 \begin{table}[ht]
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/waterHistogramLin2}
+        \includegraphics[width=\textwidth]{figs/waterHistogramLin}
         \caption{Linear}
@@
 
-\subsection{Gantry Handling}
-The gantry is there.
+Due to the meshing used in CT phantoms, the spatial domain of the problem is identical for both the deterministic and Monte Carlo problems.
+A key difference between though is the group structure used in the group-averaged deterministic cross sections.
+Figure~\ref{fig:airwarerxs} shows the continuous energy macroscopic cross section for both air and water versus the discretized data used by DOCTORS.
+Unfortunately, the datasets used in DOCTORS originate from SCALE6.2, a light water reactor analysis code.
+Therefore, their upper photon energy bound is 20 MeV which is far higher than necessary for diagnostic imaging.
+This results in the region of interesting being discretized very roughly.
+In the 19-group discretization, the entire diagnostic domain is covered by only three energy groups.
 
-\subsection{Geometry Simplification}
-The original phantom data for both the water phantom and the liver phantom are $256\times256\times64$ meshes which result in 4,194,304 total voxels.
-Though MCNP6 is capable of running such a mesh, the overhead of loading the mesh into memory and initiating the Monte Carlo solver can take on the order of hours.
-Therefore, the benchmarks were run using a simplified geometry of $64\times64\times16$ which has only 65,536 voxels.
-The smaller files can run to completion of $10^9$ particles within two hours.
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
+        \caption{19 Groups}
+        \label{fig:airwaterxs19}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_47group}
+        \caption{47 Groups}
+        \label{fig:airwaterxs47}
+    \end{subfigure}
+    \caption{Comparison of the group cross section data used by DOCTORS to the continuous cross section data used by MCNP6. The continuous data is provided by NIST !!!CITE!!!.}\label{fig:airwaterxs}
+\end{figure}
 
-\section{Benchmark}
+The detailed flux distribution from MCNP6 and the solver are not included here, they are provided in Appendix~\ref{sec:appc_benchmark}.
 
-MCNP is good.
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_comp}
+        \caption{10 - 20 MeV.}
+        \label{fig:bench_cone1_comp}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout}
+        \caption{10 - 20 MeV lineout.}
+        \label{fig:bench_cone1_lineout}
+    \end{subfigure}
 
-\section{Comparison}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_comp_g17}
+        \caption{100 - 200 keV.}
+        \label{fig:bench_cone1_comp_g17}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout_g17}
+        \caption{100 - 200 keV lineout.}
+        \label{fig:bench_cone1_lineout_g17}
+    \end{subfigure}
 
-They agree.
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_comp_g18}
+        \caption{45 - 100 keV.}
+        \label{fig:bench_cone1_comp_g18}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout_g18}
+        \caption{45 - 100 keV lineout.}
+        \label{fig:bench_cone1_lineout_g18}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_comp_g19}
+        \caption{10 - 45 keV.}
+        \label{fig:bench_cone1_comp_g19}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout_g19}
+        \caption{10 - 45 keV lineout.}
+        \label{fig:bench_cone1_lineout_g19}
+    \end{subfigure}
+    \caption{Comparison of the MCNP6 results and the raytracer.}\label{fig:waterHist}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_centerdiff}
+        \caption{}
+        \label{fig:bench_cone1_centerdiff}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/bench_cone1_centerdiff2}
+        \caption{}
+        \label{fig:bench_cone1_centerdiff2}
+    \end{subfigure}
+    \caption{Comparison}\label{fig:bench_cone1_centerdifffig}
+\end{figure}
+
+\section{3 Cone-Beam Water Benchmark}
+
+Now there are three
+
+\section{16 Cone-Beam Water Benchmark}
+
+And now 16.
+
+\section{1 Cone-beam Phantom Benchmark}
+
+It's a person!
+
+\section{16 Cone-beam Phantom Benchmark}
+
+It's a person!
 

### 2017-05-12 docprep/norris_phd_dissertation.tex
Added more info on the subsweep
Added a new chapter - results.
Worked on the intro and results

@@
 \usepackage[noprefix]{nomencl}
+\usepackage{subcaption}
+\usepackage{xr}
 \makenomenclature
@@
 \begin{ThesisAcknowledgment}
-I would like to thank these people. They are all so super cool. So much coolness. So much so that this is now over a line long.
+I would like to thank my adviser, Dr.
+Liu and all of the other graduate students for their continued support through my graduate career.
+I also appreciate the time spent by my committee members helping me identify problems and provide insight to solutions.
 
-And these funding agencies.
+I would also like to acknowledge the NRC for providing funding for my work. My work is supported by NRC grant NRC-HQ-13-G-38-0026.
 \end{ThesisAcknowledgment}
+
 \begin{ThesisFrontMatter}
@@
  % ... conclusions chapter ...
+
+\ThesisBodyChapter{Results}\input{chapresults}
+
 \ThesisBodyChapter{Conclusions}\input{chapconc}
@@
 \ThesisAppendixChapter{Legendre Polynomials and Spherical Harmonics}\input{appdx_a}
-\ThesisAppendixChapter{Two}\input{appdx_b}
+
+\ThesisAppendixChapter{Code Listings}\input{appdx_b}
+
+\ThesisAppendixChapter{Benchmark Results}\input{appdx_c}
+
 \end{ThesisAppendix}
@@
 \begin{Vita}
-Edward Norris graduated from Missouri University of Science and Technology in 2013 with a B.S. in both nuclear engineering and computer science.
+Edward Norris graduated from Missouri University of Science and Technology in 2013 with a B.S.
+in both nuclear engineering and computer science, he went on to pursue a M.S.
+in computer science and a PhD in nuclear engineering.
 
-He has worked...
+His PhD work was funded by a fellowship through the NRC.
 
-He is a member of...
+Previously he has participated in four summer internships with Sandia National Laboratories.
+While there, he served in various capacities ranging from cyber defense to radiation detection and neutron scatter simulation.
+
+Edward was awarded SOME AWARDS.
+
+Edward was also an active member of the Nuclear Science Design Team at Missouri University of Science and Technology.
+While on the team he solved radiation transport problems to facilitate the dosimetry of an inertial electrostatic confintement fusion device being constructed.
+
 \end{Vita}

### 2017-05-18 docprep/chapbib.tex
Updated a table

@@
 Connective  & 15 to 129     & 10.0 & 16.3 &  4.3 & 68.4 & 0.4 &     &      & 0.4 & 0.3 &     &     &      \\
-Bone        & 129 to 200    & 9.7  & 44.7 & 2.5  & 35.9 &     &    & 2.3  & 0.2 & 0.1 &     & 1.0 & 4.5   \\  [1ex]
+Bone 1      & 129 to 200    &  9.7 & 44.7 &  2.5 & 35.9 &     &     &  2.3 & 0.2 & 0.1 &     & 1.0 &  4.5 \\ 
+Bone 2      & 200 to 300    &  9.1 & 41.4 &  2.7 & 36.8 &     & 0.1 &  3.2 & 0.2 & 0.1 &     & 1.0 &  6.3 \\ 
+Bone 3      & 300 to 400    &  8.5 & 37.8 &  2.9 & 37.9 &     & 0.1 &  4.1 & 0.2 & 0.1 &     & 1.0 &  8.2 \\ 
+Bone 4      & 400 to 500    &  8.0 & 34.5 &  3.1 & 38.8 &     & 0.1 &  5.0 & 0.2 & 0.1 &     & 1.0 & 10.0 \\ 
+Bone 5      & 500 to 600    &  7.5 & 31.6 &  3.2 & 39.7 &     & 0.1 &  5.8 & 0.2 & 0.1 &     &     & 11.6 \\ 
+Bone 6      & 600 to 700    &  7.1 & 28.7 &  3.4 & 40.4 &     & 0.1 &  6.6 & 0.2 & 0.1 &     &     & 13.1 \\ 
+Bone 7      & 700 to 800    &  6.7 & 26.7 &  3.5 & 41.2 &     & 0.2 &  7.2 & 0.3 &     &     &     & 14.4 \\ 
+Bone 8      & 800 to 900    &  6.3 & 24.7 &  3.7 & 41.8 &     & 0.2 &  7.8 & 0.3 &     &     &     & 15.7 \\ 
+Bone 9      & 900 to 1000   &  6.0 & 22.7 &  3.8 & 42.4 &     & 0.2 &  8.4 & 0.3 &     &     &     & 16.8 \\ 
+Bone 10     & 1000 to 1100  &  5.6 & 20.7 &  3.9 & 43.0 &     & 0.2 &  8.9 & 0.3 &     &     &     & 17.9 \\ 
+Bone 11     & 1100 to 1200  &  5.3 & 18.7 &  4.0 & 43.5 &     & 0.2 &  9.4 & 0.3 &     &     &     & 18.9 \\ 
+Bone 12     & 1200 to 1300  &  5.1 & 16.7 &  4.1 & 44.0 &     & 0.2 &  9.9 & 0.3 &     &     &     & 19.8 \\
+Bone 13     & 1300 to 1400  &  4.8 & 15.7 &  4.2 & 44.4 &     & 0.2 & 10.3 & 0.3 &     &     &     & 20.7 \\
+Bone 14     & 1400 to 1500  &  4.6 & 13.7 &  4.2 & 44.9 &     & 0.2 & 10.7 & 0.3 &     &     &     & 21.5 \\  
+Bone 15     & > 1500        &  4.3 & 12.7 &  4.3 & 45.3 &     & 0.2 & 11.1 & 0.3 &     &     &     & 22.2 \\[1ex]
 \hline

### 2017-05-18 docprep/chapmeth.tex
Updated a table

@@
 
+However, user defined directions \textit{should not} be parallel to any major axis.
+Doing so may result in undefined behavior.
+Also, after normalizaiton, no two rows should have identical values for $\mu$, $\eta$, and $\xi$ such as the last two rows of Table~\ref{tab:quad_interp}.
+Also, no weights can be negative.
+
 \begin{table}[ht]
-\caption{User Defined Quadrature Format}
+\caption{User Defined Quadrature Input}
 \centering 
@@
 \hline
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\
-5.12345  & 5.12345 & 5.12345 & 5.12345 \\ [1ex]
+1        & 0      & 0      & 1 \\
+0.999    & 0.001  & 0.001  & 1 \\
+1.0      & 1.0    & 1.0    & 1.3 \\
+0.02198  & 0.2987 & .34520 & .02226 \\
+.001     & .001   & .999   & 0.999 \\
+-1       & 1      & -1     & .2 \\
+-1.2     & 1.2    & -1.2   & 0.1 \\
+1        & -1     & -1     & .2 \\ [1ex]
 \hline
 \end{tabular}
-\label{table:quad_format}
+\label{tab:quad_format}
+\end{table}
+
+\begin{table}[ht]
+\caption{User Defined Quadrature Interpretation}
+\centering 
+\begin{tabular}{c c c c}
+\hline \hline   
+$\mu$    & $\eta$ & $\xi$ & Weight\\ [0.5ex] 
+\hline
+1.0000   & 0.0000 & 0.0000 & 0.2074 \\
+0.999    & 0.001  & 0.001  & 0.2074 \\
+1.0      & 1.0    & 1.0    & 0.2696 \\
+0.02198  & 0.2987 & .34520 & 0.0207 \\
+.001     & .001   & .999   & 0.0046 \\
+1        & -1     & -1     & 0.2072 \\
+-1.2     & 1.2    & -1.2   & 0.0415 \\
+-1       & 1      & -1     & 0.0415 \\ [1ex]
+\hline
+\end{tabular}
+\label{tab:quad_interp}
 \end{table}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/airwaterxs1_19group}
-        \caption{Data}
-        \label{fig:ampx1}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
+        \caption{19 group data}
+        \label{fig:airwaterxs2_19group}
     \end{subfigure}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
-        \caption{Same data.}
-        \label{fig:ampx2}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_47group}
+        \caption{47 group data.}
+        \label{fig:airwaterxs2_47group}
     \end{subfigure}
-    \caption{Comparison to NIST data.}\label{fig:airwaterxs_19group}
+    \caption{Comparison of the group-averaged DOCTORS cross section data to continuous NIST data.}\label{fig:airwaterxs_group}
 \end{figure}
@@
 
-\subsection{Multileaf Collimator}
-How fancy.
+\section{Dose Computation}
+Dose can be computed in either one of two ways.
+The first way is to compute the spatial energy dependent fluence distribution and multiply by a fluence-to-dose conversion factor from ICRP 116 [CITE].
+The second way to compute dose is to compute the energy deposition in each cell.
 
-Talk about the MCNP generation in the comparison to results part.
+\subsection{Fluence-to-dose Conversion}
 
-\section{Dose Computation}
-Dose is computed.
+The publications of the ICRP include fluence-to-dose conversion factors for a reference phantom from various particle types and patient orientations.
+For CT systems, the rotationally symmetric (ROT) orientation is the most appropriate since the beam rotates about the patient.
+
+This methodology is very simplistic and can be computed quickly with no extra (memory) overhead. When DOCTORS finishes running, the flux is multiplied by the photon
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/icrp116_rotconv}
+  \end{center}
+  \caption{Fluence-to-dose conversion factors from ICRP 116 for photons.}
+\label{fig:icrp116_rotconv}
+\end{figure}
+
+\subsection{Energy Deposition}
+
+blah
 

### 2017-05-29 docprep/chapintro.tex
Filled in some of the gaps

@@
 
-Computed tomography (CT) is becoming increasingly pervasive in medical diagnostics because improved algorithms are giving doctors more access to patient information.
+Computed tomography (CT) is becoming increasingly pervasive in medical diagnostics as improved algorithms give doctors access to more information.
 Increased information results in faster and more accurate diagnosis.
-However, the CT operation itself gives a radiation dose to the patient which carries some associated risk.
-A more detailed measure of that risk would help doctors make informed decisions regarding whether a CT scan is warannted, and if so, which type.
-This can also help doctors estimate spatial dose distribution to the patient to ensure no specific organ received more dose than permissible.
+However, the CT operation itself delivers a radiation dose to the patient increasing risk of long term complications such as cancer [CITE].
+Currently, dose estimation relies on \textit{a priori} computation verified with a standardized benchmark.
+Once the procedure (whether diagnostic or treatment) is completed, the dose distribution is assumed to follow the empirical model.
 
-Currently, dose estimation relies on \textit{a priori} computation verified with a standardized benchmark.
 No methodology currently exists to verify that the dosimetry evaluation was accurate after the patient has undergone the procedure.
 This work proposes a methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam.
+This can also help doctors estimate spatial dose distribution to the patient to ensure no specific organ received more dose than permissible.
 
@@
 Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment.
+With further develpment, this methodology may enable real time dose computation as the treatment is administered.
+This would be greatly beneficial to both patients and the doctors administering the procedure.
 
-ESTIMATE. PHANTOM. CORRELATE.
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.38\textwidth}
+        \includegraphics[width=\textwidth]{figs/backproject1}
+        \caption{}
+        \label{fig:beamconexy}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.19\textwidth}
+        \includegraphics[width=\textwidth]{figs/backproject2}
+        \caption{}
+        \label{fig:beamfanxy}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.38\textwidth}
+        \includegraphics[width=\textwidth]{figs/backproject3}
+        \caption{}
+        \label{fig:subsweep_general3}
+    \end{subfigure}
+    \caption{(a) The original }\label{fig:backprojection}
+\end{figure}
 
-BASED ON REFERENCE PHANTOM.
+In order for the proposed methodology to be viable, the patient's CT phantom must be available.
+The phantom is reconstructed from the CT data.
+Though it is not the focus of this work, a brief summary of reconstruction algorithms is given here.
 
-ADD FIGURE
+Fundamentally, all reconstructions algorithms are based on a thin parallel beam of x-rays that roatate axially about a patient producing a sinogram.
+The inverse radon transform backprojects each row of the sinogram back to physical space..
+Filtering each line in Fourier space and then compositing all layers together recreates th eimage [CTIE].
+This methodology has been extended to fan beams [CITE] and cone beams [CITE].
+After the filtered backprojection, iterative reconstruction techniques can be applied to greatly improve the reconstruction quality [CITE].
 
-Once the procedure (whether diagnostic or treatment) is completed, the dose distribution is assumed to follow the empirical model.
+Along any particular ray, monoenergetic particles will attenuate through a homogeneus media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert} [CITE] resulting in a small fraction of the source particles reaching the detector.
+This phenomena leads to an image of the attenutating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
+With sufficient rays from varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation.
+
+\begin{equation}\label{eq:beer_lambert}
+I(x) = I_0 e^{-\mu x}
+\end{equation}
+
+Once a patient specific aphantom is obtained, the flux throughout the mesh must be computed. Computational techniques to do this cam be split into two broad categories: Monte Carlo and deterministic.
+
+Monte Carlo codes are generally known for their accuracy; they are able to sample the problem and attain arbitrarily high precision given sufficient runtime.
+They also take advantage of combinatorial geometry allowing a user to produce geometrically complex 3D structures.
+Deterministic methods are generally much faster than their Monte Carlo counterparts bu tdo not benefit from excess runtime.
+They also consume vastly more resources, particularly random-access memory (RAM).
+In regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
+A large number of spatial regions slows down th eparticle transport.
+Monte Carlo will also require a large amoutn of RAM to store the numerous voxels.
+
+The Hounsfield units (also called CT numbers) are traditionally scaled such that -1000 is dry air and water is zero.
+
+Deterministic codes can be accelerated with the use of a grpahics processin gunit (GPU).
+GPUs differ from the central processin gunit (CPU) in that  each of th emany cores performes identical instruction to all others at the same time bue each with different data.
+For example, consider a grapyscale 1024$\times$768 image.
+If some pixelwise operation is applied on a CPU, 786,432 operations must occur.
+Onboard the GPU with 1024 cores, every pixel in the entire row can be computed simultaneously reducing the number of operations to 768.
+However, issues such as copying the memory to and from the GPU and cache coherency problems can degrade performance.
+GPU tehcnology and performance is discussed more throuroughly in Section~\ref{sec:cuda}.
+
+In addition to rapidly computing th epatient dose, another goal of DOCTORS is to present th ecode in a user-friendly fashion.
+To this end, a grpahical user interface (GUI) was developed.
+The GUI utilize the Qt5 graphics framework for the window, buttons, and other necessary widgets.
+The GUI leads the user through th esteps necessary to use th ecode in an intuitive way by using colors to indicated requaired and completed steps.
+The output is then plotted graphically and the output is also sent to an ASCII text file for more advanced postprocessing.
+More details about th eGUI are included in Section~\ref{sec:gui}.
 
@@
 
-Computational mesh from a sinogram. Reconstruction.
 The DOCTORS code takes a computational reconstruction mesh of CT numbers, cross section data, and other parameters related to the solution methodology as input.
@@
 
-SHOULD BE FAST - GPU
-
-USER INTERFACE
-
 \endinput

### 2017-05-29 docprep/chapimpl.tex
Moved the AMPX parsing section to implementation

@@
 
-\section{Qt5 Framework}
+\section{Cross Section Parsing}\label{sec:xsparse}
+AMPX cross sections are stored in a binary format.
+The binary file is split into a series of records.
+Each record contains a string of binary bytes between a four byte header and four byte footer.
+The header and footer are identical and, when interpreted as a signed four byte integer, give the size in words of the record.
+Each word is a 32-bit section (4~bytes) of binary data.
+An example record is shown in Fig.~\ref{fig:ampxbytes1}.
+
+The data is stored in Big Endian format which must be converted to Little Endian for most Intel and AMD processors.
+The endianess rearrangement is shown in Fig.~\ref{fig:ampxbytes2}.
+All records are one of 12 unique types.
+Each type of record is formatted and interpreted differently.
+For example, Type 1 contains general information about the data file such as the number of nuclides, number of energy groups, and a brief description.
+Type 2 records contains a list of floating-point numbers representing the energy group boundaries.
+Figure~\ref{fig:ampxbytes1} shows a Type 1 record.
+In the case of the header bytes given in Fig.~\ref{fig:ampxbytes1} and~\ref{fig:ampxbytes2}, the bytes correspond to the integer 440 which is the length in bytes of the record (not including the header or footer).
+
+Each word of data is sequentially read, reordered, and converted to an apprpriately typed variable. Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/ampxbytes1}
+  \end{center}
+  \caption{An example record parsed. The header and footer are identical and are the number of 32-bit words in the record.}
+\label{fig:ampxbytes1}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=2.75in]{figs/ampxbytes2}
+  \end{center}
+  \caption{The byte reordering from Big Endian to Little Endian. Individual bits withing a byte are not rearranged, but the ordering of the four bytes withing a 32-bit word is reversed.}
+\label{fig:ampxbytes2}
+\end{figure}
+
+\begin{table}[ht]
+\caption{Byte Reordering}
+\centering 
+\begin{tabular}{l | c | c | c | l}
+  \hline \hline   
+  Word  & Big Endian & Little Endian & Interpreted & Notes\\ [0.5ex] % inserts table 
+  \hline
+  0   & B8 01 00 00 & 00 00 01 B8 & 440    & Header                    \\
+  1   & 8B 69 00 00 & 00 00 69 8B & 27019  & ID                        \\
+  2   & A4 01 00 00 & 00 00 01 A4 & 420    & Number of nuclides        \\
+  ... &      ---    &      ---    &    --- & ---                       \\
+  11  & 70 75 6F 43 & 43 6F 75 70 & "Coup" & First word of the title   \\
+  12  & 20 64 65 6C & 6C 65 64 20 & "led " & Second word of the title  \\
+  ... &     ---     &    ---      &  ---   & ---                       \\
+  440 & 20 20 20 20 & 20 20 20 20 & "    " & 4 spaces ending the title \\
+  441 & B8 01 00 00 & 00 00 01 B8 & 440    & Footer                    \\ 
+  [1ex]      % [1ex] adds vertical space
+  \hline
+\end{tabular}
+\label{table:reorder}
+\end{table}
+
+The header always reports the number of 8-bit bytes required to store the data.
+However, some data types, such as texttt{char}, are only one byte so each word represents multiple (4 in the case of \texttt{char}) distinct characters.
+
+DOCTORS reads AMPX fromatted binary data files. Figure~\ref{fig:ampx} shows the data format of the overall file.
+
+The first entry in each nuclide entry is the directory record.
+The directory contains general information regarding the nuclide and information necessary to parse the proceeding records.
+Following the directory, Bondarenko data, resonance parameters, neutron data, and gamma production which are not of interest in photon only problems are read.
+
+The penultimate record contains the average cross section data.
+This data is averaged over all energies and directions using Eq.~\ref{eq:groupxs}.
+The only data used in this section in DOCTORS is the total cross section values necessary for implementing the fully discretized form of the LBE given in Eq.~\ref{eq:boltz_i}.
+
+The 2D data is stored in a special format optimized for scatter matrix data.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/ampx1}
+        \caption{The AMPX file format}
+        \label{fig:ampx1}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/ampx2}
+        \caption{The structure of each individual record.}
+        \label{fig:ampx2}
+    \end{subfigure}
+    \caption{The format of the AMPX file.
+(a) The file contains a header, a list of directory records, energy group information and a list of nuclide entries.
+(b) Each nuclide entry contains multiple records; the two of concern in this work are the last two containing gamma data.}\label{fig:ampx}
+\end{figure}
+
+SHOW CARBON ELEMENTAL DATA
+
+\section{Generation of Material Cross Section Data}\label{sec:xsgen}
+
+Once the data is parsed, weighted combinations of elemental data is used to create material cross sections. The details of this computation are covered in Section~\ref{sec:
+
+WATER EXAMPLE.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
+        \caption{19 group data}
+        \label{fig:airwaterxs2_19group}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/airwaterxs2_47group}
+        \caption{47 group data.}
+        \label{fig:airwaterxs2_47group}
+    \end{subfigure}
+    \caption{Comparison of the group-averaged DOCTORS cross section data to continuous NIST data.}\label{fig:airwaterxs_group}
+\end{figure}
+
+\section{Qt5 Framework}\label{sec:qt}
 The Qt5 framework was used for implementation of the user interface.
@@
 
-\section{CUDA}
+\section{Graphical User Interface}\label{sec:gui}
+
+The GUI is cool. Lots of pictures of it.
+
+\section{CUDA}\label{sec:cuda}
 CUDA code is compoiled with the Nvidia complier nvcc.
@@
 
-\section{Hardware}
+\section{Hardware}\label{sec:hardware}
 For this work, a computer with an Intel i7-5960X 8 core (16 hyperthreads) processor with a base clock speed of 3.X GHz and an Nvidia Titan Z graphics card was used.

### 2017-05-30 docprep/chapmeth.tex
Many fixes for readability

@@
 The techniques used to compute the collided and uncollided fluxes are described in seperate sections.
-After the flux methodology is explained, the methods used for source generation and covered as well as the flux to dose conversion.
+After the flux solution methodology is explained, the methods used for source generation and flux to dose conversion are covered.
 
-\section{Collided Solution Methods}
+\section{Discrete Ordinate Methods}
 
-The discrete ordinate solver computes the collided flux distribution inside the CT mesh given an input uncollided flux, physical geometry information, and other associated solver parameters such as quadrature and energy discretization.
+The discrete ordinate solver computes the flux distribution inside the CT mesh given a source, physical geometry information, and other associated solver parameters such as quadrature and energy discretization.
+Discrete ordinates is a deterministic solution to the linear Boltzmann equation.
 
@@
 
-The time dependent gamma transport equation is
+The linear Boltzmann equation (LBE) is generally true of any particle given that outside forces (electromagnetic or graitational) are either not present or negligible and particle-particle collisions are insignificant.
+Therefore, the LBE is generally not valid for charged parpticles or systems whose particle density is on the order of the medium they are traveling trhrough.
+In such systems, particle-particle ineractions may not be negligible.
+
+In general, the LBE is valid for multiplying media (such as neutrons passing through fuel in a reactor), but the variant considered in this work omits such terms since only photons in medical systems are of concern.
+The time dependent linear Boltzmann equation for photons is
 
@@
 \begin{split}
-	&\left[ \frac{1}{v(\boldsymbol{r}), t} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
+	&\left[ \frac{1}{v(\boldsymbol{r},t)} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
 	\psi(\boldsymbol{r}, E, \hat{\Omega}, t) = \\
@@
 
-where $v$ is the velocity of the photon in the medium at $\boldsymbol{r}$, $\Sigma_t$ is the total macroscopic cross section, $\Sigma_s$ is the macroscopic scattering cross section, $S$ is an external photon source, and $\phi$ is the total photon angular flux.
-However, the transient case is not typically of interest in medical diagnostic imaging, rather the steady state is more important.
-The time dependence in Eq.
-\ref{eq:boltz_t_dep} is removed to yield
+\noindent
+where $v$ is the velocity of the photon in the medium at $\boldsymbol{r}$, $\Sigma_t$ is the total macroscopic cross section, $\Sigma_s$ is the macroscopic scattering cross section, $S$ is an external photon source, and $\psi$ is the total photon angular flux.
+
+The transient case is not typically of interest in medical diagnostic imaging, rather the steady state is more important. The time dependence in Eq. \ref{eq:boltz_t_dep} is removed to yield
 
@@
 
+The steady state form of the LBE can be readily derived from simple intuition.
+In the steady state, particle production and removal must be equal.
+In any volume, three production mechanisms are present.
+Particles can stream into the volume across a bounding surface, they can inscatter from another energy-direction component of phase space, or they can be produced directly by the external source.
+Three removal terms are also present, particles can stream out of the volume of interest by crossing a bounding surface, be absorbed and vanish entirely, or outscatter to a energy-direction componennt of phase space other than the one of interest.
+
+The downscatter and absorbtiption removal terms are typically grouped together as
+
+\begin{equation}
+\Sigma_t(\boldsymbol{r}, E) \psi(\boldsymbol{r}, E, \hat{\Omega})
+\end{equation}
+
+\noindent
+and both production and removal streaming operators are combined as well as
+
+\begin{equation}
+\hat{\Omega} \cdot \nabla \psi(\boldsymbol{r}, E, \hat{\Omega})
+\end{equation}
+
+\noindent
+and included as a removal term since they normal is defined pointing outward from the volume.
+The normal sign convention results in particles streaming out being positive and those streaming in beocming negative.
+The inscatter to a particular energy and direction is computed by integrating over all other energies and diretions
+
+\begin{equation}
+\int_{4\pi}^{} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'
+\end{equation}
+
 The following sections show how Eq.~\ref{eq:boltz} is discretized and solved computationally.
@@
 A discrete set of $N_a$ angles ($\Omega_{a}, \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space.
-Particles are transported only along these discrete directions from one voxel to adjacent voxels.
+Particles are transported only along these discrete directions from a voxel to adjacent voxels.
 
@@
 The rotation symmetrical quadrature is implemented in DOCTORS.
-There are many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers.
-This enables simple, direct comparison to other solvers.
+From amongst many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers [CITE] which enables simple, direct comparison to other solvers.
 Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.
 
-The plain text file should be formatted such that it contains four columns similar to Table~\ref{tab:quad_format} (without any headings).
+THEORY
+
+To inlcude a user defined quadrature, the plain text file should be formatted such that it contains four columns similar to Table~\ref{tab:quad_format} (without any headings).
 No checks are performed to ensure that octants are covered, this allows simple testing quadratures that utilize only a single direction to be used for testing purposes.
@@
 Likewise, the weights will be normalized such that their summation is unity.
-
 However, user defined directions \textit{should not} be parallel to any major axis.
@@
 Also, after normalizaiton, no two rows should have identical values for $\mu$, $\eta$, and $\xi$ such as the last two rows of Table~\ref{tab:quad_interp}.
-Also, no weights can be negative.
+Additionally, no weights may be negative.
+
+Values are read in as a 32 bit IEEE-754 floating point number [CITE] which ensures seven significant digits.
+Significant digits beyond this will be truncated.
+Note that Table~\ref{tab:quad_interp} rounds four digits after the decimal.
+
+The data in Table~\ref{tab:quad_format} would be read in and renormalized by DOCTORS. The normalized data would be identical to the data in Table~\ref{tab:quad_interp}.
 
@@
 -1.2     & 1.2    & -1.2   & 0.1 \\
-1        & -1     & -1     & .2 \\ [1ex]
+-1        & 1     & -1     & .2 \\ [1ex]
 \hline
@@
  1.0000  &  0.0000  &  0.0000  & 0.2074 \\
-0.999    & 0.001  & 0.001  & 0.2074 \\
-1.0      & 1.0    & 1.0    & 0.2696 \\
-0.02198  & 0.2987 & .34520 & 0.0207 \\
-.001     & .001   & .999   & 0.0046 \\
-1        & -1     & -1     & 0.2072 \\
--1.2     & 1.2    & -1.2   & 0.0415 \\
--1       & 1      & -1     & 0.0415 \\ [1ex]
+ 1.0000  &  0.0010  &  0.0010  & 0.2074 \\
+ 0.5774  &  0.5774  &  0.5774  & 0.2696 \\
+ 0.0481  &  0.6536  &  0.7553  & 0.0207 \\
+ 0.5000  & -0.5000  &  0.7071  & 0.0046 \\
+ 0.5774  & -0.5774  & -0.5774  & 0.2072 \\
+-0.5774  &  0.5774  & -0.5774  & 0.0415 \\
+-0.5774  &  0.5774  & -0.5774  & 0.0415 \\ [1ex]
 \hline
@@
 
-A given direction, $\hat{\Omega}$ is determined by its three cosine components.
+A given direction, $\hat{\Omega}$, is determined by its three cosine components.
 This relation is given in Eq.~\ref{eq:omega_cos} where $\hat{i}$, $\hat{j}$, and $\hat{k}$ are the unit directions in the $x$, $y$, and $z$ directions respectively.
@@
 Each discrete angle is associated a weight.
-The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by Eq.~\ref{eq:disc_int}.
+The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by
 
 \begin{equation} \label{eq:disc_int}
-\int_{4 \pi} f(\hat{\Omega}) d\Omega \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a
+\int_{4 \pi} f(\hat{\Omega}) d\hat{\Omega} \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a
 \end{equation}
 
-Discretizing Eq.~\ref{eq:boltz} in angular space gives Eq.~\ref{eq:boltz_a} where the angle is denoted by subscript $a$.
+Discretizing Eq.~\ref{eq:boltz} in angular space gives
 
@@
 
+\noindent
+where the subscript $a$ denotes the direction.
+
 \subsection{Energy Discretization}
@@
 Therefore, Eq.~\ref{eq:boltz_a} becomes the energy discretized Eq.~\ref{eq:boltz_e} where the group number is indexed by superscript $g$.
-The group structure is shown in Fig.~\ref{fig:energy_groups}.
 
@@
 \psi_{a}^{g}(\boldsymbol{r}) = 
+\sum_{a=0}^{N_a-1} \sum_{g'=0}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
+\end{equation}
+
+In photon based problems, upscatter is typicall negligible [CITE].
+By the time photons become low enough in energy to upscatter significantly, they are no longer of dosimetric concern.
+Removing upscatter reduces Eq.~\ref{eq:boltz_e} to
+
+\begin{equation} \label{eq:boltz_e2}
+\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t^g(\boldsymbol{r}) \right]
+\psi_{a}^{g}(\boldsymbol{r}) = 
 \sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
 \end{equation}
+
+\noindent
+which is very similar; only the starting point for the scatter summation changes.
+Computationall, though, this makes a significant different.
+This means that the highest energy group can be solved independently of all those below it.
+Once that group is solved, the next group can also be solved and so forth.
 
@@
 
+The group structure cannot be arbitrarily selected by DOCOTRS, instead it is determined by the external cross section data file.
 The energy group structure should be selected appropriately for the problem.
-
-The cross sections currently availabl eare those available from the SCALE 6.2 distribution.
+The cross sections currently available are those available from the SCALE 6.2 distribution.
 Those cross sections are designed for light water reactor analysis and are not well suited to medical physics applications.
@@
 The SCALE cross section data also contains neutron data which is useless for the CT dosimetry DOCTORS performs.
+More details of the cross section parsing and generation are included in Section~\ref{sec:xsparse} and~\ref{sec:xsgen}.
 
@@
 The weighting function can be any of a number commonly used functions.
-The following section discusses the data format used by the DOCTORS code.
 
 \begin{equation}\label{eq:groupxs}
-\sigma_G = \frac{\int_{E_1}^{E_2}f(E')\sigma(E') dE'}{\int_{E_1}^{E_2}}
+\sigma_G = \frac{\int_{E_1}^{E_2}f(E')\sigma(E') dE'}{\int_{E_1}^{E_2} dE'}
 \end{equation}
-
-\subsection{Cross Section Parsing}
-AMPX cross sections are read in binary. They are formatted like so. They are read like so. They were obtained from SCALE6.2. These files can be generated by other codes.
-
-The file is split into a series of records.
-Each record is a string of binary bytes between a four byte header and four byte footer.
-The header and footer are identical and, when interpreted as a signed four byte integer, give the size in words of the record.
-A word is a 32-bit chunk (4 bytes) of binary data.
-The data is stored in Big Endian format which must be converted to Little Endian for most Intel and AMD processors.
-An example record is shown in Fig.~\ref{fig:ampxbytes1} and the endianess rearrangement is shown in Fig.~\ref{fig:ampxbytes2}.
-Once the bytes are properly ordered, the entire string can be reinterpreted as a 32-bit integer.
-In the case of the header bytes given in Fig.~\ref{fig:ampxbytes1} and~\ref{fig:ampxbytes2}, the bytes correspond to the integer 440 which is the length in bytes of the header record.
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/ampxbytes1}
-  \end{center}
-  \caption{The bytes}
-\label{fig:ampxbytes1}
-\end{figure}
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/ampxbytes2}
-  \end{center}
-  \caption{The byte ordering.}
-\label{fig:ampxbytes2}
-\end{figure}
-
-The size always reports the number of 32-bit words required to store the data. However, some data types, such as characters, are only 8 bits so each word represents four distinct characters.
-
-DOCTORS reads AMPX fromatted binary data files.
-Figure~\ref{fig:ampx} shows the data format of the overall file.
-The header section contains information about each individual nuclide such as which reaction types are available as well as information the same such as the group structure information.
-
-The first entry in a nuclide is the directory record. The directory contains general information regarding the nuclide and information necessary to parse the proceeding records.
-
-Bondarenko data and resonance parameters.
-
-The next block of data contains the average cross section data.
-This data is averaged over all energies and directions using Eq.~\ref{eq:groupxs}.
-The only data used in this section in DOCTORS is the total cross section values necessary for implementing the fully discretized form of the LBE given later.
-
-The 2D data is stored in a special format optimized for scatter matrix data.
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/ampx1}
-        \caption{The AMPX file format}
-        \label{fig:ampx1}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/ampx2}
-        \caption{The structure of each individual record.}
-        \label{fig:ampx2}
-    \end{subfigure}
-    \caption{The format of the AMPX file.}\label{fig:ampx}
-\end{figure}
-
-The binary data file is composed of multiple records, each of one of thirteen types. Each record is formatted differently to hold different data.
-
-Once the data is parsed, weighted combinations of elemental data is used to create material cross sections.
-
-WATER EXAMPLE.
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
-        \caption{19 group data}
-        \label{fig:airwaterxs2_19group}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/airwaterxs2_47group}
-        \caption{47 group data.}
-        \label{fig:airwaterxs2_47group}
-    \end{subfigure}
-    \caption{Comparison of the group-averaged DOCTORS cross section data to continuous NIST data.}\label{fig:airwaterxs_group}
-\end{figure}
 
@@
 
-Finally, Eq.~\ref{eq:boltz_e} is discretized in space to give Eq.~\ref{eq:boltz_i} which is fully discretized.
 The problem domain is split into an evenly spaced Cartesian grid.
@@
 
-The problem domain must be a regular rectangular parallelpiped of size $D_x$, $D_y$, and $D_z$ in the $x$, $y$, and $z$ directions respectively.
-The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each direction.
+The problem domain must be a regular, rectangular parallelpiped of dimension $D_x \times D_y \times D_z$.
+The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each major direction.
 The total number of voxels, $N_V$ is easily computed as the multiple of each of the component directions given in Eq.~\ref{eq:n_v}.
@@
 
-Voxel indices are reordered into a one dimensional vector (flattened) according to the rule given in Eq.~\ref{eq:indx_flat}. Note that indices start at zero.
+Voxels can be indexed in one of two ways, eitehr be their component indices or its global index.
+Each voxel in the mesh has three components indices, one in each major direction, $ix$, $i_y$, and $i_z$ that rante from 0 to $N_x - 1$, $N_y - 1$, or $N_z-1$.
+However, this makes indexing cumbersome, so they are combined into a unique glorbal index, $i$ using Eq.~\ref{eq:n_v}.
+In general, this is known as "flattening" a matrix into a one-dimensional vector.
 
@@
 The spatial mesh is shown in Fig.~\ref{fig:spatial_disc}.
-The size of each voxel in each dimension is computed by dividing the size of the problem domain along in that dimension by the number of mesh elements along the same direction.
+The length of a voxel along a major direction is computed by dividing the size of the problem domain along that dimension by the number of mesh elements.
 Eq.~\ref{eq:mesh_x} shows this computation for the $x$ direction.
@@
 
-The index of interest is denoted as a subscript $i$ which makes the fully discretized form of the steady state Linear Boltzmann Equation given in Eq.~\ref{eq:boltz} into the fully discretized form given in Eq.~\ref{eq:boltz_i}
+The fully discretized form of the steady state LBE given in Eq.~\ref{eq:boltz} is
 
@@
 
-To the first order, a partial derivitive can be computed using Eq.~\ref{eq:deriv_1}.
+To a first order approximation, a partial derivitive can be computed using Eq.~\ref{eq:deriv_1}.
 
 \begin{equation} \label{eq:deriv_1}
-\frac{\partial f(x)}{\partial x} \approx \frac{f(x) - f(x + \Delta x)}{\Delta x}
+\frac{\partial f(x)}{\partial x} \approx \frac{f(x+\Delta x/2) - f(x + \Delta x/2)}{\Delta x}
 \end{equation}
 
-Using the definition of $\Omega$ and Eq.~\ref{eq:grad} the $\hat{\Omega} \cdot \nabla \psi$ term can be rewritten using Eq.~\ref{eq:spatial_1}
+Using the definition of $\hat{\Omega}$ and Eq.~\ref{eq:grad} the $\hat{\Omega} \cdot \nabla \psi$ term can be rewritten as
 
@@
 \begin{split}
-\Omega \cdot \nabla \psi & \approx 
-\left\langle \mu, \eta, \xi \right\rangle \cdot
-\left\langle \frac{\psi(x) - \psi(x + \Delta x)}{\Delta x},
-\frac{\psi(y + \Delta y) - \psi(y)}{\Delta y},
-\frac{\psi(z + \Delta z) - \psi(z)}{\Delta z} \right\rangle \\
+\Omega \cdot \nabla \psi 
+& = <\mu, \eta, \xi> \cdot <\frac{\partial \psi}{\partial x}, \frac{\partial \psi}{\partial y}, \frac{\partial \psi}{\partial z}> \\
+& =
+\frac{\partial \psi}{\partial x}\mu + \frac{\partial \psi}{\partial y}\eta + \frac{\partial \psi}{\partial z}\xi \\
 & \approx 
-\frac{\psi(x + \Delta x) - \psi(x)}{\Delta x} \mu + 
-\frac{\psi(y + \Delta y) - \psi(y)}{\Delta y} \eta + 
-\frac{\psi(z + \Delta z) - \psi(z)}{\Delta z} \xi
+\frac{\psi(x + \Delta x/2, y, z) - \psi(x - \Delta x/2, y, z)}{\Delta x} \mu \\
+&+ 
+\frac{\psi(x, y + \Delta y/2, z) - \psi(x, y - \Delta y/2, z)}{\Delta y} \eta \\
+&+ 
+\frac{\psi(x, y, z + \Delta z/2) - \psi(x, y, z - \Delta z/2)}{\Delta z} \xi
 \end{split}
@@
 
-Along some direction, $\Omega$, as it enters an arbitrary voxel, it must pass through either of the sides indicated in red and exit via the other as shown in Fig.~\ref{fig:gradient}.
-Fig.~\ref{fig:gradient} uses the $yz$ faces of the voxel for illustration.
-If $\eta$ is positive, $\Omega$ will enter through the $x_i$ surface and exit through the $x_{i+1}$ surface.
-If $\eta$ is negative, this order will be reversed.
-A limitation to this methodology is that $\mu$, $\eta$, and $\xi$ are assumed to be nonzero.
-This is ensured as long as none of the directions in the quadrature are parallel to a major axis.
+As a particle enters the voxel shown in Fig.~\ref{fig:gradient} along some direction, $\Omega$, as indicated in Fig.~\ref{fig:gradient}, SOMETHING.
+
+The surface flux. WRT the direction omega. This allows generalization.
 
@@
 \end{figure}%
-
-Depending on the sign of $\eta$, the first order approximation of the a derivitive along the $y$ direction is given by Eq.~\ref{eq:grad_y}.
-
-\begin{equation} \label{eq:grad_y}
-\begin{split}
-\frac{\partial f}{\partial y} &= \frac{f(x_i) - f(x_{i+1})}{\Delta x}, \quad \eta>0 \\
-\frac{\partial f}{\partial y} &= \frac{f(x_{i+1}) - f(x_{i})}{\Delta x}, \quad \eta <0 \\
-\frac{\partial f}{\partial y} &= 0, \quad \eta=0
-\end{split}
-\end{equation}
-
-Equation~\ref{eq:grad_y} can be generalized to use $f_{in}$ and $f_{out}$ instead of relying on the $x$-coordinate. This is given in Eq.~\ref{eq:grad_y2}
-
-\begin{equation} \label{eq:grad_y2}
-\begin{split}
-\frac{\partial f}{\partial y} &= \frac{f_{in} - f_{out}}{\Delta x}, \quad \eta \neq 0 \\
-\frac{\partial f}{\partial y} &= 0, \quad \eta=0
-\end{split}
-\end{equation}
 
@@
 \begin{equation} \label{eq:boltz_i2}
-\left[ 
+\begin{split}
+&\left[ 
 \frac{\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
@@
 \right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
++ \Sigma_{t,i}^g \psi_{i,a}^{g} \\
+& = 
 \sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
+\end{split}
 \end{equation}
 
-It is useful to point out that each of $\psi_{i,a,x,in}^g$, $\psi_{i,a,x,out}^g$, $\psi_{i,a,y,in}^g$, $\psi_{i,a,y,out}^g$, $\psi_{i,a,z,in}^g$, and $\psi_{i,a,z,out}^g$ are \textit{surface} flux values while $\psi_{i,a}^{g}$ is a \textit{volume averaged} flux value.
+It is useful to point out that each of $\psi_{i,a,x,in}^g$, $\psi_{i,a,x,out}^g$, $\psi_{i,a,y,in}^g$, $\psi_{i,a,y,out}^g$, $\psi_{i,a,z,in}^g$, and $\psi_{i,a,z,out}^g$ are \textit{surface} averaged flux values while $\psi_{i,a}^{g}$ is a \textit{volume} averaged flux value.
 
-In order for the discretized Linear Boltzmann Equation to be solved using Eq.~\ref{eq:boltz_i2}, the incoming flux must always be known.
-Either the boundary conditions or previously calculated values are used to determine each incoming flux.
+In order for the discretized LBE to be solved using Eq.~\ref{eq:boltz_i2}, the incoming flux must always be known.
+Either the boundary conditions or previously calculated values determine the incoming flux.
 Therefore, in order to compute the average flux in a cell, some relationship between the average flux and the outgoing surface fluxes must be known.
@@
 
-In the diamond difference approximation, it is assumed that the incoming and outgoing flux are both known.
+The diamond difference approximation, relates the incoming and outgoing surface fluxs to the cell averaged flux.
 The average flux is assumed to be the average of the two surface fluxes.
@@
 This allows Eq.~\ref{eq:dd2} to be used to make a substitution in Eq.~\ref{eq:boltz_i2} that removes the dependence on the outgoing flux from discretized equation.
-After the substitution, Eq.~\ref{eq:boltz_i3} will arise.
+After the substitution:
 
 \begin{equation} \label{eq:boltz_i3}
-\left[ 
+\begin{split}
+&\left[ 
 \frac{2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
@@
 \right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
++ \Sigma_{t,i}^g \psi_{i,a}^{g} \\
+& = 
 \sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
+\end{split}
 \end{equation}
@@
 \begin{equation} \label{eq:boltz_i4}
-\left[ 
+\begin{split}
+&\left[ 
 \frac{2\psi_{i,a}^{g}}{\Delta x} \mu + 
@@
 \right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} = 
++ \Sigma_{t,i}^g \psi_{i,a}^{g} \\
+& = 
 \sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
+\end{split}
 \end{equation}
 
-Equation~\ref{eq:boltz_i4} can be directly solved for the volume averaged flux. The result is given in Eq.~\ref{eq:boltz_i5}.
+Equation~\ref{eq:boltz_i4} can be directly solved for the volume averaged flux. The result is
 
@@
 
+Repeatedly iteration Eq.~\ref{eq:boltz_i5} will eventually yield the solution to the flux distribution regardless of the intial guess.
+Choosing an accurate guess can greatly accelerate the code and guide convergence on a more accurate solution as discussed in more detail in Section~\ref{sec:uncol}.
+Either way, some metric must be used to determine when to stop the iteration process.
+
 \subsection{Convergence Criteria}
-The typical converence criteria used to determine whether or to terminate further iteration is the maximum relative error from the $i-1^{th}$ iteration to the $i^{th}$ iteration given in Eq.~\ref{eq:conv}.
+The typical converence criteria used to determine whether or to terminate further iteration is the maximum relative error from the $i-1^{th}$ iteration to the $i^{th}$ iteration given in Eq.~\ref{eq:conv} over the phase space $\mathbb{R}$.
 Once the relative error in the parameter of interest, $\phi$ is below some threshold, $\epsilon$, the iteration stops and the solution is considered converged.
@@
 \begin{equation}\label{eq:conv}
-\frac{\phi_i - \phi_{i-1}}{\phi_i} < \epsilon
+\max_{\mathbb{R}} \left\{ \frac{\phi_i - \phi_{i-1}}{\phi_i} \right\} < \epsilon
 \end{equation}
@@
 The flux magnitude fluctuates wildly in regions of very flow flux, such as the gantry.
-In these regions, the relative uncertainty will remain much higher than in regions with smoother flux distributions.
+In these regions, the relative uncertainty remains much higher than in regions with smoother flux distributions.
+
 FIGURE XYZ SHOWS.
@@
 
-Scatter is assumed to be a function of only the scatter angle between the initial and scattered directions, $\theta_s$. Often, the cosine of the scatter angle, $\mu_s$ is used.
+Scatter is assumed to be a function of only the scatter angle $\theta_s$ between the initial, $\hat{\Omega}_a$, and scattered, $\hat{\Omega}_{a'}$, directions.
+The cosine of the scatter angle, $\mu_s$ is related by Eq.~\ref{eq:scat_cos}.
+
+\begin{equation} \label{eq:scat_cos}
+\mu_s = \Omega_a \cdot \Omega_{a'} = \cos(\theta_s) \,, \quad 0 \leq \theta_s \leq \pi
+\end{equation}
 
@@
 The use of a Legendre expansion removes the dependence on the quadrature from the data.
-This allows any quadrature to work with any dataset.
-A $l$-order Legendre polynomial is donoted by $P_l$.
-The anisotropic scatter cross section can then be rewritten as a Legendre expansion given in Eq.~\ref{eq:leg_1} where the scatter cosine is defined in Eq.~\ref{eq:scat_cos}
+This allows any quadrature to work with any cross section dataset.
+A $l$-order Legendre polynomial is denoted by $P_l$.
+Details of the Legendre polynomials can be found in Appendix~\ref{appdx:leg}.
+The anisotropic scatter cross section ise rewritten as a Legendre expansion given where the scatter cosine is defined in Eq.~\ref{eq:scat_cos} as
 
@@
 
-\begin{equation} \label{eq:scat_cos}
-\mu_s = \Omega_a \cdot \Omega_{a'}
-\end{equation}
-
-\section{Uncollided Solution Methods}
+\section{Uncollided Solution Methods}\label{sec:uncol}
 To avoid ray effect, the flux is split into collided and uncollided components as shown in Eq.~\ref{eq:totflux}.

### 2017-06-01 docprep/norris_phd_dissertation.tex
Added the rotating package for sideways tables
Added a not about future plans
Worked on the Vita

@@
 \usepackage{xr}
+\usepackage{rotating}
 \makenomenclature
@@
 Edward Norris graduated from Missouri University of Science and Technology in 2013 with a B.S.
-in both nuclear engineering and computer science, he went on to pursue a M.S.
+in both nuclear engineering and computer science, he is currently pursuing both a M.S.
 in computer science and a PhD in nuclear engineering.
+His PhD work is currently funded by a fellowship through the NRC.
 
-His PhD work was funded by a fellowship through the NRC.
+Previously he has participated in four summer internships at Sandia National Laboratories.
+While there, he served in various capacities ranging from cyber defense to radiation detection and neutron scatter simulation.
+He was largely involved with numerical simulations of radiation transport using Geant4 and MCNP6 as well as data collection.
 
-Previously he has participated in four summer internships with Sandia National Laboratories.
-While there, he served in various capacities ranging from cyber defense to radiation detection and neutron scatter simulation.
-
-Edward was awarded SOME AWARDS.
+Edward was awarded the College of Engineering and Computing PhD Scholar award in 2017 for his academic achievements.
 
@@
 
+After completing his PhD, he plans to take a one year postdoctorial fellowship position with the NNSA.
+
 \end{Vita}

### 2017-06-01 docprep/chapbib.tex
The table is now sideways
Added some MC notes

@@
 
+Dose estimations have been made using many methodologies including
+
 COMPUTATION
@@
 
-\begin{table}[ht]
+\begin{sidewaystable}[ht]
 \caption{Water Regions}
@@
 \hline \hline   
-Media       & CT Range      & H    & C    & N    & 0    & Na  & Mg  & P    & S   & Cl  & Ar  & K   & Ca \\ [0.5ex] 
+Media       & CT Range      & H    & C    & N    & O    & Na  & Mg  & P    & S   & Cl  & Ar  & K   & Ca \\ [0.5ex] 
 \hline
@@
 \label{table:ctmap}
-\end{table}
+\end{sidewaystable}
 
@@
 
+Monte Carlo codes such as TOPAS and Geant4 \citep{ref:agostinellis} are the gold standard \citep{ref:jiax}
+
+Codes such as EGS4, its improved versions EGSnrc, and EGS5 \citep{ref:nelsonw}, and PENELOPE \citep{ref:salvatf} are all Monte Carlo and have shown very good results to experimental.
+But they are developed for high energy physics.
+
 The Exnihilo thing by \citet{ref:evanst}. And DeHart wrote in \citep{ref:dehartm}
@@
 With regards to angles, stuff has been done \citep{ref:ahrensc}
+
+There are many acceleration techniques available for discrete ordinates \citep{ref:efremenkod} that can increase speed by 15-30\%.
+
+GPU methods exist for solar stuff at satellites.
+
+The geometry has been extended to Non-Uniform Rational B-Splins (NURBS) for reactor eigenvalue problems.
 
@@
 \section{GPU Acceleration}
+
+Monte Carlo dose computation for protons has been accelerated by GPU technology \citep{ref:jiax} and
+
+Monte Carlo in voxelized geometries has been done by \citep{ref:hissoinys} for monoenergetic beams.
 GPUs are cool.

### 2017-06-01 docprep/chapintro.tex
Expanded the intro

@@
 
-Computed tomography (CT) is becoming increasingly pervasive in medical diagnostics as improved algorithms give doctors access to more information.
+Computed tomography (CT) is becoming increasingly pervasive in medical diagnostics since its introduction in 1973 by \citet{ref:hounsfieldg} as improved algorithms give doctors access to more information.
 Increased information results in faster and more accurate diagnosis.
-However, the CT operation itself delivers a radiation dose to the patient increasing risk of long term complications such as cancer [CITE].
+However, the exponentially increasing usage of CT and corresponding radiation dose it delivers to the patient has raised concerns regarding potential long term radiation risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}.
+A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.
 Currently, dose estimation relies on \textit{a priori} computation verified with a standardized benchmark.
@@
 
+This work implements and analyses a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient can be computed with a full transport solution using a CT phantom.
+The methodology converts a CT mesh into a voxel phantom of materials and densities.
+Then a discrete ordinate method computes the flux throughout the phantom.
+Local energy deposition is assumed relieving the need for secondary electron transport.
+
 Another, similar, application is dose estimation of patients receiving radiation therapy.
-Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment.
+Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment [CITE].
 With further develpment, this methodology may enable real time dose computation as the treatment is administered.
@@
     \end{subfigure}
-    \caption{(a) The original }\label{fig:backprojection}
+    \caption{(a) The original phantom. (b) The sinogram produced by 360 projections. (c) The filter backprojection reconstruction.}\label{fig:backprojection}
 \end{figure}
@@
 
-Fundamentally, all reconstructions algorithms are based on a thin parallel beam of x-rays that roatate axially about a patient producing a sinogram.
-The inverse radon transform backprojects each row of the sinogram back to physical space..
-Filtering each line in Fourier space and then compositing all layers together recreates th eimage [CTIE].
+Fundamentally, all reconstructions algorithms are based on a thin parallel beam of x-rays rotating axially about a patient to produce a sinogram.
+The inverse radon transform backprojects each row of the sinogram back to physical space.
+Filtering each row in Fourier space and then compositing all layers together recreates the image [CTIE].
 This methodology has been extended to fan beams [CITE] and cone beams [CITE].
@@
 This phenomena leads to an image of the attenutating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
-With sufficient rays from varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation.
+With sufficient rays of varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation coefficient.
+The Hounsfield units (also called CT numbers) are traditionally scaled such that -1000 is dry air and water is zero [CITE].
 
@@
 
-Once a patient specific aphantom is obtained, the flux throughout the mesh must be computed. Computational techniques to do this cam be split into two broad categories: Monte Carlo and deterministic.
+Once a patient specific phantom is obtained, the flux throughout the mesh must be computed. Computational techniques to do this can be split into two broad categories: Monte Carlo and deterministic.
 
@@
 They also take advantage of combinatorial geometry allowing a user to produce geometrically complex 3D structures.
-Deterministic methods are generally much faster than their Monte Carlo counterparts bu tdo not benefit from excess runtime.
+Deterministic methods are generally much faster than their Monte Carlo counterparts but do not benefit from excess runtime.
 They also consume vastly more resources, particularly random-access memory (RAM).
-In regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
-A large number of spatial regions slows down th eparticle transport.
-Monte Carlo will also require a large amoutn of RAM to store the numerous voxels.
-
-The Hounsfield units (also called CT numbers) are traditionally scaled such that -1000 is dry air and water is zero.
+However, in regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
+A large number of spatial regions slows down the particle transport in Monte Carlo and increases the RAM required to store the numerous voxels.
 
@@
 If some pixelwise operation is applied on a CPU, 786,432 operations must occur.
-Onboard the GPU with 1024 cores, every pixel in the entire row can be computed simultaneously reducing the number of operations to 768.
-However, issues such as copying the memory to and from the GPU and cache coherency problems can degrade performance.
+Onboard a GPU with 1024 cores, every pixel in the entire row can be computed simultaneously reducing the number of operations to 768.
+However, issues such as communication to and from the GPU and cache coherency problems can degrade performance.
 GPU tehcnology and performance is discussed more throuroughly in Section~\ref{sec:cuda}.
 
-In addition to rapidly computing th epatient dose, another goal of DOCTORS is to present th ecode in a user-friendly fashion.
+In addition to rapidly computing the patient dose, another goal of DOCTORS is to present the code in a user-friendly fashion.
 To this end, a grpahical user interface (GUI) was developed.
 The GUI utilize the Qt5 graphics framework for the window, buttons, and other necessary widgets.
-The GUI leads the user through th esteps necessary to use th ecode in an intuitive way by using colors to indicated requaired and completed steps.
-The output is then plotted graphically and the output is also sent to an ASCII text file for more advanced postprocessing.
-More details about th eGUI are included in Section~\ref{sec:gui}.
+The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicated required and completed steps.
+The output is then plotted graphically as well as sent to an ASCII text file for more advanced postprocessing.
+More details about the GUI are included in Section~\ref{sec:gui}.
 
-There are two primary methodologies available to provide the transport solution: Monte Carlo and deterministic methods.
-Monte Carlo has proven to be very slow in these areas.
-Further, the 3D Cartesian mesh generated from CT reconstruction is well suited to deterministic computation.
-
-This work forcuses on the discrete ordinates method and presents an implementation specially developed for dose estimation from a medical CT scan.
-
-The DOCTORS code takes a computational reconstruction mesh of CT numbers, cross section data, and other parameters related to the solution methodology as input.
-The output is the collided and uncollided flux as well as the dose in each computational cell.
-These values can be compared to the empirical correlation results as a verification that the patient received an appropriate dose.
+BRIEF DISCUSSION OF RESULTS
 

### 2017-06-01 docprep/chapmeth.tex
Added some figures and expanded on the raytracing part

@@
 The techniques used to compute the collided and uncollided fluxes are described in seperate sections.
-After the flux solution methodology is explained, the methods used for source generation and flux to dose conversion are covered.
+After the flux solution methodology is explained, the methods used for source generation and flux-to-dose conversion are covered.
 
@@
 In general, the LBE is valid for multiplying media (such as neutrons passing through fuel in a reactor), but the variant considered in this work omits such terms since only photons in medical systems are of concern.
-The time dependent linear Boltzmann equation for photons is
+Solutions to the LBE can also be used to solve criticality eigenvalue problems or to produce adjoint parameters for other codes.
+Such solutions are not considered in this work.
 
@@
 \section{Uncollided Solution Methods}\label{sec:uncol}
-To avoid ray effect, the flux is split into collided and uncollided components as shown in Eq.~\ref{eq:totflux}.
-Equation~\ref{eq:boltz} can be rewritten as two equations solving for the uncollided flux, $\psi_u$, and collided flux, $\psi_c$ independently.
+One of the earliest problems with the discrete orinates method is the ppresence of ray effect.
+Ray effect arises due to the restriction of particle transport to discrete set of directions.
+As particles stream out of a point source, into a vacuum, the stream out along the set of direction schosen only.
+Since no scatter takes place in vacuum conditions, the praticles have no means to escape their initial direction.
+This gives rise to the phenomena observed in Fig.~FIGURE.
+This phenomena has been thrououghly studied by many [CITE].
+
+The solution to prevent ray effect is to utilize an analytical uncollided source term.
+The full flux solution is split into two subproblems.
+Equation~\ref{eq:boltz} is rewritten as two equations solving for the uncollided flux, $\psi_u$, and collided flux, $\psi_c$ independently.
 Equation~\ref{eq:unc} has no scatter term since scattered particles are not considered in the uncollided flux.
-The source term, $S_u$, in Eq.~\ref{eq:col} is the first collision source.
+The lack of a scatter term makes the uncollided flux (whose analytical solution is intuitive) computable with a raytracing algorithm.
+
+The source term, $S_u$, in Eq.~\ref{eq:col} is the first collision source. The uncollided flux is driven by an external source. The flux distribution is solved analytically.
 
@@
 
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_ex1}
+        \caption{}
+        \label{fig:rayeffect_ex1}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_ex2}
+        \caption{}
+        \label{fig:rayeffect_ex2}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_ex3}
+        \caption{}
+        \label{fig:rayeffect_ex3}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_ex4}
+        \caption{}
+        \label{fig:rayeffect_ex4}
+    \end{subfigure}
+    \caption{The process}\label{fig:rayeffect_ex}
+\end{figure}
+
+This then results in the value shown in the next figure.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_iso4annot}
+        \caption{}
+        \label{fig:rayeffect_iso4annot}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_beam4annot}
+        \caption{}
+        \label{fig:rayeffect_beam4annot}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_iso12}
+        \caption{}
+        \label{fig:rayeffect_iso12}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_beam12}
+        \caption{}
+        \label{fig:rayeffect_beam12}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_iso500}
+        \caption{}
+        \label{fig:rayeffect_iso500}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/rayeffect_beam500}
+        \caption{}
+        \label{fig:rayeffect_beam500}
+    \end{subfigure}
+    \caption{The ray effect artifact.
+The subfigures on the left (a, c, and e) show the correct solution modeled with $1/r^2$.
+The right subfigures (b, d, and f) show the corresponding solution with ray effect generated by the process described in Fig.~\ref{fig:rayeffect_ex}.
+The top row of subfigures uses 4 cells, the middle row uses 12, and the bottom row uses 500.}\label{fig:rayeffect_comp}
+\end{figure}
+
+Consider an isotropic point source in a homogeneous media. The (scalar) flux well known to be
+
+\begin{equation}
+\varphi(r, E) = \frac{S_0(E) e^{-\mu(E) r}}{4 \pi r^2}
+\end{equation}
+
+\noindent where $S_0$ is the source strength, $\mu$ is the attenuation coefficient in the material, and $r$ is the distance traveled by the photon [CITE].
+Expanding the problem to a full 3D model, the flux at $\boldsymbol{r}$ from a point source at $\boldsymbol{s}$ is
+
+\begin{equation}\label{eq:isouncol}
+\varphi(\boldsymbol{r}, E) = \frac{S_0(E) e^{-\mu(E) |\boldsymbol{r}-\boldsymbol{s}|}}{4 \pi |\boldsymbol{r}-\boldsymbol{s}|^2}.
+\end{equation}
+
+\noindent
+The only tricky step is the introduction of anisotropy. The $4 \pi$ solid angle term is generally expressed by
+
+\begin{equation}
+\int_{4 \pi}^{} \int_{0}^{\infty} S_0(E, \hat{\Omega}) dE d\hat{\Omega}
+\end{equation}
+
+\noindent
+but generally the normalization remains $4 \pi$.
+The flux is nonzero \textit{only} in the direction of travel, namely in the unit direction $\vec{r} - \vec{s}$.
+Therefore, a Kroenecker delta function (defined in Eq.~\ref{eq:kronecker}) is used to produce
+
+\begin{equation}
+\psi(\boldsymbol{r}, E, \hat{\Omega}) = \frac{S_0(E, \hat{\Omega}) e^{-\mu(E) |\boldsymbol{r}-\boldsymbol{s}|}}{4 \pi |\boldsymbol{r}-\boldsymbol{s}|^2} \delta \left( \hat{\Omega}, \frac{\boldsymbol{r} - \boldsymbol{s}}{|\boldsymbol{r} - \boldsymbol{s}|^2}\right)
+\end{equation}
+
 The uncollided flux in Eq.~\ref{eq:unc} can be analytically solved.
@@
 \begin{equation} \label{eq:kronecker}
-\delta(\vec{A}, \vec{B}) = 
+\delta(\boldsymbol{A}, \boldsymbol{B}) = 
 \begin{cases}
-1, \quad \mathrm{if} \ \vec{A}=\vec{B} \\
-0, \quad \mathrm{otherwise}
+1 \,, \quad \mathrm{if} \ \boldsymbol{A}=\boldsymbol{B} \\
+0 \,, \quad \mathrm{otherwise}
 \end{cases}
@@
 	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
-	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + \\
+	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) \\
+	&=
+	\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + \\
 	&\int_{4\pi} \int_{0}^{\infty} 
 \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_u(\boldsymbol{r}, E', \hat{\Omega}') 
-dE' d\hat{\Omega}'
+dE' d\hat{\Omega}' \\
+& = \int_{4 \pi} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'
 \end{split}
@@
 \section{Source Generation}
-There are many kinds of external sources implemented in this work.
+All photon tranpot problems considered in the DOCTORS code ar eultimately driven by an external source, namely th emedical equipment producing the photon beam.
+These sources can com ein a variety of geometric shapes and energy spectra.
+Currently, CODTORS has  anumber of built in beam typees available for anlaysis including point sources, fan beams, and cone beams.
+Fan beams and cone beams cna be arranged about a phantom to produce an axial slice.
+Currently, helical paths are not supported, though they could be added with minimal excess effort.
 
@@
 This source is defined only by its position and energy distribution.
-The center of the cell whose flux is being computed is located at position $<C_x, C_y, C_z>$ and the point source is located at position $<S_x, S_y, S_z>$.
+Given a source a $s$, the uncollided flux at the center of a cell, $<c>$ is computing using Eq.~\ref{eq:isouncol}.
+The collided flux is then computed using the discrete ordinate solution to the LBE using Eq.~\ref{eq:xxx}.
 
@@
 The raytracing algorithm used for the fan beam is identical to the point source raytracer except for one difference: before the raytracing is done, whether the ray falls within the fan beam is determined and any ray falling outside the fan beam is not traced, its uncollided flux is immediately set to zero.
+Some voxels are partially enclosed inside the beam.
 
-Some voxels would be partially enclosed inside the beam. A partial acceptance value is computed for these voxels and multiplied by the final result of the raytrace.
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{.45 \textwidth}
+        \includegraphics[width=\textwidth]{figs/shape_fanbeam}
+        \caption{}
+        \label{fig:shape_fanbeam}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{.45 \textwidth}
+        \includegraphics[width=\textwidth]{figs/shape_conebeam}
+        \caption{}
+        \label{fig:shape_conebeam}
+    \end{subfigure}
+    \caption{(a) The fanbeam is defined by $\theta$ in the polar coordinate and $\varphi$ azimuthally. (b) The cone beam is described only be $\theta$.}\label{fig:shape}
+\end{figure}
 
@@
 
-First, the ray direction is projected onto the $xy$ plane giving $<S_x, S_y, 0>$ and $C_x, C_y, 0$.
-Then the vector from the source to the geometry center and the vector from the source to the cell center are dotted to give the cosine of the angle between them as shown in Eq.~\ref{eq:phicos}.
+First, the ray direction is projected onto the $xy$ plane giving $<S_x, S_y, 0>$ and $<C_x, C_y, 0>$.
+Then the cosine of the angle between them computed using Eq.~\ref{eq:phicos}.
 Once $\zeta$ is computed, the condition given by Eq.~\ref{eq:phicoscon} is evaluated.
@@
 Treatment for the $\theta$ variable is similar to that of $\varphi$ except the polar angle is required instead of the azimuthal angle.
-This is computed using Eq.~\ref{eq:thetacos} for an arbitrary vector $\vec{\xi}$.
+The polar angle of an arbitrary vector $\vec{\xi}$ can be computed using Eq.~\ref{eq:thetacos}.
 This is used to compute the polar angle for both $\vec{S}$ and $\vec{C}$.
@@
 \begin{equation}\label{eq:thetacoscon}
-|\zeta_\theta| < \frac{\vartheta}{2}
+|\zeta_\theta| < \frac{\theta}{2}
 \end{equation}
@@
         \caption{Cone yz}
-        \label{fig:subsweep_general3}
+        \label{fig:beamconeyz}
     \end{subfigure}
@@
         \caption{Fan yz}
-        \label{fig:subsweep_general4}
+        \label{fig:beamfanyz}
     \end{subfigure}
@@
         \caption{Cone xz}
-        \label{fig:subsweep_general5}
+        \label{fig:beamconexz}
     \end{subfigure}
@@
         \caption{Fan xz}
-        \label{fig:subsweep_general6}
+        \label{fig:beamfanxz}
     \end{subfigure}
@@
 The second way to compute dose is to compute the energy deposition in each cell.
+
+Transport of secondary particles is neglected and all energy deposition is assumed to be local.
 
@@
 
-blah
+The alternative way to compute the dose is directly from the local energy deposition, $\mathcal{E}$, is by
+
+\begin{equation}
+\mathcal{E}_i = \sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g V_i \bar{E}^g F
+\end{equation}
+
+\noindent
+where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F$ is a unit conversion factor that converts MeV to Joules. The mass of the voxel, $M$ is then computed
+
+\begin{equation}
+M_i = \rho_i V_i G
+\end{equation}
+
+\noindent
+where $\rho_i$ is the density in g/cm$^3$ and G is a conversion factor to convert grams to kilograms. The dose in grays to the voxel is then
+
+\begin{equation}
+D_i = \frac{\mathcal{E}_i}{M_i} = \frac{\sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g \bar{E}^g F}{\rho_i G}
+\end{equation}
 

### 2017-06-01 docprep/chapintro.tex
Reworded sections of the methods and fixed typos in the intro

@@
 Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment [CITE].
-With further develpment, this methodology may enable real time dose computation as the treatment is administered.
+With further development, this methodology may enable real time dose computation as the treatment is administered.
 This would be greatly beneficial to both patients and the doctors administering the procedure.
@@
 
-Along any particular ray, monoenergetic particles will attenuate through a homogeneus media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert} [CITE] resulting in a small fraction of the source particles reaching the detector.
-This phenomena leads to an image of the attenutating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
+Along any particular ray, monoenergetic particles will attenuate through a homogenous media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert} [CITE] resulting in a small fraction of the source particles reaching the detector.
+This phenomena leads to an image of the attenuating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
 With sufficient rays of varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation coefficient.
@@
 
-Deterministic codes can be accelerated with the use of a grpahics processin gunit (GPU).
-GPUs differ from the central processin gunit (CPU) in that  each of th emany cores performes identical instruction to all others at the same time bue each with different data.
-For example, consider a grapyscale 1024$\times$768 image.
+Deterministic codes can be accelerated with the use of a graphics processing unit (GPU).
+GPUs differ from the central processing unit (CPU) in that each of the many cores performs identical instruction to all others at the same time but each with different data.
+For example, consider a grayscale 1024$\times$768 image.
 If some pixelwise operation is applied on a CPU, 786,432 operations must occur.
@@
 However, issues such as communication to and from the GPU and cache coherency problems can degrade performance.
-GPU tehcnology and performance is discussed more throuroughly in Section~\ref{sec:cuda}.
+GPU technology and performance is discussed more thoroughly in Section~\ref{sec:cuda}.
 
 In addition to rapidly computing the patient dose, another goal of DOCTORS is to present the code in a user-friendly fashion.
-To this end, a grpahical user interface (GUI) was developed.
+To this end, a graphical user interface (GUI) was developed.
 The GUI utilize the Qt5 graphics framework for the window, buttons, and other necessary widgets.

### 2017-06-01 docprep/chapmeth.tex
Reworded sections of the methods and fixed typos in the intro

@@
 This chapter summarizes the solution methodology employed by DOCTORS.
-The techniques used to compute the collided and uncollided fluxes are described in seperate sections.
+The techniques used to compute the collided and uncollided fluxes are described in separate sections.
 After the flux solution methodology is explained, the methods used for source generation and flux-to-dose conversion are covered.
@@
 
-The linear Boltzmann equation (LBE) is generally true of any particle given that outside forces (electromagnetic or graitational) are either not present or negligible and particle-particle collisions are insignificant.
-Therefore, the LBE is generally not valid for charged parpticles or systems whose particle density is on the order of the medium they are traveling trhrough.
-In such systems, particle-particle ineractions may not be negligible.
+The linear Boltzmann equation (LBE) is generally true of any particle given that outside forces (electromagnetic or gravitational) are either not present or negligible and particle-particle collisions are insignificant.
+Therefore, the LBE is generally not valid for charged particles or systems whose particle density is on the order of the medium they are traveling through.
+In such systems, particle-particle interactions may not be negligible.
 
@@
 
-\begin{equation} \label{eq:boltz_t_dep}
-\begin{split}
-	&\left[ \frac{1}{v(\boldsymbol{r},t)} \frac{\partial}{\partial t} + \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E, t) \right]
-	\psi(\boldsymbol{r}, E, \hat{\Omega}, t) = \\
-	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}, t) \psi(\boldsymbol{r}, E', \hat{\Omega}', t) dE' d\hat{\Omega}' + S(\boldsymbol{r}, E, \hat{\Omega}, t)
-\end{split}
+The steady state form of the LBE can be readily derived from simple intuition.
+In the steady state, particle production and removal must be equal.
+In any volume, three production mechanisms are present.
+Particles can (1) stream into the volume across a bounding surface, (2) inscatter from another energy-direction component of phase space, or (3) be produced directly by the external source.
+Three removal terms are also present, particles can (1) stream out of the volume of interest by crossing a bounding surface, (2) be absorbed and vanish entirely, or (3) outscatter to a energy-direction component of phase space other than the one of interest.
+
+The downscatter and absorption removal terms are typically grouped together using
+\begin{equation}
+\Sigma_t = \Sigma_a + \Sigma_s
 \end{equation}
+where $\Sigma_t$, $\Sigma_a$, and $\Sigma_s$ are the macroscopic total, absorption, and scatter cross sections respectively.
+The removal of particles from interactions (scatter and absorption) per volume is then
+\begin{equation}
+\Sigma_t(\boldsymbol{r}, E) \psi(\boldsymbol{r}, E, \hat{\Omega})
+\end{equation}
+where $\psi$ is the angular flux. The production and removal streaming operators are combined as 
+\begin{equation}
+\hat{\Omega} \cdot \nabla \psi(\boldsymbol{r}, E, \hat{\Omega})
+\end{equation}
+and included as a removal term since the normal is defined pointing outward from the volume.
+The normal sign convention results in particles streaming out being positive and those streaming in becoming negative.
 
-\noindent
-where $v$ is the velocity of the photon in the medium at $\boldsymbol{r}$, $\Sigma_t$ is the total macroscopic cross section, $\Sigma_s$ is the macroscopic scattering cross section, $S$ is an external photon source, and $\psi$ is the total photon angular flux.
-
-The transient case is not typically of interest in medical diagnostic imaging, rather the steady state is more important. The time dependence in Eq. \ref{eq:boltz_t_dep} is removed to yield
-
+The inscatter to a particular energy and direction is computed by integrating over all other energies and directions
+\begin{equation}
+\int_{4\pi}^{} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'.
+\end{equation}
+An external source $S$ produces particles per volume. Combining the removal terms on the left and the production terms on the right yields
 \begin{equation} \label{eq:boltz}
@@
 \end{equation}
-
-The steady state form of the LBE can be readily derived from simple intuition.
-In the steady state, particle production and removal must be equal.
-In any volume, three production mechanisms are present.
-Particles can stream into the volume across a bounding surface, they can inscatter from another energy-direction component of phase space, or they can be produced directly by the external source.
-Three removal terms are also present, particles can stream out of the volume of interest by crossing a bounding surface, be absorbed and vanish entirely, or outscatter to a energy-direction componennt of phase space other than the one of interest.
-
-The downscatter and absorbtiption removal terms are typically grouped together as
-
-\begin{equation}
-\Sigma_t(\boldsymbol{r}, E) \psi(\boldsymbol{r}, E, \hat{\Omega})
-\end{equation}
-
-\noindent
-and both production and removal streaming operators are combined as well as
-
-\begin{equation}
-\hat{\Omega} \cdot \nabla \psi(\boldsymbol{r}, E, \hat{\Omega})
-\end{equation}
-
-\noindent
-and included as a removal term since they normal is defined pointing outward from the volume.
-The normal sign convention results in particles streaming out being positive and those streaming in beocming negative.
-The inscatter to a particular energy and direction is computed by integrating over all other energies and diretions
-
-\begin{equation}
-\int_{4\pi}^{} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'
-\end{equation}
-
-The following sections show how Eq.~\ref{eq:boltz} is discretized and solved computationally.
+which is the LBE. The following sections show how Eq.~\ref{eq:boltz} is discretized and solved computationally.
 
@@
 A discrete set of $N_a$ angles ($\Omega_{a}, \quad a = 0 \ldots N_a-1$) is selected to represent continuous directional space.
-Particles are transported only along these discrete directions from a voxel to adjacent voxels.
+Particles are transported only along these discrete directions from each voxel to adjacent voxels.
 
@@
 
-THEORY
+THEORY AND MATH?
 
-To inlcude a user defined quadrature, the plain text file should be formatted such that it contains four columns similar to Table~\ref{tab:quad_format} (without any headings).
-No checks are performed to ensure that octants are covered, this allows simple testing quadratures that utilize only a single direction to be used for testing purposes.
-Values for $\mu$, $\eta$, and $\xi$ do not necessarily need to add to unity, they will be automatically normalized by DOCTORS.
+To include a user defined quadrature, the plain text file should be formatted such that it contains four columns similar to Table~\ref{tab:quad_format} (without any headings).
+No checks are performed to ensure that octants are covered, this allows simple testing quadratures that utilize only a single direction to be used.
+Values for $\mu$, $\eta$, and $\xi$ do not necessarily need to add to unity as they will be automatically normalized by DOCTORS.
 Likewise, the weights will be normalized such that their summation is unity.
@@
 Doing so may result in undefined behavior.
-Also, after normalizaiton, no two rows should have identical values for $\mu$, $\eta$, and $\xi$ such as the last two rows of Table~\ref{tab:quad_interp}.
-Additionally, no weights may be negative.
+Also, after normalization, no two rows should have identical values for $\mu$, $\eta$, and $\xi$ such as the last two rows of Table~\ref{tab:quad_interp}, nor may any weights be negative.
+The data in Table~\ref{tab:quad_format} would be read in and renormalized by DOCTORS.
+The normalized data would be identical to the data in Table~\ref{tab:quad_interp}.
 
@@
 Note that Table~\ref{tab:quad_interp} rounds four digits after the decimal.
-
-The data in Table~\ref{tab:quad_format} would be read in and renormalized by DOCTORS. The normalized data would be identical to the data in Table~\ref{tab:quad_interp}.
 
@@
 
-A given direction, $\hat{\Omega}$, is determined by its three cosine components.
-This relation is given in Eq.~\ref{eq:omega_cos} where $\hat{i}$, $\hat{j}$, and $\hat{k}$ are the unit directions in the $x$, $y$, and $z$ directions respectively.
-
+A given direction, $\hat{\Omega}$, is determined by its three cosine components: 
 \begin{equation} \label{eq:omega_cos}
@@
 \end{equation}
-
-Each discrete angle is associated a weight.
-The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by
-
-\begin{equation} \label{eq:disc_int}
-\int_{4 \pi} f(\hat{\Omega}) d\hat{\Omega} \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a
-\end{equation}
+where $\hat{i}$, $\hat{j}$, and $\hat{k}$ are the unit directions in the $x$, $y$, and $z$ directions respectively.
+Therefore, any discrete direction $\hat{\Omega}_a$ can be expressed as $<\mu, \eta, \xi>$.
+The two forms will be used interchangeably.
 
 Discretizing Eq.~\ref{eq:boltz} in angular space gives
-
 \begin{equation} \label{eq:boltz_a}
@@
 \end{equation}
-
-\noindent
-where the subscript $a$ denotes the direction.
+where the subscript $a$ denotes the direction number and $\omega_a$ is a weight associated with that direction.
+The weights are designed with the angles such that once a discrete quadrature is selected, integration over continuous space becomes integration in discrete space approximated by
+\begin{equation} \label{eq:disc_int}
+\int_{4 \pi} f(\hat{\Omega}) d\hat{\Omega} \approx \sum_{a=0}^{N_a-1} f(\hat{\Omega}_a) \omega_a = 1.
+\end{equation}
 
@@
 
-Continuous energy is discretized into $G$ groups indexed from 0 to $G-1$.
-Upscatter is assumed to be negligible.
-Therefore, Eq.~\ref{eq:boltz_a} becomes the energy discretized Eq.~\ref{eq:boltz_e} where the group number is indexed by superscript $g$.
-
+Continuous energy is discretized into $G$ groups indexed from 0 to $G-1$ as illustrated in Fig.~\ref{fig:energy_groups}. Therefore, Eq.~\ref{eq:boltz_a} becomes
 \begin{equation} \label{eq:boltz_e}
@@
 \end{equation}
-
-In photon based problems, upscatter is typicall negligible [CITE].
-By the time photons become low enough in energy to upscatter significantly, they are no longer of dosimetric concern.
-Removing upscatter reduces Eq.~\ref{eq:boltz_e} to
-
-\begin{equation} \label{eq:boltz_e2}
-\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t^g(\boldsymbol{r}) \right]
-\psi_{a}^{g}(\boldsymbol{r}) = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
-\end{equation}
-
-\noindent
-which is very similar; only the starting point for the scatter summation changes.
-Computationall, though, this makes a significant different.
-This means that the highest energy group can be solved independently of all those below it.
-Once that group is solved, the next group can also be solved and so forth.
+where the group number is indexed by superscript $g$.
 
@@
 
-The group structure cannot be arbitrarily selected by DOCOTRS, instead it is determined by the external cross section data file.
-The energy group structure should be selected appropriately for the problem.
-The cross sections currently available are those available from the SCALE 6.2 distribution.
-Those cross sections are designed for light water reactor analysis and are not well suited to medical physics applications.
-This is currently one of the greatest limitaitons of DOCOTRS.
-The SCALE cross section data also contains neutron data which is useless for the CT dosimetry DOCTORS performs.
-More details of the cross section parsing and generation are included in Section~\ref{sec:xsparse} and~\ref{sec:xsgen}.
-
 Once the energy group structure is selected, cross section data must be available.
@@
 The weighting function can be any of a number commonly used functions.
+The SCALE data libraries us a uniform function [CITE].
+\begin{equation}\label{eq:groupxs}
+\sigma_G = \frac{\int_{E_1}^{E_2}f(E')\sigma(E') dE'}{\int_{E_1}^{E_2} f(E') dE'}
+\end{equation}
 
-\begin{equation}\label{eq:groupxs}
-\sigma_G = \frac{\int_{E_1}^{E_2}f(E')\sigma(E') dE'}{\int_{E_1}^{E_2} dE'}
+In photon based problems, upscatter is typically negligible [CITE].
+By the time photons become low enough in energy to upscatter significantly, they are no longer of dosimetric concern.
+Removing upscatter reduces Eq.~\ref{eq:boltz_e} to
+\begin{equation} \label{eq:boltz_e2}
+\left[ \hat{\Omega}_a \cdot \nabla + \Sigma_t^g(\boldsymbol{r}) \right]
+\psi_{a}^{g}(\boldsymbol{r}) = 
+\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, a, a'}^{g, g'}(\boldsymbol{r}) \psi_{a'}^{g'}(\boldsymbol{r}) \omega_a + S_a^g(\boldsymbol{r})
 \end{equation}
+which is very similar; only the starting point for the scatter summation changes.
+Computationally, though, the lack of upscatter makes a significant difference.
+The highest energy group can be solved independently of all those below it.
+Once that group is solved, the next group can also be solved and so forth.
+In the presence of upscatter, the highest group depends on the solution to all groups below it so all energy groups must be solved simultaneously.
+
+The group structure cannot be arbitrarily selected by DOCOTRS, instead it is determined by the external cross section data file loaded by the user.
+The energy group structure should be selected appropriately for the problem.
+The cross sections currently included with DOCTORS are those available from the SCALE 6.2 distribution.
+Those cross sections are designed for light water reactor analysis and are not well suited to medical physics applications.
+This is currently one of the greatest limitations of DOCOTRS.
+The SCALE cross section data also contains neutron data which is useless for the CT dosimetry DOCTORS performs.
+More details of the cross section parsing and generation are included in Section~\ref{sec:xsparse} and~\ref{sec:xsgen} respectively.
 
@@
 The problem domain is split into an evenly spaced Cartesian grid.
-The grid spacing between each of the spatial dimensions does not necessarily need to be uniform, but it often is in the case of CT voxel phantoms.
+Uniform grid spacing is not required by discrete ordinate methods though it is helpful [CITE], but is necessarily the case with CT voxel phantoms.
 
-The problem domain must be a regular, rectangular parallelpiped of dimension $D_x \times D_y \times D_z$.
-The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each major direction.
-The total number of voxels, $N_V$ is easily computed as the multiple of each of the component directions given in Eq.~\ref{eq:n_v}.
-
+The problem domain for a CT voxel phantom is a regular, rectangular parallelepiped of dimension $D_x \times D_y \times D_z$, the values of which must be known based on the CT setup by the user.
+The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each major direction as shown in Fig.~\ref{fig:spatial_disc}.
+The total number of voxels, $N_V$ is easily computed from the number of bins in each major direction:
 \begin{equation} \label{eq:n_v}
-	N_V = N_x N_y N_z
+N_V = N_x N_y N_z.
 \end{equation}
-
-Voxels can be indexed in one of two ways, eitehr be their component indices or its global index.
-Each voxel in the mesh has three components indices, one in each major direction, $ix$, $i_y$, and $i_z$ that rante from 0 to $N_x - 1$, $N_y - 1$, or $N_z-1$.
-However, this makes indexing cumbersome, so they are combined into a unique glorbal index, $i$ using Eq.~\ref{eq:n_v}.
-In general, this is known as "flattening" a matrix into a one-dimensional vector.
-
-\begin{equation} \label{eq:indx_flat}
-	i = i_x (N_z + N_y) + i_y N_z + i_z
+The length of a voxel along the $x$-direction is computed by 
+\begin{equation} \label{eq:mesh_x}
+\Delta x = \frac{D_x}{N_x}.
 \end{equation}
-
-The spatial mesh is shown in Fig.~\ref{fig:spatial_disc}.
-The length of a voxel along a major direction is computed by dividing the size of the problem domain along that dimension by the number of mesh elements.
-Eq.~\ref{eq:mesh_x} shows this computation for the $x$ direction.
-Analagous equations apply in the $y$ and $z$ directions as well.
-
-\begin{equation} \label{eq:mesh_x}
-\Delta x = \frac{D_x}{N_x}
-\end{equation}
+Analogous equations apply in the $y$ and $z$ directions as well.
 
@@
 
-The fully discretized form of the steady state LBE given in Eq.~\ref{eq:boltz} is
+Voxels can be indexed in one of two ways, either by their component indices or global index.
+Each voxel in the mesh has three components indices, one in each major direction, $i_x$, $i_y$, and $i_z$ that range from 0 to $N_x - 1$, $N_y - 1$, or $N_z-1$.
+However, this makes indexing cumbersome, so they are combined into a unique global index, $i$ using Eq.~\ref{eq:indx_flat}.
+In general, this is known as "flattening" a matrix into a one-dimensional vector.
+More details about the flattening implementation used in DOCTORS is given in Section~\ref{sec:flatten}.
 
+\begin{equation} \label{eq:indx_flat}
+i = i_x (N_z + N_y) + i_y N_z + i_z
+\end{equation}
+
+Using global indexing discussed, the fully discretized the steady state LBE given is
 \begin{equation} \label{eq:boltz_i}
@@
 Recall that $\Omega_a$ can be written in vector notation using its cosine components as defined in Eq.~\ref{eq:omega_cos}.
-Also recall the definition of the gradient, given in Eq.~\ref{eq:grad}.
-
+Also recall the definition of the gradient:
 \begin{equation} \label{eq:grad}
-\nabla f(x, y, z) = \frac{\partial f(x, y, z)}{\partial x} \hat{i} + \frac{\partial f(x, y, z)}{\partial y} \hat{j} + \frac{\partial f(x, y, z)}{\partial z} \hat{k}
+\nabla f(x, y, z) = \frac{\partial f(x, y, z)}{\partial x} \hat{i} + \frac{\partial f(x, y, z)}{\partial y} \hat{j} + \frac{\partial f(x, y, z)}{\partial z} \hat{k}.
 \end{equation}
-
-To a first order approximation, a partial derivitive can be computed using Eq.~\ref{eq:deriv_1}.
-
+To a first order approximation, a partial derivative is computed as
 \begin{equation} \label{eq:deriv_1}
-\frac{\partial f(x)}{\partial x} \approx \frac{f(x+\Delta x/2) - f(x + \Delta x/2)}{\Delta x}
+\frac{\partial f(\zeta)}{\partial \zeta} \approx \frac{f(\zeta+\Delta \zeta/2) - f(\zeta + \Delta \zeta/2)}{\Delta \zeta}
 \end{equation}
-
-Using the definition of $\hat{\Omega}$ and Eq.~\ref{eq:grad} the $\hat{\Omega} \cdot \nabla \psi$ term can be rewritten as
-
+with respect to $\zeta$. Combining Eq.~\ref{eq:grad} and~\ref{eq:deriv_1}, the $\hat{\Omega} \cdot \nabla \psi$ term can be rewritten as:
 \begin{equation} \label{eq:spatial_1}
@@
 &+ 
-\frac{\psi(x, y, z + \Delta z/2) - \psi(x, y, z - \Delta z/2)}{\Delta z} \xi
+\frac{\psi(x, y, z + \Delta z/2) - \psi(x, y, z - \Delta z/2)}{\Delta z} \xi.
+\end{split}
+\end{equation}
+The flux variables $\psi_{i,a,x,in}^g$, $\psi_{i,a,x,out}^g$, $\psi_{i,a,y,in}^g$, $\psi_{i,a,y,out}^g$, $\psi_{i,a,z,in}^g$, and $\psi_{i,a,z,out}^g$ are \textit{surface} averaged flux values while $\psi_{i,a}^{g}$ is a \textit{volume} averaged flux value.
+
+The gradient is computed with respect to th edirection $\hat{\Omega}$, thus Eq.~\ref{eq:spatial_1} is valid only for the first octant ($\mu, \eta, xi > 0$).
+In other octants, a new gradient must be used.
+Rather than enumerating all eight variants, Eq.~\ref{eq:spatial_1} can be generalized to
+\begin{equation} \label{eq:deriv_2}
+\Omega \cdot \nabla \psi \approx 
+\frac{\psi_{x,out} - \psi_{x,in}}{\Delta x} \mu + 
+\frac{\psi_{y,out} - \psi_{y,in}}{\Delta y} \eta + 
+\frac{\psi_{z,out} - \psi_{z,in}}{\Delta z} \xi
+\end{equation}
+where
+\begin{equation}
+\psi_{x,out} = 
+\begin{cases}
+\psi(x+\Delta x/2, y, z) - \psi(x-\Delta x/2, y, z) \,, \quad \mu > 0 \\
+\psi(x-\Delta x/2, y, z) - \psi(x+\Delta x/2, y, z) \,, \quad \mu < 0
+\end{cases}
+\end{equation}
+and
+\begin{equation}
+\psi_{x,in} = 
+\begin{cases}
+\psi(x-\Delta x/2, y, z) - \psi(x+\Delta x/2, y, z) \,, \quad \mu > 0 \\
+\psi(x+\Delta x/2, y, z) - \psi(x-\Delta x/2, y, z) \,, \quad \mu < 0
+\end{cases}
+\end{equation}
+and $\psi_{y,out}$, $\psi_{y,in}$, $\psi_{z,out}$, and $\psi_{z,in}$ are similarly defined but with respect to $\eta$ and $\xi$ respectively.
+
+%\begin{figure}[tb]
+%  \begin{center}
+%   \includegraphics[width=3.75in]{figs/gradient}
+%  \end{center}
+%  \caption{The caption.}
+%\label{fig:gradient}
+%\end{figure}%
+
+Using the generalized approximation given in Eq.~\ref{eq:deriv_2}, Eq.~\ref{eq:boltz_i} can be rewritten as
+\begin{equation} \label{eq:boltz_i2}
+\begin{split}
+&\left[ 
+\frac{\psi_{i,a,x,out}^g - \psi_{i,a,x,in}^g}{\Delta x} \mu + 
+\frac{\psi_{i,a,y,out}^g - \psi_{i,a,y,in}^g}{\Delta y} \eta + 
+\frac{\psi_{i,a,z,out}^g - \psi_{i,a,z,in}^g}{\Delta z} \xi
+\right]
++ \Sigma_{t,i}^g \psi_{i,a}^{g} \\
+& = 
+\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g.
 \end{split}
@@
 
-As a particle enters the voxel shown in Fig.~\ref{fig:gradient} along some direction, $\Omega$, as indicated in Fig.~\ref{fig:gradient}, SOMETHING.
-
-The surface flux. WRT the direction omega. This allows generalization.
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/gradient}
-  \end{center}
-  \caption{The caption.}
-\label{fig:gradient}
-\end{figure}%
-
-The generalization of Eq.~\ref{eq:grad_y2} is applied to Eq.~\ref{eq:deriv_1} to yield Eq.~\ref{eq:deriv_2} which is generally valid for all octants.
-
-\begin{equation} \label{eq:deriv_2}
-\Omega \cdot \nabla \psi \approx 
-\frac{\psi_{x,in} - \psi_{x,out}}{\Delta x} \mu + 
-\frac{\psi_{y,in} - \psi_{y,out}}{\Delta y} \eta + 
-\frac{\psi_{z,in} - \psi_{z,out}}{\Delta z} \xi
-\end{equation}
-
-Using the approximation given in Eq.~\ref{eq:deriv_2}, Eq.~\ref{eq:boltz_i} can be rewritten as Eq.~\ref{eq:boltz_i2}.
-
-\begin{equation} \label{eq:boltz_i2}
-\begin{split}
-&\left[ 
-\frac{\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
-\frac{\psi_{i,a,y,in}^g - \psi_{i,a,y,out}^g}{\Delta y} \eta + 
-\frac{\psi_{i,a,z,in}^g - \psi_{i,a,z,out}^g}{\Delta z} \xi
-\right]
-+ \Sigma_{t,i}^g \psi_{i,a}^{g} \\
-& = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
-\end{split}
-\end{equation}
-
-It is useful to point out that each of $\psi_{i,a,x,in}^g$, $\psi_{i,a,x,out}^g$, $\psi_{i,a,y,in}^g$, $\psi_{i,a,y,out}^g$, $\psi_{i,a,z,in}^g$, and $\psi_{i,a,z,out}^g$ are \textit{surface} averaged flux values while $\psi_{i,a}^{g}$ is a \textit{volume} averaged flux value.
-
-In order for the discretized LBE to be solved using Eq.~\ref{eq:boltz_i2}, the incoming flux must always be known.
-Either the boundary conditions or previously calculated values determine the incoming flux.
+In order for the discretized LBE to be solved using Eq.~\ref{eq:boltz_i2}, both the incoming and outgoing flux must be known.
 Therefore, in order to compute the average flux in a cell, some relationship between the average flux and the outgoing surface fluxes must be known.
@@
 
-The diamond difference approximation, relates the incoming and outgoing surface fluxs to the cell averaged flux.
-The average flux is assumed to be the average of the two surface fluxes.
+The diamond difference approximation, relates the incoming and outgoing surface fluxes to the cell averaged flux.
+The cell averaged flux is assumed to be the average of any two opposite surface fluxes.
 This is assumed to be the case in all three spatial dimensions.
-This leads to Eq.~\ref{eq:dd}.
-
+This leads to:
 \begin{equation} \label{eq:dd}
@@
 \frac{\psi_{i,a,y,in}^g + \psi_{i,a,y,out}^g}{2} &= \psi_{i,a}^{g} \\
-\frac{\psi_{i,a,z,in}^g + \psi_{i,a,z,out}^g}{2} &= \psi_{i,a}^{g}
+\frac{\psi_{i,a,z,in}^g + \psi_{i,a,z,out}^g}{2} &= \psi_{i,a}^{g}.
 \end{split}
 \end{equation}
-
-Multiplying both sides of Eq.~\ref{eq:dd} and subtracting twice the outgoing surface flux from each yields Eq.~\ref{eq:dd2}.
-
+Multiplying both sides of Eq.~\ref{eq:dd} by 2 and subtracting twice the incoming surface flux from each yields:
 \begin{equation} \label{eq:dd2}
 \begin{split}
-\psi_{i,a,x,in}^g - \psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g \\
-\psi_{i,a,y,in}^g - \psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,out}^g \\
-\psi_{i,a,z,in}^g - \psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,out}^g
+\psi_{i,a,x,out}^g - \psi_{i,a,x,in}^g &= 2\psi_{i,a}^{g} - 2\psi_{i,a,x,in}^g \\
+\psi_{i,a,y,out}^g - \psi_{i,a,y,in}^g &= 2\psi_{i,a}^{g} - 2\psi_{i,a,y,in}^g \\
+\psi_{i,a,z,out}^g - \psi_{i,a,z,in}^g &= 2\psi_{i,a}^{g} - 2\psi_{i,a,z,in}^g.
 \end{split}
@@
 After the substitution:
-
 \begin{equation} \label{eq:boltz_i3}
@@
 &\left[ 
-\frac{2\psi_{i,a}^{g} - \psi_{i,a,x,out}^g}{\Delta x} \mu + 
-\frac{2\psi_{i,a}^{g} - \psi_{i,a,y,out}^g}{\Delta y} \eta + 
-\frac{2\psi_{i,a}^{g} - \psi_{i,a,z,out}^g}{\Delta z} \xi
+\frac{2\psi_{i,a}^{g} - 2\psi_{i,a,x,in}^g}{\Delta x} \mu + 
+\frac{2\psi_{i,a}^{g} - 2\psi_{i,a,y,in}^g}{\Delta y} \eta + 
+\frac{2\psi_{i,a}^{g} - 2\psi_{i,a,z,in}^g}{\Delta z} \xi
 \right]
@@
 & = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
+\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g.
 \end{split}
@@
 
-The diamond difference approximation defined in Eq.~\ref{eq:dd} can also be rearranged to solve for each outgoing flux once the volume averaged flux is computed.
-The outgoing flux is given by Eq.~\ref{eq:dd3}.
-
-\begin{equation} \label{eq:dd3}
-\begin{split}
-\psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,in}^g \\
-\psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,in}^g \\
-\psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,in}^g
-\end{split}
-\end{equation}
-
-Rearranging Eq.~\ref{eq:boltz_i3} to factor out the volume averaged flux term yields Eq.~\ref{eq:boltz_i4}.
-
+Rearranging Eq.~\ref{eq:boltz_i3} to factor out the volume averaged flux term yields:
 \begin{equation} \label{eq:boltz_i4}
@@
 \left[ 
-\frac{2\psi_{i,a,x,out}^g}{\Delta x} \mu + 
-\frac{2\psi_{i,a,y,out}^g}{\Delta y} \eta + 
-\frac{2\psi_{i,a,z,out}^g}{\Delta z} \xi
+\frac{2\psi_{i,a,x,in}^g}{\Delta x} \mu + 
+\frac{2\psi_{i,a,y,in}^g}{\Delta y} \eta + 
+\frac{2\psi_{i,a,z,in}^g}{\Delta z} \xi
 \right]
@@
 \end{equation}
-
-Equation~\ref{eq:boltz_i4} can be directly solved for the volume averaged flux. The result is
-
+which can be directly solved for the volume averaged flux:
 \begin{equation} \label{eq:boltz_i5}
@@
   \left[ 
-    \frac{2\psi_{i,a,x,out}^g}{\Delta x} \mu + 
-    \frac{2\psi_{i,a,y,out}^g}{\Delta y} \eta + 
-    \frac{2\psi_{i,a,z,out}^g}{\Delta z} \xi
+    \frac{2\psi_{i,a,x,in}^g}{\Delta x} \mu + 
+    \frac{2\psi_{i,a,y,in}^g}{\Delta y} \eta + 
+    \frac{2\psi_{i,a,z,in}^g}{\Delta z} \xi
   \right] + S_{i,a}^g
@@
   \Sigma_{t,i}^g
-}
+}.
 \end{equation}
 
-Repeatedly iteration Eq.~\ref{eq:boltz_i5} will eventually yield the solution to the flux distribution regardless of the intial guess.
+The diamond difference approximation defined in Eq.~\ref{eq:dd} can also be rearranged to solve for each outgoing flux once the volume averaged flux is computed. The outgoing flux is:
+\begin{equation} \label{eq:dd3}
+\begin{split}
+\psi_{i,a,x,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,x,in}^g \\
+\psi_{i,a,y,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,y,in}^g \\
+\psi_{i,a,z,out}^g &= 2\psi_{i,a}^{g} - \psi_{i,a,z,in}^g.
+\end{split}
+\end{equation}
+The outgoing flux values are then used as the incoming flux values for subsequent voxels.
+
+Repeatedly iteration Eq.~\ref{eq:boltz_i5} will eventually yield the solution to the flux distribution regardless of the initial guess.
 Choosing an accurate guess can greatly accelerate the code and guide convergence on a more accurate solution as discussed in more detail in Section~\ref{sec:uncol}.
@@
 \subsection{Convergence Criteria}
-The typical converence criteria used to determine whether or to terminate further iteration is the maximum relative error from the $i-1^{th}$ iteration to the $i^{th}$ iteration given in Eq.~\ref{eq:conv} over the phase space $\mathbb{R}$.
-Once the relative error in the parameter of interest, $\phi$ is below some threshold, $\epsilon$, the iteration stops and the solution is considered converged.
+The typical convergence criteria used to determine whether or to terminate further iteration is the maximum relative error from the $i-1^{th}$ iteration to the $i^{th}$ iteration given in Eq.~\ref{eq:conv} over the phase space $\mathbb{R}$.
+Once the relative error in the parameter of interest, $\phi$, is below some threshold, $\epsilon$, the iteration stops and the solution is considered converged.
 
@@
 In these cases, terminating the solution early rather than waiting for full convergence is appropriate.
-This is enforced by a second criterial.
+This is enforced by a second criterial given by Eq.~\ref{eq:conv2}.
 Whenever the number of iterations, $i$, exceeds the total permissible number of iterations, $I$, the iteration is terminated.
@@
 
-FIGURE XYZ SHOWS.
+FIGURE TO SHOW CONVERGENCE RATE.
 
 In order to measure the uncertainty due to floating-point error, another convergence criteria was added.
-Rather than looking at the relative change from iteration to iteration, the total absolute change is measured, computed by Eq.~\ref{eq:conv3}.
+Rather than looking at the relative change from iteration to iteration, the total absolute change is measured:
+\begin{equation}\label{eq:conv3}
+\epsilon = \sum_{\mathbb{R}} |\phi_i - \phi_{i-1}|.
+\end{equation}
 The termination criteria is given by Eq.~\ref{eq:conv4}.
 Computation of a value for $\epsilon$ requires two previous iteration, thus this condition cannot be met until the third iteration and thereafter.
-
-\begin{equation}\label{eq:conv3}
-\epsilon = \sum_{\vec{r}} |\phi_i - \phi_{i-1}|
-\end{equation}
 
@@
 Scatter is assumed to be a function of only the scatter angle $\theta_s$ between the initial, $\hat{\Omega}_a$, and scattered, $\hat{\Omega}_{a'}$, directions.
-The cosine of the scatter angle, $\mu_s$ is related by Eq.~\ref{eq:scat_cos}.
-
+The cosine of the scatter angle, $\mu_s$ is related by
 \begin{equation} \label{eq:scat_cos}
-\mu_s = \Omega_a \cdot \Omega_{a'} = \cos(\theta_s) \,, \quad 0 \leq \theta_s \leq \pi
+\mu_s = \Omega_a \cdot \Omega_{a'} = \cos(\theta_s) \,, \quad 0 \leq \theta_s \leq \pi.
 \end{equation}
@@
   \end{center}
-  \caption{The scatter angle.
-A photon at energy $E$ traveling in direction $\Omega$, hits a stationary atom, indicated by the green circle, and scatters into a new direction $\Omega'$ with a new energy $E'$.}
+  \caption{The scatter angle. A photon (blue) at energy $E$ traveling in direction $\Omega$, hits a stationary atom (green) and scatters into a new direction $\Omega'$ with a new energy $E'$.}
 \label{fig:scat_ang}
@@
 Details of the Legendre polynomials can be found in Appendix~\ref{appdx:leg}.
-The anisotropic scatter cross section ise rewritten as a Legendre expansion given where the scatter cosine is defined in Eq.~\ref{eq:scat_cos} as
-
+The anisotropic scatter cross section is rewritten as a Legendre expansion as
 \begin{equation} \label{eq:leg_1}
-\Sigma_{s, i, a, a'}^{g, g'} = \sum_{l=0}^L \frac{2l+1}{4 \pi}\Sigma_{s, i, l}^{g, g'} P_l(\mu_s)
+\Sigma_{s, i, a, a'}^{g, g'} = \sum_{l=0}^L \frac{2l+1}{4 \pi}\Sigma_{s, i, l}^{g, g'} P_l(\mu_s).
 \end{equation}
@@
 \section{Uncollided Solution Methods}\label{sec:uncol}
-One of the earliest problems with the discrete orinates method is the ppresence of ray effect.
-Ray effect arises due to the restriction of particle transport to discrete set of directions.
-As particles stream out of a point source, into a vacuum, the stream out along the set of direction schosen only.
-Since no scatter takes place in vacuum conditions, the praticles have no means to escape their initial direction.
-This gives rise to the phenomena observed in Fig.~FIGURE.
-This phenomena has been thrououghly studied by many [CITE].
+One of the earliest problems with the discrete ordinate methods is the presence of ray effect.
+Ray effect arises due to the restriction of particle transport to a set of discrete directions.
+To illustrate ray effect, consider a point source streaming particles isotropically into vacuum.
+Since no scatter or absorption takes place in vacuum conditions, the particles can only stream and particles must be conserved.
+However, if particles are transported along a single discrete direction only as illustrated in Fig.~\ref{fig:rayeffect_ex} ($\hat{\Omega} = <\sqrt{2}/2, \sqrt{2}/2>$ in Fig.~\ref{rayeffect_ex} and~\ref{fig:rayeffect_comp}) the phenomena observed in Fig.~\ref{fig:rayeffect_comp} is observed.
+As the discretization becomes more refined, the isotropic source becomes more heavily biased in the $\hat{\Omega}$ direction.
+This phenomena is called ray effect and has been thoroughly studied by many [CITE].
 
-The solution to prevent ray effect is to utilize an analytical uncollided source term.
-The full flux solution is split into two subproblems.
-Equation~\ref{eq:boltz} is rewritten as two equations solving for the uncollided flux, $\psi_u$, and collided flux, $\psi_c$ independently.
-Equation~\ref{eq:unc} has no scatter term since scattered particles are not considered in the uncollided flux.
-The lack of a scatter term makes the uncollided flux (whose analytical solution is intuitive) computable with a raytracing algorithm.
-
-The source term, $S_u$, in Eq.~\ref{eq:col} is the first collision source. The uncollided flux is driven by an external source. The flux distribution is solved analytically.
-
-\begin{equation} \label{eq:totflux}
-\psi = \psi_u + \psi_c
-\end{equation}
-
-\begin{equation} \label{eq:unc}
-\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-\psi_u(\boldsymbol{r}, E, \hat{\Omega}) =  S(\boldsymbol{r}, E, \hat{\Omega})
-\end{equation}
-
-\begin{equation} \label{eq:col}
-\begin{split}
-	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
-	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + S_{u}(\boldsymbol{r}, E, \hat{\Omega})
-\end{split}
-\end{equation}
+All particles must stream from one voxel to adjacent voxels.
+The streaming is considered (in 2D) along $<\sqrt{2}/2, \sqrt{2}/2>$ as shown in Fig.~\ref{fig:rayeffect_ex1}.
+All particles in the initial cell must stream outward into the two cells touching in the $\hat{\Omega}$ direction.
+The particles in each voxel split, half go to the voxel above and the other half travel to the right.
+This process is illustrated in Fig.~\ref{fig:rayeffect_ex}.
 
@@
 \end{figure}
-
-This then results in the value shown in the next figure.
 
@@
 
-Consider an isotropic point source in a homogeneous media. The (scalar) flux well known to be
-
-\begin{equation}
-\varphi(r, E) = \frac{S_0(E) e^{-\mu(E) r}}{4 \pi r^2}
+The solution to prevent ray effect is to utilize an analytical uncollided source term. The full flux solution is split into two subproblems.
+Equation~\ref{eq:boltz} is rewritten as two equations solving for the uncollided flux, $\psi_u$, and collided flux, $\psi_c$ independently:
+\begin{equation} \label{eq:totflux}
+\psi = \psi_u + \psi_c.
+\end{equation}
+Both $\psi_u$ and $\psi_c$ obey the LBE. The uncollided flux 
+\begin{equation} \label{eq:unc}
+\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+\psi_u(\boldsymbol{r}, E, \hat{\Omega}) =  S(\boldsymbol{r}, E, \hat{\Omega})
+\end{equation}
+has no scatter term since scattered particles are not considered in the uncollided flux.
+The lack of a scatter term makes the uncollided flux (whose analytical solution is derived shortly) computable with a raytracing algorithm.
+The uncollided flux is then used to compute the first collision source, $S_u$:
+\begin{equation}\label{eq:uncsrc}
+S_u = \int_{4\pi}^{} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'.
+\end{equation}
+which is then used to drive the collided flux in the same way the external source drives the uncollided flux distribution:
+\begin{equation} \label{eq:col}
+\begin{split}
+	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
+	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + S_{u}(\boldsymbol{r}, E, \hat{\Omega}).
+\end{split}
+\end{equation}
+Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~\ref{eq:col} simplifies it to
+\begin{equation} \label{eq:col2}
+\begin{split}
+	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) \\
+	&=
+	\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + \\
+	&\int_{4\pi} \int_{0}^{\infty} 
+\Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_u(\boldsymbol{r}, E', \hat{\Omega}') 
+dE' d\hat{\Omega}' \\
+& = \int_{4 \pi} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'.
+\end{split}
 \end{equation}
 
-\noindent where $S_0$ is the source strength, $\mu$ is the attenuation coefficient in the material, and $r$ is the distance traveled by the photon [CITE].
+A rigorous derivation for the uncollided flux is not provided in this work but can be obtained from [CITE].
+Instead, an intuitive approach is used.
+First, consider an isotropic point source in a homogeneous media.
+The (scalar) flux well known to be
+\begin{equation}
+\varphi(r, E) = \frac{S(E) e^{-\mu(E) r}}{4 \pi r^2}
+\end{equation}
+where $S$ is the source strength, $\mu$ is the attenuation coefficient in the material, and $r$ is the distance traveled by the photon [CITE].
 Expanding the problem to a full 3D model, the flux at $\boldsymbol{r}$ from a point source at $\boldsymbol{s}$ is
-
 \begin{equation}\label{eq:isouncol}
-\varphi(\boldsymbol{r}, E) = \frac{S_0(E) e^{-\mu(E) |\boldsymbol{r}-\boldsymbol{s}|}}{4 \pi |\boldsymbol{r}-\boldsymbol{s}|^2}.
+\varphi(\boldsymbol{r}, E) = \frac{S(E) e^{-\mu(E) |\boldsymbol{r}-\boldsymbol{s}|}}{4 \pi |\boldsymbol{r}-\boldsymbol{s}|^2}.
 \end{equation}
-
-\noindent
-The only tricky step is the introduction of anisotropy. The $4 \pi$ solid angle term is generally expressed by
-
-\begin{equation}
-\int_{4 \pi}^{} \int_{0}^{\infty} S_0(E, \hat{\Omega}) dE d\hat{\Omega}
+The only non-intuitive step is the introduction of anisotropy. The $4 \pi$ solid angle term in the denominator of Eq.~\ref{eq:isouncol} is generally expressed by
+\begin{equation}\label{eq:anisonorm}
+\int_{4 \pi}^{} \int_{0}^{\infty} S(E, \hat{\Omega}) dE d\hat{\Omega}
 \end{equation}
-
-\noindent
-but generally the normalization remains $4 \pi$.
-The flux is nonzero \textit{only} in the direction of travel, namely in the unit direction $\vec{r} - \vec{s}$.
-Therefore, a Kroenecker delta function (defined in Eq.~\ref{eq:kronecker}) is used to produce
-
-\begin{equation}
-\psi(\boldsymbol{r}, E, \hat{\Omega}) = \frac{S_0(E, \hat{\Omega}) e^{-\mu(E) |\boldsymbol{r}-\boldsymbol{s}|}}{4 \pi |\boldsymbol{r}-\boldsymbol{s}|^2} \delta \left( \hat{\Omega}, \frac{\boldsymbol{r} - \boldsymbol{s}}{|\boldsymbol{r} - \boldsymbol{s}|^2}\right)
-\end{equation}
-
-The uncollided flux in Eq.~\ref{eq:unc} can be analytically solved.
-The solution is given in Eq.~\ref{eq:uncsol} for a known isotropic point source located at position $\boldsymbol{s}$.
-The $\delta$ symbol in Eq.~\ref{eq:uncsol} is the Kronecker delta given by Eq.~\ref{eq:kronecker}.
-
-\begin{equation} \label{eq:uncsol}
-\psi_u(\boldsymbol{r}, E, \hat{\Omega}) = 
-S(\boldsymbol{s}, E)
-\frac{e^{-\sum_i \Sigma_{t,i} x_i}}{4\pi |\boldsymbol{r}-\boldsymbol{s}|^2}
-\delta\left( \hat{\Omega}, \frac{\boldsymbol{r}-\boldsymbol{s}}{|\boldsymbol{r}-\boldsymbol{s}|^2}\right)
+but DOCTORS assumes $S$ is normalized such that the integral in Eq.~\ref{eq:anisonorm} is $4 \pi$.
+The flux is nonzero \textit{only} in the direction of travel, namely in the unit direction $\boldsymbol{r} - \boldsymbol{s}$.
+Therefore, a Kronecker delta function (defined in Eq.~\ref{eq:kronecker}) is used to produce the angular flux:
+\begin{equation}\label{eq:uncsol}
+\psi(\boldsymbol{r}, E, \hat{\Omega}) = \frac{S(E, \hat{\Omega}) e^{-\mu(E) |\boldsymbol{r}-\boldsymbol{s}|}}{4 \pi |\boldsymbol{r}-\boldsymbol{s}|^2} \delta \left( \hat{\Omega}, \frac{\boldsymbol{r} - \boldsymbol{s}}{|\boldsymbol{r} - \boldsymbol{s}|^2}\right).
 \end{equation}
@@
 
-Equation~\ref{eq:uncsol} can be expanded to accomodate any number of spatially distributed sources of arbitrary directionality given in Eq.~\ref{eq:uncsol2} by integrating over the entire spatial domain, $V$.
-
+Equation~\ref{eq:uncsol} can be expanded to accomodate any number of spatially distributed sources of arbitrary directionality given in Eq.~\ref{eq:uncsol2} by integrating over the entire spatial domain, $V$
 \begin{equation} \label{eq:uncsol2}
@@
 \delta\left( \hat{\Omega}, \frac{\boldsymbol{r}-\boldsymbol{s}}{|\boldsymbol{r}-\boldsymbol{s}|^2}\right)
-d \boldsymbol{s}
+d \boldsymbol{s}.
 \end{equation}
 
-The uncollided source term, $S_u$ in Eq.~\ref{eq:col} is computed from the uncollided flux using Eq.~\ref{eq:uncsrc}.
-
-\begin{equation} \label{eq:uncsrc}
-S_u(\boldsymbol{r}, E, \hat{\Omega}) = \int_{4\pi} \int_{0}^{\infty} 
-\Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_u(\boldsymbol{r}, E', \hat{\Omega}') 
-dE' d\hat{\Omega}'
-\end{equation}
-
-Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~\ref{eq:col} gives Eq.~\ref{eq:col2} which simplifies to Eq.~\ref{eq:col3} by using Eq.~\ref{eq:totflux} to replace the collided and uncollided components with the total angular flux.
-
-\begin{equation} \label{eq:col2}
-\begin{split}
-	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) \\
-	&=
-	\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_c(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}' + \\
-	&\int_{4\pi} \int_{0}^{\infty} 
-\Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi_u(\boldsymbol{r}, E', \hat{\Omega}') 
-dE' d\hat{\Omega}' \\
-& = \int_{4 \pi} \int_{0}^{\infty} \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'
-\end{split}
-\end{equation}
-
-\begin{equation} \label{eq:col3}
-\begin{split}
-	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
-	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, \hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', \hat{\Omega}') dE' d\hat{\Omega}'
-\end{split}
-\end{equation}
+%\begin{equation} \label{eq:col3}
+%\begin{split}
+%	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
+%	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
+%	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, %\hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', %\hat{\Omega}') dE' d\hat{\Omega}'
+%\end{split}
+%\end{equation}
 
 \section{Source Generation}
-All photon tranpot problems considered in the DOCTORS code ar eultimately driven by an external source, namely th emedical equipment producing the photon beam.
-These sources can com ein a variety of geometric shapes and energy spectra.
-Currently, CODTORS has  anumber of built in beam typees available for anlaysis including point sources, fan beams, and cone beams.
-Fan beams and cone beams cna be arranged about a phantom to produce an axial slice.
-Currently, helical paths are not supported, though they could be added with minimal excess effort.
-
-\subsection{Point Source}
-The simplest source implemented is an isotropic point source.
-This source is defined only by its position and energy distribution.
-Given a source a $s$, the uncollided flux at the center of a cell, $<c>$ is computing using Eq.~\ref{eq:isouncol}.
-The collided flux is then computed using the discrete ordinate solution to the LBE using Eq.~\ref{eq:xxx}.
-
-\subsection{Fan Source}
-In addition to the values defining a point source (position $\vec{S}$ and the energy distribution) a fan source is defined by two additional parameters, $\varphi$ and $\theta$ which describe the azimuthal and polar angles subtented by the fan beam respectively.
-Figure~\ref{fig:fanbeam} shows $\varphi$ and $\theta$ in reference to the geometry setup.
-The raytracing algorithm used for the fan beam is identical to the point source raytracer except for one difference: before the raytracing is done, whether the ray falls within the fan beam is determined and any ray falling outside the fan beam is not traced, its uncollided flux is immediately set to zero.
-Some voxels are partially enclosed inside the beam.
+All photon transport problems considered in the DOCTORS code are ultimately driven by an external source, namely the medical equipment producing the photon beam.
+These sources can come in a variety of geometric shapes and energy spectra.
+Currently, DOCTORS has a number of built-in beam types available for analysis including point sources, fan beams, and cone beams.
+Fan beams and cone beams can be arranged about a phantom to produce an axial slice.
+Currently, helical paths are not supported, though they could be added with little additional effort.
+The geometry of fan beams and cone beams in DOCTORS is shown in Fig.~\ref{fig:shape}.
 
@@
 
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/fan_rejection}
-  \end{center}
-  \caption{The fan.}
-\label{fig:fan_rejection}
-\end{figure}
-
-First, the ray direction is projected onto the $xy$ plane giving $<S_x, S_y, 0>$ and $<C_x, C_y, 0>$.
-Then the cosine of the angle between them computed using Eq.~\ref{eq:phicos}.
-Once $\zeta$ is computed, the condition given by Eq.~\ref{eq:phicoscon} is evaluated.
-If the condition is met, the angle is accepted and raytraced normally, otherwise, it is rejected immediately.
-Figure~\ref{fig:phi}
+\subsection{Point Source}
+The simplest source implemented is an isotropic point source.
+This source is defined only by its position and energy distribution.
+Given a vector from the source to the center of the geometry, $\boldsymbol{s}$, and a vector from the source to the center of a cell, $\boldsymbol{c}$,the uncollided flux is computing using Eq.~\ref{eq:isouncol}.
+The collided flux is then computed using the discrete ordinate solution to the LBE given in Eq.~\ref{eq:col2}.
+The vectors are shown in Fig.~\ref{fig:phi}.
 
@@
   \end{center}
-  \caption{How $\varphi$ is computed.}
+  \caption{Typical DOCTORS phantom geometry viewed axially with the phantom.
+The beam (gray) is centered on $\boldsymbol{s}$ and the vector from the source to the voxel of interest $\boldsymbol{c}$ is used to determine whether that voxel should be accepted or rejected.
+In the case of isotropic point sources, $\boldsymbol{s}$ is undefined.}
 \label{fig:phi}
@@
 
+\subsection{Fan Source}
+In addition to the values defining a point source (position $\boldsymbol{s}$ and the energy distribution) a fan source is defined by two additional parameters, $\varphi$ and $\theta$ which describe the azimuthal and polar angles subtended by the fan beam respectively.
+Figure~\ref{fig:shape_fanbeam} shows $\varphi$ and $\theta$ with respect to the geometry setup.
+The raytracing algorithm used for the fan beam is identical to the point source raytracer except for one difference: before the raytracing is done, whether the ray falls within the fan beam is determined and any ray falling outside the fan beam is not traced, instead its uncollided flux is immediately set to zero.
+Some voxels are partially enclosed inside the beam which is not accounted for resulting in artifacts at the edges of the beam.
+
+To determine whether a ray is accepted or rejected, the ray direction is first projected onto the $xy$ plane giving $<s_x, s_y, 0>$ and $<c_x, c_y, 0>$.
+Then the cosine of the angle between them, $\zeta_\varphi$, is computed using
 \begin{equation}\label{eq:phicos}
-\cos(\zeta_\varphi) = \frac{<S_x, S_y, 0>}{\sqrt{S_x^2 + S_y^2}} \cdot \frac{<C_x, C_y, 0>}{\sqrt{C_x^2 + C_y^2}} = \frac{S_x C_x + S_y C_y}{\sqrt{S_x^2 + S_y^2} \sqrt{C_x^2 + C_y^2}}
+\cos(\zeta_\varphi) = \frac{<S_x, S_y, 0>}{\sqrt{S_x^2 + S_y^2}} \cdot \frac{<C_x, C_y, 0>}{\sqrt{C_x^2 + C_y^2}} = \frac{S_x C_x + S_y C_y}{\sqrt{S_x^2 + S_y^2} \sqrt{C_x^2 + C_y^2}}.
 \end{equation}
-
+Once $\zeta_\varphi$ is computed,
 \begin{equation}\label{eq:phicoscon}
@@
 \end{equation}
+is evaluated. If the condition is met, the angle is accepted and raytraced normally, otherwise, it is rejected immediately.
 
 Treatment for the $\theta$ variable is similar to that of $\varphi$ except the polar angle is required instead of the azimuthal angle.
-The polar angle of an arbitrary vector $\vec{\xi}$ can be computed using Eq.~\ref{eq:thetacos}.
-This is used to compute the polar angle for both $\vec{S}$ and $\vec{C}$.
-The difference between the polar angles of the two vectors is computed using Eq.~\ref{eq:thetacos2} and then the condition given in Eq.~\ref{eq:thetacoscon} is evaluated.
-If the condition is not met, the ray is rejected and no raytrace is performed.
+The polar angle of an arbitrary vector $\boldsymbol{\xi}$ can be computed using \begin{equation}\label{eq:thetacos}
+\cos(\theta_\xi) = \frac{\xi_z}{\sqrt{\xi_z^2 + \xi_y^2 + \xi_z^2}}.
+\end{equation}
+This is used to compute the polar angle for both $\boldsymbol{s}$ and $\boldsymbol{c}$. The difference between the polar angles of the two vectors 
+\begin{equation}\label{eq:thetacos2}
+\zeta_\theta = \theta_s - \theta_c
+\end{equation}
+is used to evaluate
+\begin{equation}\label{eq:thetacoscon}
+|\zeta_\theta| < \frac{\theta}{2}.
+\end{equation}
+If the condition is not met, the ray is rejected and no raytrace is performed. Figure~\ref{fig:fan_rejection} shows a simple example demonstrating cases when a voxel would be accepted or rejected.
 
-\begin{equation}\label{eq:thetacos}
-\cos(\theta_\xi) = \frac{\xi_z}{\sqrt{\xi_z^2 + \xi_y^2 + \xi_z^2}}
-\end{equation}
-
-\begin{equation}\label{eq:thetacos2}
-\zeta_\theta = \theta_S - \theta_C
-\end{equation}
-
-\begin{equation}\label{eq:thetacoscon}
-|\zeta_\theta| < \frac{\theta}{2}
-\end{equation}
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/fan_rejection}
+  \end{center}
+  \caption{The fan beam rejection. Voxels whose center lies outside the beam are rejected. This may lead to artifacts along the edges of the beam.}
+\label{fig:fan_rejection}
+\end{figure}
 
 \subsection{Multi-fan Source}
-A focal point is determined. The focal spot is at the center of each fan beam. The focal spot is always assumed to be at the center of the geometry and at the same $z$ elevation as the source point.
+The focal point about which the CT system rotates is assumed to be the center of the $xy$ plane and at the same $z$-elevation as the first user specified point.
+The focal spot defines the center of each fan beam.
 
@@
         \includegraphics[width=\textwidth]{figs/proj6}
-        \caption{$N=6$}
+        \caption{}
         \label{fig:proj6}
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
+    ~ 
     \begin{subfigure}[b]{0.2\textwidth}
         \includegraphics[width=\textwidth]{figs/proj12}
-        \caption{$N=12$}
+        \caption{}
         \label{fig:proj12}
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-    %(or a blank line to force the subfigure onto a new line)
+    ~ 
     \begin{subfigure}[b]{0.2\textwidth}
         \includegraphics[width=\textwidth]{figs/proj16}
-        \caption{$N=16$}
+        \caption{}
         \label{fig:proj16}
     \end{subfigure}
+    ~
     \begin{subfigure}[b]{0.2\textwidth}
         \includegraphics[width=\textwidth]{figs/proj24}
-        \caption{$N=24$}
+        \caption{}
         \label{fig:proj24}
     \end{subfigure}
-    \caption{Different numbers of fan beam projections.}\label{fig:fanproj}
+    \caption{Different numbers of fan beam projections ($N$). (a) $N = 6$. (b) $N = 12$. (c) $N = 16$. (d) $N = 24$.}\label{fig:fanproj}
 \end{figure}
 
+Rotation of an arbitrary point $(x,y)$ about the origin $\theta$ radians to a new position $(x', y')$ is performed by
 \begin{equation}\label{eq:xp}
@@
 \end{equation}
-
+and
 \begin{equation}\label{eq:yp}
-y' = y \cos \theta + x \sin \theta
+y' = y \cos \theta + x \sin \theta.
 \end{equation}
 
-To rotate a point $(x,y)$ about an arbitrary point $(x_f, y_f)$, Eq.~\ref{eq:xp2} and~\ref{eq:yp2} are used.
-
+Rotation about an arbitrary point (namely the focal point), $(x_f, y_f)$ requires translating the system such that the focal point is at the origin, performing the rotation given in Eq.~\ref{eq:xp} and~\ref{eq:yp}, and then translating back to the original coordinates.
+This is done by
 \begin{equation}\label{eq:xp2}
@@
 \end{equation}
-
+and
 \begin{equation}\label{eq:yp2}
-x' = (y - y_f) \cos \theta + (x - x_f) \sin \theta + y_f
+y' = (y - y_f) \cos \theta + (x - x_f) \sin \theta + y_f.
 \end{equation}
@@
 Cone sources are identical to fan beam sources except for the rejection criteria used to determine whether or not a voxel is inside the beam.
-The rejection crieria is a function of only a single parameter, $\theta$, the maximum permissible angle between the center of the beam and a ray toward the point of interest.
+The rejection criteria is a function of only a single parameter, $\theta$, the maximum permissible angle between the center of the beam, $\boldsymbol{s}$, and a ray toward the point of interest, $\boldsymbol{c}$.
 
-The actual angle is $\zeta$ which is computed similarly to Eq.~\ref{eq:phicos} without the projection onto the $xy$ plane in Eq.~\ref{eq:zetacos} can be simplified to a simple summation since both $\vec{S}$ and $\vec{C}$ are unit vectors.
-
+The angle between $\boldsymbol{s}$ and $\boldsymbol{c}$ is
 \begin{equation}\label{eq:zetacos}
-\cos(\zeta) = \frac{<S_x, S_y, S_z>}{\sqrt{S_x^2 + S_y^2 + S_z^2}} \cdot \frac{<C_x, C_y, C_z>}{\sqrt{C_x^2 + C_y^2 + C_z^2}} = S_x C_x + S_y C_y + S_z C_z
+\cos(\zeta) = \frac{<S_x, S_y, S_z>}{\sqrt{S_x^2 + S_y^2 + S_z^2}} \cdot \frac{<C_x, C_y, C_z>}{\sqrt{C_x^2 + C_y^2 + C_z^2}}
 \end{equation}
-
+which, if $\boldsymbol{s}$ and $\boldsymbol{c}$ are unit vectors, simplifies to
+\begin{equation}\label{eq:zetacos}
+\cos(\zeta) = S_x C_x + S_y C_y + S_z C_z.
+\end{equation}
+Once the angle is computed, the condition
 \begin{equation}\label{eq:zetacon}
@@
 \end{equation}
+is evaluated to determine whether the ray is accepted or rejected.
 
@@
 Dose can be computed in either one of two ways.
-The first way is to compute the spatial energy dependent fluence distribution and multiply by a fluence-to-dose conversion factor from ICRP 116 [CITE].
+The first way is to compute the fluence distribution and multiply by a fluence-to-dose conversion factor from ICRP 116 [CITE].
 The second way to compute dose is to compute the energy deposition in each cell.
-
-Transport of secondary particles is neglected and all energy deposition is assumed to be local.
+In both methods, transport of secondary particles is neglected and all energy deposition is assumed to be local.
 
@@
 For CT systems, the rotationally symmetric (ROT) orientation is the most appropriate since the beam rotates about the patient.
-
-This methodology is very simplistic and can be computed quickly with no extra (memory) overhead. When DOCTORS finishes running, the flux is multiplied by the photon
+Figure~\ref{fig:icrp116_rotconv} shows the energy dependent fluence-to-dose conversions for the ROT orientation.
 
@@
 
+This methodology is very simplistic and can be computed quickly with no extra (memory) overhead.
+When DOCTORS finishes running, the scalar flux is multiplied by the conversion factor and summed over all energy groups:
+\begin{equation}
+D_i = \sum_{g=0}^{G-1} \varphi_i^g H^g
+\end{equation}
+where $H^g$ is the conversion factor for the $g^{th}$ group.
+
 \subsection{Energy Deposition}
@@
 The alternative way to compute the dose is directly from the local energy deposition, $\mathcal{E}$, is by
-
 \begin{equation}
@@
 \end{equation}
-
-\noindent
 where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F$ is a unit conversion factor that converts MeV to Joules. The mass of the voxel, $M$ is then computed
-
 \begin{equation}
@@
 \end{equation}
-
-\noindent
 where $\rho_i$ is the density in g/cm$^3$ and G is a conversion factor to convert grams to kilograms. The dose in grays to the voxel is then
-
 \begin{equation}
-D_i = \frac{\mathcal{E}_i}{M_i} = \frac{\sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g \bar{E}^g F}{\rho_i G}
+D_i = \frac{\mathcal{E}_i}{M_i} = \frac{\sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g \bar{E}^g F}{\rho_i G}.
 \end{equation}

### 2017-06-02 docprep/chapconc.tex
Added the clinical settings conclusion

@@
 %%
-%% This is file `chapmin.tex',
-%% generated with the docstrip utility.
+%% Edward T. Norris
+%% Discrete Ordinates Computed Tomography Organ Dose Simulator (DOCTORS)
 %% 
-%% The original source files were:
+%% === Conclusion ===
 %%
-%% ths.dtx  (with options: `chapmin')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from chapmin.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file ths.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
 
- %% ... sample chapter ...
-
-Conclusions.
+This chapter summarizes the conclusions drawn from the results given in Chapter 5 and suggests further improvements to DOCTORS.
 
@@
 I concluded many things.
+
+\section{Applicability to Clinical Settings}
+With further development, DOCTORS is well positioned to become a code of significant clincal impact.
+
+Great effort was put into producing a code that is simple to use yet capable of producing powerful results quickly.
+However, usage in a clinical setting would likely require further refinement of the GUI to make the code more robust and intuitive.
+Additionally, the direct output of DOCTORS may not necessarily be of clinical importance since
 
@@
 \subsection{Temporal Extension}
+Use 4D CT to do temporal stuff as well as spatial stuff.
 
-Use 4D CT to do temporal stuff as well as spatial stuff.
+\subsection{Partial Acceptance Criteria}
+Allow it in the raytracer.
 
@@
 
+\subsection{Extension to Therapy Beams}
+High energy will be more anisotropic. Will need spherical harmonics.
+
+\subsection{Organ Identification}
+Currently, DOCTORS cannot automatically identify specific organs, thus it is only able to compute the equivalent dose via dose deposition.
+
 \endinput

### 2017-06-02 docprep/chapbib.tex
Finished draft of the thesis through the results

@@
 This chapter summarizes the literature reviewed in preparation of this work.
-The literature review is split into multiple sections, each highlighting relevant works pertaining to a particular component or section of this work.
-The first two sections cover introductory work and steps necessary for preprocessing data.
+The literature review is split into multiple sections, each highlighting relevant works pertaining to a particular component or methodology behind DOCTORS.
+The first two sections cover current dose estimation techniques and steps necessary for preprocessing data.
 Section~\ref{sec:discordlit} covers historical and recent work in discrete ordinate solution methodology, the methodology used by DOCTORS.
-The remaining sections cover GPU hardware and other implementation facets.
+The remaining sections give an overview of GPU hardware and other implementation facets.
 
 \section{CT Dose Estimation}
-The dose a patient receives form both diagnostic CT procedures and radiation therapy is of great concern to the medical community [CITE].
+As discussed in Chapter 1, the dose a patient receives from both diagnostic CT procedures and radiation therapy is of great concern to the medical community [CITE].
+However, dose quantification is not currently patient specific.
+The air kerma (measured in Gy) measures the x-ray beam intensity and can be measured directly [CITE].
+As the x-ray beam passes through the patient, it deposits energy.
+The energy deposition per unit mass is the absorbed dose (measured in Gy) which is then weighted by the radiation type to give the equivalent dose (measured in Sv).
+The equivalent dose quantifies the biological detriment to the patient from the procedure.
+The effective dose (measured in Sv) accounts for the radiosensitivity of each organ and is the most meaningful measure of risk to the patient.
+Any two procedures that result in the same effective dose to the patient carry the same acute and long-term risks.
+Any procedure that gives an effective dose to a patient has the same risk as any procedure that would give the same whole body dose.
 
-MOTIVATION
+Unfortunately, neither the absorbed dose, equivalent dose, nor effective dose to a patient can be measured directly.
+Instead, the CT Dose Index (CTDI) is measured using a pencil shaped ionization chamber at the center of  a standardized acrylic phantom placed at the scanner isocenter~\citep{ref:hudaw}.
+The CTDI quantifies the equivalent dose the patient receives from a single slice of the CT scan protocol.
+The CTDI does \textit{not} capture patient-specific papameters, but rather quantifies the amount or radiation emitted by the source.
 
-Monte Carlo codes have been used in the past to compute flux-to-dose conversion factors for a reference person [CITE ICRP].
+The CTDI does not provide the dose from an entire procedure, but rather a single slice.
+The dose from a procedure is obtained from the dose-length-product (DLP) which is the product of the CTDI and the axial scan length.
+The DLP is related to the effective dose to the patient for a particular procedure by a conversion factor that is dependent on many variables including the anatomical region scanned, the x-ray tube voltage, and the patient's age, gender, and size~\citep{ref:hudaw}.
+
+\citet{ref:lauk} found that severly obese patients received twice as much dose from CT procedures as others on average. Some patients received as much as five times the typical dose.
+
+The need for a complete understanding of the transport and scatter through the patient motivates full transport solutions.
+Typical solution modalities involve three key steps: (1) interpret the user input and prepare the solver, (2) run the transport solution to compute the scalar flux, and (3) convert the scalar flux to a dose.
+The first step is completely code dependent.
+The second and third steps of all major solvers falls into one of a few categories.
+
+Monte Carlo codes have been used in the past to compute flux-to-dose conversion factors for a reference person~\citep{ref:icrp116}.
+The values provided by~\citet{ref:icrp116} are energy dependent conversion factors, $H(E)$ that convert flux to effective dose for a particular orientation.
+The dose, $D$ is then computed
+\begin{equation}
+D = \int_0^\infty \varphi(E) H(E) dE
+\end{equation}
+where $\varphi$ is the scalar flux.
+
+Monte Carlo codes are generally known for their accuracy; they are able to sample the problem and attain arbitrarily high precision given sufficient runtime.
+They also take advantage of combinatorial geometry allowing a user to produce geometrically complex 3D structures.
+Deterministic methods are generally much faster than their Monte Carlo counterparts but do not benefit from excess runtime.
+They also consume vastly more resources, particularly random-access memory (RAM).
+However, in regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
+A large number of spatial regions slows down the particle transport in Monte Carlo and increases the RAM required to store the numerous voxels.
+
+Monte Carlo codes such as TOPAS and Geant4 \citep{ref:agostinellis} are the gold standard \citep{ref:jiax}
+
+Codes such as EGS4, its improved versions EGSnrc, and EGS5 \citep{ref:nelsonw}, and PENELOPE \citep{ref:salvatf} are all Monte Carlo and have shown very good results to experimental.
+But they are developed for high energy physics.
+
+The alternative to Monte Carlo is deterministic solutions.
+The best known solution techniques include the discrete ordinate method, the finite element method, and the method of characteristics.
+The discrete ordinate method is employed by DOCTORS and the methodology is covered independently in Section~\ref{sec:discordlit}.
+
+One of the most well known MOC codes id DeCART.
+DeCART is actually a coupled solver in that it utilizes different solution modalities for the radial plane and the axial direction.
+It uses a more complex modality in the radial plane such as MOC and then uses diffusion to transport particles axially~\citep{ref:hursinm}.
+
+The Method of Characteristics is a deterministic solution that has some advantages over discrete ordinates but is very complex to set up in 3D.
+One implementation is the TRANSMED code which was used to compute external beam therapy dose profiles accurately~\citep{ref:williamsm}.
+
+The MOC solution has been extended to the time dependent domain by \citet{ref:hoffmana}.
+
+Raytracing algorithms have been developed for numerous discret ordinate codes include RAY3D~\citep{ref:yingz} and ATTILLA~\citep{ref:wareingt} which utilizes unstructured tetrahedral meshes.
 
@@
 \section{CT Phantom Generation}
-The CT phantom is populated with CT numbers, also known as Hounsfield numbers.
+
+In order for the proposed methodology to be viable, the patient's CT phantom must be available.
+The phantom is reconstructed from the CT data.
+Though it is not the focus of this work, a brief summary of reconstruction algorithms is given here.
+
+Fundamentally, all reconstructions algorithms are based on a thin parallel beam of x-rays rotating axially about a patient to produce a sinogram.
+The inverse radon transform backprojects each row of the sinogram back to physical space.
+Filtering each row in Fourier space and then compositing all layers together recreates the image [CTIE].
+This methodology has been extended to fan beams [CITE] and cone beams [CITE].
+After the filtered backprojection, iterative reconstruction techniques can be applied to greatly improve the reconstruction quality [CITE].
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.38\textwidth}
+        \includegraphics[width=\textwidth]{figs/backproject1}
+        \caption{}
+        \label{fig:beamconexy}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.19\textwidth}
+        \includegraphics[width=\textwidth]{figs/backproject2}
+        \caption{}
+        \label{fig:beamfanxy}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.38\textwidth}
+        \includegraphics[width=\textwidth]{figs/backproject3}
+        \caption{}
+        \label{fig:subsweep_general3}
+    \end{subfigure}
+    \caption{(a) The original phantom. (b) The sinogram produced by 360 projections. (c) The filter backprojection reconstruction.}\label{fig:backprojection}
+\end{figure}
+
+Along any particular ray, monoenergetic particles will attenuate through a homogenous media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert} [CITE] resulting in a small fraction of the source particles reaching the detector.
+This phenomena leads to an image of the attenuating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
+With sufficient rays of varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation coefficient.
+The Hounsfield units (also called CT numbers) are traditionally scaled such that -1000 is dry air and water is zero [CITE].
+
+\begin{equation}\label{eq:beer_lambert}
+I(x) = I_0 e^{-\mu x}
+\end{equation}
+
+The reconstruction populates the CT mesh with CT numers, also known as Hounsfield units.
 These values are related to the attenuation coefficient of the material represented by the voxel.
-Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air CITE though some variations do exist CITE.
+Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air CITE though some variations do exist \citep{ref:plessisf} \citep{ref:sawc}.
 
-Many authors have provided correlations that map the CT number to both a material and a density.
+\citet{ref:schneideru} proposed a method to utilize different phantom materials to calibrate a HU-to-material conversion.
+He used only six different materials that descriminated between fat, water, muscle, and three densities of bone.
+From the calibration curve, he computed proton stopping powers accurate to within 1-2\%.
 
-First ones.
-Saw \citep{ref:sawc} did stuff. And \citep{ref:plessisf}. Then \citet{ref:schneideru} said stuff. And \citet{ref:kimh} did stuff with GEANT4.
+\citet{ref:plessisf} used a methodology similar to Schneider's but instead of performing experiments with phantoms, he used the ITS3 \citep{ref:halbleibj} Monte Carlo code to generate data.
+He identified 16 major tissue types in the body and classified them into seven dosimetrically equivalent categories.
+The resulting algorithm transormed HU directly into material and density values.
+Shortly thereafter, \citet{ref:schneiderw} completed a similar study but proposed considering all materials as a mixture of tissue and bone to varying degrees.
+However, the proposed algorithm fails to work well in soft tissue regions due to the presence of three components (water, fat, and muscle).
 
-Another Schneider did some stuff too \citep{ref:schneiderw}. And Vanderstraeten has a long name \citep{ref:vanderstraetenb}.
+Many of the earlier studies in HU-to-material conversions were done with high energy (typically 6-8 MeV) photon beams used for treatment \citep{ref:kimh} \citep{ref:vanderstraetenb}.
+\citet{ref:sawc} used a phantom with 17 inserts representing different dosimetric tissues.
 
-The new guy provides 19 materials. The density is measured.
-The DOCTORS code relies on the CT nuber to material mapping of \citet{ref:ottossonr} who developed a 19-group BLAH.
+\citet{ref:ottossonr} extended the methodology provided by \citet{ref:schneideru} to provide 19 dosimetric groups of materials in the diagnostic energy domain independent of the particular scanner used to generate the data.
+The DOCTORS code relies on his proposed 19-group model.
 Table~\ref{table:ctmap} shows the mapping of CT number to material composition (reproduced from \citep{ref:ottossonr}).
@@
 
-Two phantom datasets are used in this work [CITE].
-
 \section{Discrete Ordinates}\label{sec:discordlit}
 The dsicrete ordinates method dates back to CITE but remains one of the most prominent solution modalities for ration transport in use today.
+A comprehensive review of discrete ordinate methods is given by \citet{ref:lewise}.
+Additional references include [CITES].
 
-A comprehensive review of discrete ordinate methods is given by \citet{ref:lewise}. Additional references include
+The oldest major discrete ordinate method implementation is DORT which was superceded by its 3D counterpart TORT~\citep{ref:rhoadesw}.
+Since TORT, new discrete ordinate method implementations have laregly been extensions allowing more advanced computation.
 
-Monte Carlo codes such as TOPAS and Geant4 \citep{ref:agostinellis} are the gold standard \citep{ref:jiax}
+Discrete ordinate methods are almost always done on a Cartesian grid, though alternative derivations on general geometry do exist~\citep{ref:dehart}.
+Extension to unstructured tetrahedral meshes was done by ~\citet{ref:wareingt}.
 
-Codes such as EGS4, its improved versions EGSnrc, and EGS5 \citep{ref:nelsonw}, and PENELOPE \citep{ref:salvatf} are all Monte Carlo and have shown very good results to experimental.
-But they are developed for high energy physics.
+\citet{ref:waltersw} extended the discrete ordinates method to employ an adaptive collision source.
+Similarly, \citet{ref:ahrensc} wrote an algorithm that allows the quadrature to adapt to the energy group.
+Both of these methods improve the angular refinement.
+\citet{ref:ibrahima} added an adaptive mesh refinement.
+The adaptive mesh refinement allowed the code to automatically refine areas of interest where the flux was rapidly changing.
+This reduced the error in those regions reducing the computation time needed to reach a particular uncertainty level.
+An overview of other multigrid reduction schemes is given by \citet{ref:leeb}.
+\citet{ref:efremenkod} enumerates additional acceleration techniques available for discrete ordinates that can increase speed by 15-30\%.
 
-The Exnihilo thing by \citet{ref:evanst}. And DeHart wrote in \citep{ref:dehartm}
+Denovo is a massively parallel general-purpose discrete ordinate solver developed by~\citet{ref:evanst} to replace TORT.
+Denovo uses modern programming standards and utilizes the Exnihilo package for data processing and solution methods~\citep{ref:evanst2}.
 
-The Ibrahim did stuff \citep{ref:ibrahima} and Lee did stuff too \citep{ref:leeb}.
-
-With regards to angles, stuff has been done \citep{ref:ahrensc}
-
-There are many acceleration techniques available for discrete ordinates \citep{ref:efremenkod} that can increase speed by 15-30\%.
-
-GPU methods exist for solar stuff at satellites.
-
-The geometry has been extended to Non-Uniform Rational B-Splins (NURBS) for reactor eigenvalue problems.
-
-Attila had a 1st col src added by \citep{ref:wareingt}. Some rays are better than others \citep{ref:mathewsk}.
-
-Dort and Tort \citep{ref:rhoadesw}.
-
-Denovo.
+KBA DISCUSSION
 
@@
 
-Raytracing is important too. Woo came up with on \citep{ref:wooa} which was improved by \citet{ref:liuy} and \citet{ref:hel}.
+Many raytracing algorithms are based on the Bresenham line raserizing algorithm which was designed for graph plotting hardware~citep{ref:bresenhamj}.
+That algorithm was intende to produce lines acceptable for human interpretation and no guarantee is made that all pixels along the line will be identified making this class of algorithms inappropriate for dosimetric raytracing since attenuation through some voxels would be skipped.
+Some extensions have been added to it that guarantee passage through all voxels though~\citep{ref:liuy}.
+
+\citet{ref:clearyj} proposed a voxel traversal algorithm in which they initially identify the voxel containing the starting point of the ray, and transport it to the first surface of its bounding voxel it would cross.
+This is done by computing the distance traveled in order to cross each enclosing surface.
+The surface with the least distance is the first surface crossed.
+Therefore, the particle is advanced to the intersection point and considered inside the next cell, having crossed the surface.
+This process is repeated until a termination criteria is met.
+
+This work uses the raytracing algorithm proposed by~\citet{ref:wooa} which is very similar to the method proposed by~\citet{ref:clearyj} but based on the parametric equation of a line, $\boldsymbol{u} + \boldsymbol{v}t$ which makes implementation in 3D simple.
+Since its develpment, many improvements and extensions have been made for alternate geometries such as rhombic dodecahedra~\citep{ref:hel}
+
+GPU RAYTRACING
 
@@
 
-Monte Carlo dose computation for protons has been accelerated by GPU technology \citep{ref:jiax} and
+Two major GPU manufacturers control the vast majority of the market, Nvidia and AMD [CITE].
+In 2007, Nvidia released its Compute Unified Device Architecture (CUDA) language.
+CUDA allows code written in C/C++/FORTRAN to communicate with a GPU with minimal additional code.
+Therefore, CUDA was selected for the GPU acceleration of DOCTORS.
+Therefore, only Nvidia GPUs are considered in this work.
 
-Monte Carlo in voxelized geometries has been done by \citep{ref:hissoinys} for monoenergetic beams.
-GPUs are cool.
+In GPU texts, the GPU itself is referred to as a device and the CPU or other hardward running the GPU is referred to as the host.
+The device has its own instruction set which requires a special compiler.
+The device also has its own onboard memory so data must be copied from the host to the device and then back.
 
-CUDA is cool.
+Conceptually, code run on the device is split into block on a grid, each block then runs multiple concurrent computations, each on a thread.
+In hardware, threads execute on 32-128 streaming multiprocessors depending on the GPU architecture.
+
+Monte Carlo dose computation for protons has been accelerated by GPU technology \citep{ref:jiax} and Monte Carlo in voxelized geometries has been done by \citep{ref:hissoinys} for monoenergetic beams.
 

### 2017-06-02 docprep/chapintro.tex
Finished draft of the thesis through the results

@@
 %%
-%% This is file `chapmin.tex',
-%% generated with the docstrip utility.
+%% Edward T. Norris
+%% Discrete Ordinates Computed Tomography Organ Dose Simulator (DOCTORS)
 %% 
-%% The original source files were:
+%% === Introduction ===
 %%
-%% ths.dtx  (with options: `chapmin')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from chapmin.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file ths.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
 
-Computed tomography (CT) is becoming increasingly pervasive in medical diagnostics since its introduction in 1973 by \citet{ref:hounsfieldg} as improved algorithms give doctors access to more information.
+Since its introduction in 1973 by \citet{ref:hounsfieldg} computed tomography (CT) has become pervasive in medical diagnostics as improved algorithms and techniques give doctors access to higher quality information.
 Increased information results in faster and more accurate diagnosis.
-However, the exponentially increasing usage of CT and corresponding radiation dose it delivers to the patient has raised concerns regarding potential long term radiation risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}.
+However, the exponentially increasing usage of CT and corresponding dose patients receive has raised concerns regarding potential long term risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}.
+These concerns are exemplified by the push for low dose CT.
+
 A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.
+
+Though concerns have been raised in the medical community about the risks associated with CT, the techniques used to quantify dose are very simple and not patient specific.
+The dose-length-product (DLP) is used as a measure to quantify the effective dose a patient receives.
+However, the DLP must be modified by a factor accounting for may variations amongst patients such as age, gender, and size as well as the kVp used in the procedure.
 Currently, dose estimation relies on \textit{a priori} computation verified with a standardized benchmark.
-Once the procedure (whether diagnostic or treatment) is completed, the dose distribution is assumed to follow the empirical model.
 
 No methodology currently exists to verify that the dosimetry evaluation was accurate after the patient has undergone the procedure.
-This work proposes a methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam.
+This work proposes patient specific a methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam.
 This can also help doctors estimate spatial dose distribution to the patient to ensure no specific organ received more dose than permissible.
 
-This work implements and analyses a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient can be computed with a full transport solution using a CT phantom.
+This work implements and analyses a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient can be computed with a full transport solution inside a CT phantom.
 The methodology converts a CT mesh into a voxel phantom of materials and densities.
-Then a discrete ordinate method computes the flux throughout the phantom.
-Local energy deposition is assumed relieving the need for secondary electron transport.
+The user supplies information about the beam and a discrete ordinate method computes the flux throughout the phantom.
+The flux is then used to compute the dose using local energy deposition to relieve the need for secondary electron transport.
 
-Another, similar, application is dose estimation of patients receiving radiation therapy.
-Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment [CITE].
-With further development, this methodology may enable real time dose computation as the treatment is administered.
-This would be greatly beneficial to both patients and the doctors administering the procedure.
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.38\textwidth}
-        \includegraphics[width=\textwidth]{figs/backproject1}
-        \caption{}
-        \label{fig:beamconexy}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.19\textwidth}
-        \includegraphics[width=\textwidth]{figs/backproject2}
-        \caption{}
-        \label{fig:beamfanxy}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.38\textwidth}
-        \includegraphics[width=\textwidth]{figs/backproject3}
-        \caption{}
-        \label{fig:subsweep_general3}
-    \end{subfigure}
-    \caption{(a) The original phantom. (b) The sinogram produced by 360 projections. (c) The filter backprojection reconstruction.}\label{fig:backprojection}
-\end{figure}
-
-In order for the proposed methodology to be viable, the patient's CT phantom must be available.
-The phantom is reconstructed from the CT data.
-Though it is not the focus of this work, a brief summary of reconstruction algorithms is given here.
-
-Fundamentally, all reconstructions algorithms are based on a thin parallel beam of x-rays rotating axially about a patient to produce a sinogram.
-The inverse radon transform backprojects each row of the sinogram back to physical space.
-Filtering each row in Fourier space and then compositing all layers together recreates the image [CTIE].
-This methodology has been extended to fan beams [CITE] and cone beams [CITE].
-After the filtered backprojection, iterative reconstruction techniques can be applied to greatly improve the reconstruction quality [CITE].
-
-Along any particular ray, monoenergetic particles will attenuate through a homogenous media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert} [CITE] resulting in a small fraction of the source particles reaching the detector.
-This phenomena leads to an image of the attenuating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
-With sufficient rays of varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation coefficient.
-The Hounsfield units (also called CT numbers) are traditionally scaled such that -1000 is dry air and water is zero [CITE].
-
-\begin{equation}\label{eq:beer_lambert}
-I(x) = I_0 e^{-\mu x}
-\end{equation}
-
-Once a patient specific phantom is obtained, the flux throughout the mesh must be computed. Computational techniques to do this can be split into two broad categories: Monte Carlo and deterministic.
-
-Monte Carlo codes are generally known for their accuracy; they are able to sample the problem and attain arbitrarily high precision given sufficient runtime.
-They also take advantage of combinatorial geometry allowing a user to produce geometrically complex 3D structures.
-Deterministic methods are generally much faster than their Monte Carlo counterparts but do not benefit from excess runtime.
-They also consume vastly more resources, particularly random-access memory (RAM).
-However, in regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
-A large number of spatial regions slows down the particle transport in Monte Carlo and increases the RAM required to store the numerous voxels.
-
-Deterministic codes can be accelerated with the use of a graphics processing unit (GPU).
+DOCTORS leverages graphics processing units (GPUs) to accelerate the transport step.
 GPUs differ from the central processing unit (CPU) in that each of the many cores performs identical instruction to all others at the same time but each with different data.
@@
 To this end, a graphical user interface (GUI) was developed.
-The GUI utilize the Qt5 graphics framework for the window, buttons, and other necessary widgets.
+The GUI utilizes the Qt5 graphics framework for the window, buttons, and other necessary widgets.
 The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicated required and completed steps.
 The output is then plotted graphically as well as sent to an ASCII text file for more advanced postprocessing.
+Utilization of Qt5 also makes the code portable, it can be compiled on either Linux or Windows operating systems with no changes to the code.
 More details about the GUI are included in Section~\ref{sec:gui}.
@@
 
+Another, similar, application is dose estimation of patients receiving radiation therapy.
+Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment [CITE].
+With further development, this methodology may enable real time dose computation as the treatment is administered.
+This would be greatly beneficial to both patients and the doctors administering the procedure.
+
+The remainder of this work is organized into five chapters.
+Chapter 2 summarizes the existing literature pertinent to this work.
+Chapter 3 lays the mathematical foundation for each of the major components of DOCTORS.
+Chapter 4 gives a overview of the implementation strategies used to transform the mathematical framework in Chapter 3 into code that can execute on a modern computing platform.
+Chapter 5 summarizes the results obtained.
+It quantifies both the dosimetric accuracy and runtime of the DOCTORS code.
+Finally, Chapter 6 summarizes the entire work and provides some concluding remarks.
+Afterward, some appendices provide additional information that was not included in the main text but may still be helpful to a reader.
+
 \endinput

### 2017-06-03 docprep/chapimpl.tex
Deleted error
Added a section label to the vector flattening section
Updated the header
Minor fixes and added some figs

@@
 %%
-%% This is file `chapall.tex',
-%% generated with the docstrip utility.
+%% Edward T. Norris
+%% Discrete Ordinates Computed Tomography Organ Dose Simulator (DOCTORS)
 %% 
-%% The original source files were:
+%% === Implementation ===
 %%
-%% ths.dtx  (with options: `chapmin,addfig,addtbl,addbib,addglo')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from chapall.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file ths.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-
- %% ... sample chapter ...
 
@@
 
-The code is implemented in C++ and requires a C0x11 complianct compiler.
+The code is implemented in C++ and requires a C0x11 compliant compiler.
 The user interface is written using the Qt5 framework.
@@
 
-\section{Vector Flattening}
+\section{Vector Flattening}\label{sec:flatten}
 
@@
 
-SHOW CARBON ELEMENTAL DATA
+Figure~\ref{fig:sigmacomp} shows the microscopic cross section data pulled from
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/sigmacomp1}
+        \caption{}
+        \label{fig:ampx1}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/sigmacomp2}
+        \caption{}
+        \label{fig:ampx2}
+    \end{subfigure}
+    \caption{The microscopic cross section in bars for hydrogen and oxygen.
+Both subfigures show identical data, (a) shows the entire data range available in the reference data library and (b) shows only the data range applicable to DOCTORS.}\label{fig:sigmacomp}
+\end{figure}
 
@@
 
-Once the data is parsed, weighted combinations of elemental data is used to create material cross sections. The details of this computation are covered in Section~\ref{sec:
+Once the data is parsed, weighted combinations of elemental data is used to create material cross sections.
+
+MT AND MF NUMBERS
 

### 2017-06-03 docprep/chapmeth.tex
Finished draft of the thesis through the results
Minor fixes and added some figs

@@
 %%
-%% This is file `chapmin.tex',
-%% generated with the docstrip utility.
+%% Edward T. Norris
+%% Discrete Ordinates Computed Tomography Organ Dose Simulator (DOCTORS)
 %% 
-%% The original source files were:
+%% === Methods ===
 %%
-%% ths.dtx  (with options: `chapmin')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from chapmin.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file ths.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
 
@@
 
-The linear Boltzmann equation (LBE) is generally true of any particle given that outside forces (electromagnetic or gravitational) are either not present or negligible and particle-particle collisions are insignificant.
-Therefore, the LBE is generally not valid for charged particles or systems whose particle density is on the order of the medium they are traveling through.
-In such systems, particle-particle interactions may not be negligible.
+The linear Boltzmann equation (LBE) is generally true of any particle for which outside forces (electromagnetic or gravitational) are either not present or negligible and particle-particle collisions are insignificant.
+Therefore, the LBE is generally not valid for charged particles or systems whose particle density is on the order of the confining medium.
+In such systems, particle-particle interactions may not be negligible requiring the non-linear Boltzmann equation.
 
@@
 Particles can (1) stream into the volume across a bounding surface, (2) inscatter from another energy-direction component of phase space, or (3) be produced directly by the external source.
-Three removal terms are also present, particles can (1) stream out of the volume of interest by crossing a bounding surface, (2) be absorbed and vanish entirely, or (3) outscatter to a energy-direction component of phase space other than the one of interest.
+At the same time, three removal terms are present, particles can (1) stream out of the volume of interest by crossing a bounding surface, (2) be absorbed and vanish entirely, or (3) outscatter to a energy-direction component of phase space other than the one of interest.
 
-The downscatter and absorption removal terms are typically grouped together using
+The downscatter and absorption removal terms are typically grouped together as
 \begin{equation}
@@
 Given an arbitrary direction, $\Omega$, $\mu$, $\eta$, and $\xi$ are its direction cosines with respect to the $x$, $y$, and $z$ axes respectively.
-$\varphi$ is the azimuthal angle (with respect to $x$ and the d $\theta$ is the polar angle (with respect to $z$).}
+$\varphi$ is the azimuthal angle (with respect to $x$) and $\theta$ is the polar angle (with respect to $z$).}
 \label{fig:coord_sys}
@@
 
-The rotation symmetrical quadrature is implemented in DOCTORS.
-From amongst many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers [CITE] which enables simple, direct comparison to other solvers.
+The rotation symmetrical ($S_N$) quadrature is implemented in DOCTORS.
+From amongst many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and used in production discrete ordinate solvers~\citep{ref:evanst} which enables simple, direct comparison to other solvers.
 Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.
 
-THEORY AND MATH?
+ADD THEORY AND MATH OF ANGLE/WEIGHT GENERATION?
 
@@
 Also, after normalization, no two rows should have identical values for $\mu$, $\eta$, and $\xi$ such as the last two rows of Table~\ref{tab:quad_interp}, nor may any weights be negative.
-The data in Table~\ref{tab:quad_format} would be read in and renormalized by DOCTORS.
-The normalized data would be identical to the data in Table~\ref{tab:quad_interp}.
+The data in Table~\ref{tab:quad_format} would be read in and renormalized by DOCTORS resulting in data identical to  Table~\ref{tab:quad_interp}.
 
-Values are read in as a 32 bit IEEE-754 floating point number [CITE] which ensures seven significant digits.
+Values are read in as a 32-bit IEEE-754 floating point number~\citep{ref:ieee754} which ensures seven significant digits.
 Significant digits beyond this will be truncated.
@@
 where $\hat{i}$, $\hat{j}$, and $\hat{k}$ are the unit directions in the $x$, $y$, and $z$ directions respectively.
-Therefore, any discrete direction $\hat{\Omega}_a$ can be expressed as $<\mu, \eta, \xi>$.
+Therefore, any discrete direction $\hat{\Omega}_a$ can be expressed as $<\mu_a, \eta_a, \xi_a>$.
 The two forms will be used interchangeably.
 
-Discretizing Eq.~\ref{eq:boltz} in angular space gives
+Discretizing the LBE (Eq.~\ref{eq:boltz}) in angular space gives
 \begin{equation} \label{eq:boltz_a}
@@
 
-Once the energy group structure is selected, cross section data must be available.
+The user must provide cross section data in the form of an AMPX formatted binary file (Format detaisl are provided in Section~\ref{sec:xsparse}).
+The cross section data file contains group structure information.
 Group averaged cross section values from energy $E_1$ to $E_2$ are computed using Eq.~\ref{eq:groupxs} where $\sigma(E)$ is the continuous microscopic cross section.
-A weighting function, $f$, is used to weight some energies.
+A weighting function, $f$, is used to weight the energies.
 The weighting function can be any of a number commonly used functions.
@@
 \end{equation}
-which is very similar; only the starting point for the scatter summation changes.
+which nearly identical; only the starting point for the scatter summation changes.
 Computationally, though, the lack of upscatter makes a significant difference.
@@
 
-The problem domain for a CT voxel phantom is a regular, rectangular parallelepiped of dimension $D_x \times D_y \times D_z$, the values of which must be known based on the CT setup by the user.
+The problem domain for a CT voxel phantom is a regular, rectangular parallelepiped of dimension $D_x \times D_y \times D_z$, the values of which are known based on the CT setup provided by the user.
 The mesh is partitioned into $N_x$, $N_y$, and $N_z$ evenly spaced bins along each major direction as shown in Fig.~\ref{fig:spatial_disc}.
@@
 Each voxel in the mesh has three components indices, one in each major direction, $i_x$, $i_y$, and $i_z$ that range from 0 to $N_x - 1$, $N_y - 1$, or $N_z-1$.
-However, this makes indexing cumbersome, so they are combined into a unique global index, $i$ using Eq.~\ref{eq:indx_flat}.
+However, this makes indexing cumbersome, so they are combined into a unique global index, $i$ using
+\begin{equation} \label{eq:indx_flat}
+i = i_x (N_z + N_y) + i_y N_z + i_z.
+\end{equation}
 In general, this is known as "flattening" a matrix into a one-dimensional vector.
 More details about the flattening implementation used in DOCTORS is given in Section~\ref{sec:flatten}.
-
-\begin{equation} \label{eq:indx_flat}
-i = i_x (N_z + N_y) + i_y N_z + i_z
-\end{equation}
-
 Using global indexing discussed, the fully discretized the steady state LBE given is
@@
 \psi_{i,a}^{g} = 
-\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g
+\sum_{a=0}^{N_a-1} \sum_{g'=g}^{G-1} \Sigma_{s, i, a, a'}^{g, g'} \psi_{i, a'}^{g'} \omega_a + S_{i,a}^g.
 \end{equation}
@@
 \end{equation}
-To a first order approximation, a partial derivative is computed as
+To a first order approximation, a partial derivative is computed
 \begin{equation} \label{eq:deriv_1}
@@
 
-The gradient is computed with respect to th edirection $\hat{\Omega}$, thus Eq.~\ref{eq:spatial_1} is valid only for the first octant ($\mu, \eta, xi > 0$).
+The gradient is computed with respect to the direction $\hat{\Omega}$, thus Eq.~\ref{eq:spatial_1} is valid only for the first octant ($\mu, \eta, \xi > 0$).
 In other octants, a new gradient must be used.
@@
 \begin{cases}
-\psi(x+\Delta x/2, y, z) - \psi(x-\Delta x/2, y, z) \,, \quad \mu > 0 \\
-\psi(x-\Delta x/2, y, z) - \psi(x+\Delta x/2, y, z) \,, \quad \mu < 0
+\psi(x+\Delta x/2, y, z) \,, \quad \mu > 0 \\
+\psi(x-\Delta x/2, y, z) \,, \quad \mu < 0
 \end{cases}
@@
 \begin{cases}
-\psi(x-\Delta x/2, y, z) - \psi(x+\Delta x/2, y, z) \,, \quad \mu > 0 \\
-\psi(x+\Delta x/2, y, z) - \psi(x-\Delta x/2, y, z) \,, \quad \mu < 0
+\psi(x-\Delta x/2, y, z) \,, \quad \mu > 0 \\
+\psi(x+\Delta x/2, y, z) \,, \quad \mu < 0
 \end{cases}
@@
 The cell averaged flux is assumed to be the average of any two opposite surface fluxes.
-This is assumed to be the case in all three spatial dimensions.
-This leads to:
+This is assumed to be the case in all three spatial dimensions leading to:
 \begin{equation} \label{eq:dd}
@@
 The form of the left hand side of Eq.~\ref{eq:dd2} is the same as in Eq.~\ref{eq:boltz_i2}.
-This allows Eq.~\ref{eq:dd2} to be used to make a substitution in Eq.~\ref{eq:boltz_i2} that removes the dependence on the outgoing flux from discretized equation.
+This allows Eq.~\ref{eq:dd2} to be used to make a substitution in Eq.~\ref{eq:boltz_i2} that removes the dependence on the outgoing flux.
 After the substitution:
@@
 \subsection{Convergence Criteria}
-The typical convergence criteria used to determine whether or to terminate further iteration is the maximum relative error from the $i-1^{th}$ iteration to the $i^{th}$ iteration given in Eq.~\ref{eq:conv} over the phase space $\mathbb{R}$.
+The typical convergence criteria used to determine whether or not to terminate further iteration is the maximum relative error from the previous iteration to the current iteration given in Eq.~\ref{eq:conv} over the phase space $\mathbb{R}$ where $t$ is the current iteration number.
 Once the relative error in the parameter of interest, $\phi$, is below some threshold, $\epsilon$, the iteration stops and the solution is considered converged.
@@
 \begin{equation}\label{eq:conv}
-\max_{\mathbb{R}} \left\{ \frac{\phi_i - \phi_{i-1}}{\phi_i} \right\} < \epsilon
+\max_{\mathbb{R}} \left\{ \frac{\phi_t - \phi_{t-1}}{\phi_t} \right\} < \epsilon
 \end{equation}
@@
 This is enforced by a second criterial given by Eq.~\ref{eq:conv2}.
-Whenever the number of iterations, $i$, exceeds the total permissible number of iterations, $I$, the iteration is terminated.
+Whenever the number of iterations, $t$, exceeds the total permissible number of iterations, $T$, the iteration is terminated.
 
 \begin{equation}\label{eq:conv2}
-i > I
+t > T
 \end{equation}
@@
 
-FIGURE TO SHOW CONVERGENCE RATE.
+ADD FIGURE TO SHOW CONVERGENCE RATE.
 
 In order to measure the uncertainty due to floating-point error, another convergence criteria was added.
-Rather than looking at the relative change from iteration to iteration, the total absolute change is measured:
+Rather than considering the relative change from iteration to iteration, the total absolute change is measured:
 \begin{equation}\label{eq:conv3}
-\epsilon = \sum_{\mathbb{R}} |\phi_i - \phi_{i-1}|.
+\varepsilon = \sum_{\mathbb{R}} |\phi_i - \phi_{i-1}|.
 \end{equation}
 The termination criteria is given by Eq.~\ref{eq:conv4}.
-Computation of a value for $\epsilon$ requires two previous iteration, thus this condition cannot be met until the third iteration and thereafter.
+Computation of a value for $\varepsilon$ requires two previous iteration, thus this condition cannot be met until the third iteration and thereafter.
 
 \begin{equation}\label{eq:conv4}
-\epsilon_i > \epsilon_{i-1}
+\varepsilon_i > \varepsilon_{i-1}
 \end{equation}
@@
 
-The data files containing the scatter cross sections are distributed using a Legendre polynomial expansion.
+The data files containing the scatter cross sections stor the information using Legendre polynomial expansion coefficients.
 The use of a Legendre expansion removes the dependence on the quadrature from the data.
@@
 Details of the Legendre polynomials can be found in Appendix~\ref{appdx:leg}.
-The anisotropic scatter cross section is rewritten as a Legendre expansion as
+The anisotropic scatter cross section is rewritten using a Legendre expansion as
 \begin{equation} \label{eq:leg_1}
@@
 As the discretization becomes more refined, the isotropic source becomes more heavily biased in the $\hat{\Omega}$ direction.
-This phenomena is called ray effect and has been thoroughly studied by many [CITE].
+This phenomena is called ray effect and has been thoroughly studied by many \citep{ref:mathewsk} \citep{ref:tencerj}.
 
 All particles must stream from one voxel to adjacent voxels.
-The streaming is considered (in 2D) along $<\sqrt{2}/2, \sqrt{2}/2>$ as shown in Fig.~\ref{fig:rayeffect_ex1}.
+The streaming is considered (in 2D) along $<\sqrt{2}/2, \sqrt{2}/2>$ as shown in Fig.~\ref{fig:rayeffect_ex}.
 All particles in the initial cell must stream outward into the two cells touching in the $\hat{\Omega}$ direction.
@@
     \end{subfigure}
-    \caption{The process}\label{fig:rayeffect_ex}
+    \caption{(a) Particles start in the lower left cell and travel along $\Omega$.
+(b) Half travel to the cell above and half to the right.
+(c) Half of the particles in each cell move up and half to the right.
+The center cell gets particles from the cells below and left of it.
+(d) Some particles escape across the vacuum boundary.
+}\label{fig:rayeffect_ex}
 \end{figure}
@@
 First, consider an isotropic point source in a homogeneous media.
-The (scalar) flux well known to be
+The (scalar) flux is well known to be
 \begin{equation}
@@
 When DOCTORS finishes running, the scalar flux is multiplied by the conversion factor and summed over all energy groups:
-\begin{equation}
+\begin{equation}\label{eq:doseicrp}
 D_i = \sum_{g=0}^{G-1} \varphi_i^g H^g
@@
 
-The alternative way to compute the dose is directly from the local energy deposition, $\mathcal{E}$, is by
+The alternative way to compute the dose is directly from the local energy deposition. The total deposition, $\mathcal{E_{T,i}}$ for voxel $i$ is
+\begin{equation}\label{eq:dept}
+\mathcal{E}_{T,i} = \mathcal{E}_{a,i} + \mathcal{E}_{s,i}
+\end{equation} 
+where $\mathcal{E_{a,i}}$ is the energy deposition from absorption and $\mathcal{E_{s,i}}$ is the energy deposition from scatter. The absorption deposition is
+\begin{equation}\label{eq:depa}
+\mathcal{E}_{a,i} = \sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g V_i \bar{E}^g F_1
+\end{equation}
+where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F_1$ is a unit conversion factor that converts MeV to Joules.
+Equation~\ref{eq:depa} assumes total local energy deposition in the voxel of interest.
+The energy deposition from scatter is
+\begin{equation}\label{eq:deps}
+\mathcal{E}_{s,i} = \sum_{g=0}^{G-1} \sum_{h=g-1}^{G-1} \left[ \Sigma_{s,i}^{g,h} \varphi_i^g V_i (\bar{E}^g-\bar{E}^h) F_1 \right]
+\end{equation}
+making the total deposition
+\begin{equation}\label{eq:dept2}
+\mathcal{E}_{T,i} = V_i F_1 \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g-1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right]
+\end{equation}
+The mass of the voxel, $M_i$ is then computed
 \begin{equation}
-\mathcal{E}_i = \sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g V_i \bar{E}^g F
+M_i = \rho_i V_i F_2
 \end{equation}
-where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F$ is a unit conversion factor that converts MeV to Joules. The mass of the voxel, $M$ is then computed
-\begin{equation}
-M_i = \rho_i V_i G
-\end{equation}
-where $\rho_i$ is the density in g/cm$^3$ and G is a conversion factor to convert grams to kilograms. The dose in grays to the voxel is then
-\begin{equation}
-D_i = \frac{\mathcal{E}_i}{M_i} = \frac{\sum_{g=0}^{G-1} \Sigma_{a,i}^g \varphi_i^g \bar{E}^g F}{\rho_i G}.
+where $\rho_i$ is the density in g/cm$^3$ and $F_2$ is a conversion factor to convert grams to kilograms. The dose (in Gy) to the voxel is then
+\begin{equation}\label{eq:dosedep}
+D_i = \frac{\mathcal{E}_{T,i}}{M_i} = \frac{F_1}{\rho_i F_2} \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g-1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right].
 \end{equation}

### 2017-06-03 docprep/chapresults.tex
Updated the header
Minor fixes and added some figs

@@
 %%
-%% This is file `chapmin.tex',
-%% generated with the docstrip utility.
+%% Edward T. Norris
+%% Discrete Ordinates Computed Tomography Organ Dose Simulator (DOCTORS)
 %% 
-%% The original source files were:
+%% === Results ===
 %%
-%% ths.dtx  (with options: `chapmin')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from chapmin.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file ths.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
 
@@
 
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/bench_cone1_centerdiff3}
+  \end{center}
+  \caption{The 19 group center flux comparison}
+\label{fig:bench_cone1_centerdiff3}
+\end{figure}
+
+A plot is below.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/bench_cone1_centerdiff47}
+  \end{center}
+  \caption{The 47 group center flux comparison}
+\label{fig:bench_cone1_centerdiff47}
+\end{figure}
+
+The analytical point is computed as
+\begin{equation}
+\frac{d\varphi}{dE} = \frac{e^{-(\mu^g_a x_a + \mu^g_w x_w)}}{4\pi G (x_a + x_w)^2 (\Delta E)^g}
+\end{equation}
+where $\mu_a^g$ and $\mu_w^g$ are the attenuation coefficients for air and water repectively for the $g$ group, $x_a$ and $x_w$ are the pathlength through air and water respectively, and $\Delta E$ is the width of the $g$ group in MeV.
+
 \section{3 Cone-Beam Water Benchmark}
@@
 
+\section{GPU Acceleration}
+
+it's faster!
+
 \endinput

### 2017-06-05 doctors/solver.cpp
Reduced memory overhead in the legendre solver

@@ -614,8 +614,8 @@ void Solver::gsSolverIsoCPU(const Quadrature *quad, const Mesh *mesh, const XSec
         for(unsigned int i = 0; i < totalSource.size(); i++)
             totalSource[i] = 0;
 
-        for(unsigned int i = 0; i < tempFlux.size(); i++)
-            tempFlux[i] = 0.0f;
+        //for(unsigned int i = 0; i < tempFlux.size(); i++)
+        //    tempFlux[i] = 0.0f;
 
         // downscatter from the collided flux NOT including group inscatter
         for(unsigned int iie = highestEnergy; iie < ie; iie++)
@@ -641,8 +641,8 @@ void Solver::gsSolverIsoCPU(const Quadrature *quad, const Mesh *mesh, const XSec
 
         while(iterNum <= maxIterations && maxDiff > epsilon && totDiff/totDiffPre < 1.0)  // while not converged
         {
-            for(unsigned int i = 0; i < tempFlux.size(); i++)
-                tempFlux[i] = 0.0f;
+            //for(unsigned int i = 0; i < tempFlux.size(); i++)
+            //    tempFlux[i] = 0.0f;
 
             // Clear for a new sweep
             for(unsigned int i = 0; i < tempFlux.size(); i++)
@@ -951,14 +951,14 @@ void Solver::gsSolverLegendreCPU(const Quadrature *quad, const Mesh *mesh, const
     legendre.precompute(quad, solPar->pn);
 
 
-    std::vector<SOL_T> *angularFlux = new std::vector<SOL_T>(xs->groupCount() * quad->angleCount() * mesh->voxelCount());
+    std::vector<SOL_T> *cFlux = new std::vector<SOL_T>(xs->groupCount() * quad->angleCount() * mesh->voxelCount(), 0.0);
     std::vector<SOL_T> *scalarFlux = new std::vector<SOL_T>(xs->groupCount() * mesh->voxelCount(), 0.0f);
-    std::vector<SOL_T> tempFlux(mesh->voxelCount());
-    std::vector<SOL_T> preFlux(mesh->voxelCount(), -100.0f);
+    std::vector<SOL_T> tempScalarFlux(mesh->voxelCount());
+    //std::vector<SOL_T> preFlux(mesh->voxelCount(), -100.0f);
     std::vector<SOL_T> outboundFluxX(mesh->voxelCount(), -100.0f);
[…]

### 2017-06-05 docprep/norris_phd_dissertation.tex
Finished draft of the thesis through the results
Wrote the abstract

@@
 This is the text for the abstract. Here, I explain what I did and how cool I am for doing it. Because I am awesome.
+
+Computed tomography (CT) has become pervasive in medical diagnostics as improved imaging techniques and processing algorithms provide higher quality information to doctors.
+However, the exponentially increasing usage of CT has raised concerns regarding long term radiological risks.
+
 \end{ThesisAbstract}
@@
 His PhD work is currently funded by a fellowship through the NRC.
+Edward was awarded the College of Engineering and Computing PhD Scholar award in 2017 for his academic achievements.
 
@@
 
-Edward was awarded the College of Engineering and Computing PhD Scholar award in 2017 for his academic achievements.
-
 Edward was also an active member of the Nuclear Science Design Team at Missouri University of Science and Technology.
 While on the team he solved radiation transport problems to facilitate the dosimetry of an inertial electrostatic confintement fusion device being constructed.
-
-After completing his PhD, he plans to take a one year postdoctorial fellowship position with the NNSA.
 

### 2017-06-06 docprep/chapbib.tex
Updates to thesis

@@
 
-One of the most well known MOC codes id DeCART.
+One of the most well known MOC codes is DeCART.
 DeCART is actually a coupled solver in that it utilizes different solution modalities for the radial plane and the axial direction.
@@
 This methodology has been extended to fan beams [CITE] and cone beams [CITE].
-After the filtered backprojection, iterative reconstruction techniques can be applied to greatly improve the reconstruction quality [CITE].
+Besides the filtered backprojection (FBP) method, iterative reconstruction techniques can also be applied to greatly improve the image reconstruction quality [CITE].
 
@@
 \section{Discrete Ordinates}\label{sec:discordlit}
-The dsicrete ordinates method dates back to CITE but remains one of the most prominent solution modalities for ration transport in use today.
+The dsicrete ordinates method dates back to CITE but remains one of the most prominent solution modalities for radiation transport in use today.
 A comprehensive review of discrete ordinate methods is given by \citet{ref:lewise}.

### 2017-06-06 docprep/chapimpl.tex
Updates to thesis

@@
 
-The code is available from the GitHub repository. At the time of this writing, the repository is located at \texttt{https://github.com/eNorris/thesis}.
+The code is available from a GitHub Git repository.
+At the time of this writing, the repository is located at \texttt{https://github.com/eNorris/thesis} and is only available subject to special request and license agreement.
+An executable version is available.
 

### 2017-06-06 docprep/chapintro.tex
Updates to thesis

@@
 The dose-length-product (DLP) is used as a measure to quantify the effective dose a patient receives.
-However, the DLP must be modified by a factor accounting for may variations amongst patients such as age, gender, and size as well as the kVp used in the procedure.
+However, the DLP must be modified by a factor accounting for many variations amongst patients such as age, gender, and size as well as the kVp used in the procedure.
 Currently, dose estimation relies on \textit{a priori} computation verified with a standardized benchmark.

### 2017-06-06 docprep/chapmeth.tex
Updates to thesis

@@
 The weighting function can be any of a number commonly used functions.
-The SCALE data libraries us a uniform function [CITE].
+The SCALE data libraries use a uniform function [CITE].
 \begin{equation}\label{eq:groupxs}
@@
 Those cross sections are designed for light water reactor analysis and are not well suited to medical physics applications.
-This is currently one of the greatest limitations of DOCOTRS.
+This is one of the limitations of the current version of DOCOTRS.
 The SCALE cross section data also contains neutron data which is useless for the CT dosimetry DOCTORS performs.
@@
 
-The problem domain is split into an evenly spaced Cartesian grid.
-Uniform grid spacing is not required by discrete ordinate methods though it is helpful [CITE], but is necessarily the case with CT voxel phantoms.
+The problem domain is split into an evenly spaced Cartesian grid. Uniform grid spacing is not required by discrete ordinate methods though it is helpful and can speed up the computation[CITE].
 
@@
 \begin{equation} \label{eq:deriv_1}
-\frac{\partial f(\zeta)}{\partial \zeta} \approx \frac{f(\zeta+\Delta \zeta/2) - f(\zeta + \Delta \zeta/2)}{\Delta \zeta}
+\frac{\partial f(\zeta)}{\partial \zeta} \approx \frac{f(\zeta+\Delta \zeta/2) - f(\zeta - \Delta \zeta/2)}{\Delta \zeta}
 \end{equation}
@@
 The polar angle of an arbitrary vector $\boldsymbol{\xi}$ can be computed using \begin{equation}\label{eq:thetacos}
-\cos(\theta_\xi) = \frac{\xi_z}{\sqrt{\xi_z^2 + \xi_y^2 + \xi_z^2}}.
+\cos(\theta_\xi) = \frac{\xi_z}{\sqrt{\xi_x^2 + \xi_y^2 + \xi_z^2}}.
 \end{equation}
@@
 where $\bar{E}$ is the average energy (in MeV) of the $g^{th}$ group and $F_1$ is a unit conversion factor that converts MeV to Joules.
+The average group energy is assumed to be the average between the two bounding energies of the group.
 Equation~\ref{eq:depa} assumes total local energy deposition in the voxel of interest.
@@
 \begin{equation}\label{eq:dept2}
-\mathcal{E}_{T,i} = V_i F_1 \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g-1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right]
+\mathcal{E}_{T,i} = V_i F_1 \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g+1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right]
 \end{equation}
@@
 \begin{equation}\label{eq:dosedep}
-D_i = \frac{\mathcal{E}_{T,i}}{M_i} = \frac{F_1}{\rho_i F_2} \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g-1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right].
+D_i = \frac{\mathcal{E}_{T,i}}{M_i} = \frac{F_1}{\rho_i F_2} \sum_{g=0}^{G-1} \varphi_i^g \left[\Sigma_{a,i}^g \bar{E}^g + \sum_{h=g+1}^{G-1} \left\{ \Sigma_{s,i}^{g,h}  (\bar{E}^g-\bar{E}^h) \right\} \right].
 \end{equation}

### 2017-06-06 docprep/norris_phd_dissertation.tex
Updates to thesis

@@
 Computed tomography (CT) has become pervasive in medical diagnostics as improved imaging techniques and processing algorithms provide higher quality information to doctors.
-However, the exponentially increasing usage of CT has raised concerns regarding long term radiological risks.
+However, the exponentially increasing usage of CT has raised concerns regarding long term low-dose radiological risks.
+Currently, the dose to patients is computed using Monte Carlo methods and experimental tests.
+In other areas of radiation transport, deterministic codes have been shown to be much faster than Monte Carlo codes.
+
+Currently, no deterministic methodology exists to automatically generate a spatially distributed dose profile from a CT voxel phantom.
+This work proposes a new code, Discrete Ordinate CT Organ Dose Simulator (DOCTORS) which utilizes a GPU accelerated raytracer and discrete ordinate solver to compute photon flux in the patient.
+The flux is then converted to dose.
+
+The DOCTORS code was benchmarked against MCNP6 and found to have very good agreement using both a water phantom and a realistic patient phantom.
+DOCTORS was also found to be much faster than MCNP6, even more so with GPU acceleration.
+
+GPU
 

### 2017-06-08 python/ct_data_reader/main/meshsplitter.py
Added tally notification

@@ -1,6 +1,6 @@
 __author__ = 'etnc6d'
 
-filename = '/media/Storage/thesis/mcnp.gitignore/meshtax'
+filename = '/media/Storage/thesis/mcnp.gitignore/meshtaz'
 
 file = open(filename, 'r')
 #with open(filename, 'r') as file:
@@ -13,6 +13,10 @@ lines = file.readlines()
 
 for line in lines:
     l = line.lstrip()
+
+    if l.startswith("Mesh Tally Number"):
+        print(l)
+
     if l.startswith("Energy"):
         fout.close()
         indx += 1

### 2017-06-08 docprep/appdx_a.tex
Include a appendix label
Updated thesis text

@@
 %%
-%% This is file `chapmin.tex',
-%% generated with the docstrip utility.
+%% Edward T. Norris
+%% Discrete Ordinates Computed Tomography Organ Dose Simulator (DOCTORS)
 %% 
-%% The original source files were:
+%% === Appendix A ===
 %%
-%% ths.dtx  (with options: `chapmin')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from chapmin.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file ths.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
 
- %% ... sample chapter ...
+This appendix gives a brief summary of flux moments and the Legendre spherical harmonics needed to compute them.
 
-This appendix covers the math for Legendre polynomials and spherical harmonics
+\section{Flux Moments}\label{appdx:moments}
 
-\section{Legendre Polynomials}
+The flux moments are an alternate representation of the angular flux  within a system.
+The rationale behind usage of flux moments is that they map the discrete flux ($\psi$) from a function of angle to a function of $P_N$ expansion.
+This reduces the amount of memory required for computation.
+Storing $\psi$ directly requires storing $G \times N_V \times N_a$ values where $G$ is the number of groups, $N_V$ is the number of voxels, and $N_a$ is the number of angles.
+Storing the flux moments, however, requires storing only $G \times N_V \times (N^2 + 2N + 1)$ values where $N$ is the $P_N$ expansion value.
+For example, a problem requiring $S_6$ needs $6 \times 8 = 48$ directions, but a problem requiring $P_6$ needs 49 expansion coefficients which would seem approximately eqivalent, but typical problems require far more discrete directions than expansion coefficients.
+Comples problems can easily require $S_{16}$ or higher but expansions above $P_5$ are rarely encountered.
 
-They exist.
+\section{Legendre Polynomials}\label{appdx:leg}
 
-\subsection{Spherical Harmonics}
+The Legendre polynomials are solutions to:
+\begin{equation}
+\begin{split}
+P_0(\mu) &= 1 \\
+P_l(\mu) &= \frac{1}{2^l l!} \frac{d^l}{d\mu^l}(\mu^2 - 1)^l \,, \quad l \in \mathbb{N}
+\end{split}
+\end{equation}
+where $\mathbb{N}$ is the set of natural numbers (1, 2, 3...).
+Table~\ref{tab:legendre} gives the solution of the first few Legendre polynomials.
+The Legendre polynomials are orthogonal on the domain $[-1, 1]$ in that
+\begin{equation}
+\frac{1}{2} \int_{-1}^{1} P_l(\mu) P_{l'}(\mu) d\mu = 
+\begin{cases}
+\frac{1}{2l+1} \,, \quad l = l'\\
+0 \,, \quad \text{otherwise}
+\end{cases}
+\end{equation}
 
-They also exist.
+\begin{table}[ht]
+\caption{Legendre Polynomials}
+\centering 
+\begin{tabular}{c c}
+\hline \hline   
+$P_l$    & $P_l(\mu)$ \\ [0.5ex] 
+\hline
+$P_0$ & 1 \\
+$P_1$ & $\mu$ \\ 
+$P_2$ & $\frac{1}{2}(3\mu^2 - 1)$ \\
+$P_3$ & $\frac{1}{2}(5\mu^3 - 3\mu)$ \\
+$P_4$ & $\frac{1}{8}(35\mu^4 - 30x^2 + 3)$ \\
+$P_5$ & $\frac{1}{8}(63\mu^5 - 70\mu^3 + 15\mu)$ \\
+$P_6$ & $\frac{1}{16}(231\mu^6 - 315^4 + 105^2 - 5)$ \\
+$P_7$ & $\frac{1}{16}(429\mu^7 + 693\mu^5 - 315^3 + 35^3\mu)$ \\[1ex]
+\hline
+\end{tabular}
+\label{tab:legendre}
+\end{table}
 
-The real harmonics. The even/odd harmonics.
+The orthogonality of the Legendre polynomials allows an infinite series to exactly represent any function on the domain $[-1, 1]$ as
+\begin{equation}
+f(\mu) = \sum_{l=0}^{\infty} C_l P_l(\mu)
+\end{equation}
+with apprpriately selected constants, $C_l$ and is used to approximate functions arbitrarily
+\begin{equation}
+f(\mu) \approx \sum_{l=0}^{N} C_l P_l(\mu).
+\end{equation}
+The Legendre polynomials are extended into the associated Legendre polynomials which have an additional orthogonality.
 
-The addition theorem.
+\section{Associated Legendre Polynomials}\label{appdx:assoc}
+
+The associated Legendre polynomials are defined as the solution to
+\begin{equation}
+\begin{split}
+P_l^m(\mu) &= (-1)^m (1-\mu^2)^{m/2} \frac{d^m}{d\mu^m}P_l(\mu) \\
+P_l^0(\mu) &= P_l(\mu)
+\end{split}
+\end{equation}
+and exhibit the following orthogonality on $[-1, 1]$:
+\begin{equation}
+\frac{1}{2} \int_{-1}^{1} P_l^m(\mu) P_{}^{}(\mu) d\mu = 
+\begin{cases}
+\frac{1}{2l+1} \frac{(l+m)!}{(l-m)!} \,, \quad l=l' \text{ and } m=m'\\
+0\,, \quad \text{otherwise}.
+\end{cases}
+\end{equation}
+The spherical harmonics take advantage of the double orthogonality to approximate 2D distributions.
+
+\section{Spherical Harmonics}\label{appdx:spherical}
+
+The spherical harmonics are defined as:
+\begin{equation}
+Y_{lm}(\hat{\Omega}) = \sqrt{\frac{(2l+1)(l-m)!}{(l+m)!}}P_l^m(\mu)e^{e \tau m}
+\end{equation}
+where $\mu$ is the cosine of the angle between $\hat{Omega}$ and the $x$-axis and $\tau$ is the rotation about the $x$-axis with respect to the $y$-axis of $\hat{\Omega}$ projected onto the $yz$ plane which can be compuated as:
+\begin{equation}
+\tau = \frac{\eta}{\sqrt{\eta^2 + \xi^2}}.
+\end{equation}
+The spherical harmonics have the orghogonality
+\begin{equation}
+\int_{}^{}Y_{lm}(\hat{Omega})Y_{l'm'}^*(\hat{\Omega}) d\hat{\Omega}= 
+\begin{cases}
+1 \,, \quad l = l' \text{ and } m = m' \\
+0 \,, \quad \text{otherwise}
+\end{cases}
+\end{equation}
+where $Y_{lm}^*$ is the complex conjugate of $Y_{lm}$. Using the addition theorem which states
+\begin{equation}
+P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2l+1}\sum_{m=-l}^{l}Y_{lm}^*(\hat{\Omega}') Y_{lm}(\hat{\Omega})
+\end{equation}
+gives
+\begin{equation}
+P_l(\hat{\Omega} \cdot \hat{\Omega}') = P_l(\mu)P_l(\mu') + 2\sum_{m=1}^{l}\frac{(l-m)!}{(l+m)!}P_l^m(\mu) P_l^m(\mu') cos(m[\tau - \tau']).
+\end{equation}
 

### 2017-06-08 docprep/chapbib.tex
Updated thesis text

@@
 However, dose quantification is not currently patient specific.
-The air kerma (measured in Gy) measures the x-ray beam intensity and can be measured directly [CITE].
+Measuring the dose in a patient directly is impossible, intead the air kerma (measured in Gy), which measures the x-ray beam intensity, and can be measured directly [CITE].
 As the x-ray beam passes through the patient, it deposits energy.
@@
 
+INSERT MORE EXAMPLES.
 \citet{ref:lauk} found that severly obese patients received twice as much dose from CT procedures as others on average. Some patients received as much as five times the typical dose.
@@
 
+Nearly all transport solutions fall into one of two categories: Monte Carlo and deterministic.
+Monte Carlo codes are generally known for their accuracy; they are able to sample the problem and attain arbitrarily high precision given sufficient runtime.
+They also take advantage of combinatorial geometry allowing a user to produce geometrically complex 3D structures.
+Deterministic methods are generally much faster than their Monte Carlo counterparts but do not benefit from excess runtime.
+They also consume vastly more resources, particularly random-access memory (RAM).
+However, in regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
+A large number of spatial regions slows down the particle transport in Monte Carlo and increases the RAM required to store the numerous voxels.
+
+Codes such as EGS4, its improved versions EGSnrc, and EGS5 \citep{ref:nelsonw}, and PENELOPE \citep{ref:salvatf} are early Monte Carlo that have shown very good comparison to experimental results.
+However, these codes were developed primarily for shower simulations in high energy physics.
+
+Monte Carlo codes such as TOPAS [CITE], Geant4 \citep{ref:agostinellis}, and MCNP6 [CITE] are the current gold standard in computational dosimetry \citep{ref:jiax}.
+This work compares results to those obtained from MCNP6.
+MCNP6 is a Monte Carlo code developed for general purpose particle transport and is well validated.
+MCNP6 was selected for comparison because it is so well validated and the input files can be easily generated procedurally.
+Codes whose input such as Geant4 do not lend themselves easily to procedural code generation are difficult to benchmark precisely against for arbitrary computational phantoms.
+
+The alternative to Monte Carlo is deterministic solutions.
+The best known solution techniques include the discrete ordinate method, the finite element method, and the method of characteristics.
+The discrete ordinate method is employed by DOCTORS and the methodology is covered independently in Section~\ref{sec:discordlit}.
+Other deterministic methods are summarized here.
+
+The method of characteristics solves the characteristic form of the Boltzmann transport equation which is defined along a single direction (called the characteristic).
+Along that direction, the 3-dimensional gradient operator found in the full Boltzmann equation (see Sec.~\ref{sec:lbe}) becomes a 1-dimensional derivitive.
+This reduction gives rise to an analytical solution along the characteristic [CITE].
+Sufficient characteristic solutions solved simulateously gives the solution over the entire problem domain.
+Currently, though, MOC has not been extended to full 3D, instead, hybrid methods such as the NEM-$S_N$ technique employed by DeCART have been used.
+DeCART uses 2D MOC to solve planar slices of reactor transport problems and links the slices using diffusion~\citep{ref:hursinm}.
+Another MOC implementation is the TRANSMED code which was used to compute external beam therapy dose profiles accurately~\citep{ref:williamsm}.
+Method of characteristic solutions have been extended to the time dependent domain by \citet{ref:hoffmana}.
+
+However the energy dependent flux is computed, it must afterward be converted to dose.
 Monte Carlo codes have been used in the past to compute flux-to-dose conversion factors for a reference person~\citep{ref:icrp116}.
@@
 where $\varphi$ is the scalar flux.
+The weakness of this method is that even though the flux may be computed to be patient specific, the weighting factors are evaluated using a standard reference man phantom.
 
-Monte Carlo codes are generally known for their accuracy; they are able to sample the problem and attain arbitrarily high precision given sufficient runtime.
-They also take advantage of combinatorial geometry allowing a user to produce geometrically complex 3D structures.
-Deterministic methods are generally much faster than their Monte Carlo counterparts but do not benefit from excess runtime.
-They also consume vastly more resources, particularly random-access memory (RAM).
-However, in regularly structures meshes, deterministic methods can greatly outperform Monte Carlo.
-A large number of spatial regions slows down the particle transport in Monte Carlo and increases the RAM required to store the numerous voxels.
-
-Monte Carlo codes such as TOPAS and Geant4 \citep{ref:agostinellis} are the gold standard \citep{ref:jiax}
-
-Codes such as EGS4, its improved versions EGSnrc, and EGS5 \citep{ref:nelsonw}, and PENELOPE \citep{ref:salvatf} are all Monte Carlo and have shown very good results to experimental.
-But they are developed for high energy physics.
-
-The alternative to Monte Carlo is deterministic solutions.
-The best known solution techniques include the discrete ordinate method, the finite element method, and the method of characteristics.
-The discrete ordinate method is employed by DOCTORS and the methodology is covered independently in Section~\ref{sec:discordlit}.
-
-One of the most well known MOC codes is DeCART.
-DeCART is actually a coupled solver in that it utilizes different solution modalities for the radial plane and the axial direction.
-It uses a more complex modality in the radial plane such as MOC and then uses diffusion to transport particles axially~\citep{ref:hursinm}.
-
-The Method of Characteristics is a deterministic solution that has some advantages over discrete ordinates but is very complex to set up in 3D.
-One implementation is the TRANSMED code which was used to compute external beam therapy dose profiles accurately~\citep{ref:williamsm}.
-
-The MOC solution has been extended to the time dependent domain by \citet{ref:hoffmana}.
-
-Raytracing algorithms have been developed for numerous discret ordinate codes include RAY3D~\citep{ref:yingz} and ATTILLA~\citep{ref:wareingt} which utilizes unstructured tetrahedral meshes.
-
-Dose estimations have been made using many methodologies including
-
-COMPUTATION
+Alternatively, codes can compute the dose deposition directly during the transport solution.
+The simplest example of this is in Monte Carlo codes which tally the energy deposited by each collision rather than track length in voxels which estimates the absorbed dose.
+Since the particle type that caused the scatter is always, known, the equivalent dose is simple to compute.
+However, in order to compute the effective dose, which is organ weighted, the organ to which the voxel of interest belongs must be known.
+In general, this is a very difficult problem with no fully automated solutions without significant additional information about human anatomy and the scan procedure employed.
+Therefore, this dose estimation technique is rarely employed in practice.
 
@@
 These values are related to the attenuation coefficient of the material represented by the voxel.
-Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air CITE though some variations do exist \citep{ref:plessisf} \citep{ref:sawc}.
+Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air [CITE] though some variations do exist \citep{ref:plessisf} \citep{ref:sawc}.
 
@@
 \section{Discrete Ordinates}\label{sec:discordlit}
-The dsicrete ordinates method dates back to CITE but remains one of the most prominent solution modalities for radiation transport in use today.
+The dsicrete ordinates method dates back to [CITE] but remains one of the most prominent solution modalities for radiation transport in use today.
 A comprehensive review of discrete ordinate methods is given by \citet{ref:lewise}.
@@
 The oldest major discrete ordinate method implementation is DORT which was superceded by its 3D counterpart TORT~\citep{ref:rhoadesw}.
-Since TORT, new discrete ordinate method implementations have laregly been extensions allowing more advanced computation.
+Since TORT, new discrete ordinate method implementations have laregly been extensions allowing more advanced computation or updates to modernize the code.
 
@@
 
-KBA DISCUSSION
+To overcome ray effect artifacts (discussed more thouroughly in Sec.~\ref{sec:uncol}), discrete ordinate solutions often employ a raytracing algorithm to compute the uncollided flux.
+The uncollided flux is then computed using the full discrete ordinate method.
+Raytracing algorithms have been developed for numerous discret ordinate codes include RAY3D~\citep{ref:yingz} and ATTILLA~\citep{ref:wareingt} which utilizes unstructured tetrahedral meshes.
+
+INSERT KBA DISCUSSION
 
@@
 Many raytracing algorithms are based on the Bresenham line raserizing algorithm which was designed for graph plotting hardware~citep{ref:bresenhamj}.
-That algorithm was intende to produce lines acceptable for human interpretation and no guarantee is made that all pixels along the line will be identified making this class of algorithms inappropriate for dosimetric raytracing since attenuation through some voxels would be skipped.
+That algorithm was intended to produce lines acceptable for human interpretation and no guarantee is made that all pixels along the line will be identified making this class of algorithms inappropriate for dosimetric raytracing since attenuation through some voxels would be skipped.
 Some extensions have been added to it that guarantee passage through all voxels though~\citep{ref:liuy}.
 
-\citet{ref:clearyj} proposed a voxel traversal algorithm in which they initially identify the voxel containing the starting point of the ray, and transport it to the first surface of its bounding voxel it would cross.
-This is done by computing the distance traveled in order to cross each enclosing surface.
+\citet{ref:clearyj} proposed a voxel traversal algorithm in which he initially identifies the voxel containing the starting point of the ray, and transport it to the first surface of its bounding voxel it would cross.
+This is done by computing the distance traveled to cross each enclosing surface.
 The surface with the least distance is the first surface crossed.
@@
 This work uses the raytracing algorithm proposed by~\citet{ref:wooa} which is very similar to the method proposed by~\citet{ref:clearyj} but based on the parametric equation of a line, $\boldsymbol{u} + \boldsymbol{v}t$ which makes implementation in 3D simple.
-Since its develpment, many improvements and extensions have been made for alternate geometries such as rhombic dodecahedra~\citep{ref:hel}
+Since its develpment, many improvements and extensions have been made for alternate geometries such as rhombic dodecahedra~\citep{ref:hel} for volume rendering but the core algorithm remains state of the art for floating-point arithmetic.
+On systems that are highly efficient at computing integer arithmetic, alternatives exist which are faster~\citep{ref:liuy} but they do not outperform on hardware optimized for floating-point operations.
 
-GPU RAYTRACING
+INSERT SECTION ON GPU RAYTRACING
 
 \section{GPU Acceleration}
+
+A graphics processing unit (GPU) is similar to a central processing unit (CPU)in that it executes instructions to process data.
+However, it differs in that a CPU may have up up to 22 physical cores, each of which can operate independently, a GPU has up to thousands of cores, but they cannot operate independently.
+Instead, cores execute simulaneously in groups called warps.
+All cores in a warp execute the same instruction at the same time as all others.
 
@@
 Conceptually, code run on the device is split into block on a grid, each block then runs multiple concurrent computations, each on a thread.
-In hardware, threads execute on 32-128 streaming multiprocessors depending on the GPU architecture.
+In hardware, threads execute in a warp (32 to 128 streaming multiprocessors depending on the GPU architecture).
 

### 2017-06-08 docprep/chapconc.tex
Updated thesis text

@@
 
-I concluded many things.
+The raytracer is very fast compared to the discrete ordinate solution which computes the collided flux.
+The uncollided flux quantifies the source distribution from the medical system.
+This uniquely enables discrete ordinate solutions to rapidly characterize complex systems.
+Multiple cone beams and fan beams are possible as well as many other, more complex beam shapes.
+
+Anisotropy cannot be ignored.
 
@@
 \subsection{Partial Acceptance Criteria}
-Allow it in the raytracer.
+In the current version of doctors, the raytracer is very accurate with respect to MCNP6 in predicting the uncollided flux, except for along the periphery of the beam.
+Voxels are currently either completely inside the beam or completely outside of it as determined by its isocenter.
+A more sophisticated technique whereby a voxel cah be partially accepted in the beam would remove these artifacts.
 
 \subsection{Multileaf Collimation}
-Collimators!
+Some medical systems currently utilize a multileaf collimator [CITE] to deliver dose to a specific organ or region of the body. As the
 

### 2017-06-08 docprep/chapimpl.tex
Updated thesis text

@@
 This chapter covers the details of the implementation of the different components of the DOCTORS code base.
-
 The code is available from a GitHub Git repository.
@@
 
-Data is stored as \texttt{std::vector<>} objects.
+Data is stored in large, multi-dimensional arrays.
+For example, the CT voxel phantom can be intuitively stored as a 3D array of floating precision values which can be easily indexed.
+However, DOCTORS is written such that it is extensible to GPUs.
+GPUs are not optimized for data stored in a multi-dimensional fashion, but rather for data stored as a single 1-dimensional array.
+To emulate the 3-dimensional storage arrangement of the data, indexing arithmetic is used.
+
+On the CPU, data is stored as C++ \texttt{std::vector<T>} objects where the \texttt{T} template parameter can be either \texttt{float} or \texttt{double} depending on the precision (32 bits or 64 bits respectively) needed.
+The storage space available to a \texttt{std::vector<T>} object is effectively unlimited, bounded only by the host CPU's physical RAM.
+However, other effects such as memory fragmentation can limit the practical size of a single, continuous \texttt{std::vector<T>} object significantly.
+To emulate indexing of a $N_x \times N_y$ matrix, a single \texttt{std::vector<T>} is created and resized to $N_x \times N_y$ elements.
+The global index, $i$, in the flattened array is then
+\begin{equation}
+i = i_x N_y + i_y.
+\end{equation}
+This pattern extends to multi-dimensional matrices. For example, a $N_x \times N_y$ matrix would be globally indexed as
+\begin{equation}
+i = i_x N_y N_z + i_y N_z + i_z.
+\end{equation}
 
@@
   \end{center}
-  \caption{The indexing scheme.}
+  \caption{Conversion from 3D indexing to 1D "flattened" indexing.}
 \label{fig:indx_ex}
@@
 AMPX cross sections are stored in a binary format.
-The binary file is split into a series of records.
-Each record contains a string of binary bytes between a four byte header and four byte footer.
-The header and footer are identical and, when interpreted as a signed four byte integer, give the size in words of the record.
-Each word is a 32-bit section (4~bytes) of binary data.
+Figure~\ref{fig:ampx} shows the data format of the overall file.
+The binary file is split into a sequence of records.
+The header section of the file has overall information about the file including the number of nuclides stored, their energy group structure, and a comment string describing file.
+Following the header section is a list of directory records.
+Each directory has information about a specific nuclide including the cross section reaction types available, temperatures for which the cross sections were evaluated, the number of Legendre expansion coefficients stored in the scatter matrix, and other data.
+Following the directories, more records follow that describe the energy structure used in the file.
+There is one such record for each particle type.
+The libraries used in the current version of DOCTORS have both neutron and photon cross section data stored.
+
+The file header section, directories, and energy group boundaries describe the structural information necessary to read the following nuclide records.
+Nuclides are listed in the same order their directories are found after the header setion.
+Each nuclide contains numerous records.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/ampx1}
+        \caption{}
+        \label{fig:ampx1}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/ampx2}
+        \caption{}
+        \label{fig:ampx2}
+    \end{subfigure}
+    \caption{The format of the AMPX file.
+(a) The file contains a header, a list of directory records, energy group information and a list of nuclide entries.
+(b) Each nuclide entry contains multiple records; the two of concern in this work are the last two containing gamma data.}\label{fig:ampx}
+\end{figure}
+
+Each record contains a string of data bytes between a four byte header and four byte footer.
+The header and footer are identical and, when interpreted as a signed four byte integer, give the size in bytes of the record.
 An example record is shown in Fig.~\ref{fig:ampxbytes1}.
@@
 The endianess rearrangement is shown in Fig.~\ref{fig:ampxbytes2}.
-All records are one of 12 unique types.
-Each type of record is formatted and interpreted differently.
+Each record is one of 12 unique types; each type of record is formatted and interpreted differently.
 For example, Type 1 contains general information about the data file such as the number of nuclides, number of energy groups, and a brief description.
@@
 
-Each word of data is sequentially read, reordered, and converted to an apprpriately typed variable. Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.
+Each byte of data is sequentially read, reordered, and converted to an apprpriately typed variable. Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.
 
@@
   \end{center}
-  \caption{An example record parsed. The header and footer are identical and are the number of 32-bit words in the record.}
+  \caption{An example record parsed. The header and footer are identical and equal to the number of bytes in the record.}
 \label{fig:ampxbytes1}
@@
   \end{center}
-  \caption{The byte reordering from Big Endian to Little Endian. Individual bits withing a byte are not rearranged, but the ordering of the four bytes withing a 32-bit word is reversed.}
+  \caption{The byte reordering from Big Endian to Little Endian. Individual bits within a byte are not rearranged, but the ordering of the four bytes withing a 32-bit word is reversed.}
 \label{fig:ampxbytes2}
@@
 
-DOCTORS reads AMPX fromatted binary data files. Figure~\ref{fig:ampx} shows the data format of the overall file.
-
 The first entry in each nuclide entry is the directory record.
@@
 The 2D data is stored in a special format optimized for scatter matrix data.
+The format concists of a sequence of "magic numbers" interpersed within a list of data points.
+A magic number is a nine digit number IIIJJJKKK.
+The first three digits, III, are the group number of the highest energy group to scatter to the energy of interest.
+The next three digits, JJJ, are the group number of the lowest energy to scatter.
+The final three digits, KKK, are the group number of the sink energy to which particles are scattering.
+Given a magic number \texttt{magic}, the three values can be computed in Listing~\ref{lst:magic}.
+Note that the group numbers are indexed from 1 to $G$ instead of 0 to $G-1$ as required by DOCTORS.
+This is corrected by subtracting 1 while doing the indexing arithmetic.
 
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/ampx1}
-        \caption{The AMPX file format}
-        \label{fig:ampx1}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/ampx2}
-        \caption{The structure of each individual record.}
-        \label{fig:ampx2}
-    \end{subfigure}
-    \caption{The format of the AMPX file.
-(a) The file contains a header, a list of directory records, energy group information and a list of nuclide entries.
-(b) Each nuclide entry contains multiple records; the two of concern in this work are the last two containing gamma data.}\label{fig:ampx}
-\end{figure}
+\begin{listing}
+\begin{minted}[frame=lines,linenos]{c++}
+# Given magic
+KKK = magic % 1000
+JJJ = (magic % 1000000 - KKK) /1000
+III = (magic - JJJ - KKK) / 1000000
 
-Figure~\ref{fig:sigmacomp} shows the microscopic cross section data pulled from
+src = JJJ
+while src >= III
+	data = READ_NEXT_BINARY_FLOAT()
+	xs[(src - 1)*G + KKK - 1] = data
+	src = src - 1
+\end{minted}
+\caption{Computation of the source and sink groups from the magic number and the subsequent data parsing.}\label{lst:magic}
+\end{listing}
+
+Figure~\ref{fig:sigmacomp} shows the microscopic cross section data pulled from the 200-neutron/47-gamma group data file currently used by DOCTORS for hydrogen and oxygen.
+The data is compared to reference data pulled directly from the ENDF/B-VII.1 photoatomic (MT=501) data library~\citep{ref:cullend} accessible through the Sigma database~\citep{ref:sigma}.
 
@@
 Once the data is parsed, weighted combinations of elemental data is used to create material cross sections.
+Cross section data for photons always uses the naturally ocurring since photo-atomic reactions are not sensitive to the nuclear differences between isotopes.
+The compositions of N materials are assumed to be given as weight fractions, $w_i$ subject to
+\begin{equation}
+sum_{i = 1}^N w_i = 1
+\end{equation}
+and is converted to an atom fraction, $a_i$:
+\begin{equation}
+a_i = \frac{w_i}{M m_i}
+\end{equation}
+where
+\begin{equation}
+M = \sum_{i = 1}^N \frac{w_i}{m_i}
+\end{equation}
+and $m_i$ is the element's atomic weight.
 
-MT AND MF NUMBERS
+In the solution employed by DOCTORS, only the total and scattering cross sections are required.
+The AMPX data files however, support arbitrary reaction types and have up to a dozen or more reactions for photo-atomic reactions alone for elemental datasets.
+The reaction types are identified by their MT designation.
+The total, inelastic, and elastic scatter cross sections are MT 501, 502, and 504 respectively.
+The scatter cross section used by DOCTORS is the sum of the elastic and inelastic cross sections.
 
-WATER EXAMPLE.
+The cross section of compound materials are computed as a weighted summation of their components. For example, the cross section of water for a particular reaction would be
+\begin{equation}
+\sigma_{H_2 O} = \frac{2 \sigma_H + \sigma_O}{3}.
+\end{equation}
 
@@
 The Qt5 framework was used for implementation of the user interface.
-Qt5 enables asynchronous calls through its ssignal/slot mechanism.
+Qt5 enables asynchronous calls through its signal/slot mechanism.
+Signals and slots are special functions that have additional processing performed by the meta-object-compiler (MOC).
 A signal can be emitted which will execute all connected slots.
-Signals and slots can be connected manually by the user except for ones automatically generated by Qt.
-Qt prvides a user interface for building user interfaces.
+Signals and slots are connected manually by the user except special ones automatically generated by Qt5.
+
+Qt5 prvides a user interface for building user interfaces.
 Components such a sbuttons, drop boxes, radio buttons, etc.
-are offered to the user.
-Qt makes extensive use of polymorphism.
-All user components inherit from the base \texttt{QWidget} class which inherits from \texttt{QObject}.
-Any class that utilizes the signal/sloot mechanism must extend the \texttt{QOjbect} class.
+are made available to the user.
+Through extensive use of polymorphism, Qt simplifies the addition of graphical interactive elements called widgets.
+All widgets inherit from the base \texttt{QWidget} class which inherits from \texttt{QObject}.
+Any class that utilizes the signal/slot mechanism must extend the \texttt{QOjbect} class.
+As an example, a button can be created in the Qt5 user interface and named \texttt{button1}.
+This object will be accessible as \texttt{ui->button1}.
+When the user clicks on the button, the signal \texttt{button1.clicked()} will automatically be emitted.
+The user can connect this signal to any slot with an identical number arguments of the same type.
+The \texttt{button1.clicked()} signal can be connected to the slot \texttt{doStuff1()} but not \texttt{doStuff2(int)} since the arguments are not compatible.
 
@@
 \begin{listing}
-\begin{minted}[frame=lines,linenos]{python}
-import numpy as np
-def myfcn(hello,world):
-   return hello*world
+\begin{minted}[frame=lines,linenos]{c++}
+class MainWindow : public QWindow
+{
+	// Constructor
+	MainWindow(QObject *parent);
+
+	// Other parts of the class
+
+public slot:
+	void doSomethingLong();
+}
+
+void MainWindow::doSomethingLong()
+{
+	// Execute a long piece of code
+}
+
+MainWindow::MainWindow(QObject *parent)
+{
+	// Initializations
+
+	// When a button named button1 which was created in the Qt5 creator interface is clicked, the clicked() signal is automatically emitted which executes the doSomethingLong() function
+	MainWindow.connect(ui->button1, clicked(), this, doSomethingLong())
+}
 \end{minted}
-\caption{Blah blah blah.}\label{lst:sync1}
+\caption{A long function causes the user interface to block.}\label{lst:sync1}
 \end{listing}
@@
 
+\begin{listing}
+\begin{minted}[frame=lines,linenos]{c++}
+class MainWindow : public QWindow
+{
+	// Constructor
+	MainWindow(QObject *parent);
+
+	// Other parts of the class
+
+public slot:
+	void doSomethingLong();
+}
+
+void MainWindow::doSomethingLong()
+{
+	// Execute a long piece of code
+}
+
+MainWindow::MainWindow(QObject *parent)
+{
+	// Initializations
+
+	// When a button named button1 which was created in the Qt5 creator interface is clicked, the clicked() signal is automatically emitted which executes the doSomethingLong() function
+	MainWindow.connect(ui->button1, clicked(), this, doSomethingLong())
+}
+\end{minted}
+\caption{A long function causes the user interface to block.}\label{lst:sync1}
+\end{listing}
+
 \begin{figure}[tb]
@@
 
-The GUI is cool. Lots of pictures of it.
+INSERT PICTURES OF THE GUI
 

### 2017-06-08 docprep/chapintro.tex
Updated thesis text

@@
 Increased information results in faster and more accurate diagnosis.
-However, the exponentially increasing usage of CT and corresponding dose patients receive has raised concerns regarding potential long term risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}.
+However, the exponentially increasing usage of CT and dose patients receive from it has raised concerns of potential long term risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}.
 These concerns are exemplified by the push for low dose CT.
 
+INSERT MORE LOW DOSE CT STUFF
 A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.
@@
 No methodology currently exists to verify that the dosimetry evaluation was accurate after the patient has undergone the procedure.
-This work proposes patient specific a methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam.
-This can also help doctors estimate spatial dose distribution to the patient to ensure no specific organ received more dose than permissible.
+This work proposes a patient specific methodology by which a patient's CT reconstruction is used to compute the dose received from the radiation beam.
+This can also help doctors estimate spatial dose distribution in the patient to ensure no specific organ received more dose than permissible.
 
-This work implements and analyses a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient can be computed with a full transport solution inside a CT phantom.
+This work implements and analyzes a new computer code system called Discrete Ordinate CT ORgan dose Simulator (DOCTORS) by which the dose to a patient is computed using a full transport solution inside a CT phantom.
 The methodology converts a CT mesh into a voxel phantom of materials and densities.
@@
 To this end, a graphical user interface (GUI) was developed.
-The GUI utilizes the Qt5 graphics framework for the window, buttons, and other necessary widgets.
+The GUI was built using the Qt5 graphics framework for the windows, buttons, and other necessary widgets.
 The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicated required and completed steps.
@@
 
-BRIEF DISCUSSION OF RESULTS
+INSERT BRIEF DISCUSSION OF RESULTS
 

### 2017-06-08 docprep/chapmeth.tex
Updated thesis text

@@
 
-ADD THEORY AND MATH OF ANGLE/WEIGHT GENERATION?
+The $S_N$ quadrature ensures 90$\deg$ symmetry about any axis.
+In order to ensure the rotational symmetry, the $S_N$ quadrature has a single degree of freedom.
+The $LQ_N$ quadrature (which is the specific $S_N$ implemented in DOCTORS) uses the remaining degree of freedom to minimize error caused when integrating Legendre polynomials since that class of polynomials are used to represent anisotropy in scattering (see Sec.~\ref{sec:aniso}).
+For a more through discussion of quadrature generation, see \citet{ref:lewise}.
 
@@
 
-\subsection{Anisotropy Treatment}
+\subsection{Anisotropy Treatment}\label{sec:aniso}
 
@@
 \end{equation}
-
-%\begin{equation} \label{eq:col3}
-%\begin{split}
-%	&\left[ \hat{\Omega} \cdot \nabla + \Sigma_t(\boldsymbol{r}, E) \right]
-%	\psi_c(\boldsymbol{r}, E, \hat{\Omega}) = \\
-%	&\int_{4 \pi} \int_0^\infty \Sigma_s(\boldsymbol{r}, E' \rightarrow E, %\hat{\Omega}' \rightarrow \hat{\Omega}) \psi(\boldsymbol{r}, E', %\hat{\Omega}') dE' d\hat{\Omega}'
-%\end{split}
-%\end{equation}
 
@@
 
-The alternative way to compute the dose is directly from the local energy deposition. The total deposition, $\mathcal{E_{T,i}}$ for voxel $i$ is
+The alternative way to compute the dose is directly from the local energy deposition. The total deposition, $\mathcal{E}_{T,i}$ for voxel $i$ is
 \begin{equation}\label{eq:dept}
@@
 \end{equation} 
-where $\mathcal{E_{a,i}}$ is the energy deposition from absorption and $\mathcal{E_{s,i}}$ is the energy deposition from scatter. The absorption deposition is
+where $\mathcal{E}_{a,i}$ is the energy deposition from absorption and $\mathcal{E}_{s,i}$ is the energy deposition from scatter. The absorption deposition is
 \begin{equation}\label{eq:depa}
@@
 \begin{equation}\label{eq:deps}
-\mathcal{E}_{s,i} = \sum_{g=0}^{G-1} \sum_{h=g-1}^{G-1} \left[ \Sigma_{s,i}^{g,h} \varphi_i^g V_i (\bar{E}^g-\bar{E}^h) F_1 \right]
+\mathcal{E}_{s,i} = \sum_{g=0}^{G-1} \sum_{h=g+1}^{G-1} \left[ \Sigma_{s,i}^{g,h} \varphi_i^g V_i (\bar{E}^g-\bar{E}^h) F_1 \right]
 \end{equation}

### 2017-06-08 docprep/chapresults.tex
Minor changes in notation
Updates to thesis
Updated thesis text

@@
 Two computational phantoms were used for verification of the DOCTORS code.
+The first phantom is a $256 \times 256 \times 64$ voxel mesh of a 35 cm water phantom.
+The water phantom is in an acrylic casing and the entire mesh measures $50 \times 50 \times 12.5$ cm.
+Each voxel is then a cube with sides of 1.95 mm.
+The mesh has a total of 4194304 voxels.
+The second phantom is of the same dimensions and resolution, but of a realistic human phantom.
+The mesh emulates a CT scan of a patient's midsection where the liver would be located.
 
@@
 
-The analytical point is computed as
+The analytical points are computed
 \begin{equation}
-\frac{d\varphi}{dE} = \frac{e^{-(\mu^g_a x_a + \mu^g_w x_w)}}{4\pi G (x_a + x_w)^2 (\Delta E)^g}
+\left(\frac{d\varphi}{dE}\right)^g = \frac{e^{-(\mu^g_a x_a + \mu^g_w x_w)}}{4\pi G (x_a + x_w)^2 (\Delta E)^g}
 \end{equation}
 where $\mu_a^g$ and $\mu_w^g$ are the attenuation coefficients for air and water repectively for the $g$ group, $x_a$ and $x_w$ are the pathlength through air and water respectively, and $\Delta E$ is the width of the $g$ group in MeV.
+Dividing by the number of groups maintains consistency with the computational codes which uniformly distribute particles across all groups.
 
-\section{3 Cone-Beam Water Benchmark}
+In MCNP6, the effective dose was computed using a FMESH4 tally with associated DE and DF cards to weight the flux.
+The most accurate tally to compute the absorbed dose with is a F6 tally, however, those cannot be spatially distributed via a FMESH tally.
+Therefore, a F4 tally was used with a multiplier designed to emulate a F6 tally as closely as possible.
+The comparison is shown in Fig.~\ref{fig:mcnp_f4vsf6}.
 
-Now there are three
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/mcnp_f4vsf6}
+  \end{center}
+  \caption{Comparison of the absorbed dose computed with F4 and F6 tallies in MCNP6.
+All eight data points were located at the center of the phantom and the two computation methods were within 2\% of each other.}
+\label{fig:mcnp_f4vsf6}
+\end{figure}
+
+The effective dose and absorbed doses are given in Fig.~\ref{fig:b2_dose}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_effdose}
+        \caption{}
+        \label{fig:b2_effdose}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_absdose}
+        \caption{}
+        \label{fig:b2_absdose}
+    \end{subfigure}
+    \caption{(a) The effective dose computed using ICRP 116 fluence-to-dose conversion factors. (b) The absorbed dose.}\label{fig:b2_dose}
+\end{figure}
 
@@
 
-And now 16.
-
-\section{1 Cone-beam Phantom Benchmark}
-
-It's a person!
+INSERT THE RESULTS FROM THE 16 BEAM BENCHMARK
 
@@
 
-It's a person!
+INSERT THE RESULTS FROM THE 16 BEAM PHANTOM BENCHMARK
 
@@
 
-it's faster!
+INSERT THE RESULTS FROM THE GPU ACCELERATION
 

### 2017-06-09 docprep/chapimpl.tex
Updated the implementation chapter

@@
 This chapter covers the details of the implementation of the different components of the DOCTORS code base.
+
 The code is available from a GitHub Git repository.
@@
 
-Data is stored in large, multi-dimensional arrays.
-For example, the CT voxel phantom can be intuitively stored as a 3D array of floating precision values which can be easily indexed.
-However, DOCTORS is written such that it is extensible to GPUs.
-GPUs are not optimized for data stored in a multi-dimensional fashion, but rather for data stored as a single 1-dimensional array.
-To emulate the 3-dimensional storage arrangement of the data, indexing arithmetic is used.
-
-On the CPU, data is stored as C++ \texttt{std::vector<T>} objects where the \texttt{T} template parameter can be either \texttt{float} or \texttt{double} depending on the precision (32 bits or 64 bits respectively) needed.
-The storage space available to a \texttt{std::vector<T>} object is effectively unlimited, bounded only by the host CPU's physical RAM.
-However, other effects such as memory fragmentation can limit the practical size of a single, continuous \texttt{std::vector<T>} object significantly.
-To emulate indexing of a $N_x \times N_y$ matrix, a single \texttt{std::vector<T>} is created and resized to $N_x \times N_y$ elements.
-The global index, $i$, in the flattened array is then
-\begin{equation}
-i = i_x N_y + i_y.
-\end{equation}
-This pattern extends to multi-dimensional matrices. For example, a $N_x \times N_y$ matrix would be globally indexed as
-\begin{equation}
-i = i_x N_y N_z + i_y N_z + i_z.
-\end{equation}
+Data is stored as \texttt{std::vector<>} objects.
 
@@
   \end{center}
-  \caption{Conversion from 3D indexing to 1D "flattened" indexing.}
+  \caption{The indexing scheme.}
 \label{fig:indx_ex}

### 2017-06-12 doctors/ctdatamanager.cpp
Updates to solver and GUI

@@ -170,9 +170,14 @@ Mesh *CtDataManager::ctNumberToHumanPhantom(Mesh *mesh)
             qDebug() << "EXPLODE!";
         }
 
-        if(mesh->zoneId[i] == 0)
+        if(mesh->zoneId[i] == 0) // Air
         {
             mesh->density[i] = 0.001225f;  // force air density because it is very sensitive to miscalibrations
+        }
+
+        if(mesh->zoneId[i] == 19) // Aluminum
+        {
+            mesh->density[i] = 2.70f;
         }
 
         mesh->atomDensity[i] = mesh->density[i] * atomPerG[mesh->zoneId[i]] * 1.0E-24f;  // g/cc * @/g * cm^2/b = @/cm-b
@@ -182,6 +187,14 @@ Mesh *CtDataManager::ctNumberToHumanPhantom(Mesh *mesh)
             qDebug() << "Got zero density";
         }
     }
+
+    Histogram h(0, 70000, 10000);
+    for(int i = 0; i < mesh->voxelCount(); i++)
+        h.insert(mesh->ct[i]);
+    std::ofstream fout;
+    fout.open("./ctunits.dat");
+    fout << h;
+    qDebug() << "Finished histogram";
 
     return mesh;
 }

### 2017-06-12 doctors/doseconverter.h

@@ -121,10 +121,10 @@ std::vector<T> DoseConverter::doseEdep(const XSection &xs, const Mesh &mesh, con
 
     for(unsigned int ir = 0; ir < mesh.voxelCount(); ir++)
     {
-        if(ir == 32*64*16 + 32*16 + 8)
-        {
-            qDebug() << "Hi";
-        }
+        //if(ir == 32*64*16 + 32*16 + 8)
+        //{
+        //    qDebug() << "Hi";
+        //}
         int zid = mesh.zoneId[ir];
         for(unsigned int ie = 0; ie < xs.groupCount(); ie++)
         {

### 2017-06-12 doctors/mainwindow.cpp

@@ -75,6 +75,7 @@ MainWindow::MainWindow(QWidget *parent):
     MainWindow::m_goodPalette = new QPalette();
     m_goodPalette->setColor(QPalette::Text, Qt::green);
 
+    // Construct the dialogs
     outputDialog = new OutputDialog(this);
     geomDialog = new GeomDialog(this);
     quadDialog = new QuadDialog(this);
@@ -85,6 +86,7 @@ MainWindow::MainWindow(QWidget *parent):
     m_configSelectDialog->setAcceptMode(QFileDialog::AcceptOpen);
     m_configSelectDialog->setFileMode(QFileDialog::ExistingFile);
 
+    // Construct the data readers
     m_parser = new AmpxParser;
     m_solver = new Solver;
     m_xs = new XSection;
@@ -127,6 +129,10 @@ MainWindow::MainWindow(QWidget *parent):
 
     connect(this, SIGNAL(signalLaunchRaytracerHarmonic(const Quadrature*,const Mesh*,const XSection*, const SolverParams*,const SourceParams*)), m_solver, SLOT(raytraceHarmonic(const Quadrature*,const Mesh*,const XSection*,const SolverParams*,const SourceParams*)));
     connect(this, SIGNAL(signalLaunchSolverHarmonic(const Quadrature*,const Mesh*,const XSection*,const SolverParams*, const SourceParams*,const std::vector<RAY_T>*)), m_solver, SLOT(gsSolverHarmonic(const Quadrature*,const Mesh*,const XSection*,const SolverParams*,const SourceParams*,const std::vector<RAY_T>*)));
+
+    connect(this, SIGNAL(signalLaunchRaytracerKlein(const Quadrature*,const Mesh*,const XSection*, const SolverParams*,const SourceParams*)), m_solver, SLOT(raytraceKlein(const Quadrature*,const Mesh*,const XSection*,const SolverParams*,const SourceParams*)));
+    connect(this, SIGNAL(signalLaunchSolverKlein(const Quadrature*,const Mesh*,const XSection*,const SolverParams*, const SourceParams*,const std::vector<RAY_T>*)), m_solver, SLOT(gsSolverKlein(const Quadrature*,const Mesh*,const XSection*,const SolverParams*,const SourceParams*,const std::vector<RAY_T>*)));
+
 
[…]

### 2017-06-12 doctors/mainwindow.h

@@ -41,6 +41,7 @@ public:
     const static unsigned int ISOTROPIC = 0;
     const static unsigned int LEGENDRE = 1;
     const static unsigned int HARMONIC = 2;
+    const static unsigned int KLEIN = 3;
 
 private:
     Ui::MainWindow *ui;
@@ -125,6 +126,9 @@ signals:
 
     void signalLaunchRaytracerHarmonic(const Quadrature *quad, const Mesh *mesh, const XSection *xs,  const SolverParams *solPar, const SourceParams *srcPar);
     void signalLaunchSolverHarmonic(const Quadrature *quad, const Mesh *mesh, const XSection *xs,  const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
+
+    void signalLaunchRaytracerKlein(const Quadrature *quad, const Mesh *mesh, const XSection *xs,  const SolverParams *solPar, const SourceParams *srcPar);
+    void signalLaunchSolverKlein(const Quadrature *quad, const Mesh *mesh, const XSection *xs,  const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
 
     void signalDebugHalt(std::vector<float>);
     void signalBeginXsParse(QString);

### 2017-06-12 doctors/materialutils.cpp

@@ -509,7 +509,8 @@ const std::vector<int> MaterialUtils::hounsfieldRangePhantom19{
     1400,
     1500,
     3000,  // 19
-    1000000,  // 20
+    65000,  // 20 (Al gantry)
+    1000000,  // 21
 };
 
 const std::vector<int> MaterialUtils::hounsfieldRangePhantom19Elements{
@@ -524,30 +525,32 @@ const std::vector<int> MaterialUtils::hounsfieldRangePhantom19Elements{
     17,    // Chlorine
     18,    // Argon
     19,    // Potassium
-    20     // Calcium
+    20,    // Calcium
+    13     // Aluminum (for gantry)
 };
 
 const std::vector<std::vector<float> > MaterialUtils::hounsfieldRangePhantom19Weights{
-    std::vector<float> {0.000f, 0.000f, 0.757f, 0.232f, 0.000f, 0.000f, 0.000f, 0.000f, 0.000f, 0.013f, 0.000f, 0.000f}, // Air 1
-    std::vector<float> {0.103f, 0.105f, 0.031f, 0.749f, 0.002f, 0.000f, 0.002f, 0.003f, 0.003f, 0.000f, 0.002f, 0.000f}, // Lung 2
-    std::vector<float> {0.112f, 0.508f, 0.012f, 0.364f, 0.001f, 0.000f, 0.000f, 0.001f, 0.001f, 0.000f, 0.000f, 0.000f}, // Adipose/adrenal 3
-    std::vector<float> {0.100f, 0.163f, 0.043f, 0.684f, 0.004f, 0.000f, 0.000f, 0.004f, 0.003f, 0.000f, 0.000f, 0.000f}, // Small intestine 4
-    std::vector<float> {0.097f, 0.447f, 0.025f, 0.359f, 0.000f, 0.000f, 0.023f, 0.002f, 0.001f, 0.000f, 0.001f, 0.045f}, // Bone 5
-    std::vector<float> {0.091f, 0.414f, 0.027f, 0.368f, 0.000f, 0.001f, 0.032f, 0.002f, 0.001f, 0.000f, 0.001f, 0.063f}, // Bone 6
[…]

### 2017-06-12 doctors/solver.cpp

@@ -105,6 +105,30 @@ void Solver::gsSolverHarmonic(const Quadrature *quad, const Mesh *mesh, const XS
     }
 }
 
+void Solver::raytraceKlein(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar,  const SourceParams *srcPar)
+{
+    if(solPar->gpu_accel)
+    {
+        raytraceKleinGPU(quad, mesh, xs, solPar, srcPar);
+    }
+    else
+    {
+        raytraceKleinCPU(quad, mesh, xs, solPar, srcPar);
+    }
+}
+
+void Solver::gsSolverKlein(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux)
+{
+    if(solPar->gpu_accel)
+    {
+        gsSolverKleinGPU(quad, mesh, xs, solPar, srcPar, uflux);
+    }
+    else
+    {
+        gsSolverKleinCPU(quad, mesh, xs, solPar, srcPar, uflux);
+    }
+}
+
 /************************************************************************************************
  * ========================================= CPU Code ========================================= *
  ************************************************************************************************/
@@ -250,6 +274,11 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *, const Mesh *mes
                 for(unsigned int xIndxStart = 0; xIndxStart < mesh->xElemCt; xIndxStart++)  // For every voxel
                 {
                     RAY_T acceptance = 1.0;
+
+                    //if(xIndxStart == 32 && yIndxStart == 32 && zIndxStart == 8)
+                    //{
+                    //    qDebug() << "Stop here please";
+                    //}
 
                     //if(zIndxStart == 7 && yIndxStart == 32 && xIndxStart == 31)
                     //    qDebug() << "Here";
@@ -422,7 +451,7 @@ std::vector<RAY_T> *Solver::basicRaytraceCPU(const Quadrature *, const Mesh *mes
                             z += tmin*zcos;
                             if(ycos >= 0)
                             {
-                                if(yBoundIndx = mesh->yNodeCt-1)
[…]

### 2017-06-12 doctors/solver.h

@@ -49,6 +49,9 @@ public slots:
     void raytraceHarmonic(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar);
     void gsSolverHarmonic(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
 
+    void raytraceKlein(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar);
+    void gsSolverKlein(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
+
     // CPU versions
     void raytraceIsoCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar);
     void gsSolverIsoCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
@@ -59,6 +62,9 @@ public slots:
     void raytraceHarmonicCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar);
     void gsSolverHarmonicCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
 
+    void raytraceKleinCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar);
+    void gsSolverKleinCPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
+
     // The GPU versions
     void raytraceIsoGPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar);
     void gsSolverIsoGPU(const Quadrature *quad, const Mesh *mesh, const XSection *xs, const SolverParams *solPar, const SourceParams *srcPar, const std::vector<RAY_T> *uflux);
[…]

### 2017-06-12 doctors/xsection.cpp

@@ -5,9 +5,11 @@
 
 #include <QDebug>
 #include <QMessageBox>
+#include <fstream>
 
 #include "xs_reader/ampxparser.h"
 #include "materialutils.h"
+#include "quadrature.h"
 
 XSection::XSection() : m_elements(NULL), m_weights(NULL), m_groups(0), m_matsLoaded(0)
 {
@@ -36,15 +38,31 @@ float XSection::totXs1d(const int matid, const int g) const
 
 float XSection::scatxs2d(const int matid, const int gSrcIndx, const int gSinkIndx, const int n) const
 {
+    //qDebug() << "This function is now depricated until flux moments are implemented";
+
     if(matid > m_matsLoaded || gSrcIndx >= m_groups || gSinkIndx >= m_groups || n > m_pn)
     {
-        qDebug() << "XSection::scatxs2d(): 39: Illegal index value";
+        qDebug() << "Illegal index value";
         qDebug() << "Violated matid > m_matsLoaded || gSrcIndx >= m_groups || gSinkIndx >= m_groups || n > m_pn";
         qDebug() <<  matid << "> " << m_matsLoaded << " || " << gSrcIndx << " >= " << m_groups << " || " << gSinkIndx << " >= " << m_groups << " || " << n << " > " << m_pn;
         return -1.0;
     }
     const int pnCount = m_pn+1;
     return m_scat2d[matid*m_groups*m_groups*pnCount + gSrcIndx*m_groups*pnCount + gSinkIndx*pnCount + n];
+}
+
+float XSection::scatxsKlein(const int matid, const int gSrcIndx, const int gSinkIndx, const int aSrcIndx, const int aSinkIndx) const
+{
+    if(matid > m_matsLoaded || gSrcIndx >= m_groups || gSinkIndx >= m_groups || aSrcIndx >= m_angles || aSinkIndx >= m_angles)
+    {
+        QString errmsg = QString("Indexing subscript went out of bounds in scatxsKlein()");
+        //errmsg += "The table has room for " + QString::number(m_mats) + " materials.";
+        QMessageBox::warning(NULL, "Indexing Error", errmsg, QMessageBox::Close);
+        //qDebug() <<  matid << "> " << m_matsLoaded << " || " << gSrcIndx << " >= " << m_groups << " || " << gSinkIndx << " >= " << m_groups << " || " << n << " > " << m_pn;
+        return -1.0;
+    }
+    //const int pnCount = m_pn+1;
[…]

### 2017-06-12 doctors/xsection.h

@@ -8,6 +8,7 @@
 #include <vector>
 
 class AmpxParser;
+class Quadrature;
 
 class XSection
 {
@@ -17,6 +18,7 @@ public:
 
     std::vector<float> m_tot1d;
     std::vector<float> m_scat1d;
+    std::vector<float> m_scatKlein;
 
     const std::vector<int> *m_elements;
     const std::vector<std::vector<float> > *m_weights;
@@ -31,11 +33,14 @@ public:
     float scatXs1d(const int matid, const int g) const;
     float totXs1d(const int matid, const int g) const;
     float scatxs2d(const int matid, const int gSource, const int gSink, const int n) const;
+    float scatxsKlein(const int matid, const int gSource, const int gSink, const int aSource, const int aSink) const;
 
     bool allocateMemory(const unsigned int materialCount, const unsigned int groupCount, const unsigned int PnCount);
     bool allocateMemory(const unsigned int groupCount, const unsigned int PnCount);
     bool addMaterial(const std::vector<int> &z, const std::vector<float> &w, const AmpxParser *p);
     bool addAll(AmpxParser *parser);
+
+    bool buildKleinTable(Quadrature *quad);
 
     bool setElements(const std::vector<int> &elem, const std::vector<std::vector<float> > &wt);
 
@@ -44,6 +49,7 @@ private:
     int m_mats;
     int m_pn;
     int m_matsLoaded;
+    int m_angles;
 
     const static unsigned int MT_GAMMA_TOTAL_INTERACTION = 501;
     const static unsigned int MT_GAMMA_COHERENT_SCATTER = 502;

### 2017-06-12 docprep/chapbib.tex
Updates

@@
 \section{CT Dose Estimation}
-As discussed in Chapter 1, the dose a patient receives from both diagnostic CT procedures and radiation therapy is of great concern to the medical community [CITE].
+As discussed in Chapter 1, the dose a patient receives from both diagnostic CT procedures and radiation therapy is of concern to the medical community.
 However, dose quantification is not currently patient specific.
-Measuring the dose in a patient directly is impossible, intead the air kerma (measured in Gy), which measures the x-ray beam intensity, and can be measured directly [CITE].
+Measuring the dose in a patient directly is impossible, intead the air kerma (measured in Gy), which measures the x-ray beam intensity, and can be measured directly with ionization chambers~\citep{ef:wolbarsta}.
 As the x-ray beam passes through the patient, it deposits energy.
@@
 
-Monte Carlo codes such as TOPAS [CITE], Geant4 \citep{ref:agostinellis}, and MCNP6 [CITE] are the current gold standard in computational dosimetry \citep{ref:jiax}.
+Monte Carlo codes such as TOPAS~\citep{ref:perlj}, Geant4~\citep{ref:agostinellis}, and MCNP6~\citep{ref:mcnp6} are the current gold standard in computational dosimetry~\citep{ref:jiax}.
 This work compares results to those obtained from MCNP6.
@@
 Along that direction, the 3-dimensional gradient operator found in the full Boltzmann equation (see Sec.~\ref{sec:lbe}) becomes a 1-dimensional derivitive.
-This reduction gives rise to an analytical solution along the characteristic [CITE].
+This reduction gives rise to an analytical solution along the characteristic~\citep{ref:askewj}.
 Sufficient characteristic solutions solved simulateously gives the solution over the entire problem domain.
@@
 The inverse radon transform backprojects each row of the sinogram back to physical space.
-Filtering each row in Fourier space and then compositing all layers together recreates the image [CTIE].
-This methodology has been extended to fan beams [CITE] and cone beams [CITE].
-Besides the filtered backprojection (FBP) method, iterative reconstruction techniques can also be applied to greatly improve the image reconstruction quality [CITE].
+Filtering each row in Fourier space and then compositing all layers together recreates the image~\citep{ref:mersereaur}.
+This methodology has been extended to fan beams~\citep{ref:bessong} and cone beams~\citep{ref:turbellh}.
+Besides the filtered backprojection (FBP) method, iterative reconstruction techniques can also be applied to greatly improve the image reconstruction quality~\citep{ref:pontanaf}.
 
@@
 
-Along any particular ray, monoenergetic particles will attenuate through a homogenous media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert} [CITE] resulting in a small fraction of the source particles reaching the detector.
+Along any particular ray, monoenergetic particles will attenuate through a homogenous media according to the Beer-Lambert law given in Eq.~\ref{eq:beer_lambert}~\citep{ref:lamarshj} resulting in a small fraction of the source particles reaching the detector.
 This phenomena leads to an image of the attenuating media but also results in the detector receiving information about the attenuation coefficient, $\mu$, along that ray.
 With sufficient rays of varying directionality, the entire object can be quantified with respect to its spatially distributed attenuation coefficient.
-The Hounsfield units (also called CT numbers) are traditionally scaled such that -1000 is dry air and water is zero [CITE].
 
@@
 These values are related to the attenuation coefficient of the material represented by the voxel.
-Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air [CITE] though some variations do exist \citep{ref:plessisf} \citep{ref:sawc}.
+Traditionally, a CT number of zero corresponds to the attenuation of water while -1000 corresponds to dry air though some variations do exist \citep{ref:plessisf,ref:sawc}.
 
@@
 \section{Discrete Ordinates}\label{sec:discordlit}
-The dsicrete ordinates method dates back to [CITE] but remains one of the most prominent solution modalities for radiation transport in use today.
+The dsicrete ordinates method dates back to~\citet{ref:chandrasekahar} who used it for radiation transport in atmospheres, yet it remains one of the most prominent solution modalities for radiation transport in use today.
 A comprehensive review of discrete ordinate methods is given by \citet{ref:lewise}.
-Additional references include [CITES].
+Additional historical references include~\citet{ref:carlsonb}~\citet{ref:lanthropk}.
 
@@
 
-Discrete ordinate methods are almost always done on a Cartesian grid, though alternative derivations on general geometry do exist~\citep{ref:dehart}.
+Discrete ordinate methods are almost always done on a Cartesian grid, though alternative derivations on general geometry do exist~\citep{ref:dehartm}.
 Extension to unstructured tetrahedral meshes was done by ~\citet{ref:wareingt}.

### 2017-06-12 docprep/chapconc.tex
Updates

@@
 
-\section{First conclusion}
+\section{Accuracy}
 
@@
 
-Anisotropy cannot be ignored.
+The collided flux suffers from the isotropic approximation made to simplify the code.
 
@@
 Additionally, the direct output of DOCTORS may not necessarily be of clinical importance since
+
+\section{GPU Speedup}
+
+The GPU acceleration algorithm was found to speed the DOCTORS code up by a factor of ~40x.
 
@@
 
+\subsection{Anisotropy Treatment}
+A more sophisticated anisotropy treatment would be appropriate since the collided flux was found to have issues regarding its behavior.
+
 \subsection{Group Structure}
@@
 
-NJOY or AMPX.
+INSERT SECTION ABOUT NJOY or AMPX.
 
 \subsection{Temporal Extension}
-Use 4D CT to do temporal stuff as well as spatial stuff.
+An advantage of the raytracer is that it is very fast.
+Therefore, it can characterize many beams and integrate them temporally easily.
+This would allow DOCTORS to model more complex beam shapes and scan protocols accurately.
 
@@
 \subsection{Extension to Therapy Beams}
-High energy will be more anisotropic. Will need spherical harmonics.
+DOCTORS can likely be extended to high energy therapeutic beams for clinical treatment.
+However, higher energy photons will scatter more anisotropicly.
+This will require spherical harmonics and high order $P_l$ expansion tables in the cross section data.
 
@@
 Currently, DOCTORS cannot automatically identify specific organs, thus it is only able to compute the equivalent dose via dose deposition.
+If specific organs could be identified, tissue specific weighting factors could be applied resulting in the effective dose to each organ which would be of greater clinical significance.
+However, such identification currently requires additional input about the scan protocol and extensive knowledge about the human anatomy not currently integrated into DOCTORS.
 

### 2017-06-12 docprep/chapimpl.tex
Updates

--- 2017-06-08 docprep/chapimpl.tex
@@
 GPU computation is performed via CUDA which must be compiled seperatly by Nvidia's proprietary compiler, \texttt{nvcc}.
+
+INSERT NOTE ABOUT CODE LISTING FORMATTING
 
@@
 In the case of the header bytes given in Fig.~\ref{fig:ampxbytes1} and~\ref{fig:ampxbytes2}, the bytes correspond to the integer 440 which is the length in bytes of the record (not including the header or footer).
-
-Each byte of data is sequentially read, reordered, and converted to an apprpriately typed variable. Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.
+Each byte of data is sequentially read, reordered, and converted to an apprpriately typed variable.
+Table~\ref{tab:reorder} shows the converstion for the record given in Fig.~\ref{fig:ampxbytes1}.
 
@@
   \begin{center}
-   \includegraphics[width=3.75in]{figs/ampxbytes1}
+   \includegraphics[width=4.75in]{figs/ampxbytes1}
   \end{center}
@@
   \begin{center}
-   \includegraphics[width=2.75in]{figs/ampxbytes2}
+   \includegraphics[width=4.75in]{figs/ampxbytes2}
   \end{center}
@@
   ... &      ---    &      ---    &    --- & ---                       \\
-  11  & 70 75 6F 43 & 43 6F 75 70 & "Coup" & First word of the title   \\
-  12  & 20 64 65 6C & 6C 65 64 20 & "led " & Second word of the title  \\
+  11  & 70 75 6F 43 & 43 6F 75 70 & "Coup" & First four \texttt{char} of the title   \\
+  12  & 20 64 65 6C & 6C 65 64 20 & "led " & Second four \texttt{char} of the title  \\
   ... &     ---     &    ---      &  ---   & ---                       \\
-  440 & 20 20 20 20 & 20 20 20 20 & "    " & 4 spaces ending the title \\
-  441 & B8 01 00 00 & 00 00 01 B8 & 440    & Footer                    \\ 
+  110 & 20 20 20 20 & 20 20 20 20 & "    " & 4 spaces ending the title \\
+  111 & B8 01 00 00 & 00 00 01 B8 & 440    & Footer                    \\ 
   [1ex]      % [1ex] adds vertical space
@@
 \end{tabular}
-\label{table:reorder}
+\label{tab:reorder}
 \end{table}
@@
 \begin{minted}[frame=lines,linenos]{c++}
-# Given magic
+magic = READ_NEXT_BINARY_INT()
 KKK = magic % 1000
@@
 \begin{equation}
-sum_{i = 1}^N w_i = 1
+\sum_{i = 1}^N w_i = 1
 \end{equation}
@@
 \begin{minted}[frame=lines,linenos]{c++}
-class MainWindow : public QWindow
+class MainWindow : public QMainWindow
 {
@@
 
-public slot:
+protected slots:
 	void doSomethingLong();
@@
 
-	// When a button named button1 which was created in the Qt5 creator interface is clicked, the clicked() signal is automatically emitted which executes the doSomethingLong() function
-	MainWindow.connect(ui->button1, clicked(), this, doSomethingLong())
+	// When a button named button1 which was created in the Qt5
+	creator interface is clicked, the clicked() signal is 
+	automatically emitted which executes the doSomethingLong()
+	function
+	connect(ui->button1, clicked(), this, 
+		doSomethingLong());
 }
@@
 
-The code in Listing~\ref{lst:sync1} is updated to use the Qt signal/slot mechanism.
+The code in Listing~\ref{lst:sync1} is updated to use the Qt signal/slot mechanism whose code is given in Listing~\ref{lst:async1}.
 
@@
 {
+	// Member variables
+	QThread workerThread;
+	Worker *worker;
+
 	// Constructor
@@
 
-	// Other parts of the class
-
-public slot:
-	void doSomethingLong();
-}
-
-void MainWindow::doSomethingLong()
-{
-	// Execute a long piece of code
+protected slots:
+	handleResult();
 }
@@
 	// Initializations
+	worker = new Worker;
 
-	// When a button named button1 which was created in the Qt5 creator interface is clicked, the clicked() signal is automatically emitted which executes the doSomethingLong() function
-	MainWindow.connect(ui->button1, clicked(), this, doSomethingLong())
+	worker.moveToThread(&workerThread);
+
+	// Set up the thread connections
+	connect(&workerThread, finished(), worker, deleteLater());
+	connect(this, begin(), worker, doSomethingLong());
+	connect(worker, done(), this, handleResult());
 }
+
+class Worker : public QObject
+{
+private signals:
+	void done();
+
+public slots:
+	void soSomethingLong();
+}
+
+Worker::doSomethingLong()
+{
+	// Do stuff...
+	emit done();
+}
+
 \end{minted}
-\caption{A long function causes the user interface to block.}\label{lst:sync1}
+\caption{Signals and slots enable a long function to be called without blocking the user interface.}\label{lst:async1}
 \end{listing}
@@
 CUDA code is compoiled with the Nvidia complier nvcc.
-Qt uses the gcc compiler and its MOC generator for meta code.
+Qt5 uses the gcc compiler and its MOC generator for meta code.
 In order to connect CUDA code to the Qt MOC, the CUDA code is compiled by nvcc to produce a .o file.
-Qt then compiles all other files into corresponding.o files.
+Qt5 then compiles all other files into corresponding.o files.
 The linker than automactically picks up all of the .o files gneerated by nvcc.
-The final result is an executable that has a Qt generarted user interface that can communicat with an Nvidia graphics card through the CUDA language.
+The final result is an executable that has a Qt5 generarted user interface that can communicat with an Nvidia graphics card through the CUDA language.
 
@@
         \includegraphics[width=\textwidth]{figs/subsweep_cube1}
-        \caption{Subsweep 0 ($S=0$)}
+        \caption{}
         \label{fig:subsweep_cube1}
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
+    ~ 
     \begin{subfigure}[b]{0.45\textwidth}
         \includegraphics[width=\textwidth]{figs/subsweep_cube2}
-        \caption{Subsweep 1 ($S=1$)}
+        \caption{}
         \label{fig:subsweep_cube2}
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-    %(or a blank line to force the subfigure onto a new line)
+    ~ 
     \begin{subfigure}[b]{0.45\textwidth}
         \includegraphics[width=\textwidth]{figs/subsweep_cube3}
-        \caption{Subsweep 6 ($S=6$)}
+        \caption{}
         \label{fig:subsweep_cube3}
@@
         \includegraphics[width=\textwidth]{figs/subsweep_cube4}
-        \caption{Subsweep 6 ($S=6$) with labels}
+        \caption{) with labels}
         \label{fig:subsweep_cube4}
@@
 Each subsweep must complete before those after it.
-Each voxel within a subsweep can be solved in parallel with all others in its subsweep.}\label{fig:subsweep_cube}
+Each voxel within a subsweep can be solved in parallel with all others in its subsweep.
+(a) Subsweep 0 ($S=0$) (b) Subsweep 1 ($S=1$) (c) Subsweep 6 ($S=6$) (d)Subsweep 6 ($S=6$) with labels.}\label{fig:subsweep_cube}
 \end{figure}

### 2017-06-12 docprep/chapintro.tex
Updates

@@
 Increased information results in faster and more accurate diagnosis.
-However, the exponentially increasing usage of CT and dose patients receive from it has raised concerns of potential long term risks~\citep{ref:brennerd} \citep{ref:einsteina1} \citep{ref:abramsh} \citep{ref:einsteina2} \citep{ref:mccolloughc} \citep{ref:yul}.
+However, the exponentially increasing usage of CT and dose patients receive from it has raised concerns of potential long term risks~\citep{ref:brennerd,ref:einsteina1,ref:abramsh,ref:einsteina2,ref:mccolloughc,ref:yul}.
 These concerns are exemplified by the push for low dose CT.
 
-INSERT MORE LOW DOSE CT STUFF
-A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.
+%INSERT MORE LOW DOSE CT STUFF
+%A study by \citet{ref:kovalchiks} found that using a low-dose screening method as opposed to CT radiography reduced lung cancer mortality in high risk patients by 20\%.
 
@@
 Another, similar, application is dose estimation of patients receiving radiation therapy.
-Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment [CITE].
+Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment~\citep{ref:pant}.
 With further development, this methodology may enable real time dose computation as the treatment is administered.

### 2017-06-12 docprep/chapmeth.tex
Updates

@@
 
-The $S_N$ quadrature ensures 90$\deg$ symmetry about any axis.
+The $S_N$ quadrature ensures 90 degree symmetry about any axis.
 In order to ensure the rotational symmetry, the $S_N$ quadrature has a single degree of freedom.
@@
 For a more through discussion of quadrature generation, see \citet{ref:lewise}.
+Figure~\ref{fig:quad} shows the simples three quadratures implemented in DOCTORS.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/quad2}
+        \caption{}
+        \label{fig:quad2}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/quad4}
+        \caption{}
+        \label{fig:quad4}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/quad6}
+        \caption{}
+        \label{fig:quad6}
+    \end{subfigure}
+    \caption{Some various $S_N$ quadratures. The black dots indicate the location of a ray. (a) $S_2$ (b) $S_4$ (c) $S_6$ }\label{fig:quad}
+\end{figure}
 
@@
 
+INSERT WHY NOT USE FLUX MOMENTS
+
 Scatter is assumed to be a function of only the scatter angle $\theta_s$ between the initial, $\hat{\Omega}_a$, and scattered, $\hat{\Omega}_{a'}$, directions.
@@
 
+An alternate anisotropy treatment is to use the Klein-Nishina formula~\citep{ref:shultisj}. The Klein-Nishina formula describes the scattering cross section as
+\begin{equation}
+\sigma(E, \mu) \propto q^3 + q\mu^2
+\end{equation}
+where
+\begin{equation}
+q(E, \mu) = \frac{1}{1 + (E/E_0)(1-\mu)}
+\end{equation}
+
 \section{Uncollided Solution Methods}\label{sec:uncol}
@@
 Since no scatter or absorption takes place in vacuum conditions, the particles can only stream and particles must be conserved.
-However, if particles are transported along a single discrete direction only as illustrated in Fig.~\ref{fig:rayeffect_ex} ($\hat{\Omega} = <\sqrt{2}/2, \sqrt{2}/2>$ in Fig.~\ref{rayeffect_ex} and~\ref{fig:rayeffect_comp}) the phenomena observed in Fig.~\ref{fig:rayeffect_comp} is observed.
+However, if particles are transported along a single discrete direction only as illustrated in Fig.~\ref{fig:rayeffect_ex} ($\hat{\Omega} = <\sqrt{2}/2, \sqrt{2}/2>$ in Fig.~\ref{fig:rayeffect_ex} and~\ref{fig:rayeffect_comp}) the phenomena observed in Fig.~\ref{fig:rayeffect_comp} is observed.
 As the discretization becomes more refined, the isotropic source becomes more heavily biased in the $\hat{\Omega}$ direction.
@@
     \begin{subfigure}[b]{0.2\textwidth}
-        \includegraphics[width=\textwidth]{figs/proj24}
+        \includegraphics[width=\textwidth]{figs/proj64}
         \caption{}
@@
     \end{subfigure}
-    \caption{Different numbers of fan beam projections ($N$). (a) $N = 6$. (b) $N = 12$. (c) $N = 16$. (d) $N = 24$.}\label{fig:fanproj}
+    \caption{Different numbers of fan beam projections ($N$). (a) $N = 6$. (b) $N = 12$. (c) $N = 16$. (d) $N = 64$.}\label{fig:fanproj}
 \end{figure}
@@
 The rejection criteria is a function of only a single parameter, $\theta$, the maximum permissible angle between the center of the beam, $\boldsymbol{s}$, and a ray toward the point of interest, $\boldsymbol{c}$.
-
 The angle between $\boldsymbol{s}$ and $\boldsymbol{c}$ is
@@
 is evaluated to determine whether the ray is accepted or rejected.
+Figure~\ref{fig:beamfancone} shows a comparison of the cone beam and fan beam shapes.
+The most noticable difference is in the $xz$ plne where the cone beam is circular, but the fan beam produces a rectangular cross section.
 
@@
     \end{subfigure}
-    \caption{The beam shapes.}\label{fig:beamfancone}
+    \caption{Comparison of a cone beam (a), (c), and (e) to a fan beam (b), (d), and (f).
+The cone beam is 20 degrees in both the $xy$ and $yz$ planes but the fant beam is different.
+The cross section of the cone beam is cicular (e), but the fan beam cross section is rectangular (f).}\label{fig:beamfancone}
 \end{figure}
@@
 Dose can be computed in either one of two ways.
-The first way is to compute the fluence distribution and multiply by a fluence-to-dose conversion factor from ICRP 116 [CITE].
+The first way is to compute the fluence distribution and multiply by a fluence-to-dose conversion factor from ICRP~116~\citep{ref:icrp116}.
 The second way to compute dose is to compute the energy deposition in each cell.
@@
 Figure~\ref{fig:icrp116_rotconv} shows the energy dependent fluence-to-dose conversions for the ROT orientation.
+ICRP 116 distributes the conversion factors as point-wise data which must be converted to group data.
+This is performed using linear interpolation between the ICRP 116 data points.
 
@@
   \end{center}
-  \caption{Fluence-to-dose conversion factors from ICRP 116 for photons.}
+  \caption{Fluence-to-dose conversion factors from ICRP 116 for photons and the group averaged conversion factors for the 47-group data library in DOCTORS.}
 \label{fig:icrp116_rotconv}
@@
 \end{equation}
-where $\rho_i$ is the density in g/cm$^3$ and $F_2$ is a conversion factor to convert grams to kilograms. The dose (in Gy) to the voxel is then
+where $\rho_i$ is the density in g/cm$^3$ and $F_2$ is a conversion factor to convert grams to kilograms. The absorbed dose (in Gy) to the voxel is then
 \begin{equation}\label{eq:dosedep}

### 2017-06-12 docprep/chapresults.tex
Updates

@@
 %%
+
+This chapter summarizes the computational results obtained from DOCTORS and the validation to MCNP6.
+The MCNP6 reference compared to is automatically generated by DOCTORS to guarantee that the geometry and source parameters are identical.
+This process is described in detail in Section~\ref{sec:mcnpgen}.
 
@@
 
-Some preprocessing was done to each phantom before it was used in DOCTORS to provide results.
-
-The results from DOCTORS were compared to analagous results from MCNP6.
-The MCNP6 reference compared to is automatically generated by DOCTORS to guarantee that the geometry and source parameters are identical.
-This process is described in detail in Section~\ref{sec:mcnpgen}.
-
-First subtract 1024 to account for the unsigned 16-bit int data structure.
+Hounsfield units range from -1000 for air to 0 for water and higher for more radiopaque materials.
+However, data is stored as an 16-bit \texttt{unsigned int} in the CT voxel phantoms.
+Therefore, the data has 1024 ($2^{10}$) added to it to avoid rolling negative values over.
+The first step is to reinterpret the \texttt{unsigned int} values as \texttt{int} values and subtract 1024.
 
@@
 
-\subsection{Gantry Handling}
-In some phantom models, a gantry is present. The gantry poses a challenge since it is a highly attenuating material relative to the patient which reduces the available flux.
-
-The CT number to material conversion developed by~\cite{ref:ottossonr} which is implemented in this work does not have an equivalent for the gantry region.
-
 \subsection{Geometry Simplification}
@@
   \end{center}
-  \caption{The small one}
+  \caption{Raw CT number data for the $64 \times 64 \times 16$ water phantom.}
 \label{fig:waterPhantom64}
@@
 
-\section{Cone-Beam Water Benchmark}
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/phantomwater_64}
+        \caption{}
+        \label{fig:waterHistLin}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/phantomwater_256}
+        \caption{}
+        \label{fig:waterHistLog}
+    \end{subfigure}
+    \caption{The (a) full water phantom and (b) simplified phantom after the CT numbers are converted into materials.}\label{fig:waterHist}
+\end{figure}
 
-The first benchmark, the cone with 19 groups is given in Fig. 1. The cross section data for the water and air regions are shown in Fig.~\ref{fig:airwater19}.
+\section{Cone-Beam Water Benchmark}\label{sec:cone1}
 
-The water phantom is a 35 cm diameter water cylinder that stands 12.5 inches tall. The water is encased in a 0.5 cm thick plastic container.
+The first benchmark, is a 8 degree cone targeting a water phantom.
+The water phantom is a 35 cm diameter water cylinder that stands 12.5 inches tall.
+The water is encased in a 0.5 cm thick plastic container.
+Each voxel in the mesh is assigned one of the materials given in Table~\ref{tab:watermats}.
+Note that the CT range column is of the raw data before 1024 is subtracted.
+Once the artifacts are removed, the histogram is plotted in Figure~\ref{fig:waterHist}.
+The logarithmic scaled histogram shows the low frequency voxels between -800 and -100 the are difficult to identify as water or air.
+These voxels are around the periphery of
 
@@
 \hline
-Water     & $0 \leq x < 67$ & 1581199 & 37.70 \\
-Container & $67 \leq x < 600$ & 81502 & 1.94 \\
-Air       & $600 \leq x < 1080$ & 1650372  & 39.35 \\
-Artifact  & $1080 \leq x < 65535$ & 881231 & 21.01 \\  [1ex]
+Water     & $0 \geq x > -67$ & 1581199 & 37.70 \\
+Container & $-67 \geq x > -600$ & 81502 & 1.94 \\
+Air       & $-600 \geq x > -1080$ & 1650372  & 39.35 \\
+Artifact  & $-1080 \geq x > -65535$ & 881231 & 21.01 \\  [1ex]
 \hline
 \end{tabular}
-\label{table:nonlin}
+\label{tab:watermats}
 \end{table}
 
-Once the artifacts are removed, the histogram is plotted.
-Figure~\ref{fig:waterHist} shows two versions of the water histogram, \ref{fig:waterHistLin} on a linear scale and \ref{fig:waterHistLog} on a logarithmic scale.
+One of the materials listed in Table~\ref{tab:watermats} is the "Artifact" material.
+These voxels apppear in the corners of the reconstruction and are artifacts caused by the reconstruction algorithm.
+The corners are not sampled by each projection as the CT system rotates about the phantom.
+During reconstruction, this results in those regions being asigned proportionally higher CT numbers.
 
@@
         \includegraphics[width=\textwidth]{figs/waterHistogramLin}
-        \caption{Linear}
+        \caption{}
         \label{fig:waterHistLin}
@@
         \includegraphics[width=\textwidth]{figs/waterHistogramLog}
-        \caption{Logarithmic}
+        \caption{}
         \label{fig:waterHistLog}
     \end{subfigure}
-    \caption{The histogram.}\label{fig:waterHist}
+    \caption{The histogram of the CT numbers in the water phantom on a (a) linear scale and (b) logarithmic scale.}\label{fig:waterHist}
 \end{figure}
@@
 A key difference between though is the group structure used in the group-averaged deterministic cross sections.
-Figure~\ref{fig:airwarerxs} shows the continuous energy macroscopic cross section for both air and water versus the discretized data used by DOCTORS.
+Figure~\ref{fig:airwaterxs} shows the continuous energy macroscopic cross section for both air and water versus the discretized data used by DOCTORS.
 Unfortunately, the datasets used in DOCTORS originate from SCALE6.2, a light water reactor analysis code.
@@
         \includegraphics[width=\textwidth]{figs/airwaterxs2_19group}
-        \caption{19 Groups}
+        \caption{}
         \label{fig:airwaterxs19}
@@
         \includegraphics[width=\textwidth]{figs/airwaterxs2_47group}
-        \caption{47 Groups}
+        \caption{}
         \label{fig:airwaterxs47}
     \end{subfigure}
-    \caption{Comparison of the group cross section data used by DOCTORS to the continuous cross section data used by MCNP6. The continuous data is provided by NIST !!!CITE!!!.}\label{fig:airwaterxs}
+    \caption{Comparison of the group cross section data used by DOCTORS to the continuous cross section data used by MCNP6 for (a) 19 groups and (b) 47 groups.
+The continuous data is provided by NIST~\citep{ref:hubbellj}.}\label{fig:airwaterxs}
 \end{figure}
 
-The detailed flux distribution from MCNP6 and the solver are not included here, they are provided in Appendix~\ref{sec:appc_benchmark}.
+\subsection{Uniform beam}
 
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_comp}
-        \caption{10 - 20 MeV.}
-        \label{fig:bench_cone1_comp}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout}
-        \caption{10 - 20 MeV lineout.}
-        \label{fig:bench_cone1_lineout}
-    \end{subfigure}
+This benchmark uses a "uniform" beam in which source particles are generated equiprobably in all groups.
+Therefore, the energy distribution $S_0(E)$, is not uniform, but rather $S_0(E)/\Delta E$ is.
+The goal of this benchmark is to determine the impact the energy discretization has on the results.
+The 19-group data set is very coarse in the diagnostic energy range; this benchmark is designed to determine whether or not this data set can yield meaningful results.
 
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_comp_g17}
-        \caption{100 - 200 keV.}
-        \label{fig:bench_cone1_comp_g17}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout_g17}
-        \caption{100 - 200 keV lineout.}
-        \label{fig:bench_cone1_lineout_g17}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_comp_g18}
-        \caption{45 - 100 keV.}
-        \label{fig:bench_cone1_comp_g18}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout_g18}
-        \caption{45 - 100 keV lineout.}
-        \label{fig:bench_cone1_lineout_g18}
-    \end{subfigure}
-
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_comp_g19}
-        \caption{10 - 45 keV.}
-        \label{fig:bench_cone1_comp_g19}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/bench_cone1_lineout_g19}
-        \caption{10 - 45 keV lineout.}
-        \label{fig:bench_cone1_lineout_g19}
-    \end{subfigure}
-    \caption{Comparison of the MCNP6 results and the raytracer.}\label{fig:waterHist}
-\end{figure}
+Figure~\ref{fig:bench_cone1_centerdifffig} shows the results of the comparison of the group averaged uncollided flux values between MCNP6 and DOCTORS for the 19-group approximation.
+The plot shows the uncollided flux averaged over the center 8 voxels in the phantom.
+Only the uncollided flux is used in this benchmark because it is not dependent on the additional solver parameters such as quadrature and anisotropy treatment but instead has an exact solution.
+The agreement is within about 2\% for energy groups above 300 keV but falls off rapidly in the diagnostic energy range.
 
@@
     \end{subfigure}
-    \caption{Comparison}\label{fig:bench_cone1_centerdifffig}
+    \caption{Ratio of the MCNP6 uncollided flux to the DOCTORS uncollided flux for each group.
+(a) The entire range of data.
+(b) The same data scaled to show the range of interest.
+}\label{fig:bench_cone1_centerdifffig}
 \end{figure}
+
+To explore the reason for the discrepancy in the diagnostic range, the finely meshed uncollided flux in energy space is examined in Fig.~\ref{fig:bench_cone1_centerdiff3}.
+The uncollided flux at lower energy drops off rapicly at the low energy side of each group in the diagnostic regime.
+This is caused by the significant downscatter that is not accurately modeled with such wide energy groups.
+To address this, the 47-group data file was also run.
+Similar results are plotted in Fig.~\ref{fig:bench_cone1_centerdiff47}.
 
@@
   \begin{center}
-   \includegraphics[width=3.75in]{figs/bench_cone1_centerdiff3}
+   \includegraphics[width=\textwidth]{figs/bench_cone1_centerdiff3}
   \end{center}
-  \caption{The 19 group center flux comparison}
+  \caption{The 19 group center flux comparison for a uniform beam.}
 \label{fig:bench_cone1_centerdiff3}
@@
 
-A plot is below.
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/bench_cone1_centerdiff47}
-  \end{center}
-  \caption{The 47 group center flux comparison}
-\label{fig:bench_cone1_centerdiff47}
-\end{figure}
-
-The analytical points are computed
+As an additional verification that DOCTORS is computing correct results, analytical data points were added. The analytical points are computed
 \begin{equation}
@@
 
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=\textwidth]{figs/bench_cone1_centerdiff47}
+  \end{center}
+  \caption{The 47 group center flux comparison for a uniform beam.}
+\label{fig:bench_cone1_centerdiff47}
+\end{figure}
+
+The results shown in Fig.~\ref{fig:bench_cone1_centerdiff47} agree with the MCNP6 uncollided flux distribution much more accurately than for the 19-group approximation.
+Therefore, all results after this use the 47-group approximation exclusively.
+Ideally, a more refined dataset with \textit{only} photon data in the diagnostic range would be used, but currently, no such data set is available in DOCTORS.
+
+\subsection{Diagnostic Pulse}
+Once the 19 and 47 group datasets were compared using the uniform energy distribution, a second benchmark was created.
+This benchmark uses a 1-group pulse of photons in group 41 (70-75 keV).
+Using a single group pulse is not representative of a diagnostic x-ray beam, but it makes interpretation of the results simpler.
+Since all uncollided flux is in a single group, the downscatter and within-group inscatter can be differentiated explicitly which would not be otherwise possible.
+This facilitates debugging of code and validation against MCNP6.
+
+Figure~\ref{fig:b2_unc} shows the uncollided flux distribution for both MCNP6 and DOCTORS.
+Qualitatively, the distributions appear to agree very well except for the artifacts around the edges of the beam in the MCNP6 results.
+
+%The detailed flux distribution from MCNP6 and the solver are not included here, they are provided in Appendix~\ref{sec:appc_benchmark}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_dr_unc}
+        \caption{}
+        \label{fig:bench_cone1_centerdiff}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_mc_unc}
+        \caption{}
+        \label{fig:bench_cone1_centerdiff2}
+    \end{subfigure}
+    \caption{Comparison of the uncollided flux in (a) DOCTORS and (b) MCNP6. }\label{fig:b2_unc}
+\end{figure}
+
+Figure~\ref{fig:b2_col} shows a comparison of the MCNP6 collided flux and the DOCTORS collided flux.
+The DOCTORS flux exhibits noticable ray effect artifacts but good qualitative agreement otherwise with respect to the shape of the flux distribution.
+A vertical lineout of Fig.~\ref{fig:b2_col} is shown in Fig.~\ref{fig:b2_col_comp}.
+The lack of anisotropy results in more particles streaming out to the side of the beam and backscattering that expected.
+This causes a buildup when the beam enters the phantom and a rapid decrease as more particles scatter than predicted by MCNP6.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_mc_col}
+        \caption{}
+        \label{fig:b2_mc_col}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_dr_col}
+        \caption{}
+        \label{fig:b2_dr_col}
+    \end{subfigure}
+    \caption{The in-group collided flux in group 41 (70-75 keV)}\label{fig:b2_col}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/b2_col_comp}
+  \end{center}
+  \caption{Comparison of the in-group scatter in group 41 using an isotropic assumption.}
+\label{fig:b2_col_comp}
+\end{figure}
+
+Figure~\ref{fig:b2_aniso} shows lineouts for the in-group scatter with the uncollided flux (group~41) and the purely downscattered flux in group~42.
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_aniso_g41}
+        \caption{}
+        \label{fig:b2_aniso_g41}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_aniso_g42}
+        \caption{}
+        \label{fig:b2_aniso_g42}
+    \end{subfigure}
+    \caption{Linout showing the anisotropy. (a) Group 41 (70-75 keV). (b) Group 42 (60-70 keV). }\label{fig:b2_aniso}
+\end{figure}
+
+Figure~\ref{fig:b2_aniso} shows lineouts for the in-group scatter with the uncollided flux (group~41) and the purely downscattered flux in group~42.
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_quad_g41}
+        \caption{}
+        \label{fig:b2_quad_g41}
+    \end{subfigure}
+    ~ 
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{figs/b2_quad_g42}
+        \caption{}
+        \label{fig:b2_quad_g42}
+    \end{subfigure}
+    \caption{Linout showing the effect of quadrature. (a) Group 41 (70-75 keV). (b) Group 42 (60-70 keV). }\label{fig:b2_quad}
+\end{figure}
+
 In MCNP6, the effective dose was computed using a FMESH4 tally with associated DE and DF cards to weight the flux.
@@
 The comparison is shown in Fig.~\ref{fig:mcnp_f4vsf6}.
+The F6 and F4 tallies agreed within 2\% at the center of the phantom, which validates that the spatially distributed FMESH with a FM multiplier is a good approximation of the energy depositied in a voxel.
 
@@
 
-The effective dose and absorbed doses are given in Fig.~\ref{fig:b2_dose}.
+The effetive dose and equivalent doese in DOCTORS are computed using Eq.~\ref{eq:doseicrp} and~\ref{eq:dosedep} respectively. The effective dose and absorbed doses are given in Fig.~\ref{fig:b2_dose}.
 
@@
 
-INSERT THE RESULTS FROM THE 16 BEAM BENCHMARK
+This benchmark is design to test the multi-cone beam projection.
+Sixteen cone beams are simultaneously used to generate the data.
+Each cone beam fires the 1-group diagnostic pulse energy spectra described in Sec.~\ref{sec:cone1}.
 
-\section{16 Cone-beam Phantom Benchmark}
+\section{64 Cone-beam Phantom Benchmark}
 
@@
 
+\subsection{Patient Table}
+In some phantom models, a table the patient lies on is present.
+The gantry poses a challenge since it is a highly attenuating material relative to the patient which reduces the available flux.
+Further, the gantry is not mentioned in the literature which only consider the phantom.
+The CT number to material conversion developed by~\citet{ref:ottossonr} which is implemented in this work does not have an equivalent for the gantry region.
+
+All voxels whose CT number is above the highest bone value given by~\citet{ref:ottossonr} are assumed to be part of the gantry. The gantry is assumed to be aluminum with a density of 2.70 g/cm$^{3}$.
+
+Since the triangular artifacts were still seen in the 16-beam water phantom, the number of beams used in this benchmark was increased to 64.
+Also, each beam was moved further from the isocenter of the phantom.
+The total diameter between the detector and x-ray tube is 1 meter, thus the coordinate of the first projection is at $y$ = -25 cm.
+The cone beam was broadened to 30 degrees to more closely match the CT data.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/phantom19_64}
+        \caption{}
+        \label{fig:waterHistLin}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/phantom19_256}
+        \caption{}
+        \label{fig:waterHistLog}
+    \end{subfigure}
+    \caption{The (a) full phantom and (b) simplified phantom after the CT numbers are converted into materials.}\label{fig:waterHist}
+\end{figure}
+
 \section{GPU Acceleration}
 
-INSERT THE RESULTS FROM THE GPU ACCELERATION
+Table~\ref{tab:cpu} summarizes the runtime required for the CPU-only version of DOCTORS and Table~\ref{tab:gpu} summarizes the analogous runtime results for the GPU accelerated version.
+
+\begin{table}[ht]
+\caption{CPU Runtime (msec)}
+\centering 
+\begin{tabular}{l c r r}
+\hline \hline   
+Data Type & Method & $64 \times 64 \times 16$ & $256 \times 256 \times 64$\\ [0.5ex] 
+\hline
+\texttt{float}  & Raytracer & 824.9   &  \\
+                & Solver    & 15,541  &  \\
+\texttt{double} & Raytracer & 933.6   & 179,613 \\
+                & Solver    & 14,348   & 2,408,000 \\  [1ex]
+\hline
+\end{tabular}
+\label{tab:cpu}
+\end{table}
+
+\begin{table}[ht]
+\caption{GPU Runtime (msec)}
+\centering 
+\begin{tabular}{l c r r}
+\hline \hline   
+Data Type & Method & $64 \times 64 \times 16$ & $256 \times 256 \times 64$\\ [0.5ex] 
+\hline
+\texttt{float}  & Raytracer & 362.8   &  10,267 \\
+                & Solver    & 2,417.9  &  71,619 \\
+\texttt{double} & Raytracer & 362.6   & 13,452 \\
+                & Solver    & 4,554.5   & 61,355 \\  [1ex]
+\hline
+\end{tabular}
+\label{tab:cpu}
+\end{table}
 

### 2017-06-12 python/ct_data_reader/main/distribution_plotter.py
Updated file names

@@ -3,7 +3,7 @@ import matplotlib.pyplot as pyplot
 
 __author__ = 'etnc6d'
 
-with open("/media/Storage/thesis/build-doctors-Desktop_Qt_5_4_1_GCC_64bit-Debug/ctdist.dat", 'r') as f:
+with open("/media/Storage/thesis/doctors/ctunits.dat", 'r') as f:
     a = numpy.loadtxt(f)
 
 x = a[:-1, 0]
@@ -13,13 +13,17 @@ x -= shift
 overflow = a[-1, 1]
 
 pyplot.figure()
-pyplot.semilogy(x, v, 'r', [0, 0], [1, 1E5], 'k--', [-1000, -1000], [1, 1E5], 'k--')
-pyplot.title("HU Frequency (Liver)")
-pyplot.xlabel("Houndsfield Number")
-pyplot.ylabel("Count")
-pyplot.axis([-shift, 3000-shift, 0, 100000])
+pyplot.plot(x, v, 'b', [0, 0], [1, 1E5], 'k--')
+#pyplot.title("HU Frequency (Liver)")
+pyplot.xlabel("CT Number")
+pyplot.ylabel("Frequency")
+pyplot.axis([-shift, 50000-shift, 0, 100000])
 
 print("Overflows: " + str(overflow))
+
+sm = sum(v)
+f = sm/(256*256*64)
+print("sum = " + str(sm) + " = " + str(f))
 
 v = pyplot.cm.viridis(numpy.linspace(0, 256, 256, dtype=numpy.int))
 r = [x[0] for x in v]
@@ -29,7 +33,7 @@ b = [x[2] for x in v]
 for r, g, b in zip(r, g, b):
     # brushes.push_back(QBrush(QColor::fromRgbF(0.0000,    1.0000,    1.0000)));
     s = "brushes.push_back(QBrush(QColor::fromRgbF(" + str(r) + ",    " + str(g) + ",    " + str(b) + ")));"
-    print(s)
+    #print(s)
 
 #print(pyplot.cm.viridis(numpy.linspace(0, 256, 256, dtype=numpy.int)))
 

### 2017-06-12 python/ct_data_reader/main/meshsplitter.py

@@ -1,6 +1,6 @@
 __author__ = 'etnc6d'
 
-filename = '/media/Storage/thesis/mcnp.gitignore/meshtaz'
+filename = '/media/Storage/thesis/mcnp.gitignore/meshtan'
 
 file = open(filename, 'r')
 #with open(filename, 'r') as file:

### 2017-06-13 doctors/globals.h
Added a paragraph about flux moments

@@ -10,7 +10,7 @@
 #undef U16_T
 #endif
 
-#define SOL_T float
-#define RAY_T float
+#define SOL_T double
+#define RAY_T double
 
 #endif // GLOBALS_H

### 2017-06-13 docprep/norris_phd_dissertation.tex
Updated thesis text
Added some TODO

@@
 \begin{ThesisAbstract}
-This is the text for the abstract. Here, I explain what I did and how cool I am for doing it. Because I am awesome.
-
 Computed tomography (CT) has become pervasive in medical diagnostics as improved imaging techniques and processing algorithms provide higher quality information to doctors.
@@
 
-The DOCTORS code was benchmarked against MCNP6 and found to have very good agreement using both a water phantom and a realistic patient phantom.
-DOCTORS was also found to be much faster than MCNP6, even more so with GPU acceleration.
+The DOCTORS code was benchmarked against MCNP6 and found to have good qualitative agreement using both a water phantom and a realistic patient phantom.
+DOCTORS was also found to be much faster than MCNP6; MCNP takes hours to compute flux profiles that take less than a minute using DOCTORS.
 
-GPU
+A GPU algorithm was implemented that speeds up the DOCTORS code by a factor of 39.
 
@@
 I also appreciate the time spent by my committee members helping me identify problems and provide insight to solutions.
+I would also like to thank my committee members for their guidance throughout my PhD work.
 

### 2017-07-05 docprep/appdx_a.tex
Fixed typos and updated LaTeX template

@@
 For example, a problem requiring $S_6$ needs $6 \times 8 = 48$ directions, but a problem requiring $P_6$ needs 49 expansion coefficients which would seem approximately eqivalent, but typical problems require far more discrete directions than expansion coefficients.
-Comples problems can easily require $S_{16}$ or higher but expansions above $P_5$ are rarely encountered.
+Complex problems can easily require $S_{16}$ or higher but expansions above $P_5$ are rarely encountered.
 
@@
 \end{equation}
+holds.
 
@@
 \end{equation}
-where $\mu$ is the cosine of the angle between $\hat{Omega}$ and the $x$-axis and $\tau$ is the rotation about the $x$-axis with respect to the $y$-axis of $\hat{\Omega}$ projected onto the $yz$ plane which can be compuated as:
+where $\mu$ is the cosine of the angle between $\hat{\Omega}$ and the $x$-axis and $\tau$ is the rotation about the $x$-axis with respect to the $y$-axis of $\hat{\Omega}$ projected onto the $yz$ plane which can be compuated as:
 \begin{equation}
@@
 \begin{equation}
-\int_{}^{}Y_{lm}(\hat{Omega})Y_{l'm'}^*(\hat{\Omega}) d\hat{\Omega}= 
+\int_{}^{}Y_{lm}(\hat{\Omega})Y_{l'm'}^*(\hat{\Omega}) d\hat{\Omega}= 
 \begin{cases}

### 2017-07-05 docprep/chapbib.tex
Fixed typos and updated LaTeX template

@@
 However, dose quantification is not currently patient specific.
-Measuring the dose in a patient directly is impossible, intead the air kerma (measured in Gy), which measures the x-ray beam intensity, and can be measured directly with ionization chambers~\citep{ef:wolbarsta}.
+Measuring the dose in a patient directly is impossible, instead the air kerma (measured in Gy), which measures the x-ray beam intensity, and can be measured directly with ionization chambers~\citep{ref:wolbarsta}.
 As the x-ray beam passes through the patient, it deposits energy.
@@
 The CTDI quantifies the equivalent dose the patient receives from a single slice of the CT scan protocol.
-The CTDI does \textit{not} capture patient-specific papameters, but rather quantifies the amount or radiation emitted by the source.
+The CTDI does \textit{not} capture patient-specific papameters, but rather quantifies the amount of radiation emitted by the source.
 
@@
 The DLP is related to the effective dose to the patient for a particular procedure by a conversion factor that is dependent on many variables including the anatomical region scanned, the x-ray tube voltage, and the patient's age, gender, and size~\citep{ref:hudaw}.
-
-INSERT MORE EXAMPLES.
-\citet{ref:lauk} found that severly obese patients received twice as much dose from CT procedures as others on average. Some patients received as much as five times the typical dose.
+For example, \citet{ref:lauk} found that severly obese patients received twice as much dose from CT procedures as others on average.
+Some patients received as much as five times the typical dose.
 
@@
 The DOCTORS code relies on his proposed 19-group model.
-Table~\ref{table:ctmap} shows the mapping of CT number to material composition (reproduced from \citep{ref:ottossonr}).
+Table~\ref{table:ctmap} shows the mapping of CT number to material composition (reproduced from \citet{ref:ottossonr}).
 
@@
 
-INSERT KBA DISCUSSION
-
 \section{Raytracing}
 
-Many raytracing algorithms are based on the Bresenham line raserizing algorithm which was designed for graph plotting hardware~citep{ref:bresenhamj}.
+Many raytracing algorithms are based on the Bresenham line raserizing algorithm which was designed for graph plotting hardware~\citep{ref:bresenhamj}.
 That algorithm was intended to produce lines acceptable for human interpretation and no guarantee is made that all pixels along the line will be identified making this class of algorithms inappropriate for dosimetric raytracing since attenuation through some voxels would be skipped.
@@
 
-INSERT SECTION ON GPU RAYTRACING
-
 \section{GPU Acceleration}
@@
 A graphics processing unit (GPU) is similar to a central processing unit (CPU)in that it executes instructions to process data.
-However, it differs in that a CPU may have up up to 22 physical cores, each of which can operate independently, a GPU has up to thousands of cores, but they cannot operate independently.
+However, it differs in that a CPU may have up to 22 physical cores, each of which can operate independently, a GPU has up to thousands of cores, but they cannot operate independently.
 Instead, cores execute simulaneously in groups called warps.
@@
 
-Two major GPU manufacturers control the vast majority of the market, Nvidia and AMD [CITE].
+Two major GPU manufacturers control the vast majority of the market, Nvidia and AMD.
 In 2007, Nvidia released its Compute Unified Device Architecture (CUDA) language.
@@
 
-Monte Carlo dose computation for protons has been accelerated by GPU technology \citep{ref:jiax} and Monte Carlo in voxelized geometries has been done by \citep{ref:hissoinys} for monoenergetic beams.
+%Monte Carlo dose computation for protons has been accelerated by GPU technology \citep{ref:jiax} and Monte Carlo in voxelized geometries has been done by \citep{ref:hissoinys} for monoenergetic beams.
 

### 2017-07-05 docprep/chapconc.tex
Added some TODO
Fixed typos and updated LaTeX template

@@
 
-The collided flux suffers from the isotropic approximation made to simplify the code.
+Qualitatively, the most of the trends identified by DOCTORS agreed very well with MCNP6.
+The largest discrepancy was in the attenuation of particles through the patient region.
+However, since the raytracer is very accurate with respect to MCNP6, the difference is likely with the discrete ordinate methodology rather that cross section related.
+The key driver of this discrepancy is believed to be due to poor treatment of the angular transport either by way of the quadrature or anisotropy treatment.
+
+This indicates that DOCTORS is better suited for problems that use a broad, distributed source as opposed to a narrow, directionally biased source.
+Such directional sources, require much higher quadratures to accurately characterize the beam due to ray effect.
 
@@
 With further development, DOCTORS is well positioned to become a code of significant clincal impact.
+Though it is still lacking in medical diagnostic quality accuracy, it is easily extensible and capable of characterization of complex beams making it ideal for some types of clinical dosimetry.
 
@@
 However, usage in a clinical setting would likely require further refinement of the GUI to make the code more robust and intuitive.
-Additionally, the direct output of DOCTORS may not necessarily be of clinical importance since
+Additionally, the direct output of DOCTORS may not necessarily be of clinical importance since radiologists and technitians are more interested in compliance with regulations and ensuring patient safety.
+Therefore, a bulk report of overall numbers would likely be of more benfit than the entire spatial flux and dose distribution in a clinical setting.
 
@@
 
-The GPU acceleration algorithm was found to speed the DOCTORS code up by a factor of ~40x.
+The GPU algorithm was implemented in CUDA and is very straightforward.
+At each subsweep, the global index of all voxels that can be computed in parallel are determined and a kernel is launched that executes those voxels.
+Unsurprisingly, this algorithms scales much better on larger problems where more time is spent in the computational execution of the kernal as opposed to the overhead operations of launching the kernel and memory transfer.
+
+The GPU acceleration algorithm was found to speed the DOCTORS code up by a factor of ~40x for large problems and only a factor of 3-6x for smaller problems.
+The improved accuracy of double precision arithmetic was found to outweigh the speed improvement from faster calculations.
 
@@
 The other changes are additions to DOCTORS that would add new capabilities to the software.
-
 In addition to specific modifications, some additional, broader future goals can also be identified.
 
-\subsection{Anisotropy Treatment}
+\subsection{Anisotropy treatment}
 A more sophisticated anisotropy treatment would be appropriate since the collided flux was found to have issues regarding its behavior.
+Alternatively, adding a spherical harmonics solver may help the anisotropy treatment as well, or at least reduce its memory footprint and runtime.
+This would require moving the solution from discrete angle space to flux moment space which can greatly complicate the debugging.
+However, once the correctness of the overall algorithm is shown, this step should be relatively simple.
 
-\subsection{Group Structure}
-
+\subsection{Group structure}
 The cross section data currently used by DOCTORS are taken from SCALE6.2.
 While these cross sections have been found to be sufficient to produce flux distributions in medical CT imaging, a more refined group structure designed for medical applications would be worth investigation.
-
-This work can be done using either NJOY [CITE] or the newly released AMPX [CITE] code.
+Also, the data distribtuion from SCALE6.2 contains potentially export controlled information since it also includes nuclear reactor materials.
+A medical purposed cross section library would alleviate this problem and allow the code to be freely released with a dataset.
+In the current version of DOCTORS, the user is required to obtian the cross section data files independently.
+This work can be done using either NJOY or the newly released AMPX code.
 Either code has the capability to collapse an ENDF formatted data file into a group structure of the user's choosing in the format DOCTORS reads.
 
-INSERT SECTION ABOUT NJOY or AMPX.
-
-\subsection{Temporal Extension}
+\subsection{Therapy extension}
 An advantage of the raytracer is that it is very fast.
@@
 This would allow DOCTORS to model more complex beam shapes and scan protocols accurately.
+This would be particularly useful for medical treatment systems that use a multileaf collimator.
 
-\subsection{Partial Acceptance Criteria}
+Some treatment systems employ a multileaf collimator to continuously shape the beam, resulting in a large dose deposition only at the area of interest.
+DOCTORS can likely be extended to high energy therapeutic beams for clinical treatment.
+However, higher energy photons will scatter more anisotropicly.
+This will require careful analysis of the angular treatment used.
+
+\subsection{Partial acceptance criteria}
 In the current version of doctors, the raytracer is very accurate with respect to MCNP6 in predicting the uncollided flux, except for along the periphery of the beam.
@@
 
-\subsection{Multileaf Collimation}
-Some medical systems currently utilize a multileaf collimator [CITE] to deliver dose to a specific organ or region of the body. As the
-
-\subsection{Extension to Therapy Beams}
-DOCTORS can likely be extended to high energy therapeutic beams for clinical treatment.
-However, higher energy photons will scatter more anisotropicly.
-This will require spherical harmonics and high order $P_l$ expansion tables in the cross section data.
-
-\subsection{Organ Identification}
+\subsection{Organ identification}
 Currently, DOCTORS cannot automatically identify specific organs, thus it is only able to compute the equivalent dose via dose deposition.

### 2017-07-05 docprep/chapimpl.tex
Fixed typos and updated LaTeX template

@@
 At the time of this writing, the repository is located at \texttt{https://github.com/eNorris/thesis} and is only available subject to special request and license agreement.
-An executable version is available.
+An executable version can also be made available upon request.
+An executable-only public release is planned for the future.
 
@@
 
-INSERT NOTE ABOUT CODE LISTING FORMATTING
+Unless noted otherwise, all code listings are formatted as C++ code though some parts of the code may be omitted.
+Omissions include typical "boiler plate" code that accompanies all code such as the \texttt{\#include} statements and any \texttt{using namespace} commands.
+Additionally, the \texttt{main()} function and others are omitted, showing only the code relevant to the discussion.
 
@@
 \end{equation}
-This pattern extends to multi-dimensional matrices. For example, a $N_x \times N_y$ matrix would be globally indexed as
+This pattern extends to multi-dimensional matrices. For example, a $N_x \times N_y \times N_z$ matrix would be globally indexed as
 \begin{equation}
@@
 AMPX cross sections are stored in a binary format.
+These files are \textit{not} distributed with DOCTORS but must be obtained from SCALE6.2 or another code.
 Figure~\ref{fig:ampx} shows the data format of the overall file.
 The binary file is split into a sequence of records.
-The header section of the file has overall information about the file including the number of nuclides stored, their energy group structure, and a comment string describing file.
+The header section of the file has overall information about the file including the number of nuclides stored, their energy group structure, and a comment string describing the file.
 Following the header section is a list of directory records.
@@
 The header always reports the number of 8-bit bytes required to store the data.
-However, some data types, such as texttt{char}, are only one byte so each word represents multiple (4 in the case of \texttt{char}) distinct characters.
+However, some data types, such as \texttt{char}, are only one byte so each word represents multiple (4 in the case of \texttt{char}) distinct characters.
 
@@
 Qt5 prvides a user interface for building user interfaces.
-Components such a sbuttons, drop boxes, radio buttons, etc.
+Components such as buttons, drop boxes, radio buttons, etc.
 are made available to the user.
-Through extensive use of polymorphism, Qt simplifies the addition of graphical interactive elements called widgets.
+Through extensive use of polymorphism, Qt5 simplifies the addition of graphical interactive elements called widgets.
 All widgets inherit from the base \texttt{QWidget} class which inherits from \texttt{QObject}.
@@
 
-Listing~\ref{lst:sync1} gives a pseudocode snippet that has a long function that will block the user interface.
-A corresponding dequence diagram is given in Fig.~\ref{fig:sync1}.
+Listing~\ref{lst:sync1} gives a C++ snippet that has a long function that will block the user interface.
+A corresponding sequence diagram is given in Fig.~\ref{fig:sync1}.
 When the user interacts with the GUI, the writer object begins executing the \texttt{doWrite()} function.
-Until this function completes, the UI thread will be busy an dunable to handle additionaly user interaction or updates.
+Until this function completes, the UI thread will be busy and unable to handle additional user interaction or updates.
 This results in the GUI becoming unresponsive and potentially issuing a warning to the user from the operating system.
+The code in Listing~\ref{lst:sync1} is updated to use the Qt signal/slot mechanism whose code is given in Listing~\ref{lst:async1}.
 
@@
 	// When a button named button1 which was created in the Qt5
-	creator interface is clicked, the clicked() signal is 
-	automatically emitted which executes the doSomethingLong()
-	function
+	//   creator interface is clicked, the clicked() signal is 
+	//   automatically emitted which executes the doSomethingLong()
+	//   function
 	connect(ui->button1, clicked(), this, 
@@
 \end{listing}
-
-\begin{figure}[tb]
-  \begin{center}
-   \includegraphics[width=3.75in]{figs/writer_sync}
-  \end{center}
-  \caption{Sequence diagram for the synchronous call.}
-\label{fig:sync1}
-\end{figure}%
-
-The code in Listing~\ref{lst:sync1} is updated to use the Qt signal/slot mechanism whose code is given in Listing~\ref{lst:async1}.
 
@@
   \begin{center}
+   \includegraphics[width=3.75in]{figs/writer_sync}
+  \end{center}
+  \caption{Sequence diagram for the synchronous call.}
+\label{fig:sync1}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
    \includegraphics[width=3.75in]{figs/writer_async}
@@
 \label{fig:async1}
-\end{figure}%
+\end{figure}
 
@@
 
-INSERT PICTURES OF THE GUI
+The data display plots that show materials and other geometry as well as the flux output use an implementation of the built-in \texttt{QGraphicsView} object to render a matrix of \texttt{QGraphicsRectItem}s on the screen.
+The data to be drawn on the screen is populated and some metrics such as the dataset minimum and maximum are computed.
+The data value at each point is evaluated to determine which bin it belongs to.
+
+The first tab is the geometry input, shown in Fig.~\ref{fig:gui_tab1}.
+In the current version of DOCTORS, the only supported format is unsigned 16-bit binary files whose size is known \textit{a priori}.
+When the user clicks on the "Open" button, the user is prompted with a dialog for selection of a binary file.
+The file is read in as a list of $N_x \times N_y \times N_z$ CT numbers.
+As soon as the data is read in and some checks are performed, the user is prompted for selection of a CT-number-to-material conversion.
+Currently, two conversions are included in DOCTORS.
+The first is a water/air only conversion and the second is a more complex conversion to a series of dosimetrically equivalent materials representative of a human patient.
+Once the user specifies a conversion, the geometry explorer becomes available which allows the user to visualize the geometry file to verify that it was read and interpreted correctly.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/gui3}
+  \end{center}
+  \caption{The geometry selection tab. The user must first specify the dimensionality of the voxel phantom and then load it from a binary file.}
+\label{fig:gui_tab1}
+\end{figure}
+
+Some screenshots of the geometry explorer are shown in Fig.~\ref{fig:gui_geom}.
+The geometry viewer can show slices through the voxel phantom along all three major planes.
+The depth of the slice viewed is changed by moving the scroll bar in the center of the viewer up or down.
+As the scroll bar moves, the number at its bottom is updated accordingly to indicate the depth (in voxels) of the slice.
+In addition to the material, the geometry viewer can render the physical density (in g/cm$^3$) and the atom density (in atom/b-cm) as well as the raw CT number used to generate the material and density.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/gui1}
+        \caption{}
+        \label{fig:gui1}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/gui4}
+        \caption{}
+        \label{fig:gui4}
+    \end{subfigure}
+    \caption{The geometry explorer. (a) The material viewer which shows which material each voxel was converted to. (b) The atom density plot.}\label{fig:gui_geom}
+\end{figure}
+
+The next three tabs identify the cross section dataset to use for the material generation, the quadrature, and the solution anisotropy treatment used.
+All three tabs are relatively strightforward and contain few widgets.
+The final tab, shown in Fig.~\ref{fig:gui_src} defines the source.
+The user is presented with a number of different source geometries to choose from.
+Each type is described more fully in Sec.~\ref{sec:source_gen}.
+Clicking on the "Energy Distribution" button will reveal the popup shown in Fig.~\ref{fig:gui_spect}.
+This popup cannot be fully populated until \textit{after} the cross section is loaded in the "Cross Section" tab since the cross section data file defines the energy group structure.
+
+The energy distribution dialog, shown in Fig.~\ref{fig:gui_spect}, can plot the $x$ axis on either a linear or logarithmic scale.
+The user can click and drag to "paint" a spectrum to generate quick, qualitative spectra suiting the user's needs.
+Alternatively, the user can select from any of the preset spectra in the dropdown menu.
+Currently, only the uniform beam and diagnostic pulse distributions described later are present.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/gui2}
+  \end{center}
+  \caption{The source tab. The user must select a geometry type and enter all applicable values as well as define the source energy spectrum.}
+\label{fig:gui_src}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/gui6}
+  \end{center}
+  \caption{The spectrum energy distribution.
+The spectrum can be displayed on either a linear or logarithmic scale on the $x$ axis.
+The user can either click and drag on the spectrum to "paint" or select a preset.}
+\label{fig:gui_spect}
+\end{figure}
+
+Once all necessary data is loaded, the "Launch" button becomes active and will remain so until the user creates a conflicting set of input that would prevent the solver from being able to run.
+When the user clicks on the launch button, the raytracer begins running asynchronously and the output display shown in Fig.~\ref{fig:gui_out} opens automatically.
+When the raytracer completes, the output dialog is updated with the flux information and the solver begins executing asynchronously.
+As each iteration completes, the output dialog is progressively updated so that the user can monitor the evolution of the solution.
+
+The output dialog has controls very similar to the geometry explorer.
+The user can freely choose the plane to slice through the solution and can plot the solution on either a linear or logarithmic scale.
+The "Distribution" radio buttons allows the user to choose whether the uncollided, collided, or total flux is shown.
+The "Level Local Scale" checkbox allows the user to control whether the current slice colormap is scaled to the contents of the viewed slice or the contents of the entire voxel phantom.
+Checking the "Level Local Scale" box will ensure that the viewed data is normalized to include only the viewed data.
+Unchecking the box will scale the viewed data based on the range of the entire voxel phantom.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/gui5}
+  \end{center}
+  \caption{The output dialog. The controls are similar to those found in the geometry viewer.}
+\label{fig:gui_out}
+\end{figure}
+
+In the current version of DOCTORS, dosimetric output is only written to an output file and cannot be visualized graphically.
+The output format is designed to be human readable and easily read by any major programming language.
+Listing~\ref{lst:output_fmt} gives the code to read the file format where the \texttt{prod()} algorithm is given in Listing~\ref{lst:prod}.
+Note that \texttt{LIST} may be any appropriately typed list object such as a \texttt{std::vector} and a \texttt{LIST2D} is a list of lists or vector of vectors.
+
+\begin{listing}
+\begin{minted}[frame=lines,linenos]{c++}
+// Declare variables
+int N;
+LIST S;
+LIST2D dims;
+LIST data;
+
+// Read the number of dimensions
+READ_INT(N);
+
+// Read the size of each dimension
+for(int i = 0; i < N; i++)
+	READ_INT(S[i]);
+
+// Read the axis labels
+for(int i = 0; i < N; i++)
+	for(int j = 0; j < S[i]; j++)
+		READ_FLOAT(dims[i][j]);
+
+// Read the data
+int dataElem = prod(S);
+for(i = 0; i < dataElem; i++)
+	READ_FLOAT(data[i]);
+\end{minted}
+\caption{Algorithm to read the output format used by DOCTORS.}\label{lst:output_fmt}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[frame=lines,linenos]{c++}
+int prod(LIST p)
+{
+	int product = 1;
+	for(int i = 0; i < p.SIZE(); i++)
+		product = product * p[i];
+	return product;
+}
+\end{minted}
+\caption{The \texttt{prod()} algorithm}\label{lst:prod}
+\end{listing}
 
 \section{CUDA}\label{sec:cuda}
-CUDA code is compoiled with the Nvidia complier nvcc.
-Qt5 uses the gcc compiler and its MOC generator for meta code.
-In order to connect CUDA code to the Qt MOC, the CUDA code is compiled by nvcc to produce a .o file.
-Qt5 then compiles all other files into corresponding.o files.
-The linker than automactically picks up all of the .o files gneerated by nvcc.
-The final result is an executable that has a Qt5 generarted user interface that can communicat with an Nvidia graphics card through the CUDA language.
+CUDA code is compiled with the Nvidia complier \texttt{nvcc}.
+Qt5 uses the \texttt{gcc} compiler and its MOC generator for meta code.
+In order to connect CUDA code to the Qt MOC, the CUDA code is compiled by \texttt{nvcc} to produce an object (\texttt{.o}) file.
+Qt5 then compiles all other files into corresponding object files.
+The linker then automactically picks up all object files generated by \texttt{nvcc}.
+The final result is an executable that has a Qt5 generated user interface that can communicate with an Nvidia graphics card through the CUDA language.
 
-The sweep through the mesh is different.
-
-\begin{table}[ht]
-\caption{Subsweep Indices}
-\centering 
-\begin{tabular}{l | c | c | c | c}
-  \hline \hline   
-  i  & ix & iy & iz & ix+iy+iz \\ [0.5ex] % inserts table 
-  \hline
-  0  & 4 & 0 & 0 & 4\\
-  1  & 3 & 1 & 0 & 4\\
-  2  & 3 & 0 & 1 & 4\\
-  3  & 2 & 2 & 0 & 4\\
-  4  & 2 & 1 & 1 & 4\\
-  5  & 2 & 0 & 2 & 4\\
-  6  & 1 & 3 & 0 & 4\\
-  7  & 1 & 2 & 1 & 4\\
-  8  & 1 & 1 & 2 & 4\\
-  9  & 1 & 0 & 3 & 4\\
-  10 & 0 & 4 & 0 & 4\\
-  11 & 0 & 3 & 1 & 4\\
-  12 & 0 & 2 & 2 & 4\\
-  13 & 0 & 1 & 3 & 4\\
-  14 & 0 & 0 & 4 & 4\\ [1ex]      % [1ex] adds vertical space
-  \hline
-\end{tabular}
-\label{table:subsweep}
-\end{table}
-
-We first split the sweep process into $N_L$ subsweeps.
-For a $N_x \times N_x \times N_x$ mesh, the subsweep process is straightforward.
-Some of the initial subsweeps are shown for a $8 \times 8 \times 8$ mesh in Fig.~\ref{fig:subsweep_cube}.
-This is valid for any direction, $\hat{\Omega}$, for which all three of its direction cosines ($\mu$, $\xi$, and $\eta$) are positive.
+In order to accelerate the discrete ordinate solution on a GPU, the concurrent tasks at any given moment must be known.
+In the single-threaded version of the code, the solver sweeps through the voxels one by one in a pre-determined fashion based on the direction.
+At the very beginning, the input flux to a single voxel is known from its boundary conditions.
+However, once that voxel's flux is computed, all three of its outgoing flux values enable three voxels to be computed independently of each other.
+After those three, six can be computed.
+Each layer of voxels whose flux can be computed independently is called a "subsweep." Figure~\ref{fig:subsweep_cube} illustrates some subsweeps through a cubic mesh.
 
@@
 
-In general, blah blah.
+In order to parallelize the sweep through the mesh, the global index of each voxel that can be computed in the $S$ subsweep must be known.
+When those indices are known, each voxel in subsweep $S$ can be solved independently by a CUDA kernel.
+This task is greatly simplified by noticing that the $x$, $y$, and $z$ indices of all voxels in the fourth subsweep, shown in Fig.~\ref{fig:subsweep_cube} all sum to 4 as shown in Table~\ref{table:subsweep}.
+
+\begin{table}[ht]
+\caption{Subsweep Indices}
+\centering 
+\begin{tabular}{l | c | c | c | c}
+  \hline \hline   
+  i  & ix & iy & iz & ix+iy+iz \\ [0.5ex] % inserts table 
+  \hline
+  0  & 4 & 0 & 0 & 4\\
+  1  & 3 & 1 & 0 & 4\\
+  2  & 3 & 0 & 1 & 4\\
+  3  & 2 & 2 & 0 & 4\\
+  4  & 2 & 1 & 1 & 4\\
+  5  & 2 & 0 & 2 & 4\\
+  6  & 1 & 3 & 0 & 4\\
+  7  & 1 & 2 & 1 & 4\\
+  8  & 1 & 1 & 2 & 4\\
+  9  & 1 & 0 & 3 & 4\\
+  10 & 0 & 4 & 0 & 4\\
+  11 & 0 & 3 & 1 & 4\\
+  12 & 0 & 2 & 2 & 4\\
+  13 & 0 & 1 & 3 & 4\\
+  14 & 0 & 0 & 4 & 4\\ [1ex]      % [1ex] adds vertical space
+  \hline
+\end{tabular}
+\label{table:subsweep}
+\end{table}
+
+The special case of the cube shown in Fig.~\ref{fig:subsweep_cube} can be extended to a more general case illustrated in Fig.~\ref{fig:subsweep_general}.
+Notice that the aforementioned intuition that $i_x + i_y + i_z = S$ is not necessarily true.
+The final subsweep ($S = 14$) contains a single voxel even though many combinations of three integers will sum to 14.
+The additional constraint is
+\begin{equation}\label{eq:constraints}
+\begin{split}
+i_x &< N_x \\
+i_y &< N_y \\
+i_z &< N_z.
+\end{split}
+\end{equation}
 
@@
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
+    ~ 
     \begin{subfigure}[b]{0.2\textwidth}
@@
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-    %(or a blank line to force the subfigure onto a new line)
+    ~ 
     \begin{subfigure}[b]{0.2\textwidth}
@@
 
-The number of parallel tasks, $P$, that can be done on subsweep $S$ of an $N_x \times N_y \times N_z$ mesh is given by Eq.~\ref{eq:taskspersub}.
-
+The constraints listed in Eq.~\ref{eq:constraints} can be used to compute the total number of parallel tasks in any subsweep.
+The number of parallel tasks, $P$, that can be done on subsweep $S$ of an $N_x \times N_y \times N_z$ mesh is given by
 \begin{equation}\label{eq:taskspersub}
@@
 \end{equation}
-
-\begin{equation}
+where $C_S$, $L_x$, $L_y$, $L_z$, $G_{xy}$, $G_{yz}$, and $G_{xz}$ are defined by Equations \ref{eq:cs}-\ref{eq:dxz}.
+\begin{equation}\label{eq:cs}
 C_S = \frac{(S+1)(S+2)}{2}
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
-
-\begin{equation}
+\begin{equation}\label{eq:dxz}
 d_{xz} = \max(S+1-N_x - Nz, 0)
@@
 However, in practice, this results in a race condition on the GPU hardware.
+As such, parallelization in angular space is not implemented in the current version of DOCTORS.
 
-DELETE BELOW HERE.
+\section{MCNP6 Generation}\label{sec:mcnpgen}
+DOCTORS has the capability to generate MCNP6 input files from the CT mesh data and source specification provided by the user.
+The input file is procedurally generated using the information provided to DOCTORS.
+If insufficient information is provided, an error message is generated and no output file is made.
+Note that the output is always named "mcnp\_out.inp" so it will \textit{override} existing files.
+To save a file, it must be manually renamed to avoid accidental deletion!
 
-The number of elements before the 4$^{th}$ level is 1 + 2 + 3 = 6.
-This is simply the sum of all integers less than the level number given by Eq.~\ref{eq:levelsum}.
-This is conincidentally the index of the first element of that level.
-However, there is a special case of $N_0 = 1$.
+The geometry is the largest section of the MCNP6 input file.
+The dimensions of the phantom mesh are used to generate planar surfaces at the appropriate locations.
+Those surfaces are then arranged into cells defining each voxel.
+The density information is pulled directly from the internally generated data built during the CT-number to material conversion.
 
-\begin{equation} \label{eq:levelsum}
-N_L = \sum_{i=1}^{L} = \frac{L(L+1)}{2}
-\end{equation}
-
-Given an index, $i$, the level to which it belongs can be computed by substituting $i$ for $N_L$ in Eq.~\ref{eq:levelsum} and solving for $L$. The index value is then floored.
-
-\begin{equation} \label{eq:levelquadratic}
-i = \frac{L(L+1)}{2} \rightarrow L^2 + L - 2i = 0
-\end{equation}
-
-\begin{equation} \label{eq:levelquadsol1}
-L = \left \lfloor{\frac{-1 \pm \sqrt{1+8i}}{2}} \right \rfloor
-\end{equation}
-
-Since the index must be a positive value, the $\pm$ sign can be removed from Eq.~\ref{eq:levelquadsol1} yielding Eq.~\ref{eq:levelquadsol2}
-
-\begin{equation} \label{eq:levelquadsol2}
-L = \left \lfloor{\frac{-1 + \sqrt{1+8i}}{2}} \right \rfloor
-\end{equation}
-
-The maximum index of any level in a the $N$ subsweep is $N$.
-Each subsweep adds an additional level.
-Therefore, the ix, iy, and iz indices can be computed using Eq.~\ref{eq:levelix},~\ref{eq:leveliy}, and~\ref{eq:leveliz} respectively.
-
-\begin{equation} \label{eq:levelix}
-ix = N - N_L
-\end{equation}
-
-\begin{equation} \label{eq:leveliy}
-iy = L + N_L - i
-\end{equation}
-
-\begin{equation} \label{eq:leveliz}
-iz = L + i - N_L
-\end{equation}
-
-\section{MCNP Generation}\label{sec:mcnpgen}
-DOCTORS has the capability to generate MCNP6 input files from the CT mesh data and source specification provided by the user.
+The source is the most difficult component to generate.
+Currently, all source types except fan beams can be generated automatically.
+The difficulty in generating fan beams comes from an intrinsic limitation of MCNP6.
+MCNP6 can only model cone beams through a biasing mechanism; it cannot model other kinds of beams.
+In practice, this is rarely a major limitation since physical collimators can be added to a input model to generate the desired beam shape.
+However, since DOCTORS already has a geometry mesh, addition of collimators is very difficult since the collimator becomes "smeared" across voxels.
 
 \section{Hardware}\label{sec:hardware}
-For this work, a computer with an Intel i7-5960X 8 core (16 hyperthreads) processor with a base clock speed of 3.X GHz and an Nvidia Titan Z graphics card was used.
+For this work, a computer with an Intel i7-5960X 8 core (16 hyperthreads) processor with a base clock speed of 3.5 GHz and an Nvidia Titan Z graphics card was used.
 Currently, if the problem requires more memory than is available on the GPU, the problem can still be solved, but much more time will be required to to copy overhead between the CPU and GPU.
@@
 
-CUDA is Nvidia only so AMD is not considered.
+Only Nvidia GPUs are recognized by DOCTORS since it relies on the CUDA interface which is a Nvidia proprietary product.
+There are no limitations on the CPU except that it be Little-Endian based (nearly all major CPUs meet this requirment).
+Big-Endian hardware would fail to parse the data files correctly.
 

### 2017-07-05 docprep/chapintro.tex
Fixed typos and updated LaTeX template

@@
 
-Since its introduction in 1973 by \citet{ref:hounsfieldg} computed tomography (CT) has become pervasive in medical diagnostics as improved algorithms and techniques give doctors access to higher quality information.
+Since its introduction in 1973 by \citet{ref:hounsfieldg}, computed tomography (CT) has become pervasive in medical diagnostics as improved algorithms and techniques give doctors access to higher quality information.
 Increased information results in faster and more accurate diagnosis.
@@
 The GUI was built using the Qt5 graphics framework for the windows, buttons, and other necessary widgets.
-The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicated required and completed steps.
+The GUI leads the user through the steps necessary to use the code in an intuitive way by using colors to indicate required and completed steps.
 The output is then plotted graphically as well as sent to an ASCII text file for more advanced postprocessing.
@@
 
-INSERT BRIEF DISCUSSION OF RESULTS
+Overall, good qualitative agreement was found between the reference code, MCNP6, and DOCTORS.
+The primary source of error is believed to be the angular treatment of the transport which is causing the flux inside the patient to be underestimated while the flux at the very periphery is overestimated.
 
-Another, similar, application is dose estimation of patients receiving radiation therapy.
+In the future, DOCOTRS may be extended to other application domains such as dose estimation of patients receiving radiation therapy.
 Often, before the procedure, a time dependent, 4D CT scan of the patient is taken so that radiologists can account for breathing patterns during administration of the treatment~\citep{ref:pant}.

### 2017-07-05 docprep/chapmeth.tex
Added a paragraph about flux moments
Fixed typos and updated LaTeX template

@@
 
-\subsection{The Boltzmann Equation}
+\subsection{The Boltzmann equation}\label{sec:lbe}
 
@@
 
-\subsection{Angular Discretization}
+\subsection{Angular discretization}
 The coordinate system used in DOCTORS is shown in Fig.~\ref{fig:coord_sys}.
@@
 The $LQ_N$ quadrature (which is the specific $S_N$ implemented in DOCTORS) uses the remaining degree of freedom to minimize error caused when integrating Legendre polynomials since that class of polynomials are used to represent anisotropy in scattering (see Sec.~\ref{sec:aniso}).
-For a more through discussion of quadrature generation, see \citet{ref:lewise}.
-Figure~\ref{fig:quad} shows the simples three quadratures implemented in DOCTORS.
+For a more thorough discussion of quadrature generation, see \citet{ref:lewise}.
+Figure~\ref{fig:quad} shows the simplest three quadratures implemented in DOCTORS.
 
@@
 
-\subsection{Energy Discretization}
+\subsection{Energy discretization}
 
@@
 
-The user must provide cross section data in the form of an AMPX formatted binary file (Format detaisl are provided in Section~\ref{sec:xsparse}).
+The user must provide cross section data in the form of an AMPX formatted binary file (Format details are provided in Section~\ref{sec:xsparse}).
 The cross section data file contains group structure information.
@@
 The weighting function can be any of a number commonly used functions.
-The SCALE data libraries use a uniform function [CITE].
+The SCALE data libraries use a uniform function.
 \begin{equation}\label{eq:groupxs}
@@
 
-In photon based problems, upscatter is typically negligible [CITE].
+In photon based problems, upscatter is typically negligible.
 By the time photons become low enough in energy to upscatter significantly, they are no longer of dosimetric concern.
@@
 
-\subsection{Spatial Discretization}
+\subsection{Spatial discretization}
 
-The problem domain is split into an evenly spaced Cartesian grid. Uniform grid spacing is not required by discrete ordinate methods though it is helpful and can speed up the computation[CITE].
+The problem domain is split into an evenly spaced Cartesian grid. Uniform grid spacing is not required by discrete ordinate methods though it is helpful and can speed up the computation.
 
@@
 More details about the flattening implementation used in DOCTORS is given in Section~\ref{sec:flatten}.
-Using global indexing discussed, the fully discretized the steady state LBE given is
+Using global indexing discussed, the fully discretized steady state LBE is
 \begin{equation} \label{eq:boltz_i}
@@
 
-\subsection{Diamond Difference Approximation}
+\subsection{Diamond difference approximation}
 
@@
 
-\subsection{Convergence Criteria}
+\subsection{Convergence criteria}
 The typical convergence criteria used to determine whether or not to terminate further iteration is the maximum relative error from the previous iteration to the current iteration given in Eq.~\ref{eq:conv} over the phase space $\mathbb{R}$ where $t$ is the current iteration number.
@@
 
-ADD FIGURE TO SHOW CONVERGENCE RATE.
+%ADD FIGURE TO SHOW CONVERGENCE RATE.
 
@@
 
-\subsection{Anisotropy Treatment}\label{sec:aniso}
-
-INSERT WHY NOT USE FLUX MOMENTS
+\subsection{Anisotropy treatment}\label{sec:aniso}
 
@@
 
-The data files containing the scatter cross sections stor the information using Legendre polynomial expansion coefficients.
+Typically, at this point, discrete ordinate methods will describe the discrete angular flux, $\psi$ using flux moments, $\phi$.
+The motivation for using flux moments is that they improve computational performance and reduce memory overhead.
+
+The data files containing the scatter cross sections store the information using Legendre polynomial expansion coefficients.
 The use of a Legendre expansion removes the dependence on the quadrature from the data.
@@
 An alternate anisotropy treatment is to use the Klein-Nishina formula~\citep{ref:shultisj}. The Klein-Nishina formula describes the scattering cross section as
-\begin{equation}
+\begin{equation}\label{eq:klein}
 \sigma(E, \mu) \propto q^3 + q\mu^2
@@
 \begin{equation}
-q(E, \mu) = \frac{1}{1 + (E/E_0)(1-\mu)}
+q(E, \mu) = \frac{1}{1 + (E/E_0)(1-\mu)}.
 \end{equation}
+Figure~\ref{fig:klein} shows a comparison of the Klein-Nishina scattering cross section as a function of scatter angle versus the computed cross sections in DOCTORS using the Klein-Nishina formula given in Eq.~\ref{eq:klein}.
+The slight discrepancy between the two is due to the postprocessing used to merge data from the DOCTORS output.
+The underlying trends are identical.
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/klein}
+  \end{center}
+  \caption{Comparison of the Klein-Nishina formula to the computed cross section values in DOCTORS using the Klein-Nishina anisotropy treatment.}
+\label{fig:klein}
+\end{figure}
 
@@
 
-A rigorous derivation for the uncollided flux is not provided in this work but can be obtained from [CITE].
+A rigorous derivation for the uncollided flux is not provided in this work.
 Instead, an intuitive approach is used.
@@
 
-\section{Source Generation}
+\section{Source Generation}\label{sec:source_gen}
 All photon transport problems considered in the DOCTORS code are ultimately driven by an external source, namely the medical equipment producing the photon beam.
@@
 
-\subsection{Point Source}
+\subsection{Point source}
 The simplest source implemented is an isotropic point source.
@@
 
-\subsection{Fan Source}
+\subsection{Fan source}
 In addition to the values defining a point source (position $\boldsymbol{s}$ and the energy distribution) a fan source is defined by two additional parameters, $\varphi$ and $\theta$ which describe the azimuthal and polar angles subtended by the fan beam respectively.
@@
 
-\subsection{Multi-fan Source}
+\subsection{Multi-fan source}
 The focal point about which the CT system rotates is assumed to be the center of the $xy$ plane and at the same $z$-elevation as the first user specified point.
@@
 
-\subsection{Cone Source and Multi-cone Source}
+\subsection{Cone source and multi-cone source}
 Cone sources are identical to fan beam sources except for the rejection criteria used to determine whether or not a voxel is inside the beam.
@@
 
-\subsection{Fluence-to-dose Conversion}
+\subsection{Fluence-to-dose conversion}
 
@@
 
-\subsection{Energy Deposition}
+\subsection{Energy deposition}
 

### 2017-07-05 docprep/chapresults.tex
Fixed typos and updated LaTeX template

@@
 
-\subsection{Artifact Removal}
+\subsection{Artifact removal}
 At the periphery of the computational mesh, artifacts can appear as a byproduct of the reconstruction algorithm used.
-Figure~\ref{fig:phantomWaterUnmodified} shows the artifacts in question in a $sy$ slice at the lowest ($z = 0$) level in the phantom.
+Figure~\ref{fig:phantomWaterUnmodified} shows the artifacts in question in a $xy$ slice at the lowest ($z = 0$) level in the phantom.
 Two hundred sixty one (261) artifacts were detected in the phantom.
@@
 The water phantom in Fig.~\ref{fig:phantomWaterModified} can be broken into four regions: the water phantom, the container, the air, and the corner artifacts.
-The water phantom is the centermost region of the slice
+The water phantom is the centermost region of the slice.
 
@@
 
-\subsection{Geometry Simplification}
+\subsection{Geometry simplification}
 The original phantom data for both the water phantom and the liver phantom are $256\times256\times64$ meshes which result in 4,194,304 total voxels.
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/phantomwater_64}
+        \includegraphics[width=\textwidth]{figs/phantomwater_256}
         \caption{}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/phantomwater_256}
+        \includegraphics[width=\textwidth]{figs/phantomwater_64}
         \caption{}
@@
 The logarithmic scaled histogram shows the low frequency voxels between -800 and -100 the are difficult to identify as water or air.
-These voxels are around the periphery of
+These voxels are around the periphery of the container.
 
@@
 
-\subsection{Diagnostic Pulse}
+\subsection{Diagnostic pulse}
 Once the 19 and 47 group datasets were compared using the uniform energy distribution, a second benchmark was created.
@@
 This causes a buildup when the beam enters the phantom and a rapid decrease as more particles scatter than predicted by MCNP6.
+Figure~\ref{fig:b2_aniso} shows lineouts for the in-group scatter with the uncollided flux (group~41) and the purely downscattered flux in group~42.
+Figure~\ref{fig:b2_aniso} shows lineouts for the in-group scatter with the uncollided flux (group~41) and the purely downscattered flux in group~42.
 
@@
 
-Figure~\ref{fig:b2_aniso} shows lineouts for the in-group scatter with the uncollided flux (group~41) and the purely downscattered flux in group~42.
 \begin{figure}
@@
     \end{subfigure}
-    \caption{Linout showing the anisotropy. (a) Group 41 (70-75 keV). (b) Group 42 (60-70 keV). }\label{fig:b2_aniso}
+    \caption{Lineout showing the effect of anisotropy. (a) Group 41 (70-75 keV). (b) Group 42 (60-70 keV). }\label{fig:b2_aniso}
 \end{figure}
 
-Figure~\ref{fig:b2_aniso} shows lineouts for the in-group scatter with the uncollided flux (group~41) and the purely downscattered flux in group~42.
 \begin{figure}
@@
     \end{subfigure}
-    \caption{Linout showing the effect of quadrature. (a) Group 41 (70-75 keV). (b) Group 42 (60-70 keV). }\label{fig:b2_quad}
+    \caption{Lineout showing the effect of quadrature. (a) Group 41 (70-75 keV). (b) Group 42 (60-70 keV). }\label{fig:b2_quad}
 \end{figure}
 
-In MCNP6, the effective dose was computed using a FMESH4 tally with associated DE and DF cards to weight the flux.
+In MCNP6, the effective dose was computed using a FMESH tally with associated DE and DF cards to weight the flux.
 The most accurate tally to compute the absorbed dose with is a F6 tally, however, those cannot be spatially distributed via a FMESH tally.
@@
 
-This benchmark is design to test the multi-cone beam projection.
+This benchmark is designed to test the multi-cone beam projection.
 Sixteen cone beams are simultaneously used to generate the data.
 Each cone beam fires the 1-group diagnostic pulse energy spectra described in Sec.~\ref{sec:cone1}.
+Overall, the raytracer shows excellent agreement but the collided flux has more error than expected.
+
+The uncollided flux, which should be nearly identical between MCNP6 and DOCTORS is shown in Fig.~\ref{fig:b3_unc}.
+The relative error between the two and a vertical lineout through the center are shown in Fig.~\ref{fig:b3_comp}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_unc_ray}
+        \caption{}
+        \label{fig:b3_dr_unc}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_unc_mc}
+        \caption{}
+        \label{fig:b3_mc_unc}
+    \end{subfigure}
+    \caption{Comparison of the uncollided flux in (a) DOCTORS and (b) MCNP6 for the 16 cone-beam water benchmark problem.}\label{fig:b3_unc}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_unc_comp}
+        \caption{}
+        \label{fig:b3_unc_comp}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_unc_lineout}
+        \caption{}
+        \label{fig:b3_unc_lineout}
+    \end{subfigure}
+    \caption{Comparison of the uncollided flux in (a) DOCTORS and (b) MCNP6 for the 16 cone-beam water benchmark problem.}\label{fig:b3_comp}
+\end{figure}
+
+The collided flux is shown in Fig.~\ref{fig:b3_col}.
+The MCNP6 collided flux is much more uniform across the patient.
+The uniformity is seen in both the side-by-side comparison of the flux distributions shown in Fig.~\ref{fig:b3_col} and the vertical lineout shown in Fig.~\ref{fig:b3_col_comp_g41}.
+DOCTORS overestimates the flux at the periphery and underestimates the flux at the center of the patient.
+The trend is effectively an overestimation of the attenuation through the patient by DOCTORS.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_col_s6kg41}
+        \caption{}
+        \label{fig:b3_dr_col}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_col_mc_g41}
+        \caption{}
+        \label{fig:b3_mc_col}
+    \end{subfigure}
+    \caption{Comparison of the collided flux in (a) DOCTORS and (b) MCNP6 for the 16 cone-beam water benchmark problem.}\label{fig:b3_col}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/b3_col_comp_g41}
+  \end{center}
+  \caption{Vertical lineout of the collided flux for the 16 cone-beam water benchmark problem.}
+\label{fig:b3_col_comp_g41}
+\end{figure}
+
+Knowing that the flux profiles vary, the dose profiles can also be compared; a similar discrepancy appears in the dose profile generated from both dose computation methodologies used.
+Qualitatively, the dose profiles in both MCNP6 and DOCTORS are very similar, though the DOCTORS dose is overestimated by roughly a factor of two.
+This is believed to indicate that the scatter distribution at lower energies is being overestimated.
+This overestimation is in turn caused by the angular treatment of the transport.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_edep_lineout}
+        \caption{}
+        \label{fig:b3_edep_lineout}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b3_icrp_lineout}
+        \caption{}
+        \label{fig:b3_icrp_lineout}
+    \end{subfigure}
+    \caption{Comparison of the dose to the water phantom using (a) the energy deposition model and (b) the ICRP fluence-to-dose conversion model.}\label{fig:b3_dose}
+\end{figure}
+
+The angular transport is affected by two parameteres: the quadrature and the anisotropy treatment.
+Since the anisotropy treatment was explored in the previous benchmark and no significant changes were found be increasing the $P_N$ expansion or changing the anisotropy methodology altogether, the problem likely lies with the quadrature.
+Higher quadratures may result in significantly improved flux distributions since lower quadratures will cause particles to preferentially leave the beam if insufficient angular directions are within the beam.
+This then motivates the final benchmark problem.
 
@@
 
-INSERT THE RESULTS FROM THE 16 BEAM PHANTOM BENCHMARK
-
-\subsection{Patient Table}
-In some phantom models, a table the patient lies on is present.
-The gantry poses a challenge since it is a highly attenuating material relative to the patient which reduces the available flux.
-Further, the gantry is not mentioned in the literature which only consider the phantom.
-The CT number to material conversion developed by~\citet{ref:ottossonr} which is implemented in this work does not have an equivalent for the gantry region.
-
-All voxels whose CT number is above the highest bone value given by~\citet{ref:ottossonr} are assumed to be part of the gantry. The gantry is assumed to be aluminum with a density of 2.70 g/cm$^{3}$.
-
-Since the triangular artifacts were still seen in the 16-beam water phantom, the number of beams used in this benchmark was increased to 64.
+The final benchmark problem uses 64 cone beams surrounding a full patient phantom.
+Since the triangular artifacts were still seen in the 16-beam water phantom, the number of beams used in this benchmark was increased to 64 cone beam projections.
 Also, each beam was moved further from the isocenter of the phantom.
@@
 The cone beam was broadened to 30 degrees to more closely match the CT data.
+The energy distribution was not changed.
+This benchmark problem is much more realistic and provides a more distributed uncollided source which is expected to help mitigate some of the problems identified in the water benchmarks.
+
+In some phantom models, a table the patient lies on is present.
+The table poses a challenge since it is a highly attenuating material relative to the patient which reduces the available flux.
+Further, the table is not mentioned in the literature which only consider the phantom.
+The CT number to material conversion developed by~\citet{ref:ottossonr} which is implemented in this work does not have an equivalent for the table region.
+
+All voxels whose CT number is above the highest bone value given by~\citet{ref:ottossonr} are assumed to be part of the table. The table is assumed to be aluminum with a density of 2.70 g/cm$^{3}$.
 
@@
 
+As with the previous benchmark problem, the raytracer provides very good agreement (well within 1-2\% for most voxels) between MCNP6 and DOCTORS.
+Because the "corner artifacts" cannot be removed in the full phantom model, those artifacts show up in the flux profile.
+Since those regions are only air, the artifacts' impact on the overall results are expected to be negligible.
+Otherwise, the uncollided has no apparent artifacts such as the triangle artifacts in the previous benchmark problem.
+Figure~\ref{fig:b4_unc_diffs} shows that the profile between the two are effectively identical.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b4_unc_dr}
+        \caption{}
+        \label{fig:b4_unc_dr}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b4_unc_mc}
+        \caption{}
+        \label{fig:b4_unc_mc}
+    \end{subfigure}
+    \caption{Comparison of the uncollided flux in (a) DOCTORS and (b) MCNP6 for the 64 cone-beam phantom benchmark problem.}\label{fig:b4_unc}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b4_unc_diff}
+        \caption{}
+        \label{fig:b4_unc_diff}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b4_unc_lineout}
+        \caption{}
+        \label{fig:b4_unc_diff_lineout}
+    \end{subfigure}
+    \caption{The (a) relative difference between the MCNP6 and DOCTORS uncollided distribution and a (b) vertical lineout down the center.}\label{fig:b4_unc_diffs}
+\end{figure}
+
+Similar to the water phantom benchmarks, DOCTORS underestimates the dose on the interior of the patient.
+Again, this may be due to either the anisotropy treatment or the quadrature, further investigation will refeal the culprit.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b4_col_s6kg47}
+        \caption{}
+        \label{fig:b4_col_s6kg47}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/b4_col_mc_g47}
+        \caption{}
+        \label{fig:b4_col_mc_g47}
+    \end{subfigure}
+    \caption{Comparison of the uncollided flux in (a) DOCTORS and (b) MCNP6 for the 64 cone-beam phantom benchmark problem.}\label{fig:b4_col}
+\end{figure}
+
+\begin{figure}[tb]
+  \begin{center}
+   \includegraphics[width=3.75in]{figs/b4_col_comp}
+  \end{center}
+  \caption{Comparison of the collided flux for the 64 cone-beam phantom benchmark problem.}
+\label{fig:b4_col_comp}
+\end{figure}
+
 \section{GPU Acceleration}
@@
 Table~\ref{tab:cpu} summarizes the runtime required for the CPU-only version of DOCTORS and Table~\ref{tab:gpu} summarizes the analogous runtime results for the GPU accelerated version.
+The speedup of the GPU over the CPU is given in Table~\ref{tab:speedup}.
+From those tables, a number of conclusions can be drawn.
+
+The first, and most obvious, conclusion is that the speedup for large problems is much greater than the speedup for small problems.
+The GPU failed to accelerate the $64 \times 64 \times 16$ problems by more than a factor of a few which hardly warrants any acceleration at all, instead further CPU optimization would likely benefit more.
+Larger problems showed significant speedups though due to the reduced proportion of time spent in overhead operations.
+The most time consuming part of the GPU computation is launching a new kernel.
+Each kernel is itself very simple and can execute rapidly so small problems that have kernels with fewer parallel tasks do not perform as well.
+
+The second, and more surprising, conclusion is that the speedup varied very little for single and double precision floating arithmetic.
+In fact, in most cases, the double precision outperformed the single precision!
+This result is unexpected in general, but even more so on a GPU where single precision is vastly superior.
+The higher precision allows the double precision computations to converge more rapidly reducing the total number of iteraions necessary.
+Since a large portion of the overall time spent is in overhead operations, the time savings of using single precision are actually outweighed by the time  savings of reducing the convergence iterations.
 
@@
 \hline
-\texttt{float}  & Raytracer & 824.9   &  \\
-                & Solver    & 15,541  &  \\
-\texttt{double} & Raytracer & 933.6   & 179,613 \\
-                & Solver    & 14,348   & 2,408,000 \\  [1ex]
+\texttt{float}  & Raytracer & 820   &  125,000 \\
+                & Solver    & 15,500  &  2,700,000 \\
+\texttt{double} & Raytracer & 930   & 180,000 \\
+                & Solver    & 14,300   & 2,400,000 \\  [1ex]
 \hline
@@
 \hline
-\texttt{float}  & Raytracer & 362.8   &  10,267 \\
-                & Solver    & 2,417.9  &  71,619 \\
-\texttt{double} & Raytracer & 362.6   & 13,452 \\
-                & Solver    & 4,554.5   & 61,355 \\  [1ex]
+\texttt{float}  & Raytracer & 360   &  10,300 \\
+                & Solver    & 2,420  &  71,600 \\
+\texttt{double} & Raytracer & 363   & 13,500 \\
+                & Solver    & 4,500   & 61,400 \\  [1ex]
 \hline
 \end{tabular}
-\label{tab:cpu}
+\label{tab:gpu}
+\end{table}
+
+\begin{table}[ht]
+\caption{Speedup}
+\centering 
+\begin{tabular}{l c r r}
+\hline \hline   
+Data Type & Method & $64 \times 64 \times 16$ & $256 \times 256 \times 64$\\ [0.5ex] 
+\hline
+\texttt{float}  & Raytracer & 2.3   & 12.2 \\
+                & Solver    & 6.4  &  37.6 \\
+\texttt{double} & Raytracer & 2.6   & 13.5 \\
+                & Solver    & 3.2   & 39.1 \\  [1ex]
+\hline
+\end{tabular}
+\label{tab:speedup}
 \end{table}

### 2017-07-05 docprep/norris_phd_dissertation.tex
Fixed typos and updated LaTeX template

@@
 \doublespacing
+\usepackage[authoryear,sort]{natbib}
 \usepackage{threeparttable}
@@
 
-\ThesisAdviser{Xin Liu}
+\ThesisAdvisor{Xin Liu}
 
@@
 
-\memberone{Ayodeji Alajo}
-\membertwo{Hyoung Koo Lee}
-\memberthree{Gary Mueller}
-\memberfour{Fikret Ercal}
+%\memberone{Ayodeji Alajo}
+%\membertwo{Hyoung Koo Lee}
+%\memberthree{Gary Mueller}
+%\memberfour{Fikret Ercal}
+
+\ThesisCommittee{Ayodeji Alajo\\ Hyoung Koo Lee\\ Gary Mueller \\ Fikret Ercal}
 %\memberfive{Prandtl}
@@
 
-A GPU algorithm was implemented that speeds up the DOCTORS code by a factor of 39.
+A GPU algorithm was implemented that speeds up the DOCTORS code by a factor of up to nearly 40 for large problems.
+GPU acceleration was found to benefit smaller problems much less.
+Speedup was seen in both single precision and  double precision problems.
 
@@
 Liu and all of the other graduate students for their continued support through my graduate career.
-I also appreciate the time spent by my committee members helping me identify problems and provide insight to solutions.
+I appreciate the time they spent helping me identifying problems and providing insight to solutions.
 I would also like to thank my committee members for their guidance throughout my PhD work.
@@
 
-\ThesisAppendixChapter{Code Listings}\input{appdx_b}
+%\ThesisAppendixChapter{Code Listings}\input{appdx_b}
 
-\ThesisAppendixChapter{Benchmark Results}\input{appdx_c}
+%\ThesisAppendixChapter{Benchmark Results}\input{appdx_c}
 
@@
 \begin{ThesisBibliography}{BIBLIOGRAPHY}
-\bibliographystyle{plainnat}
+\bibliographystyle{unsrtnat}
+ %\bibliographystyle{ieeetr}
+ %\bibliographystyle{plainnat}
 \bibliography{refs}
